### 2017-03-08 portal-forecast.Rmd

---
title: "Portal Forecast"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Total Abundance Forecast

This is the forecast for next month's sampling of rodents at Portal.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
source('forecast-viz.R', echo = FALSE)
obs = download_observations()
source('https://raw.githubusercontent.com/weecology/PortalDataSummaries/master/RodentAbundances.R', echo = FALSE)
obs_data = abundance()
obs_data$total = rowSums(select(obs_data, -period))
new_moons = read.csv('~/PortalData/Rodents/moon_dates.csv')
obs_data_newmoon = inner_join(obs_data, new_moons, by = c("period" = "Period"))
for_data = read.csv("predictions/2016-8-31Allforecasts.csv")

forecast_viz(obs_data = obs_data_newmoon,
             obs_date_col_name = "NewMoonNumber",
             obs_val_col_name = "total",
             for_data = for_data,
             for_date_col_name = "NewMoonNumber",
             for_val_col_name = "estimate",
             for_model_name = "NegBinom Time Series",
             for_lowerpi_col_name = "LowerPI",
             for_upperpi_col_name = "UpperPI",
             start_newmoon = 300)
```

## Species-Level Forecasts

```{r speciesforecasts, echo=FALSE, message=FALSE, warning=FALSE}
source("forecast-viz.R")
source("forecast_tools.R")
source("actual_census_counts.R")
source("https://raw.githubusercontent.com/weecology/PortalDataSummaries/master/RodentAbundances.R")

data = compile_forecasts()
ensemble = make_ensemble(data)
plot_sp_predicts(ensemble, 'All', lead_time = 1)
```

### 2017-03-15 portal-forecast.Rmd
Add most abundant species time-series forecasts

@@
 knitr::opts_chunk$set(echo = TRUE)
+library(dplyr)
 ```
@@
 obs_data_newmoon = inner_join(obs_data, new_moons, by = c("period" = "Period"))
-for_data = read.csv("predictions/2016-8-31Allforecasts.csv")
+for_data = read.csv("predictions/2017-02-24Allforecasts.csv")
 
@@
 ensemble = make_ensemble(data)
-plot_sp_predicts(ensemble, 'All', lead_time = 1)
+sp_predictions = get_sp_predicts(ensemble, 'All', lead_time = 1)
+plot_data(sp_predictions, 'All')
 ```
+```{r highabund_ts_forecasts, echo=FALSE, message=FALSE, warning=FALSE}
+most_abund_sp = sp_predictions %>% 
+  filter(species != 'total') %>% 
+  arrange(desc(estimate)) %>% 
+  head(3) %>% 
+  select(species)
+
+for (species in as.vector(most_abund_sp$species)) {
+  species_forecast = forecast_viz(obs_data = obs_data_newmoon,
+               obs_date_col_name = "NewMoonNumber",
+               obs_val_col_name = species,
+               for_data = for_data,
+               for_date_col_name = "NewMoonNumber",
+               for_val_col_name = "estimate",
+               for_model_name = "NegBinom Time Series",
+               for_lowerpi_col_name = "LowerPI",
+               for_upperpi_col_name = "UpperPI",
+               start_newmoon = 300)
+  plot(species_forecast)
+}
+```

### 2017-03-29 portal-forecast.Rmd
Use ensembles for website forecasts

@@
 library(dplyr)
+source("forecast_tools.R")
 ```
@@
 obs_data_newmoon = inner_join(obs_data, new_moons, by = c("period" = "Period"))
-for_data = read.csv("predictions/2017-02-24Allforecasts.csv")
+#for_data = read.csv("predictions/2017-02-24Allforecasts.csv")
+
+for_data = compile_forecasts() %>%
+  make_ensemble() %>%
+  dplyr::filter(level == 'All', date == max(date))
+
+most_recent_forecast = max(for_data$date)
+for_data = filter(for_data, date == most_recent_forecast)
+
+for_data$model = "ensemble"
 
@@
              for_val_col_name = "estimate",
-             for_model_name = "NegBinom Time Series",
+             for_model_name = "ensemble",
              for_lowerpi_col_name = "LowerPI",
@@
                for_val_col_name = "estimate",
-               for_model_name = "NegBinom Time Series",
+               for_model_name = "ensemble",
                for_lowerpi_col_name = "LowerPI",

### 2017-04-26 portal-forecast.Rmd

@@
 library(dplyr)
+library(PortalDataSummaries)
 source("forecast_tools.R")
@@
 obs = download_observations()
-source('https://raw.githubusercontent.com/weecology/PortalDataSummaries/master/RodentAbundances.R', echo = FALSE)
-obs_data = abundance()
+obs_data = PortalDataSummaries::abundance()
 obs_data$total = rowSums(select(obs_data, -period))
@@
 source("actual_census_counts.R")
-source("https://raw.githubusercontent.com/weecology/PortalDataSummaries/master/RodentAbundances.R")
 

### 2017-04-26 portal-forecast.Rmd
Automatically update website page when making forecasts

@@
 source('forecast-viz.R', echo = FALSE)
-obs = download_observations()
 obs_data = PortalDataSummaries::abundance()

### 2017-04-26 PortalForecasts.R

@@ -4,6 +4,7 @@ library(lubridate)
 library(dplyr)
 library(magrittr)
 library(testit)
+library(rmarkdown)
 source('forecast_tools.R')
 
 #Period 203/NewMoonNumber 217 will be when the training data timeseries
@@ -236,3 +237,7 @@ forecastall <- function(abundances, level, weather_data, weatherforecast, CI_lev
 ######Run Models########################################################
 allforecasts=forecastall(all,"All",weather_data,weathermeans)
 controlsforecasts=forecastall(controls,"Controls",weather_data,weathermeans)
+
+######Update Website####################################################
+rmarkdown::render('portal-forecast.Rmd', output_format = "html_document",
+                  output_file = 'index.html', output_dir = 'docs')

### 2017-04-26 portal-forecast.Rmd

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-source('forecast-viz.R', echo = FALSE)
 obs_data = PortalDataSummaries::abundance()
@@
 ```{r speciesforecasts, echo=FALSE, message=FALSE, warning=FALSE}
-source("forecast-viz.R")
 source("forecast_tools.R")

### 2017-04-29 portal-forecast.Rmd
Update website

@@
 for_data = compile_forecasts() %>%
-  make_ensemble() %>%
-  dplyr::filter(level == 'All', date == max(date))
+  dplyr::filter(level == 'All', model == 'Ensemble', date == max(date))
 
@@
 data = compile_forecasts()
-ensemble = make_ensemble(data)
+ensemble = dplyr::filter(data, level == 'All', model == 'Ensemble', date == max(date))
 sp_predictions = get_sp_predicts(ensemble, 'All', lead_time = 1)

### 2017-04-29 forecast_tools.R

@@ -286,7 +286,7 @@ compile_forecasts=function(forecast_folder='./predictions', verbose=FALSE){
       }
     }
   }
-
+all_forecasts$date=as.Date(all_forecasts$date)
   return(all_forecasts)
 }
 

### 2017-05-02 portal-forecast.Rmd
Update file paths to use FullPath()

@@
 obs_data$total = rowSums(select(obs_data, -period))
-new_moons = read.csv('~/PortalData/Rodents/moon_dates.csv')
+new_moon_file = PortalDataSummaries::FullPath('PortalData/Rodents/moon_dates.csv', 
+                                                  '~')
+new_moons = read.csv(new_moon_file)
 obs_data_newmoon = inner_join(obs_data, new_moons, by = c("period" = "Period"))

### 2017-05-02 do_hindcasts.R

@@ -7,10 +7,13 @@ library(PortalDataSummaries)
 #' like the sampling period in period_info just happened. 
 #' @param period_info data.frame One row data.frame output from new_moons
 #' @return None
-backdate_observed_data=function(period_info){
-  rodents_filename  = '~/PortalData/Rodents/Portal_rodent.csv'
-  new_moon_filename = '~/PortalData/Rodents/moon_dates.csv'
-  weather_filename  = '~/PortalData/Weather/Portal_weather.csv'
+backdate_observed_data=function(period_info, base_path='~'){
+  rodents_filename  = PortalDataSummaries::FullPath('PortalData/Rodents/Portal_rodent.csv', 
+                                                    base_path)
+  new_moon_filename = PortalDataSummaries::FullPath('PortalData/Rodents/moon_dates.csv', 
+                                                    base_path)
+  weather_filename  = PortalDataSummaries::FullPath('PortalData/Weather/Portal_weather.csv', 
+                                                    base_path)
   
   period_info$CensusDate = as.Date(period_info$CensusDate)
   period_info$month      = lubridate::month(period_info$CensusDate)
@@ -41,8 +44,11 @@ initial_time_periods=381:458
 #Get the latest to obtain the latest new moon info
 PortalDataSummaries::download_observations()
 
+base_path = '~'
+new_moon_file = PortalDataSummaries::FullPath('PortalData/Rodents/moon_dates.csv', 
+                                                  base_path)
 
-new_moons=read.csv('~/PortalData/Rodents/moon_dates.csv')
+new_moons=read.csv(new_moon_file)
 
 #Period number, year, and  month of sampling will be used to backdate the observation data
 initial_times_info = new_moons %>%

### 2017-05-04 portal-forecast.Rmd

@@
 source("forecast_tools.R")
-source("actual_census_counts.R")
 
@@
 sp_predictions = get_sp_predicts(ensemble, 'All', lead_time = 1)
-plot_data(sp_predictions, 'All')
+plot_data(sp_predictions)
 ```

### 2017-06-27 report.Rmd
Add website pages (#83)

---
title: "Report - June 2017"
---

Note the forecast for D. spectabalis this month.

```{r speciesforecasts, echo=FALSE, message=FALSE, warning=FALSE}
source("forecast_tools.R")

data = compile_forecasts()
ensemble = dplyr::filter(data, level == 'All', model == 'Ensemble', date == max(date))
sp_predictions = get_sp_predicts(ensemble, 'All', lead_time = 1)
ds_forecast = forecast_viz(obs_data = obs_data_newmoon,
               obs_date_col_name = "NewMoonNumber",
               obs_val_col_name = 'DS',
               for_data = for_data,
               for_date_col_name = "NewMoonNumber",
               for_val_col_name = "estimate",
               for_model_name = "ensemble",
               for_lowerpi_col_name = "LowerPI",
               for_upperpi_col_name = "UpperPI",
               start_newmoon = 300)
  plot(ds_forecast)

```

There is an increasing change that we will see a spectab in the coming months. Does this mean Stephanie will return?

### 2017-07-26 index.Rmd

--- 2017-05-04 portal-forecast.Rmd
@@
 title: "Portal Forecast"
-output: html_document
 ---
@@
 library(dplyr)
-library(PortalDataSummaries)
+library(portalr)
 source("forecast_tools.R")
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-obs_data = PortalDataSummaries::abundance()
+obs_data = portalr::abundance()
 obs_data$total = rowSums(select(obs_data, -period))
-new_moon_file = PortalDataSummaries::FullPath('PortalData/Rodents/moon_dates.csv', 
+new_moon_file = portalr::FullPath('PortalData/Rodents/moon_dates.csv', 
                                                   '~')

### 2017-08-11 index.Rmd
Change references to column names to match new standard in PortalData

@@
 new_moons = read.csv(new_moon_file)
-obs_data_newmoon = inner_join(obs_data, new_moons, by = c("period" = "Period"))
+obs_data_newmoon = inner_join(obs_data, new_moons, by = c("period" = "period"))
 #for_data = read.csv("predictions/2017-02-24Allforecasts.csv")
@@
 forecast_viz(obs_data = obs_data_newmoon,
-             obs_date_col_name = "NewMoonNumber",
+             obs_date_col_name = "newmoonnumber",
              obs_val_col_name = "total",
              for_data = for_data,
-             for_date_col_name = "NewMoonNumber",
+             for_date_col_name = "newmoonnumber",
              for_val_col_name = "estimate",
@@
   species_forecast = forecast_viz(obs_data = obs_data_newmoon,
-               obs_date_col_name = "NewMoonNumber",
+               obs_date_col_name = "newmoonnumber",
                obs_val_col_name = species,
                for_data = for_data,
-               for_date_col_name = "NewMoonNumber",
+               for_date_col_name = "newmoonnumber",
                for_val_col_name = "estimate",

### 2017-08-11 PortalForecasts.R

@@ -14,7 +14,7 @@ portalr::download_observations()
 moons = get_moon_data()
 
 #Beginning and end of the forecast timeperiod
-most_recent_newmoon = moons$NewMoonNumber[which.max(moons$Period)]
+most_recent_newmoon = moons$newmoonnumber[which.max(moons$period)]
 first_forecast_newmoon=most_recent_newmoon+1
 last_forecast_newmoon=first_forecast_newmoon + 11
 forecast_newmoons = first_forecast_newmoon:last_forecast_newmoon
@@ -26,7 +26,7 @@ weather_data = get_weather_data(moons, rodent_data$all, first_forecast_newmoon, 
 
 #Get only relevent columns now that this is isn't needed to subset weather.
 rodent_data$all = rodent_data$all %>%
-  select(-NewMoonDate,-CensusDate,-Period,-Year,-Month)
+  select(-newmoondate,-censusdate,-period,-year,-month)
 
 #tscount::tsglm() will not model a timeseries of all 0's. So for those species, which are
 #ones that just haven't been observed in a while, make a forecast of all 0's.

### 2017-08-11 forecast_tools.R

@@ -51,7 +51,7 @@ make_ensemble=function(all_forecasts, models_to_use=NA, CI_level = 0.9){
   weighted_estimates = all_forecasts %>%
     mutate(model_var = ((UpperPI - estimate)/CI_quantile)^2) %>%
     left_join(weights, by=c('date','model','currency','level','species','fit_start_newmoon','fit_end_newmoon','initial_newmoon')) %>%
-    group_by(date, NewMoonNumber, forecastmonth, forecastyear,level, currency, species, fit_start_newmoon, fit_end_newmoon, initial_newmoon) %>%
+    group_by(date, newmoonnumber, forecastmonth, forecastyear,level, currency, species, fit_start_newmoon, fit_end_newmoon, initial_newmoon) %>%
     summarise(ensemble_estimate = sum(estimate*weight), 
               weighted_ss = sum(weight * (estimate - ensemble_estimate)^2) ,
               ensemble_var   = sum(model_var * weight) + weighted_ss / (n()*sum(weight)-1),
@@ -78,17 +78,17 @@ get_sp_predicts = function(data, lvl, lead_time) {
                                                             ""), format = "%m/%Y")) %>% transform(date = as.Date(date, "%Y-%m-%d"))
   data1 = filter(data, level == lvl,
                  date == max(as.Date(date)))
-  target_moon = min(data1$NewMoonNumber) + (lead_time - 1)
-  data2 = filter(data1, NewMoonNumber == target_moon)
+  target_moon = min(data1$newmoonnumber) + (lead_time - 1)
+  data2 = filter(data1, newmoonnumber == target_moon)
 }
 
 plot_data = function(data) {
   newmoons_table = read.csv(
     text = getURL(
       "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
-  target_moon=unique(data$NewMoonNumber)
-  period_code = dplyr::filter(newmoons_table, newmoons_table$NewMoonNumber == target_moon) %>%
-    dplyr::select(Period) %>%
+  target_moon=unique(data$newmoonnumber)
+  period_code = dplyr::filter(newmoons_table, newmoons_table$newmoonnumber == target_moon) %>%
+    dplyr::select(period) %>%
     as.integer()
   sp_predict = ggplot(data,
                       aes(
@@ -138,7 +138,7 @@ calculate_forecast_error = function(observations, forecasts, error_metric='MSE',
 
   if(!forecast_is_valid(forecasts)) stop('Forecast dataframe not valid')
 
[…]

### 2017-08-11 models/naive02.R

@@ -5,7 +5,7 @@ naive02=function(abundances,forecast_date,forecast_months,forecast_years,forecas
 
 model=forecast(auto.arima(abundances$total,lambda = 0),h=num_forecast_months,level=CI_level,fan=T)
 
-forecasts02=data.frame(date=forecast_date, forecastmonth=forecast_months, forecastyear=forecast_years, NewMoonNumber=forecast_newmoons,
+forecasts02=data.frame(date=forecast_date, forecastmonth=forecast_months, forecastyear=forecast_years, newmoonnumber=forecast_newmoons,
                        currency="abundance", model="AutoArima", level=level, species="total", estimate=model$mean,
                        LowerPI=model$lower[,which(model$level==CI_level*100)], UpperPI=model$upper[,which(model$level==CI_level*100)])
 forecasts02[sapply(forecasts02, is.ts)] <- lapply(forecasts02[sapply(forecasts02, is.ts)],unclass)

### 2017-08-11 models/neg_binom_ts.R

@@ -25,7 +25,7 @@ neg_binom_ts=function(abundances,forecast_date,forecast_months,forecast_years,fo
       model_aic = tryCatch(summary(model)$AIC, error = function(x) {1e6})
     }
     newpred=data.frame(date=rep(forecast_date,num_forecast_months), forecastmonth=forecast_months, forecastyear=forecast_years,
-                       NewMoonNumber=forecast_newmoons, currency="abundance", model=rep("NegBinom Time Series",num_forecast_months),
+                       newmoonnumber=forecast_newmoons, currency="abundance", model=rep("NegBinom Time Series",num_forecast_months),
                        level=level, species=rep(s,num_forecast_months), estimate=pred$pred,
                        LowerPI=pred$interval[,1], UpperPI=pred$interval[,2])
     allforecasts=rbind(allforecasts,newpred)

### 2017-08-11 models/pois_env_ts.R

@@ -10,15 +10,15 @@ pois_env_ts=function(abundances,weather_data,weathermeans,forecast_date,forecast
   allaic=data.frame()
 
 #List of candiate environmental covariate models
-model_covariates = list(c('MaxTemp','MeanTemp','Precipitation','NDVI'),
-                        c('MaxTemp','MinTemp','Precipitation','NDVI'),
-                        c('MinTemp','MaxTemp','MeanTemp','Precipitation'),
-                        c('Precipitation','NDVI'),
-                        c('MinTemp','NDVI'),
-                        c('MinTemp'),
-                        c('MaxTemp'),
-                        c('MeanTemp'),
-                        c('Precipitation'),
+model_covariates = list(c('maxtemp','meantemp','precipitation','NDVI'),
+                        c('maxtemp','mintemp','precipitation','NDVI'),
+                        c('mintemp','maxtemp','meantemp','precipitation'),
+                        c('precipitation','NDVI'),
+                        c('mintemp','NDVI'),
+                        c('mintemp'),
+                        c('maxtemp'),
+                        c('meantemp'),
+                        c('precipitation'),
                         c('NDVI'))
 
 for(s in species) {
@@ -58,7 +58,7 @@ for(s in species) {
     }
   }
   newpred = data.frame(date=rep(forecast_date,num_forecast_months), forecastmonth=forecast_months, forecastyear=forecast_years,
-                       NewMoonNumber=forecast_newmoons, currency="abundance", model=rep("Poisson Env",num_forecast_months),
+                       newmoonnumber=forecast_newmoons, currency="abundance", model=rep("Poisson Env",num_forecast_months),
                        level=level, species=rep(s,num_forecast_months), estimate=pred$pred,
                        LowerPI=pred$interval[,1],UpperPI=pred$interval[,2])
   allforecasts = rbind(allforecasts,newpred)

### 2017-08-24 report.Rmd

@@
 ds_forecast = forecast_viz(obs_data = obs_data_newmoon,
-               obs_date_col_name = "NewMoonNumber",
+               obs_date_col_name = "newmoonnumber",
                obs_val_col_name = 'DS',
                for_data = for_data,
-               for_date_col_name = "NewMoonNumber",
+               for_date_col_name = "newmoonnumber",
                for_val_col_name = "estimate",

### 2017-09-06 index.Rmd
change xaxis on prediction plots from new moon numbers to dates

@@
 obs_data_newmoon = inner_join(obs_data, new_moons, by = c("period" = "period"))
+obs_data_newmoon$censusdate = as.Date(obs_data_newmoon$censusdate)
 #for_data = read.csv("predictions/2017-02-24Allforecasts.csv")
@@
 for_data$model = "ensemble"
+for_data$forecastdate = as.Date(paste(for_data$forecastyear,for_data$forecastmonth,'15',sep='-'))
 
 forecast_viz(obs_data = obs_data_newmoon,
-             obs_date_col_name = "newmoonnumber",
+             obs_date_col_name = "censusdate",
              obs_val_col_name = "total",
              for_data = for_data,
-             for_date_col_name = "newmoonnumber",
+             for_date_col_name = "forecastdate",
              for_val_col_name = "estimate",

### 2017-09-20 index.Rmd

@@
   species_forecast = forecast_viz(obs_data = obs_data_newmoon,
-               obs_date_col_name = "newmoonnumber",
+               obs_date_col_name = "censusdate",
                obs_val_col_name = species,
                for_data = for_data,
-               for_date_col_name = "newmoonnumber",
+               for_date_col_name = "forecastdate",
                for_val_col_name = "estimate",

### 2017-09-20 profiles.Rmd
move pictures around and attempt to add rodent picture column to table

---
title: "Our Rodent Species"
---
<br>

```{r echo=FALSE, message=FALSE, warning=FALSE}

text_tbl <- data.frame(
  Rodents = c("![]('rodent_pictures/color_pictures/BA.png')",
              "![]('rodent_pictures/color_pictures/PB.png')",
              "![]('rodent_pictures/color_pictures/PH.png')",
              "![]('rodent_pictures/color_pictures/PI.png')",
              "![]('rodent_pictures/color_pictures/PP.png')",
              "![]('rodent_pictures/color_pictures/DM.png')",
              "![]('rodent_pictures/color_pictures/DO.png')",
              "![]('rodent_pictures/color_pictures/DS.png')",
              "![]('rodent_pictures/Rodent_silhouettes/5-White-throatedwoodrat.png')",
              "![]('rodent_pictures/Rodent_silhouettes/6-NorthernGrasshoppermouse.png')",
              "![]('rodent_pictures/Rodent_silhouettes/7-SouthernGrasshoppermouse.png')",
              "![]('rodent_pictures/color_pictures/PF.png')",
              "![]('rodent_pictures/color_pictures/PE.png')",
              "![]('rodent_pictures/color_pictures/PL.png')",
              "![]('rodent_pictures/color_pictures/PM.png')",
              "![]('rodent_pictures/Rodents_silhouettes/16-Fulvouseharvestmouse.png')",
              "![]('rodent_pictures/Rodents_silhouettes/17-WesternHarvestmouse.png')",
              "![]('rodent_pictures/Rodents_silhouettes/18-Plainsharvestmouse.png')",
              "![]('rodent_pictures/Rodents_silhouettes/19-Tawny-belliedcottonrat.png')",
              "![]('rodent_pictures/Rodents_silhouettes/20-HispidCottonRat.png')",
              "![]('rodent_pictures/Rodents_silhouettes/21-Yellownosedcottonrat.png)"), 
  Scientific.Name = c("_Baiomys taylori_", 
                      "_Chaetodipus baileyi_", 
                      "_Chaetodipus hispidus_",
                      "_Chaetodipus intermedius_",
                      "_Chaetodipus penicillatus_",
                      "_Dipodomys merriami_", 
                      "_Dipodomys ordii_", 
                      "_Dipodomys spectabilis_", 
                      "_Neotoma albigula_", 
                      "_Onychomys leucogaster_", 
                      "_Onychomys torridus_",
                      "_Perognathus flavus_",
                      "_Peromyscus eremicus_",
                      "_Peromyscus leucopus_",
                      "_Peromyscus maniculatus_",
                      "_Reithrodontomys fulvescens_",
                      "_Reithrodontomys megalotis_",
                      "_Reithrodontomys montanus_",
                      "_Sigmodon fulviventer_",
                      "_Sigmodon hispidus_",
                      "_Sigmodon ochorognathus_"),
  Common.Name = c("northern pygmy mouse", 
                  "Bailey's pocket mouse",
                  "hispid pocket mouse",
                  "rock pocket mouse",
                  "desert pocket mouse",
                  "Merriam's kangaroo rat", 
                  "Ord's kangaroo rat", 
                  "banner-tailed kangaroo rat", 
                  "white-throated woodrat", 
                  "northern grasshopper/scorpion mouse", 
                  "southern grasshopper/scorpion mouse",
                  "silky pocket mouse",
                  "cactus mouse",
                  "white-footed mouse",
                  "deer mouse",
                  "fulvous harvest mouse",
                  "western harvest mouse",
                  "plains harvest mouse",
                  "tawny-bellied cotton rat",
                  "white-bellied cotton rat",
                  "yellow-nosed cotton rat"),
  Description = c(
  "Yes, apparently southeastern Arizona counts as 'northern' for this genus. One of two competitors for title of 'smallest Portal rodent,' _B. taylori_ is currently an infrequent but welcome visitor to the site. As it is primarily a grassland species, individuals caught on the site rarely stay for long. Interestingly, the roughly 8-gram _B. taylori_ is often found in association with cotton rats (_Sigmodon_), which can weight in excess of 100 grams.", 
  "Bailey's pocket mouse made its first appearance on the site in the mid-1990s. It is essentially a giant version of its congeneric relative, the desert pocket mouse (_C. penicillatus_). While close in size to Merriam's and Ord's kangaroo rats, it is an inferior competitor and, therefore, is found almost solely on kangaroo rat exclosure plots.",
  "The hispid pocket mouse looks like a cactus mouse in pocket mouse form; it's the size of _C. baileyi_ with a beautiful lateral orange stripe. _C. hispidus_ has historically been the least frequently recorded species at the site. In the past few months, however, we've caught one or two each trapping period. Nevertheless, it remains a very transient speices; each individual caught this year has been caught once and then never seen again.",
  "_C. intermedius_ is another pocket mouse, though very rarely found at the site as it prefers rockier habitats; it is very similar to _C. penicillatus_ except in habitat preference.",
  "Although _C. penicillatus_ had been found at the site since the beginning of the study, it didn't become such an important player in the system until the 2000s. Now, this small pocket mouse is one of the main seasonal drivers of total rodent abundance at the site. The population explodes in the summers with known individuals, new adults, and brand new babies alike; they are ubiquitous across the site, regardless of plot type. In the winters, _C. penicillatus_ enters a type of torpor (not quite hibernation), creating a large seasonal fluctuation in abundance.  ",
  "Kangaroo rats, with their giant hind feet and namesake bipedal movement, are some of the most beloved species at the site. Of the three species in the _Dipodomys_ genus, _D. merriami_ is currently the most abundant and, presumably, dominant in the system.",
  "Ord's kangaroo rat is nearly indistinguishable from Merriam's kangaroo rat except for a tiny fifth toe on its hind feet that the Merriams lack. _D. ordii_ can be thought of as _D. merriami's_ kid brother who's always aspiring to be like his brother but can't quite keep up; a grass-loving species, _D. ordii_ is often just trailing _D. merriami_ in abundance.",
  "The largest of the three kangaroo rats at the site, _D. spectabilis_ (fondly referred to as 'spectabs' by those in the know) is known for its striking tail. It has by far the largest feet on the site at a whopping 48 mm average and often weighs in at over 100 grams, nearly twice the size of the other kangaroo rats. Spectabs were once running the show at Portal. As the site became shrubbier, however, the reign of _D. spectabilis_ came to a slow end in the 1990's. Since then, a few individuals have popped up here and there but haven't stuck around, often heading for grassier pastures.",
  "Along with being by far the largest rodent species at the site (my, what big teeth you have...), _Neotoma albigula_ also has a big reputation---roughly 200 grams worth. Woodrats are quite different ecologically from the rest of the species at the site. Unlike the majority of our rodents which are granivorous, _N. albigula_ is primarily a foliovore. (The resident woodrat in our ramada, however, seems to eat just about anything.) Woodrats---also commonly known as packrats---build middens, or large nests, out of sticks and any 'interesting' debris they can get their paws on. While their numbers at the site aren't prolific, research assistants know exactly which kangaroo rat exclosure plots this species tends to inhabit and secretly wish they'd brought leather gloves with them.",
  "Amongst smammal enthusiasts such as ourselves, the carnivorous grasshopper mice (_Onychomys_ genus) are legendary. Their name(s) and [this video](https://www.youtube.com/watch?v=ohd_mSIWTXk) basically say it all. In addition to eating insects and scorpions, they have been known to eat venomous centipedes and even other mice. Both species also have a distinctive high-pitched 'howl' when defending their territory. Mighty though they are, both species, and _O. leucogaster_ in particular, have hilariously short tails. For unknown reasons, the northern grasshopper mouse is less common at the site than the southern.",
  "Aside from overall range, the southern species differs little from the northern (_O. leucogaster_) except for a slightly longer tail and somewhat more elongated appearance. This species is more abundant at the site and can be found in all plot types.",
  "_Perognathus flavus_ rivals only _Baiomys taylorii_ for the title of smallest rodent at our site, weighing between 5 and 10 grams. Tiny, beautiful, and immensely soft, this aptly named silky pocket mouse is a favorite of previous and current research assistants alike. Its relatively low abundance and recent rarity makes it an especially exciting catch.",
  "Often having a bold lateral orange stripe, the cactus mouse is one of the most colorful and striking mice at Portal. It is also the most commonly found _Peromyscus_ species at the site, usually found in kangaroo rat exclosures and incidentally in the full removal plots. Unlike the white-footed mouse and deer mouse, however, it is mainly restricted to the deserts in the American Southwest and northern Mexico.",
  "This _Peromyscus_ species, while very common throughout the majority of the United States, is not all that common at Portal. _Peromyscus_ are good climbers and, therefore, frequently end up in the full rodent exclosure plots.",
  "Like _P. leucopus_, _P. maniculatus_ is nearly ubiquitous throughout the continental United States but is not a particularly common resident at the site. The desert variants of these two species are challenging to distinguish between, adding a fun challenge to trapping.",
  "The fulvous harvest mouse is an infrequent visitor to the site. It is the largest---and the rarest---harvest mouse species found at the site. Like many of the _Reithrodontomys_ caught at the site, it is most often found in kangaroo rat exclosures and full rodent removals.",
  "_R. megalotis_ is the most common harvest mouse found at Portal. Harvest mice abundance tends to remain fairly low throughout the site, though there is sometimes an increase in the winter when the desert pocket mice (_C. penicillatus_) are in torpor.",
  "While very similar to _R. megalotis_, _R. montanus_ is notably smaller than the other harvest mice at the site. It is less commonly found at the site than _R. megalotis_ but plays a similar role ecologically.",
  "A favorite of many Portal research assistants, rodents in the _Sigmodon_ genus are best described as 'squishy like cotton.' Okay, so maybe that's not the most descriptive identifier, but their size and coloration make them an easily identifiable genus. _S. fulviventer_ is one of the two more frequently seen cotton rat species at the site. It is distinguishable by the yellow (rather than white) fur on its belly, as indicated by its descriptive common name. ",
  "_S. hispidus_, with a white rather than a tawny belly, is the other cotton rat species that somewhat frequently has been caught at the site. The _Sigmodons_ seem to come through the site in waves; some years are boom years while others are bust years. What is driving these population cycles? How do _Sigmodons_ move through the regional landscape? We don't know but sure would like to find out.",
  "_S. ochorognathus_ almost seems like an urban (erm...rural) legend nowadays. A few individuals were caught in the 1990s, but none have been seen since."
  )
)

knitr::kable(text_tbl)

```

### 2017-09-20 report.Rmd
added code to plot the most recent data from the repo against the most recent forecast prior to that data collection.  there are still some issues, right now this will only work when June 2017 is the most recent data

@@
+---
+title: "Report - September 2017"
+---
+Here's the most recent data (blue) compared to the forecast performed right before data collection (black, with error bars)
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+source("forecast_tools.R")
+
+# get most recent rodent data
+rodents = abundance("repo", shape='flat')
+period_code = max(rodents$period)
+observed = dplyr::filter(rodents, period == period_code)
+
+# get new moon number associated with this period
+newmoons_table = read.csv(text = getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
+moon_code = dplyr::filter(newmoons_table,period==period_code) %>% select(newmoonnumber) %>% as.integer()
+
+# get most recent forecast prior to data collection -- WIP!!!!!!!! there's an issue up
+data = read.csv('predictions/2017-06-23Allforecasts.csv',na.strings = '')
+data = plyr::rename(data,c('NewMoonNumber' = 'newmoonnumber'))
+
+ensemble = dplyr::filter(data, level == 'All', model == 'Ensemble', newmoonnumber==moon_code)
+sp_predictions = get_sp_predicts(ensemble, 'All', lead_time = 1)
+
+joined_data = left_join(sp_predictions, observed, by = "species")
+joined_data[is.na(joined_data)] = 0
+joined_data[joined_data$species=='total','abundance'] = sum(joined_data$abundance,na.rm=T)
+joined_data[joined_data$species=='total','period'] = period_code
+
+sp_predict = plot_species_forecast(sp_predictions)  
+sp_predict = sp_predict + geom_point(data = joined_data, mapping = aes(x = abundance, y = species),
+             color = "blue")
+plot(sp_predict)
+```
+
 ---

### 2017-09-27 report.Rmd
removed old code from report.Rmd

@@
 source("forecast_tools.R")
+library(portalr)
 
@@
 ```
-
----
-title: "Report - June 2017"
----
-
-Note the forecast for D. spectabalis this month.
-
-```{r speciesforecasts, echo=FALSE, message=FALSE, warning=FALSE}
-source("forecast_tools.R")
-
-data = compile_forecasts()
-ensemble = dplyr::filter(data, level == 'All', model == 'Ensemble', date == max(date))
-sp_predictions = get_sp_predicts(ensemble, 'All', lead_time = 1)
-ds_forecast = forecast_viz(obs_data = obs_data_newmoon,
-               obs_date_col_name = "newmoonnumber",
-               obs_val_col_name = 'DS',
-               for_data = for_data,
-               for_date_col_name = "newmoonnumber",
-               for_val_col_name = "estimate",
-               for_model_name = "ensemble",
-               for_lowerpi_col_name = "LowerPI",
-               for_upperpi_col_name = "UpperPI",
-               start_newmoon = 300)
-  plot(ds_forecast)
-
-```
-
-There is an increasing change that we will see a spectab in the coming months. Does this mean Stephanie will return?

### 2017-09-27 index.Rmd
putting plot of most recent data against forecast on website (#111)
* changed name of plot_data function to plot_species_forecast for clarity

* added code to plot the most recent data from the repo against the most recent forecast prior to that data collection.  there are still some issues, right now this will only work when June 2017 is the most recent data

* removed old code from report.Rmd

@@
 sp_predictions = get_sp_predicts(ensemble, 'All', lead_time = 1)
-plot_data(sp_predictions)
+sp_predict = plot_species_forecast(sp_predictions)
+plot(sp_predict)
 ```

### 2017-09-27 forecast_tools.R

@@ -82,7 +82,13 @@ get_sp_predicts = function(data, lvl, lead_time) {
   data2 = filter(data1, newmoonnumber == target_moon)
 }
 
-plot_data = function(data) {
+#' this is the 'second plot on the 'Species-level Forecast' on the 'Current Forecast' page on the website
+#' 
+#' 
+#' @param data
+#' @return sp_predict is a plot object -- plot(sp_predict) displays it
+#' 
+plot_species_forecast = function(data) {
   newmoons_table = read.csv(
     text = getURL(
       "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
@@ -111,19 +117,10 @@ plot_data = function(data) {
     ylab("Species") +
     xlab("Abundance") +
     scale_y_discrete(breaks = reorder(data$species,data$estimate),labels = reorder(species_names$scientificname,species_names$estimate))
-  if(!is.na(period_code)){
-    rodents = abundance("repo", shape='flat')
-    observed = dplyr::filter(rodents, period == period_code)
-    joined_data = left_join(data, observed, by = "species")
-    joined_data[is.na(joined_data)] = 0
-    joined_data[joined_data$species=='total','abundance'] = sum(joined_data$abundance,na.rm=T)
-    joined_data[joined_data$species=='total','period'] = period_code
-    sp_predict = sp_predict +
-      geom_point(data = joined_data, mapping = aes(x = abundance, y = species),
-                 color = "blue")
-  }
-  plot(sp_predict)
+  
+  return(sp_predict)
 }
+
 
 
 #' Compares forecasts to observations over different lead times.

### 2017-09-27 profiles.Rmd

@@
 text_tbl <- data.frame(
-  Rodents = c("![]('rodent_pictures/color_pictures/BA.png')",
-              "![]('rodent_pictures/color_pictures/PB.png')",
-              "![]('rodent_pictures/color_pictures/PH.png')",
-              "![]('rodent_pictures/color_pictures/PI.png')",
-              "![]('rodent_pictures/color_pictures/PP.png')",
-              "![]('rodent_pictures/color_pictures/DM.png')",
-              "![]('rodent_pictures/color_pictures/DO.png')",
-              "![]('rodent_pictures/color_pictures/DS.png')",
-              "![]('rodent_pictures/Rodent_silhouettes/5-White-throatedwoodrat.png')",
-              "![]('rodent_pictures/Rodent_silhouettes/6-NorthernGrasshoppermouse.png')",
-              "![]('rodent_pictures/Rodent_silhouettes/7-SouthernGrasshoppermouse.png')",
-              "![]('rodent_pictures/color_pictures/PF.png')",
-              "![]('rodent_pictures/color_pictures/PE.png')",
-              "![]('rodent_pictures/color_pictures/PL.png')",
-              "![]('rodent_pictures/color_pictures/PM.png')",
-              "![]('rodent_pictures/Rodents_silhouettes/16-Fulvouseharvestmouse.png')",
-              "![]('rodent_pictures/Rodents_silhouettes/17-WesternHarvestmouse.png')",
-              "![]('rodent_pictures/Rodents_silhouettes/18-Plainsharvestmouse.png')",
-              "![]('rodent_pictures/Rodents_silhouettes/19-Tawny-belliedcottonrat.png')",
-              "![]('rodent_pictures/Rodents_silhouettes/20-HispidCottonRat.png')",
-              "![]('rodent_pictures/Rodents_silhouettes/21-Yellownosedcottonrat.png)"), 
   Scientific.Name = c("_Baiomys taylori_", 

### 2017-09-27 evaluation.Rmd
add evaluation page

---
title: "Evaluation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(dplyr)
library(portalr)
source("forecast_tools.R")
```

## How have the models done  historically?

These graphs show errors as a function of lead time. The lead time is the number of months into the future that forecast is made. The error values are an average of all forecast errors using observations since 2010. Note that this currently uses hindcasts of the prior observations.

**RMSE**: Root mean square error, this is a metric used to evaluate the point estimate of a forecast.  
**Coverage**: This is the percentage of observations which fell within the 95% confidence interval of a forecast. Ideally this would be equal to 0.95. If it's higher than 0.95 the forecasts intervals are too wide, if it's lower then the forecast intervals are too narrow.

```{r, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=9, fig.height=15}

new_moon_file = portalr::FullPath('PortalData/Rodents/moon_dates.csv', '~')
new_moons = read.csv(new_moon_file)

species_of_interest = c('BA','DM','DO','PP','OT','NA')
species_names = portalr::FullPath('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
  read.csv(stringsAsFactors=FALSE) %>%
  select(species = speciescode, full_species_name = scientificname)

#Fix neotoma, and add a total entry
species_names$species[species_names$full_species_name=='Neotoma albigula'] <- "NA"
species_names = species_names %>%
  mutate(species = ifelse(full_species_name=='Neotoma albigula', 'NA', species)) %>%
  add_row(species='total', full_species_name='Total Rodents')

species_abundance = portalr::abundance(shape='flat', level='Site') %>%
  rename(actual = abundance) %>%
  mutate(level='All',currency='abundance') %>%
  left_join(new_moons, by='period') %>%
  filter(species %in% species_of_interest)

total_abundance = portalr::abundance(shape='flat', level='Site') %>%
  group_by(period) %>%
  summarise(actual=sum(abundance)) %>%
  ungroup() %>%
  mutate(level='All',currency='abundance',species='total') %>%
  left_join(new_moons, by='period')

observation_data = species_abundance %>%
  bind_rows(total_abundance)

#Get the all the forecasts made during observation period
forecast_data = compile_forecasts(use_hindcasts = TRUE)
forecast_errors = calculate_forecast_error(observation_data, forecast_data, error_metric = 'RMSE') %>%
  filter(error_value < 200) %>% #Drop RMSE greater than this because it throws off all the graphs
  bind_rows(calculate_forecast_error(observation_data, forecast_data, error_metric = 'coverage'))

forecast_errors = forecast_errors %>%
  left_join(species_names, by='species')

ggplot(forecast_errors, aes(x=lead_time, y=error_value, group=model, color=model)) +
  geom_point()+
  geom_line() +
  geom_hline(yintercept = 0.95) +
  labs(x='Lead Time (New Moons)') +
  facet_wrap(full_species_name~error_metric, scales = 'free_y', ncol=2) + 
  theme_bw() +
  labs(y = "Error Value", colour = "Model")

```

### 2017-09-27 profiles.Rmd
Fix links to rodent images

@@
 text_tbl <- data.frame(
+  Rodents = c("![](../rodent_pictures/color_pictures/BA.png)",
+              "![](../rodent_pictures/color_pictures/PB.png)",
+              "![](../rodent_pictures/color_pictures/PH.png)",
+              "![](../rodent_pictures/color_pictures/PI.png)",
+              "![](../rodent_pictures/color_pictures/PP.png)",
+              "![](../rodent_pictures/color_pictures/DM.png)",
+              "![](../rodent_pictures/color_pictures/DO.png)",
+              "![](../rodent_pictures/color_pictures/DS.png)",
+              "![](../rodent_pictures/Rodent_silhouettes/5-White-throatedwoodrat.png)",
+              "![](../rodent_pictures/Rodent_silhouettes/6-NorthernGrasshoppermouse.png)",
+              "![](../rodent_pictures/Rodent_silhouettes/7-SouthernGrasshoppermouse.png)",
+              "![](../rodent_pictures/color_pictures/PF.png)",
+              "![](../rodent_pictures/color_pictures/PE.png)",
+              "![](../rodent_pictures/color_pictures/PL.png)",
+              "![](../rodent_pictures/color_pictures/PM.png)",
+              "![](../rodent_pictures/Rodent_silhouettes/16-Fulvouseharvestmouse.png)",
+              "![](../rodent_pictures/Rodent_silhouettes/17-WesternHarvestmouse.png)",
+              "![](../rodent_pictures/Rodent_silhouettes/18-Plainsharvestmouse.png)",
+              "![](../rodent_pictures/Rodent_silhouettes/19-Tawny-belliedcottonrat.png)",
+              "![](../rodent_pictures/Rodent_silhouettes/20-HispidCottonRat.png)",
+              "![](../rodent_pictures/Rodent_silhouettes/21-Yellownosedcottonrat.png)"), 
   Scientific.Name = c("_Baiomys taylori_", 

### 2017-09-27 profiles.Rmd

--- 2017-09-27 profiles.Rmd
@@
 text_tbl <- data.frame(
+  Rodents = c("![](rodent_pictures/color_pictures/BA.png)",
+              "![](rodent_pictures/color_pictures/PB.png)",
+              "![](rodent_pictures/color_pictures/PH.png)",
+              "![](rodent_pictures/color_pictures/PI.png)",
+              "![](rodent_pictures/color_pictures/PP.png)",
+              "![](rodent_pictures/color_pictures/DM.png)",
+              "![](rodent_pictures/color_pictures/DO.png)",
+              "![](rodent_pictures/color_pictures/DS.png)",
+              "![](rodent_pictures/Rodent_silhouettes/5-White-throatedwoodrat.png)",
+              "![](rodent_pictures/Rodent_silhouettes/6-NorthernGrasshoppermouse.png)",
+              "![](rodent_pictures/Rodent_silhouettes/7-SouthernGrasshoppermouse.png)",
+              "![](rodent_pictures/color_pictures/PF.png)",
+              "![](rodent_pictures/color_pictures/PE.png)",
+              "![](rodent_pictures/color_pictures/PL.png)",
+              "![](rodent_pictures/color_pictures/PM.png)",
+              "![](rodent_pictures/Rodent_silhouettes/16-Fulvouseharvestmouse.png)",
+              "![](rodent_pictures/Rodent_silhouettes/17-WesternHarvestmouse.png)",
+              "![](rodent_pictures/Rodent_silhouettes/18-Plainsharvestmouse.png)",
+              "![](rodent_pictures/Rodent_silhouettes/19-Tawny-belliedcottonrat.png)",
+              "![](rodent_pictures/Rodent_silhouettes/20-HispidCottonRat.png)",
+              "![](rodent_pictures/Rodent_silhouettes/21-Yellownosedcottonrat.png)"), 
   Scientific.Name = c("_Baiomys taylori_", 

### 2017-09-28 index.Rmd

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-obs_data = portalr::abundance()
+obs_data = portalr::abundance("repo")
 obs_data$total = rowSums(select(obs_data, -period))
-new_moon_file = portalr::FullPath('PortalData/Rodents/moon_dates.csv', 
-                                                  '~')
-new_moons = read.csv(new_moon_file)
+new_moons = read.csv(text = RCurl::getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
+
 obs_data_newmoon = inner_join(obs_data, new_moons, by = c("period" = "period"))

### 2017-10-05 index.Rmd
added 'ylabel' argument to forecast_viz function, to display species names on plot instead of sp codes

@@
              for_upperpi_col_name = "UpperPI",
-             start_newmoon = 300)
+             start_newmoon = 300,
+             ylabel = 'Total Abundance')
 ```
@@
 
-for (species in as.vector(most_abund_sp$species)) {
+# load in rodent species table to get scientific names to display on plots
+species_table = read.csv(
+  text = RCurl::getURL(
+      "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_species.csv"),stringsAsFactors = F,na.strings = '')
+most_abund_sp_names = filter(species_table,speciescode %in% most_abund_sp$species) %>% select(speciescode,scientificname)
+
+for (n in seq(dim(most_abund_sp_names)[1])) {
   species_forecast = forecast_viz(obs_data = obs_data_newmoon,
                obs_date_col_name = "censusdate",
-               obs_val_col_name = species,
+               obs_val_col_name = most_abund_sp_names$speciescode[n],
                for_data = for_data,
@@
                for_upperpi_col_name = "UpperPI",
-               start_newmoon = 300)
+               start_newmoon = 300,
+               ylabel = most_abund_sp_names$scientificname[n])
   plot(species_forecast)

### 2017-10-05 forecast_tools.R

@@ -324,17 +324,26 @@ all_forecasts$date=as.Date(all_forecasts$date)
 #' Visualize a time-series forecast
 #' Plots the observed time-series and the 1-step forecasts within it
 #' Plots the forecast time-series along with the prediction interval for future observations
-#' obs_data is a data.frame
-#' date_col_name is a string with the name for the date column
-#' val_col_name is a string with the name for the column of the value being forecast
+#' @param obs_data is a data.frame (observed data)
+#' @param obs_date_col_name is a string: name of the date column from obs_data
+#' @param obs_val_col_name is a string: name of the column of the value being forecast
+#' @param for_data is a data.frame (forecast data)
+#' @param for_date_col_name is a string: name of the date column from for_data
+#' @param for_val_col_name is a string: name of the column of value being forecast, from for_data
+#' @param for_model_name is a string: name of the model to be used from model column in for_data
+#' @param for_lowerpi_col_name is a string: name of the column of the lower confidence interval from for_data
+#' @param for_upperpi_col_name is a string: name of the column of the upper confidence interval from for_data
+#' @param start_newmoon is numeric: first new moon number to be plotted
+#' @param ylabel is a string: title for y-axis
 forecast_viz <- function(obs_data, obs_date_col_name, obs_val_col_name, for_data,
                          for_date_col_name, for_val_col_name, for_model_name,
-                         for_lowerpi_col_name, for_upperpi_col_name, start_newmoon){
+                         for_lowerpi_col_name, for_upperpi_col_name, start_newmoon,
+                         ylabel){
   for_data_sub = filter(for_data, species == obs_val_col_name, model == for_model_name)
   obs_data_sub = filter(obs_data, newmoonnumber >= start_newmoon)
   ggplot(obs_data_sub, aes_string(x = obs_date_col_name)) +
     geom_ribbon(data = for_data_sub, mapping = aes_string(x = for_date_col_name, ymin = for_lowerpi_col_name, ymax = for_upperpi_col_name), fill = "lightblue") +
     geom_line(aes_string(y = obs_val_col_name)) +
     geom_line(data = for_data_sub, mapping = aes_string(x = for_date_col_name, y = for_val_col_name), color = "blue") +
-    theme(axis.title.x=element_blank())
+    labs(x='',y=ylabel)
 }

### 2017-11-14 report.Rmd

@@
 ---
-title: "Report - September 2017"
+title: "Report - Latest Rodent Data"
 ---
@@
 rodents = abundance("repo", shape='flat')
-period_code = max(rodents$period)
-observed = dplyr::filter(rodents, period == period_code)
+most_recent_period = max(rodents$period)
+observed = dplyr::filter(rodents, period == most_recent_period)
 
@@
 newmoons_table = read.csv(text = getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
-moon_code = dplyr::filter(newmoons_table,period==period_code) %>% select(newmoonnumber) %>% as.integer()
+newmoons_table$censusdate = as.Date(newmoons_table$censusdate,format='%Y-%m-%d')
+moon_code = dplyr::filter(newmoons_table,period==most_recent_period) %>% select(newmoonnumber) %>% as.integer()
+most_recent_date = dplyr::filter(newmoons_table,period==most_recent_period) %>% dplyr::select(censusdate)
 
-# get most recent forecast prior to data collection -- WIP!!!!!!!! there's an issue up
-data = read.csv('predictions/2017-06-23Allforecasts.csv',na.strings = '')
-data = plyr::rename(data,c('NewMoonNumber' = 'newmoonnumber'))
+# get most recent forecast prior to data collection
+files = data.frame(names = list.files(path = './predictions', pattern = '*Allforecasts.csv'))
+files$filedates = as.Date(substr(files$names,1,10),format='%Y-%m-%d')
+latest_forecast = dplyr::filter(files,filedates<most_recent_date) %>% tail(1)
+
+data = read.csv(paste('predictions/',latest_forecast$names,sep=''),na.strings = '')
 

### 2017-11-15 report.Rmd
missed case of changing period_code to most_recent_period

@@
 joined_data[joined_data$species=='total','abundance'] = sum(joined_data$abundance,na.rm=T)
-joined_data[joined_data$species=='total','period'] = period_code
+joined_data[joined_data$species=='total','period'] = most_recent_period
 

### 2017-12-01 evaluation.Rmd
Fix paths in site files, delete temporary files after forecastall

@@
 library(portalr)
-source("forecast_tools.R")
+source("tools/forecast_tools.R")
 ```

### 2017-12-01 index.Rmd
Fix paths in site files, delete temporary files after forecastall

@@
 ```{r speciesforecasts, echo=FALSE, message=FALSE, warning=FALSE}
-source("forecast_tools.R")
+source("tools/forecast_tools.R")
 

### 2017-12-01 report.Rmd
Fix paths in site files, delete temporary files after forecastall

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-source("forecast_tools.R")
+source("tools/forecast_tools.R")
 library(portalr)
@@
 # get most recent forecast prior to data collection
-files = data.frame(names = list.files(path = './predictions', pattern = '*Allforecasts.csv'))
+files = data.frame(names = list.files(path = './predictions', pattern = '*forecasts.csv'))
 files$filedates = as.Date(substr(files$names,1,10),format='%Y-%m-%d')

### 2017-12-01 PortalForecasts.R

@@ -9,6 +9,6 @@ sapply( list.files("models", full.names=TRUE), source ) ###Temporary, while only
 #####Collect all forecast results and save to predictions directory######
 cat("Compiling forecasts", "\n")
 newforecasts=forecastall(Sys.Date())
-
+unlink("tmp/*")
 ######Update Website#####################################################
 rmarkdown::render_site()

### 2017-12-01 PortalHindcasts.R

@@ -70,9 +70,9 @@ for(this_newmoon in initial_time_newmoons){
   
   #####Run all models########################  
   cat("Running models", "\n")
+  dir.create("tmp")
   sapply( list.files("models", full.names=TRUE), source )
-  
   ####Compile all hindcasts into one file
   allhindcasts=forecastall(forecast_date, filename_suffix)
-
+  unlink("tmp/*")
 }

### 2018-01-20 manuscript/portal_iterative_forecasting_ms.Rmd
Add manuscript to repository. It was originally drafted on sharelatex.

---
title: "Automated iterative forecasting for the Portal Project"
csl: csl.csl
fontsize: 12pt
output:
  html_document: default
  pdf_document:
    keep_tex: yes
header-includes:
- \usepackage{times}
- \usepackage{setspace}
- \usepackage{booktabs}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
bibliography: refs.bib
---
\raggedright

# Introduction

Forecasting the future state of ecological systems is important for management, conservation, and evaluation of our fundamental understanding of ecology [@clark2001; @tallis2006; @diaz2015; @dietze2017]. Since Clark et al. [-clark2001] called for a more central role of forecasting in ecology, an increasing number of ecological forecasts are being published. However, most of these forecasts are made once, published, and never assessed or updated. Without assessment, we have limited information on how much confidence to place in our predictions; without regular updates, forecasts lack the most up-to-date information as conditions change [@dietze2018]. This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions. For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].

With the goal of making ecological forecasting more dynamic and responsive, Dietze et al [-dietze2018] recently called for an increase in iterative near-term forecasting. Iterative near-term forecasting means making forecasts for the near future and making these forecasts repeatedly through a cycle of forecast evaluation, integration of updated data, and generation of new forecasts. This approach to forecasting has a number of advantages. Because forecasts are made 'near-term'---daily to annual forecasting instead of multi-decadal---predictions can be assessed more quickly and frequently, leading to more rapid model improvements [@tredennick2016; @dietze2018]. Because the forecasts are made repeatedly through time, new data can be integrated with each new forecast cycle. This iterative approach to forecasting allows any changes in the state of the system that have occurred since the previous forecast to be incorporated and accounted for [@dietze2018]. Iterative near-term forecasting has the potential to promote rapid improvement in the state of ecological forecasting by quickly identifying how models are failing, facilitating rapid testing of improved models, and incorporating updated data so models run with the most up-to-date information on the system available. While use of iterative near-term forecasting is often contextualized as a management tool, this approach to model testing can also be used to improve our basic understanding of ecological systems. For example, alternative mechanistic models can be competed to see which model provides the best forecasts for near-term dynamics, thus providing insights into the relative importance of different processes driving dynamics of ecological systems [@dietze2018]. Whether deployed for basic or applied uses, iterative near-term forecasting incorporates a more dynamic interplay between models, predictions, and data that is clearly needed to improve ecological forecasting and our understanding of ecological systems more broadly.

Because iterative near-term forecasting requires a dynamic interplay of models, predictions, and data, Dietze et al [-dietze2018] highlight approaches to data management, model construction and evaluation, and cyberinfrastructure that are necessary to effectively implement this type of forecasting (Box 1). Data to be used for iterative near-term forecasting needs to be widely accessible, which requires data to be released quickly under open licenses [@dietze2018; @vargas2017] and structured so that it can be used easily by a variety of researchers and in multiple modeling approaches [@borer2009; @strasser2011]. Models need to be able to deal with uncertainty, in both the predictors and the predictions, to properly convey uncertainty in the resulting forecasts. Multiple models should be compared to assess which models are performing best and to allow for combining models to form ensemble predictions. Ensuring that data and models are regularly updated and new forecasts are made requires cyberinfrastructure to automate data processing, model fitting, prediction, model evaluation, forecast visualization, and archiving. In combination, these approaches should allow forecasts to be easily rerun and evaluated as new data becomes available [Box 1; @dietze2018].

While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement; few of their recommendations are in widespread use in ecology today. We examined what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going long-term (~40 year) ecological study that collects high-frequency data on desert rodent abundances. We constructed our forecasting pipeline with the goal of being able to forecast rodent abundances and evaluate our predictions on a monthly basis. In this paper, we discuss our approach for creating this iterative near-term forecasting pipeline, the challenges we encountered, the tools we used, and the lessons we learned that may help others to create their own iterative forecasting systems.

# System Background

Iterative forecasting requires data that is collected repeatedly, and it benefits most from data that is collected frequently, as this provides more opportunities for updating model results and assessing (and potentially improving) model performance [Box 1; @dietze2018]. The Portal Project is a long-term ecological study situated in the Chihuahuan Desert (2 km north and 6.5 km east of Portal, Arizona, US). Researchers have been continuously collecting data at the site since 1977, including data on the abundance of rodent and plant species (monthly and twice yearly, respectively) and climactic factors such as air temperature and precipitation (daily). The site consists of 24 50m x 50m experimental plots. Each plot contains 49 permanently marked trapping stations laid out in a 7 x 7 grid, and all plots are trapped with Sherman live traps for one night each month. For all rodents caught during a trapping session, information on species identity, size, and reproductive condition is collected, and new individuals are given identification tags. We use the data from the control plots at this site, where rodent populations are not experimentally manipulated. This data on rodent populations is high-frequency, uses consistent trapping methodology, and has an extended time-series (469 monthly samples and counting), making this study an ideal case for near-term iterative forecasting.

# Implementing an automated iterative forecasting system

Implementation of iterative forecasting requires the regular rebuilding of models with new raw data as it becomes available and the presentation of those forecasts in usable forms; in our case, this occurs monthly. Doing this in an efficient and maintainable way relies on developing an automated pipeline to handle the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presentation of the forecasts (Figure 1). We implemented this pipeline using a "continuous analysis" framework [*sensu* @beaulieu2017] that automatically processes the most up-to-date data, refits the models, makes new forecasts, archives the forecasts, and updates a website with analysis of current and previous forecasts. In this section we describe our approach to streamlining and automating the multiple components of the forecasting pipeline.

![Figure 1. Forecasting pipeline](figs/pipeline_fig.png)

## Continuous Analysis Framewor

A core component of iterative near-term forecasting is the regular rerunning of forecasting pipeline. This can be conducted manually - with a human making sure all code is run and all tables and files are updated - or automatically by having the computer conduct those tasks. We chose to have the computer run our pipeline and employed "continuous analysis" [*sensu* @beaulieu2017] to drive the automation of both the full pipeline and a number of its individual components. Continuous analysis uses a set of tools originally designed for software development called "continuous integration" (CI). CI combines computing environments for running code with monitoring systems to identify changes in data or code. Essentially, CI is a computer helper whose job is to watch the pipeline and, when it sees a change in the code or data, it runs all the tasks needed to ensure that the forecasting pipeline runs from beginning to end. This is useful for iterative near-term forecasting because it does not rely on humans to remember to create forecasts when new models or data are added. These tools are common in the area of software development where they are used to automate software testing and integrate work by multiple software developers working on the same software. However, these tools can be used for any computational task that needs to be regularly repeated or run after changes to code or data [@beaulieu2017]. Because of the widespread use of CI in software development, several CI services already exist. Our forecasting pipeline currently runs on a publicly available continuous integration service (Travis CI; https://travis-ci.org/) that is free for open source projects (up to a limited amount of computing time); alternative services that can be used to run code on local or cloud-based computational infrastructure are available, such as Drone (http://try.drone.io/) [@beaulieu2017]. As detailed below, we use CI to quality check data, test code using "unit tests" [@wilson2014], build models, make forecasts, and publicly present and archive the results (Figure 2).

![Figure 2. Continuous integration system.](figs/contanaly_fig.png)

In addition to automatically running software pipelines, the other key component of "continuous analysis" is making sure that the pipelines will continue to run even as software dependencies change [@beaulieu2017]. Many of us have experienced the frustrations that can occur when software updates (e.g., changes in R package versions) create errors in previously functional code. We experienced this issue when a package one of our models relies on, `tscount` [liboschik2015], was temporarily removed from CRAN (the R package repository) and, therefore, could not be installed in the usual way. This broke our forecasting pipeline because we could no longer run models that used that package. To minimize issues with changes in software dependencies, we follow Beaulieu and Green's [-@beaulieu2017] recommendation to use software containers. Software containers are standalone packages that contain copies of everything you need to run some piece of software. Once created, a software container is basically a time capsule, it contains all the software dependencies in the exact state used to develop and run the software. If those dependencies change (or disappear) in the wider world, they still exist, unchanged, in your container. We use an existing platform, Docker, to store an exact image of the complete software environment for running the forecasts. Docker also allows a specified set of packages to be used consistently across different computer and server environments. Using containers allows us to update to new package versions after testing and making any necessary changes to the data processing and analysis code. We use a container created by the Rocker project which is a Docker image with many important R packages (i.e. tidyverse) pre-installed  [@boettiger2017]. We use add our code and dependencies to this existing Rocker image to create a software container for our forecasting pipeline. In combination, the automated running of the pipeline (continuous integration) and the guarantee it will not stop working unexpectedly due to software dependencies (via a software container) allows continuous analysis to serve as the glue that connects all stages of the forecasting pipeline.

## Data Collection, Entry, and Processing

Iterative forecasting benefits from frequently updated data on the state of the system so that state changes can be quickly incorporated into new forecasts [@dietze2018]. Frequent data collection and rapid entry and processing of data are both important for providing the most up-to-date data for forecasting. Since our data is collected monthly, ensuring that the models have access to the newest data requires a data latency period of less than 1 month from collection to availability for modeling. To accomplish this, we automated components of the data processing and quality assurance/quality control (QA/QC) process to reduce the time needed to add new data to the database.

New data is double-entered into Excel using the "data validation" feature. The two versions are then compared in an R script to control for errors in data entry. Quality control (QC) checks written in R using the `testthat` R package [@wickham2011] are run on the data to test for validity and consistency both within the new data (e.g. sexual characteristics of an animal match M/F designation) and between the new and archived data (e.g. species and sex are consistent for recaptures of the same animal based on tag number). The local use of the QC scripts to flag problematic data greatly reduces the time spent error-checking and ensures that the quality of data is consistent. The data is then submitted to a GitHub-based data repository. GitHub is a software development tool for managing computer code development, but we have also found it useful for data management. Changes to data can be tracked through version control, and additions and changes to the data can be monitored through pull requests (notifications that someone has a modified version of the data to contribute). QA/QC checks are automatically rerun on the submitted data using continuous integration to ensure that these checks have been run and that no avoidable errors reach the official version of the dataset.

We also automated the updating of supplementary data tables, including information on weather and trapping history, that were previously updated manually. As soon as new field data is merged into the repository, continuous integration updates all supplementary files. Weather data is automatically fetched from our cellular-connected weather station, cleaned, and appended to the weather data table. Supplementary data tables related to trapping history are updated based on the data added to the main data tables. Using CI for this ensures that all supplementary data tables are always up-to-date with the core data.

## Data Sharing

The Portal Project has a long history of making its data publicly available, which means that anyone can use it for forecasting or other projects. Historically the publication of the data was conducted through data papers, the most common approach in ecology; however, this approach caused years of data latency. Recently, the project has switched to posting data directly to a public GitHub repository with a CC0 license. This immediate posting reduces that data latency to less than one month and, therefore, makes meaningful iterative near-term forecasting possible for not only our group but other interested parties, as well.

## Data Manipulation

Once data is available, it needs to be processed into a form appropriate for modeling. In many ecological datasets, this requires not only simple data manipulation but also a good understanding of the dataset to allow data to be aggregated appropriately. These data manipulation steps are often conducted using custom one-off code to convert the data into the desired form [@morris2013], but this approach has several limitations. First, each researcher must develop their own data manipulation code, which is inefficient and can result in different decisions by different researchers about the details of data cleaning and aggregation. Subtle differences in data processing decisions have lead to confusion when reproducing results for the Portal data in the past. Second, this kind of code is rarely robust to changes in data structure and location. Based on our experience developing and maintaining the Data Retriever [@morris2013; @senyondo2017], these kinds of changes are common. Finally, this kind of code is generally poorly tested, which can lead to errors based on mistakes in data manipulation. To avoid these issues for the Portal Project data, the Portal team has been developing an R package (portalR; http://github.com/weecology/portalr) for acquiring the data and handling common data cleaning and aggregation tasks. As a result, our modeling and forecasting code only needs to install this package and run the data manipulation and summary functions to get the appropriate data. The package undergoes thorough automated unit tests to ensure that data manipulations are achieving the desired results. Having data manipulation code maintained in a separate package that focuses on consistently providing properly summarized forms of the most recent data has made maintaining the forecasting code itself much more straightforward.

## Modeling and Forecasting

Ideally, iterative near-term forecasting involves regularly refitting a variety of different models. New models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach. We use CI to refit the models and make new forecasts each time the modeling code changes and when new data becomes available. We use a plugin infrastructure to allow new models to be easily added to the system. Each model is a script that takes in standardized inputs and returns standardized outputs (Figure 3). All models are run by the main forecasting code at each update, and the standardized outputs are combined to store the results of the different models' forecasts. A weighted ensemble model is then added with weights based on how well individual models fit the training data. This plugin infrastructure allows flexibility in all aspects of the modeling process, making it easier to explore new modeling approaches, and allows new models that fit the data well to immediately improve the ensemble forecast.

![Figure 3. Demonstration of plugin infrastructure where each model script uses data provided by the core forecasting code and returns its forecasts in a predefined structure.](figs/plugin_fig.gif)

To allow flexibility in what the models predict, we designed a forecast output format that allows us to store relatively generic forecasts. Each forecast output file contains the date being forecast, the collection date of the data used for fitting the models, the date the forecast was made, the state variable being forecast (e.g., rodent biomass, the abundance of a species), and the forecast value and associated uncertainty of that forecast (Figure 3). This allows a variety of different forecasts to be stored in a common format and may serve as a useful starting point for developing a standard for storing ecological forecasts more generally.

## Evaluation and Hindcasting

Forecasts are currently evaluated using root mean square error (RMSE) to evaluate point forecasts and coverage to evaluate uncertainty. We plan to add additional metrics in the future. In addition to evaluating the actual forecasts, we also use hindcasting (forecasting on already collected data) to gain additional insight into the methods that work best for forecasting this system. For example a model is fit using rodent observations up to June 2005, then used to make a forecast 12 months out to May 2006. The observations of that 12 month period can immediately be used to evaluate the model. Hindcasting can be conducted using all months from the beginning of the study, thus allowing model comparison of large numbers of hindcasts and giving insight into which models make the best forecasts. It can also be used to quickly evaluate new models instead of waiting for an adequate amount of data to accumulate.

## Archiving

Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2017; @harris2017]. This archiving serves as a form of pre-registration for model predictions, helping facilitate unbiased interpretation of model performance. We explored three major repositories for archiving our forecasts: FigShare, Zenodo, and Open Science Framework. While all three repositories allowed for easy manual submissions (i.e., a human uploading files after each forecast), automating this process was substantially more difficult. Various combinations of repositories, APIs and associated R packages had issues with: 1) integrating authorization with continuous integration; 2) automatically making archived files public; 3) adding new files to an existing location; or 4) automatically permanently archiving the files. Our eventual solution was to leverage the GitHub-Zenodo integration (https://guides.github.com/activities/citable-code/) and automatically push forecasts to the GitHub repository from the CI server. The GitHub-Zenodo integration is designed to automatically create versioned archives of GitHub repositories. There is an existing one-time process for linking our forecasting repository on  GitHub with Zenodo. Once this link is created, each time a new forecast is created, our pipeline adds the new forecasts to the GitHub repository and uses the GitHub API to create a new "release" of our repository. This triggers the Zenodo-GitHub integration, which automatically archives the resulting forecasts and the code that generated them under a top-level DOI that refers to all archived forecasts ([https://doi.org/10.5281/zenodo.833438](https://doi.org/10.5281/zenodo.833438)). Through this process, we automatically archive every forecast made with a documented history of the archive. While this approach is functional because everything in the repository is archived when a new forecast is made, these archives are complicated, making it more complicated than necessary to find and access the forecasting results.

## Presentation

In addition to archiving the results of each forecast, we present them on a website that displays monthly rodent forecasts, model evaluation metrics, monthly reports, and information about the study site (Figure 4; [http://portal.naturecast.org](http://portal.naturecast.org/)). The website includes a graphical presentation of the most recent month's forecasts (including uncertainty) and compares the latest data to the previous forecasts. Information on the the species and the field site targeted to a general audience are also included. The site is built using Rmarkdown [@rmarkdown], which naturally integrates into the pipeline, and is automatically updated after each forecast. The `knitr` R package [@xie2015] compiles the code into HTML, which is then published using Github Pages (https://pages.github.com/). The files for the website are stored in a subdirectory of the forecasting repository. As a result, the website is also archived automatically as part of the forecast archiving.

![Figure 4. Screen capture of the homepage of the Portal Forecasting website; http://portal.naturecast.org](figs/website_fig.png)

# Discussion

Following the recommendations of Dietze et al [-dietze2018], we developed an automated iterative forecasting system to support repeated forecasting of an ecological system. Our forecasting system automatically acquires and processes the newest data, refits the models, makes new forecasts, publicly archives those forecasts, and presents both the current forecast and information on how previous forecasts performed. Every week our forecasting system generates a new set of forecasts with no human intervention, except for the entry of new field data. This ensures that forecasts based on the most recent data are always available and allows us to rapidly assess the performance of multiple forecasting models for a number of different states of the system, including the abundances of individual species and community-level variables such as total abundance. To create this iterative near-term forecasting system, we used R to process data and conduct analyses, and we leveraged already existing services (i.e. GitHub, Travis, Docker) for more complicated cyberinfrastructure tasks. Thus, our approach to developing iterative near-term forecasting infrastructure provides an example for how short-term ecological forecasting systems can be initially developed.

We designed this forecasting system with the goal of making it relatively easy to build, maintain, and extend. We used existing technology for both running the pipeline and building individual components, which allowed the system to be built relatively cheaply in terms of both time and money. This included the use of tools like Docker for reproducibility, the Travis CI continuous integration system for automatically running the pipeline, Rmarkdown and `knitr` for generating the website, and the already existing integration between Github and Zenodo  to archive the forecasts. By using this "continuous analysis" approach [@beaulieu2017], where analyses are automatically rerun when changes are made to data, models, or associated code, we have reduced the time required by scientists to run and maintain the forecasting pipeline. To make the system extensible so that new models could be easily incorporated, we use a plugin-based infrastructure so that adding a new model to the system is as easy as adding a single file to the 'models' folder in our repository (Figure 3). This should substantially lower the barriers to other scientists contributing models to this forecasting effort. We also automatically archive the resulting forecasts publicly so that, as new data is collected, the performance of these forecasts can be assessed by both us and other researchers. This serves as a form of "pre-registration" by providing a quantitative record of the forecast before the data being predicted were collected.

While building this system was facilitated by the use of existing technological solutions, there were still a number of challenges in making existing tools work for automated iterative forecasting. Continuous integration is designed primarily for running automated tests on software, not for running a coordinated forecasting pipeline. As a result, extra effort was sometimes necessary to figure out how to get these systems to work properly in non-standard situations, like running code that was not part of a software package. In addition, hosted continuous integration solutions, like Travis, provide only limited computational resources. As the number and complexity of the models we fit has grown, we have had to continually invest effort in reducing our total compute time so we can stay within these limits. Finally, we found no satisfactory existing solution for archiving our results. All approaches we tried had limitations when it came to automatically generating publicly versioned archives of forecasts on a repeated basis. Overall, we found existing technology to be sufficient to the task, but it required greater expertise and a greater investment of time than is ideal. Tool development to reduce the effort required for scientists to set up their own short-term forecasting systems would clearly be useful, but our efforts show that it is still currently possible for scientists using existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems. By developing these systems so that they already produce automated iterative forecasts, it will be easier to convert these systems into fully operationalized forecasting systems that are relied on for decision making [@dietze2018, other paper by operationalization co-authors].

Because of the breadth of expertise needed to set up our forecasting pipeline, our effort required a team with diverse skills and perspectives, ranging from software development to field site expertise. It is rare to find such breadth within a single research group, and our system was developed as a collaboration between the lab collecting the data and a computational ecology lab. When teams have a breadth of expertise, communication can be challenging. We found a shared base of knowledge related to both the field research and fundamental computational skills was important for the success of the group. Everyone on the team had received training in fundamental data management and computing skills through a combination of university courses, Software and Data Carpentry workshops [@teal2016], and informal lab training efforts. In addition, everyone was broadly familiar with the study site and methods of data collection, and most team members had participated in field work at the site on multiple occasions. This provided a shared set of knowledge and vocabulary that actively facilitated interdisciplinary interactions. Given the current state of existing tools for forecasting, forecasting teams will need people with significant experience in working with continuous integration and APIs, which means interdisciplinary teams will generally be required for creating these pipelines until tool development improves.

We developed this infrastructure for automatically making iterative forecasts with the goals of making accurate forecasts for this well-studied system, learning what methods work well for ecological forecasting more generally, and improving our understanding of the processes driving ecological dynamics. The most obvious application of automated iterative ecological forecasting is for speeding up development of forecasting models by providing the most recent data available to models and by quickly iterating to improve the models used for forecasting. By learning what works best for forecasting in this and other ecological systems, we will better understand what the best approaches are for ecological forecasting more generally. By designing the pipeline so that it can forecast many different aspects of the ecological community, we also hope to learn about what aspects of ecology are more forecastable. Finally, automated forecasting infrastructures like this one also provide a core foundation for faster scientific inquiry more broadly because new models can quickly be applied to data and compared to existing models. The forecasting infrastructure does the time-consuming work of data processing, data integration, and model assessment, allowing new research to focus on the models being developed and the inferences about the system that can be drawn from them [@dietze2018]. We plan to use this pipeline to drive future research into understanding the processes that govern the dynamics of individual populations and the community as a whole. By regularly running different models for population and community dynamics, a near-term iterative pipeline such as ours should also make it possible to rapidly detect changes in how the system is operating, which should allow the rapid identification of ecological transitions or even possibly allow them to be prevented [@carpenter example]. By building an automated iterative near-term forecasting infrastructure we can improve our ability to forecast natural systems, our understanding of the biology driving ecological dynamics, and detect or even predict changes in system state that are important for conservation and management.

# Acknowledgements

This research was supported by the National Science Foundation through grant 1622425 to S.K.M. Ernest and by the Gordon and Betty Moore Foundation’s Data-Driven Discovery Initiative through grant GBMF4563 to E.P. White. We thank all of the graduate students, postdocs, and volunteers who have collected the Portal Project over the last 40 years and the developers of all of the software and tools that made this project possible.

# Box 1. Key practices for automated iterative near-term ecological forecasting

A list of some of the key practices developed by Dietze et al [-dietze2018] for facilitating iterative near-term ecological forecasting and discussion of why these practices are important.

## Data

**1. Frequent data collection**

Frequent data collection allows models to be regularly updated and forecasts to be frequently evaluated [@dietze2018]. Depending on the system being studied, this frequency could range from sub-daily to annual, but typically the more frequently the data is collected the better.

**2. Rapid data release under open licenses**

Data should be released as quickly as possible (low latency) under open licenses so that forecasts can be made frequently and data can be accessed by a community of forecasters [@dietze2018; @vargas2017].

**3. Best practices in data structure**

To reduce the time and effort needed to incorporate this data into models, best practices in data structure need to be employed for managing and storing collected data to ensure it is easy to integrate into other systems (interoperability) [@borer2009; @strasser2011; @white2013].

## Models

**4. Focus on uncertainty**

Understanding the uncertainty of forecasts is crucial to interpreting the forecasts and understanding their utility. Models used for forecasting should be probabilistic to properly quantify uncertainty and to convey how this uncertainty increases through time. Evaluation of forecast models should include assessment of how accurately they quantify uncertainty as well as point estimates. This can be done using "proper and local" scores [@hooten2015].

**5. Compare forecasts to simple baselines**

Understanding how much information is present in a forecast requires comparing its accuracy to simple baselines to see if the models yield improvements over the naive expectation that the system is static [@harris2018].

**6. Compare and combine multiple modeling approaches**

To quickly learn about the best approaches to forecasting different aspects of ecology, multiple modeling approaches should be compared for forecasting tasks [@harris2018]. Different modeling approaches should also be combined into ensemble models, which are known to outperform single models for many forecasting and prediction tasks [@weigel2008].

## Cyberinfrastructure

In addition to improvements in data and models, iterative near-term forecasting requires improved infrastructure and approaches to support continuous model development and iterative forecasting [@dietze2018].

**7. Best practices in software development**

Best practices should be followed in the development of scientific software and modeling to make it easier to maintain, integrate into pipelines, and build on by other researchers. Key best practices include using open licenses, good documentation, version control, and cross-platform support [@wilson2014; @hampton2015].

**8. Support easy inclusion of new models**

To facilitate the comparison and ensembling of different modeling approaches, code for fitting models and making forecasts should be easily extensible, allowing models developed by different groups to be easily integrated into a single framework [@dietze2018].

**9. Automated end-to-end reproducibility**

Iteratively making forecasts requires that acquiring the newest data, refitting the models, and making new forecasts is simple. Ideally, this should be done automatically without requiring human intervention. Therefore, the process of making forecasts should emphasize end-to-end reproducibility, including data, models, and evaluation [@stodden2014], to allow the forecasts to be easily rerun as new data becomes available [@dietze2018]. Ideally, the entire forecasting pipeline will be rerun automatically as new data becomes available.

**10. Publicly archive forecasts**

Forecasts should be openly archived to demonstrate that the forecasts were made without knowledge of the outcomes (i.e., as a form of pre-registration *sensu*) and to allow the community to assess and compare the performance of different forecasting approaches both now and in the future [@mcgill2012; @tredennick2016; @dietze2018; @harris2018]. Ideally, the forecasts and evaluation of their performance should be automatically posted publicly in a manner that is understandable by both interested scientists and other stakeholders.

### 2018-01-20 manuscript/portal_iterative_forecasting_ms.Rmd
Add missing references.

@@
 output:
-  html_document: default
   pdf_document:
     keep_tex: yes
+  html_document: default
 header-includes:
@@
 
-Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2017; @harris2017].
+Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2017; @harris2018].
 This archiving serves as a form of pre-registration for model predictions, helping facilitate unbiased interpretation of model performance.
@@
 We found a shared base of knowledge related to both the field research and fundamental computational skills was important for the success of the group.
-Everyone on the team had received training in fundamental data management and computing skills through a combination of university courses, Software and Data Carpentry workshops [@teal2016], and informal lab training efforts.
+Everyone on the team had received training in fundamental data management and computing skills through a combination of university courses, Software and Data Carpentry workshops [@teal2015], and informal lab training efforts.
 In addition, everyone was broadly familiar with the study site and methods of data collection, and most team members had participated in field work at the site on multiple occasions.

### 2018-01-20 manuscript/portal_iterative_forecasting_ms.Rmd
Switch to Methods in Ecology & Evoluation csl.

@@
 title: "Automated iterative forecasting for the Portal Project"
-csl: csl.csl
+csl: mee.csl
 fontsize: 12pt

### 2018-01-20 manuscript/portal_iterative_forecasting_ms.Rmd
Add References header

@@
 output:
+  html_document: default
   pdf_document:
     keep_tex: yes
-  html_document: default
 header-includes:
@@
 Ideally, the forecasts and evaluation of their performance should be automatically posted publicly in a manner that is understandable by both interested scientists and other stakeholders.
+
+# References

### 2018-01-22 manuscript/portal_iterative_forecasting_ms.Rmd

@@
 
-## Continuous Analysis Framewor
+## Continuous Analysis Framework
 
 A core component of iterative near-term forecasting is the regular rerunning of forecasting pipeline.
-This can be conducted manually - with a human making sure all code is run and all tables and files are updated - or automatically by having the computer conduct those tasks.
+This can be conducted manually---with a human making sure all code is run and all tables and files are updated---or automatically by having the computer conduct those tasks.
 We chose to have the computer run our pipeline and employed "continuous analysis" [*sensu* @beaulieu2017] to drive the automation of both the full pipeline and a number of its individual components.

### 2018-01-22 evaluation.Rmd
move everything to eval page and fine tune

--- 2017-09-27 evaluation.Rmd
@@
 knitr::opts_chunk$set(
-	echo = TRUE,
+	echo = FALSE,
 	message = FALSE,
@@
 library(portalr)
-source("forecast_tools.R")
+source("tools/forecast_tools.R")
+
+species_of_interest = c('BA','DM','DO','PP','OT','NA','total')
+species_names = portalr::FullPath('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
+  read.csv(stringsAsFactors=FALSE) %>%
+  select(species = speciescode, full_species_name = scientificname)
+
+#Fix neotoma, and add a total entry
+species_names$species[species_names$full_species_name=='Neotoma albigula'] <- "NA"
+species_names = species_names %>%
+  mutate(species = ifelse(full_species_name=='Neotoma albigula', 'NA', species)) %>%
+  add_row(species='total', full_species_name='Total Rodents')
+
+```
+
+## How have the models recently?
+
+This graph shows the forecast error of some recent forecasts for the most recent sampling dates.
+This is shown for all model types in the columns and several common rodent species (and total rodents) in the rows.
+These forecasts are for the entire site.
+```{r recent_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=15, fig.height=15}
+library(ggrepel)
+# The length of time to look backward
+prior_new_moons = 12
+
+new_moon_file = portalr::FullPath('PortalData/Rodents/moon_dates.csv', '~')
+new_moons = read.csv(new_moon_file)
+most_recent_newmoon = max(new_moons$newmoonnumber)
+earliest_newmoon_to_graph = most_recent_newmoon - prior_new_moons
+
+species_abundance = portalr::abundance(shape='flat', level='Site') %>%
+  rename(actual = abundance) %>%
+  mutate(level='All',currency='abundance') %>%
+  left_join(new_moons, by='period') %>%
+  filter(species %in% species_of_interest)
+
+total_abundance = portalr::abundance(shape='flat', level='Site') %>%
+  group_by(period) %>%
+  summarise(actual=sum(abundance)) %>%
+  ungroup() %>%
+  mutate(level='All',currency='abundance',species='total') %>%
+  left_join(new_moons, by='period')
+
+observation_data = species_abundance %>%
+  bind_rows(total_abundance) %>% 
+  filter(newmoonnumber %in% earliest_newmoon_to_graph:most_recent_newmoon)
+
+#Get the all the forecasts made during observation period
+forecast_data = compile_forecasts(use_hindcasts = FALSE) %>%
+  filter(newmoonnumber %in% earliest_newmoon_to_graph:most_recent_newmoon)
+
+# Round so that large decimal place differences do not cause
+# similar forecasts to be seen as distinct. 
+forecast_data$estimate = round(forecast_data$estimate, 3)
+forecast_data$LowerPI = round(forecast_data$LowerPI, 3)
+forecast_data$UpperPI = round(forecast_data$UpperPI, 3)
+
+# Keep 1 forecast per unique initial_newmoon, when there are multiple forecasts
+# using the same initial_newmoon, use the first one.
+forecast_dates_to_keep = forecast_data %>%
+  select(date, initial_newmoon) %>%
+  distinct() %>%
+  group_by(initial_newmoon) %>%
+  filter(date == min(date)) %>%
+  ungroup() %>%
+  mutate(keep='yes')
+
+forecast_data = forecast_data %>%
+  left_join(forecast_dates_to_keep, by=c('date','initial_newmoon')) %>%
+  filter(keep=='yes') %>%
+  select(-keep)
+##################################
+
+forecast_errors = forecast_data %>%
+  left_join(observation_data, by=c('species','level','currency','newmoonnumber')) %>% 
+  mutate(rmse = sqrt((estimate - actual)^2))
+
+# Sometimes there are model runs on the same day and with all the same info,
+# this gets rid of those
+forecast_errors = forecast_errors %>%
+  distinct()
+
+# Drop any entries that don't have an observation
+# (ie. a rainy sample period)
+forecast_errors = forecast_errors %>%
+  filter(!is.na(actual))
+
+# Filter to common species and apply the full name
+forecast_errors = forecast_errors %>%
+  filter(species %in% species_of_interest) %>%
+  left_join(species_names, by='species') %>%
+  select(-species) %>%
+  rename(species=full_species_name)
+
+# Add a newline to the longer species names
+forecast_errors$species = with(forecast_errors, ifelse(nchar(species)>15, stringr::str_replace(species,' ','\n'), species))
+
+###################################
+forecast_date_labels = forecast_errors %>%
+  group_by(date, model, species, censusdate) %>%
+  summarize(y_position = mean(rmse)) %>%
+  ungroup() %>%
+  distinct() %>%
+  filter(!is.na(censusdate)) %>%
+  group_by(date, model, species) %>%
+  filter(as.Date(censusdate) == min(as.Date(censusdate))) %>%
+  ungroup() %>%
+  distinct() %>%
+  mutate(label_text = paste0('Issued: ',date))
+
+####################################
+# Setup a nice palette for when prior issue dates become greater than 9
+getPalette = colorRampPalette(RColorBrewer::brewer.pal(9, "Set1"))
+#Randomize pallette so nearby regions are not similar colors
+set.seed(2)
+large_color_palette = sample(getPalette(40))
+#####################################
+ggplot(forecast_errors, aes(x=censusdate, y=rmse, group=as.character(date), color=as.character(date))) +
+  geom_point(size=4)+
+  geom_line(size=1.5) +
+  # label individual lines if there gets to be a lot of them. remove the legend if this is turned on.
+  #geom_label_repel(data=forecast_date_labels, aes(label=label_text, y=y_position),
+  #                 fill='grey90',size=4,nudge_y = 40, force=5,
+  #                 segment.alpha = 0.3, segment.size = 0.8) +
+  scale_color_brewer(palette='Dark2') +
+  #scale_color_manual(values=large_color_palette) +
+  facet_grid(species~model) + 
+  theme_bw() +
+  theme(axis.text.x = element_text(angle=45, hjust=0.9, debug = FALSE),
+        axis.text = element_text(size=12),
+        axis.title = element_text(size=20),
+        strip.text = element_text(size=12),
+        legend.text = element_text(size=15),
+        legend.title = element_text(size=18)) +
+  labs(x='Sample Date',y='RMSE', color='Forecast Issue\nDate')
+
 ```
@@
 
-```{r, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=9, fig.height=15}
+```{r hindcast_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=9, fig.height=15}
 
@@
 new_moons = read.csv(new_moon_file)
-
-species_of_interest = c('BA','DM','DO','PP','OT','NA')
-species_names = portalr::FullPath('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
-  read.csv(stringsAsFactors=FALSE) %>%
-  select(species = speciescode, full_species_name = scientificname)
-
-#Fix neotoma, and add a total entry
-species_names$species[species_names$full_species_name=='Neotoma albigula'] <- "NA"
-species_names = species_names %>%
-  mutate(species = ifelse(full_species_name=='Neotoma albigula', 'NA', species)) %>%
-  add_row(species='total', full_species_name='Total Rodents')
 

### 2018-01-22 recent_data_eval.R

@@ -1,126 +1,0 @@
-library(dplyr)
-library(portalr)
-library(ggrepel)
-source("tools/forecast_tools.R")
-
-# The length of time to look backward
-prior_new_moons = 12
-
-new_moon_file = portalr::FullPath('PortalData/Rodents/moon_dates.csv', '~')
-new_moons = read.csv(new_moon_file)
-most_recent_newmoon = max(new_moons$newmoonnumber)
-earliest_newmoon_to_graph = most_recent_newmoon - prior_new_moons
-
-species_of_interest = c('BA','DM','DO','PP','OT','NA')
-species_names = portalr::FullPath('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
-  read.csv(stringsAsFactors=FALSE) %>%
-  select(species = speciescode, full_species_name = scientificname)
-
-#Fix neotoma, and add a total entry
-species_names$species[species_names$full_species_name=='Neotoma albigula'] <- "NA"
-species_names = species_names %>%
-  mutate(species = ifelse(full_species_name=='Neotoma albigula', 'NA', species)) %>%
-  add_row(species='total', full_species_name='Total Rodents')
-
-species_abundance = portalr::abundance(shape='flat', level='Site') %>%
-  rename(actual = abundance) %>%
-  mutate(level='All',currency='abundance') %>%
-  left_join(new_moons, by='period') %>%
-  filter(species %in% species_of_interest)
-
-total_abundance = portalr::abundance(shape='flat', level='Site') %>%
-  group_by(period) %>%
-  summarise(actual=sum(abundance)) %>%
-  ungroup() %>%
-  mutate(level='All',currency='abundance',species='total') %>%
-  left_join(new_moons, by='period')
-
-observation_data = species_abundance %>%
-  bind_rows(total_abundance) %>% 
-  filter(newmoonnumber %in% earliest_newmoon_to_graph:most_recent_newmoon)
-
-#Get the all the forecasts made during observation period
-forecast_data = compile_forecasts(use_hindcasts = FALSE) %>%
-  filter(newmoonnumber %in% earliest_newmoon_to_graph:most_recent_newmoon)
-
-forecast_data$estimate = round(forecast_data$estimate, 3)
-forecast_data$LowerPI = round(forecast_data$LowerPI, 3)
-forecast_data$UpperPI = round(forecast_data$UpperPI, 3)
-
-# Scale down the number of forecasts presented. Since a new one is issued every week
[…]

### 2018-01-22 evaluation.Rmd
remove ggrepel and add note for it

@@
 ```{r recent_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=15, fig.height=15}
-library(ggrepel)
 # The length of time to look backward
@@
   geom_line(size=1.5) +
-  # label individual lines if there gets to be a lot of them. remove the legend if this is turned on.
+  # label individual issue dates. This is easier to look at  if there 
+  # gets to be a lot of them. Must install ggrepel and remove the legend if this is turned on.
   #geom_label_repel(data=forecast_date_labels, aes(label=label_text, y=y_position),

### 2018-01-23 manuscript/portal_iterative_forecasting_ms.Rmd

--- 2018-01-20 manuscript/portal_iterative_forecasting_ms.Rmd
@@
 Doing this in an efficient and maintainable way relies on developing an automated pipeline to handle the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presentation of the forecasts (Figure 1).
-We implemented this pipeline using a "continuous analysis" framework [*sensu* @beaulieu2017] that automatically processes the most up-to-date data, refits the models, makes new forecasts, archives the forecasts, and updates a website with analysis of current and previous forecasts.
-In this section we describe our approach to streamlining and automating the multiple components of the forecasting pipeline.
+To implement the pipeline outlined in Figure 1, we used a "continuous analysis" framework [*sensu* @beaulieu2017] that automatically processes the most up-to-date data, refits the models, makes new forecasts, archives the forecasts, and updates a website with analysis of current and previous forecasts.
+In this section we describe our approach to streamlining and automating the multiple components of the forecasting pipeline and the tools and infrastructure we employed to execute each stage of the pipeline.
 
-![Figure 1. Forecasting pipeline](figs/pipeline_fig.png)
+![Figure 1.
+Stages of the forecasting pipeline.
+To go from raw data to forecast presentation involves a number of stages, each of which required unique tasks, tools and infrastructure.
+While each stage has unique tasks associated with it, the stages of our pipeline are also interdependent.
+Our pipeline is primarily a linear structure where outputs from one stage form the inputs for the subsequent stage ](figs/pipeline_fig.png)
 
-## Continuous Analysis Framewor
+## Continuous Analysis Framework
 
-A core component of iterative near-term forecasting is the regular rerunning of forecasting pipeline.
-This can be conducted manually - with a human making sure all code is run and all tables and files are updated - or automatically by having the computer conduct those tasks.
+A core component of iterative near-term forecasting is the regular rerunning of the forecasting pipeline.
+This can be conducted manually---with a human making sure all code is run and all tables and files are updated---or automatically by having the computer conduct those tasks.
 We chose to have the computer run our pipeline and employed "continuous analysis" [*sensu* @beaulieu2017] to drive the automation of both the full pipeline and a number of its individual components.
@@
 
-![Figure 2. Continuous integration system.](figs/contanaly_fig.png)
+![Figure 2.
+Continuous integration system.
+Each box denotes the core infrastructure used for each stage of the forecasting pipeline denoted in Figure 1.
+Continuous integration, denoted here with the Travis icon (a girl wearing safety glasses and hardhat), plugs into our pipeline at every stage.
+Travis triggers the code involved in major events that involve integration across stages of the pipeline, such as taking the output from the forecasting stage (purple box) to create an updated presentation (rose box).
+Travis also runs tasks within a stage, such as conducting tests to make sure code changes have not introduced errors (icons on arrows originating and ending on the same box)](figs/contanaly_fig.png)
 
@@
 Since our data is collected monthly, ensuring that the models have access to the newest data requires a data latency period of less than 1 month from collection to availability for modeling.
-To accomplish this, we automated components of the data processing and quality assurance/quality control (QA/QC) process to reduce the time needed to add new data to the database.
+To accomplish this, we automated components of the data processing and quality assurance/quality control (QA/QC) process to reduce the time needed to add new data to the database (Figures 1 and 2).
 
@@
 Historically the publication of the data was conducted through data papers, the most common approach in ecology; however, this approach caused years of data latency.
-Recently, the project has switched to posting data directly to a public GitHub repository with a CC0 license.
-This immediate posting reduces that data latency to less than one month and, therefore, makes meaningful iterative near-term forecasting possible for not only our group but other interested parties, as well.
+Recently, the project has switched to posting data directly to a public GitHub repository with a CC0 license (Figure 1).
+This immediate posting reduces that data latency to less than one month and, therefore, makes meaningful iterative near-term forecasting possible for not only our group but other interested parties as well.
 
@@
 
-Once data is available, it needs to be processed into a form appropriate for modeling.
+Once data is available, it needs to be processed into a form appropriate for modeling (Figure 1).
 In many ecological datasets, this requires not only simple data manipulation but also a good understanding of the dataset to allow data to be aggregated appropriately.
@@
 To avoid these issues for the Portal Project data, the Portal team has been developing an R package (portalR; http://github.com/weecology/portalr) for acquiring the data and handling common data cleaning and aggregation tasks.
-As a result, our modeling and forecasting code only needs to install this package and run the data manipulation and summary functions to get the appropriate data.
+As a result, our modeling and forecasting code only needs to install this package and run the data manipulation and summary functions to get the appropriate data (Figure 2).
 The package undergoes thorough automated unit tests to ensure that data manipulations are achieving the desired results.
@@
 
-Ideally, iterative near-term forecasting involves regularly refitting a variety of different models.
+Ideally, iterative near-term forecasting involves regularly refitting a variety of different models (Figure 1).
 New models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach.
-We use CI to refit the models and make new forecasts each time the modeling code changes and when new data becomes available.
+We use CI to refit the models and make new forecasts each time the modeling code changes and when new data becomes available (Figure 2).
 We use a plugin infrastructure to allow new models to be easily added to the system.
@@
 
-![Figure 3. Demonstration of plugin infrastructure where each model script uses data provided by the core forecasting code and returns its forecasts in a predefined structure.](figs/plugin_fig.gif)
+![Figure 3.
+Demonstration of plugin infrastructure where each model script (represented here as AR1.R) uses data provided by the core forecasting code (represented here as rodent.csv) and returns its forecasts in a predefined structure that is consistent across models (represented here as 2017_12_08forecasts.csv) to allow cross-model evaluations.](figs/plugin_fig.gif)
 
@@
 This allows a variety of different forecasts to be stored in a common format and may serve as a useful starting point for developing a standard for storing ecological forecasts more generally.
-
-## Evaluation and Hindcasting
 
@@
 
-Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2017; @harris2018].
+Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2017; @harris2018](Figure 1).
 This archiving serves as a form of pre-registration for model predictions, helping facilitate unbiased interpretation of model performance.
@@
 
-![Figure 4. Screen capture of the homepage of the Portal Forecasting website; http://portal.naturecast.org](figs/website_fig.png)
+![Figure 4.
+Screen capture of the homepage of the Portal Forecasting website; http://portal.naturecast.org.
+This site contains information on the most current forecasts, evaluation of forecast performance, and general information about the species being forecast.](figs/website_fig.png)
 
@@
 
-Following the recommendations of Dietze et al [-dietze2018], we developed an automated iterative forecasting system to support repeated forecasting of an ecological system.
+Following the recommendations of Dietze et al [-dietze2018], we developed an automated iterative forecasting system (Figures 1 and 2) to support repeated forecasting of an ecological system.
 Our forecasting system automatically acquires and processes the newest data, refits the models, makes new forecasts, publicly archives those forecasts, and presents both the current forecast and information on how previous forecasts performed.

### 2018-01-26 manuscript/portal_iterative_forecasting_ms.Rmd
Juniper's edits to manuscript
Changes made to the manuscript in response to Juniper's comments and edits

@@
 Because forecasts are made 'near-term'---daily to annual forecasting instead of multi-decadal---predictions can be assessed more quickly and frequently, leading to more rapid model improvements [@tredennick2016; @dietze2018].
-Because the forecasts are made repeatedly through time, new data can be integrated with each new forecast cycle.
-This iterative approach to forecasting allows any changes in the state of the system that have occurred since the previous forecast to be incorporated and accounted for [@dietze2018].
+Because the forecasts are made repeatedly through time, new data can be integrated with each new iteration of the forecast cycle.
+This iterative approach to forecasting allows any changes in the state of the system that have occurred since the previous forecast to be incorporated and accounted for, thereby improving the accuracy of future forecasts [@dietze2018].
 Iterative near-term forecasting has the potential to promote rapid improvement in the state of ecological forecasting by quickly identifying how models are failing, facilitating rapid testing of improved models, and incorporating updated data so models run with the most up-to-date information on the system available.
-While use of iterative near-term forecasting is often contextualized as a management tool, this approach to model testing can also be used to improve our basic understanding of ecological systems.
-For example, alternative mechanistic models can be competed to see which model provides the best forecasts for near-term dynamics, thus providing insights into the relative importance of different processes driving dynamics of ecological systems [@dietze2018].
+While use of iterative near-term forecasting is often discussed in the context of policy and management [@clark2001; @Luo2011; @petchey2015], this approach to model testing can also be used to improve our basic understanding of ecological systems [@dietze2017].
+For example, alternative mechanistic models can be competed against each other to see which model provides the best forecasts for near-term dynamics, thus providing insights into the relative importance of different processes driving dynamics of ecological systems [@dietze2018].
 Whether deployed for basic or applied uses, iterative near-term forecasting incorporates a more dynamic interplay between models, predictions, and data that is clearly needed to improve ecological forecasting and our understanding of ecological systems more broadly.
@@
 Data to be used for iterative near-term forecasting needs to be widely accessible, which requires data to be released quickly under open licenses [@dietze2018; @vargas2017] and structured so that it can be used easily by a variety of researchers and in multiple modeling approaches [@borer2009; @strasser2011].
-Models need to be able to deal with uncertainty, in both the predictors and the predictions, to properly convey uncertainty in the resulting forecasts.
-Multiple models should be compared to assess which models are performing best and to allow for combining models to form ensemble predictions.
+Models need to be able to deal with uncertainty, in the predictors and the predictions, to properly convey uncertainty in the resulting forecasts [@diniz-filho2009].
+Multiple models should be developed, both to assess which models are performing best [@dietze2018] and to facilitate combining models to form ensemble predictions which tend to form better than single models [@araujo2007; @diniz-filho2009].
 Ensuring that data and models are regularly updated and new forecasts are made requires cyberinfrastructure to automate data processing, model fitting, prediction, model evaluation, forecast visualization, and archiving.
@@
 While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement; few of their recommendations are in widespread use in ecology today.
-We examined what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going long-term (~40 year) ecological study that collects high-frequency data on desert rodent abundances.
+We examined what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going long-term (~40 year) ecological study that collects high-frequency data on desert rodent abundances [@brown1998; @Ernest2008].
 We constructed our forecasting pipeline with the goal of being able to forecast rodent abundances and evaluate our predictions on a monthly basis.
-In this paper, we discuss our approach for creating this iterative near-term forecasting pipeline, the challenges we encountered, the tools we used, and the lessons we learned that may help others to create their own iterative forecasting systems.
+In this paper, we discuss our approach for creating this iterative near-term forecasting pipeline, the challenges we encountered, the tools we used, and the lessons we learned so that others can create their own iterative forecasting systems.
 
@@
 The Portal Project is a long-term ecological study situated in the Chihuahuan Desert (2 km north and 6.5 km east of Portal, Arizona, US).
-Researchers have been continuously collecting data at the site since 1977, including data on the abundance of rodent and plant species (monthly and twice yearly, respectively) and climactic factors such as air temperature and precipitation (daily).
+Researchers have been continuously collecting data at the site since 1977, including data on the abundance of rodent and plant species (monthly and twice yearly, respectively) and climactic factors such as air temperature and precipitation (daily)[@brown1998; @ernest2009; @ernest2016].
 The site consists of 24 50m x 50m experimental plots.
@@
 For all rodents caught during a trapping session, information on species identity, size, and reproductive condition is collected, and new individuals are given identification tags.
-We use the data from the control plots at this site, where rodent populations are not experimentally manipulated.
 This data on rodent populations is high-frequency, uses consistent trapping methodology, and has an extended time-series (469 monthly samples and counting), making this study an ideal case for near-term iterative forecasting.
@@
 Implementation of iterative forecasting requires the regular rebuilding of models with new raw data as it becomes available and the presentation of those forecasts in usable forms; in our case, this occurs monthly.
-Doing this in an efficient and maintainable way relies on developing an automated pipeline to handle the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presentation of the forecasts (Figure 1).
+Rebuilding models in an efficient and maintainable way relies on developing an automated pipeline to handle the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presenting of the forecasts (Figure 1).
 To implement the pipeline outlined in Figure 1, we used a "continuous analysis" framework [*sensu* @beaulieu2017] that automatically processes the most up-to-date data, refits the models, makes new forecasts, archives the forecasts, and updates a website with analysis of current and previous forecasts.
@@
 While each stage has unique tasks associated with it, the stages of our pipeline are also interdependent.
-Our pipeline is primarily a linear structure where outputs from one stage form the inputs for the subsequent stage ](figs/pipeline_fig.png)
+Our pipeline is primarily a linear structure where outputs from one stage form the inputs for the subsequent stage.
+All stages require code written in R [@R2017] ](figs/pipeline_fig.png).
 
@@
 Many of us have experienced the frustrations that can occur when software updates (e.g., changes in R package versions) create errors in previously functional code.
-We experienced this issue when a package one of our models relies on, `tscount` [liboschik2015], was temporarily removed from CRAN (the R package repository) and, therefore, could not be installed in the usual way.
+We experienced this issue when a package one of our models relies on, `tscount` [@liboschik2015], was temporarily removed from CRAN (the R package repository) and, therefore, could not be installed in the usual way.
 This broke our forecasting pipeline because we could no longer run models that used that package.
-To minimize issues with changes in software dependencies, we follow Beaulieu and Green's [-@beaulieu2017] recommendation to use software containers.
-Software containers are standalone packages that contain copies of everything you need to run some piece of software.
+To minimize issues with changes in software dependencies, we follow Beaulieu and Greene's [-@beaulieu2017] recommendation to use software containers.
+Software containers are standalone packages that contain copies of everything you need to run some piece of software, including the operating system.
 Once created, a software container is basically a time capsule, it contains all the software dependencies in the exact state used to develop and run the software.
 If those dependencies change (or disappear) in the wider world, they still exist, unchanged, in your container.
-We use an existing platform, Docker, to store an exact image of the complete software environment for running the forecasts.
+We use an existing platform, Docker [@Merkel2014], to store an exact image of the complete software environment for running the forecasts.
 Docker also allows a specified set of packages to be used consistently across different computer and server environments.
@@
 
-Iterative forecasting benefits from frequently updated data on the state of the system so that state changes can be quickly incorporated into new forecasts [@dietze2018].
-Frequent data collection and rapid entry and processing of data are both important for providing the most up-to-date data for forecasting.
+Iterative forecasting benefits from frequently updated data so that state changes can be quickly incorporated into new forecasts [@dietze2018].
+Frequent data collection and rapid processing are both important for providing timely forecasts.
 Since our data is collected monthly, ensuring that the models have access to the newest data requires a data latency period of less than 1 month from collection to availability for modeling.
@@
 
-New data is double-entered into Excel using the "data validation" feature.
+New data is double-entered into Microsoft Excel using the "data validation" feature.
 The two versions are then compared in an R script to control for errors in data entry.
-Quality control (QC) checks written in R using the `testthat` R package [@wickham2011] are run on the data to test for validity and consistency both within the new data (e.g.
-sexual characteristics of an animal match M/F designation) and between the new and archived data (e.g.
-species and sex are consistent for recaptures of the same animal based on tag number).
+Quality control (QC) checks written in R using the `testthat` R package [@wickham2011] are run on the data to test for validity and consistency both within the new data and between the new and archived data.
 The local use of the QC scripts to flag problematic data greatly reduces the time spent error-checking and ensures that the quality of data is consistent.
 The data is then submitted to a GitHub-based data repository.
-GitHub is a software development tool for managing computer code development, but we have also found it useful for data management.
+GitHub (https://github.com/) is a software development tool for managing computer code development, but we have also found it useful for data management.
 Changes to data can be tracked through version control, and additions and changes to the data can be monitored through pull requests (notifications that someone has a modified version of the data to contribute).
@@
 The Portal Project has a long history of making its data publicly available, which means that anyone can use it for forecasting or other projects.
-Historically the publication of the data was conducted through data papers, the most common approach in ecology; however, this approach caused years of data latency.
+Historically the publication of the data was conducted through data papers [@ernest2009,@ernest2016], the most common approach in ecology; however, this approach caused years of data latency.
 Recently, the project has switched to posting data directly to a public GitHub repository with a CC0 license (Figure 1).
-This immediate posting reduces that data latency to less than one month and, therefore, makes meaningful iterative near-term forecasting possible for not only our group but other interested parties as well.
+This immediate posting reduces that data latency to less than one month and, therefore, makes meaningful iterative near-term forecasting possible for not only our group but other interested parties, as well.
 
@@
 Once data is available, it needs to be processed into a form appropriate for modeling (Figure 1).
-In many ecological datasets, this requires not only simple data manipulation but also a good understanding of the dataset to allow data to be aggregated appropriately.
-These data manipulation steps are often conducted using custom one-off code to convert the data into the desired form [@morris2013], but this approach has several limitations.
-First, each researcher must develop their own data manipulation code, which is inefficient and can result in different decisions by different researchers about the details of data cleaning and aggregation.
-Subtle differences in data processing decisions have lead to confusion when reproducing results for the Portal data in the past.
+For many ecological datasets, this requires not only simple data manipulation but also a good understanding of the data to allow data to facilitate appropriate aggregation.
+Data manipulation steps are often conducted using custom one-off code to convert the raw data into the desired form [@morris2013], but this approach has several limitations.
+First, each researcher must develop and maintain their own data manipulation code, which is inefficient and can result in different researchers producing different versions of the data for the same task.
+Subtle differences in data processing decisions have led to confusion when reproducing results for the Portal data in the past.
 Second, this kind of code is rarely robust to changes in data structure and location.
@@
 Each forecast output file contains the date being forecast, the collection date of the data used for fitting the models, the date the forecast was made, the state variable being forecast (e.g., rodent biomass, the abundance of a species), and the forecast value and associated uncertainty of that forecast (Figure 3).
-This allows a variety of different forecasts to be stored in a common format and may serve as a useful starting point for developing a standard for storing ecological forecasts more generally.
+Our goal is to store a variety of different forecasts in a common format with the hope of this being a useful starting point for developing a standard for storing ecological forecasts more generally.
 
@@
 In addition to evaluating the actual forecasts, we also use hindcasting (forecasting on already collected data) to gain additional insight into the methods that work best for forecasting this system.
-For example a model is fit using rodent observations up to June 2005, then used to make a forecast 12 months out to May 2006.
+For example, a model is fit using rodent observations up to June 2005, then used to make a forecast 12 months out to May 2006.
 The observations of that 12 month period can immediately be used to evaluate the model.
-Hindcasting can be conducted using all months from the beginning of the study, thus allowing model comparison of large numbers of hindcasts and giving insight into which models make the best forecasts.
+Hindcasting can be conducted using data that was already been collected, thus allowing model comparison of large numbers of hindcasts and giving insight into which models make the best forecasts without needing to wait for new data to be collected.
 It can also be used to quickly evaluate new models instead of waiting for an adequate amount of data to accumulate.
@@
 
-Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2017; @harris2018](Figure 1).
+Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2017; @harris2018] (Figure 1).
 This archiving serves as a form of pre-registration for model predictions, helping facilitate unbiased interpretation of model performance.
-We explored three major repositories for archiving our forecasts: FigShare, Zenodo, and Open Science Framework.
+We explored three major repositories for archiving our forecasts: FigShare (https://figshare.com/), Zenodo (https://zenodo.org/), and Open Science Framework (https://osf.io/).
 While all three repositories allowed for easy manual submissions (i.e., a human uploading files after each forecast), automating this process was substantially more difficult.
-Various combinations of repositories, APIs and associated R packages had issues with: 1) integrating authorization with continuous integration; 2) automatically making archived files public; 3) adding new files to an existing location; or 4) automatically permanently archiving the files.
+Various combinations of repositories, APIs (i.e., interfaces that allow automated ways interacting with these websites) and associated R packages had issues with: 1) integrating authorization with continuous integration; 2) automatically making archived files public; 3) adding new files to an existing location; or 4) automatically permanently archiving the files.
 Our eventual solution was to leverage the GitHub-Zenodo integration (https://guides.github.com/activities/citable-code/) and automatically push forecasts to the GitHub repository from the CI server.
@@
 Through this process, we automatically archive every forecast made with a documented history of the archive.
-While this approach is functional because everything in the repository is archived when a new forecast is made, these archives are complicated, making it more complicated than necessary to find and access the forecasting results.
+While this approach is functional because everything in the repository is archived when a new forecast is made, these archives have complicated folder structures, making it more harder for users to find and access the forecasting results.
+Finding simpler approaches to automated archiving would increase the transparency of the forecase results and their ease of use by external researchers.
 
@@
 We designed this forecasting system with the goal of making it relatively easy to build, maintain, and extend.
-We used existing technology for both running the pipeline and building individual components, which allowed the system to be built relatively cheaply in terms of both time and money.
+We used existing technology for both running the pipeline and building individual components, which allowed is to build the system relatively cheaply in terms of both time and money.
 This included the use of tools like Docker for reproducibility, the Travis CI continuous integration system for automatically running the pipeline, Rmarkdown and `knitr` for generating the website, and the already existing integration between Github and Zenodo  to archive the forecasts.
 By using this "continuous analysis" approach [@beaulieu2017], where analyses are automatically rerun when changes are made to data, models, or associated code, we have reduced the time required by scientists to run and maintain the forecasting pipeline.
-To make the system extensible so that new models could be easily incorporated, we use a plugin-based infrastructure so that adding a new model to the system is as easy as adding a single file to the 'models' folder in our repository (Figure 3).
+To make the system extensible so that new models could be easily incorporated, we used a plugin-based infrastructure so that adding a new model to the system is as easy as adding a single file to the 'models' folder in our repository (Figure 3).
 This should substantially lower the barriers to other scientists contributing models to this forecasting effort.
-We also automatically archive the resulting forecasts publicly so that, as new data is collected, the performance of these forecasts can be assessed by both us and other researchers.
-This serves as a form of "pre-registration" by providing a quantitative record of the forecast before the data being predicted were collected.
+We also automatically archive the resulting forecasts publicly so thatthe performance of these forecasts can be assessed by both us and other researchers as new data is collected.
+This serves as a form of pre-registration of expected results () by providing a quantitative record of the forecast before the data being predicted were collected.
 
@@
 Overall, we found existing technology to be sufficient to the task, but it required greater expertise and a greater investment of time than is ideal.
-Tool development to reduce the effort required for scientists to set up their own short-term forecasting systems would clearly be useful, but our efforts show that it is still currently possible for scientists using existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems.
-By developing these systems so that they already produce automated iterative forecasts, it will be easier to convert these systems into fully operationalized forecasting systems that are relied on for decision making [@dietze2018, other paper by operationalization co-authors].
+Tool development to reduce the effort required for scientists to set up their own short-term forecasting systems would clearly be useful, but our efforts show that it is possible for scientists to use existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems.
+By developing these systems so that they are producing automated iterative forecasts, it will be easier to convert these systems into fully operationalized forecasting systems that are relied on for decision making [@dietze2018, other paper by operationalization co-authors].
 
@@
 This provided a shared set of knowledge and vocabulary that actively facilitated interdisciplinary interactions.
-Given the current state of existing tools for forecasting, forecasting teams will need people with significant experience in working with continuous integration and APIs, which means interdisciplinary teams will generally be required for creating these pipelines until tool development improves.
+Given the current state of tools for forecasting, forecasting teams will need people with significant experience in working with continuous integration and APIs, which means interdisciplinary teams will generally be required for creating these pipelines until tool development improves.
 
@@
 The forecasting infrastructure does the time-consuming work of data processing, data integration, and model assessment, allowing new research to focus on the models being developed and the inferences about the system that can be drawn from them [@dietze2018].
-We plan to use this pipeline to drive future research into understanding the processes that govern the dynamics of individual populations and the community as a whole.
+We plan to use this pipeline to drive future research into understanding the processes that govern the dynamics of individual populations and the community as a whole at this long-term study site.
 By regularly running different models for population and community dynamics, a near-term iterative pipeline such as ours should also make it possible to rapidly detect changes in how the system is operating, which should allow the rapid identification of ecological transitions or even possibly allow them to be prevented [@carpenter example].

### 2018-02-01 manuscript/portal_iterative_forecasting_ms.Rmd
Edits from Hao

@@
 
-Forecasting the future state of ecological systems is important for management, conservation, and evaluation of our fundamental understanding of ecology [@clark2001; @tallis2006; @diaz2015; @dietze2017].
-Since Clark et al.
-[-clark2001] called for a more central role of forecasting in ecology, an increasing number of ecological forecasts are being published.
+Forecasting the future state of ecological systems is important for management, conservation, and evaluation of our basic understanding of how ecology operates [@clark2001; @tallis2006; @diaz2015; @dietze2017].
+In 2001, Clark et al.
+[-clark2001] called for a more central role of forecasting in ecology, and since then an increasing number of ecological forecasts are being published.
 However, most of these forecasts are made once, published, and never assessed or updated.
@@
 This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions.
+More regular updating and assessment will advance ecological forecasting as a field by speeding up the identification of the best models for forecasting specific issues in specific ecosystems and contribute overall to our understanding of how to best design forecasting approaches for ecology.
 For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].
@@
 Stages of the forecasting pipeline.
-To go from raw data to forecast presentation involves a number of stages, each of which required unique tasks, tools and infrastructure.
+To go from raw data to forecast presentation involves a number of stages, each of which requires unique tasks, tools and infrastructure.
 While each stage has unique tasks associated with it, the stages of our pipeline are also interdependent.
 Our pipeline is primarily a linear structure where outputs from one stage form the inputs for the subsequent stage.
-All stages require code written in R [@R2017] ](figs/pipeline_fig.png).
+All stages require code written in R.
+](figs/pipeline_fig.png).
 
@@
 
-A core component of iterative near-term forecasting is the regular rerunning of the forecasting pipeline.
-This can be conducted manually---with a human making sure all code is run and all tables and files are updated---or automatically by having the computer conduct those tasks.
+A core aspect of iterative near-term forecasting is the regular rerunning of the forecasting pipeline.
 We chose to have the computer run our pipeline and employed "continuous analysis" [*sensu* @beaulieu2017] to drive the automation of both the full pipeline and a number of its individual components.
@@
 CI combines computing environments for running code with monitoring systems to identify changes in data or code.
-Essentially, CI is a computer helper whose job is to watch the pipeline and, when it sees a change in the code or data, it runs all the tasks needed to ensure that the forecasting pipeline runs from beginning to end.
+Essentially, CI is a computer helper whose job is to watch the pipeline and, when it sees a change in the code or data, it runs all the computer scripts needed to ensure that the forecasting pipeline runs from beginning to end.
 This is useful for iterative near-term forecasting because it does not rely on humans to remember to create forecasts when new models or data are added.
-These tools are common in the area of software development where they are used to automate software testing and integrate work by multiple software developers working on the same software.
+These tools are common in the area of software development where they are used to automate software testing and integrate work by multiple developers working on the same code base.
 However, these tools can be used for any computational task that needs to be regularly repeated or run after changes to code or data [@beaulieu2017].
-Because of the widespread use of CI in software development, several CI services already exist.
-Our forecasting pipeline currently runs on a publicly available continuous integration service (Travis CI; https://travis-ci.org/) that is free for open source projects (up to a limited amount of computing time); alternative services that can be used to run code on local or cloud-based computational infrastructure are available, such as Drone (http://try.drone.io/) [@beaulieu2017].
+Our forecasting pipeline currently runs on a publicly available continuous integration service (Travis CI; https://travis-ci.org/) that is free for open source projects (up to a limited amount of computing time).
+Because of the widespread use of CI in software development, alternative services that can run code on local or cloud-based computational infrastructure also exist (e.g., Drone (http://try.drone.io/)) [@beaulieu2017].
 As detailed below, we use CI to quality check data, test code using "unit tests" [@wilson2014], build models, make forecasts, and publicly present and archive the results (Figure 2).
@@
 This broke our forecasting pipeline because we could no longer run models that used that package.
-To minimize issues with changes in software dependencies, we follow Beaulieu and Greene's [-@beaulieu2017] recommendation to use software containers.
+To make our pipeline robust to changes in external software dependencies, we follow Beaulieu and Greene's [-@beaulieu2017] recommendation to use software containers.
 Software containers are standalone packages that contain copies of everything you need to run some piece of software, including the operating system.
@@
 tidyverse) pre-installed  [@boettiger2017].
-We use add our code and dependencies to this existing Rocker image to create a software container for our forecasting pipeline.
+We add our code and dependencies to this existing Rocker image to create a software container for our forecasting pipeline.
 In combination, the automated running of the pipeline (continuous integration) and the guarantee it will not stop working unexpectedly due to software dependencies (via a software container) allows continuous analysis to serve as the glue that connects all stages of the forecasting pipeline.
@@
 The two versions are then compared in an R script to control for errors in data entry.
-Quality control (QC) checks written in R using the `testthat` R package [@wickham2011] are run on the data to test for validity and consistency both within the new data and between the new and archived data.
+Quality control (QC) checks using the `testthat` R package [@wickham2011] are run on the data to test for validity and consistency both within the new data and between the new and archived data.
 The local use of the QC scripts to flag problematic data greatly reduces the time spent error-checking and ensures that the quality of data is consistent.
-The data is then submitted to a GitHub-based data repository.
+The data is then uploaded to the GitHub-based Portal Data repository.
 GitHub (https://github.com/) is a software development tool for managing computer code development, but we have also found it useful for data management.
-Changes to data can be tracked through version control, and additions and changes to the data can be monitored through pull requests (notifications that someone has a modified version of the data to contribute).
+On GitHub, changes to data can be tracked through the Git version control system which logs all changes made to any files in the repository.
+All updates to data are processed through "pull requests", which are notifications that someone has a modified version of the data to contribute.
 QA/QC checks are automatically rerun on the submitted data using continuous integration to ensure that these checks have been run and that no avoidable errors reach the official version of the dataset.
@@
 Historically the publication of the data was conducted through data papers [@ernest2009,@ernest2016], the most common approach in ecology; however, this approach caused years of data latency.
-Recently, the project has switched to posting data directly to a public GitHub repository with a CC0 license (Figure 1).
+Recently, the project has switched to posting data directly to a public GitHub repository with a CC0 (i.e.
+no rights reserved) license (Figure 1).
 This immediate posting reduces that data latency to less than one month and, therefore, makes meaningful iterative near-term forecasting possible for not only our group but other interested parties, as well.
@@
 
-Ideally, iterative near-term forecasting involves regularly refitting a variety of different models (Figure 1).
-New models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach.
+Iterative near-term forecasting involves regularly refitting a variety of different models (Figure 1).
+Additionally, any new models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach.
 We use CI to refit the models and make new forecasts each time the modeling code changes and when new data becomes available (Figure 2).
 We use a plugin infrastructure to allow new models to be easily added to the system.
-Each model is a script that takes in standardized inputs and returns standardized outputs (Figure 3).
-All models are run by the main forecasting code at each update, and the standardized outputs are combined to store the results of the different models' forecasts.
-A weighted ensemble model is then added with weights based on how well individual models fit the training data.
-This plugin infrastructure allows flexibility in all aspects of the modeling process, making it easier to explore new modeling approaches, and allows new models that fit the data well to immediately improve the ensemble forecast.
+Our approach is to treat each model as interchangable black boxes - we feed them all the same inputs and we expect them to generate the same structure for model outputs (Figure 3).
+Each iteration of the forecasting cycle, all models are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts.
+This structure allows weighted ensemble models to be produced based on how well individual models fit the training data.
+Our hope is that by requiring all models to take the same input (though models may differ in how they use that input) and provide the same outputs, we've gained the capacity to add and compare very different types of models that we may not currently be able to foresee and design around.
+Our infrastrcture should allow easy addition of more complicated models and more sophisticated approaches to creating ensembles even though to date our models and ensemble-creating approaches are extremely basic.
 
 ![Figure 3.
-Demonstration of plugin infrastructure where each model script (represented here as AR1.R) uses data provided by the core forecasting code (represented here as rodent.csv) and returns its forecasts in a predefined structure that is consistent across models (represented here as 2017_12_08forecasts.csv) to allow cross-model evaluations.](figs/plugin_fig.gif)
+Demonstration of plugin infrastructure.
+All model scripts (represented here by the example AR1.R) are housed in a single folder.
+Each model script uses data provided by the core forecasting code (represented here by rodent.csv) and returns its forecast outputs in a predefined structure that is consistent across models (represented here by the example 2017_12_08forecasts.csv).
+Outputs from all models run on a particular date are combined into the same file (i.e.
+2017_12_08forecasts.csv) to allow cross-model evaluations.
+Model outpus files are housed in a folder containing all forecast outputs from all previous dates to facilitate our arhciving process and and forecast assessment activities ](figs/plugin_fig.gif)
 
-To allow flexibility in what the models predict, we designed a forecast output format that allows us to store relatively generic forecasts.
+We designed our forecast output format to to store relatively generic forecasts information to allow flexibility in what the models predict (i.e.
+species abundances, biomass).
 Each forecast output file contains the date being forecast, the collection date of the data used for fitting the models, the date the forecast was made, the state variable being forecast (e.g., rodent biomass, the abundance of a species), and the forecast value and associated uncertainty of that forecast (Figure 3).
@@
 
-Forecasts are currently evaluated using root mean square error (RMSE) to evaluate point forecasts and coverage to evaluate uncertainty.
+Our forecasts are currently evaluated using root mean square error (RMSE) to evaluate point forecasts and coverage to evaluate uncertainty.
 We plan to add additional metrics in the future.
@@
 Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2017; @harris2018] (Figure 1).
-This archiving serves as a form of pre-registration for model predictions, helping facilitate unbiased interpretation of model performance.
+Archiving serves as a form of pre-registration for model predictions, helping facilitate unbiased interpretation of model performance.
+We wanted our archives to not only be accessible to others but to also be a permanent record whose continued existence was not dependent on any particular researcher or funding stream.
+Storing forecast outputs in our forecasting GitHub repository was not sufficient for archival purposes because repositories can be deleted.
 We explored three major repositories for archiving our forecasts: FigShare (https://figshare.com/), Zenodo (https://zenodo.org/), and Open Science Framework (https://osf.io/).
@@
 Tool development to reduce the effort required for scientists to set up their own short-term forecasting systems would clearly be useful, but our efforts show that it is possible for scientists to use existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems.
-By developing these systems so that they are producing automated iterative forecasts, it will be easier to convert these systems into fully operationalized forecasting systems that are relied on for decision making [@dietze2018, other paper by operationalization co-authors].
+Forecasting systems such as ours can begin the process of making iterative forecasts while we wait for more integrated and easier to deploy platforms for iterative forecasting to be developed.
+Once more integrated platforms exist, it will be easier to convert current systems into fully operationalized forecasting systems that are relied on for decision making [@dietze2018, other paper by operationalization co-authors].
 

### 2018-02-02 manuscript/portal_iterative_forecasting_ms.Rmd
Fix typo

@@
 Models need to be able to deal with uncertainty, in the predictors and the predictions, to properly convey uncertainty in the resulting forecasts [@diniz-filho2009].
-Multiple models should be developed, both to assess which models are performing best [@dietze2018] and to facilitate combining models to form ensemble predictions which tend to form better than single models [@araujo2007; @diniz-filho2009].
+Multiple models should be developed, both to assess which models are performing best [@dietze2018] and to facilitate combining models to form ensemble predictions which tend to perform better than single models [@araujo2007; @diniz-filho2009].
 Ensuring that data and models are regularly updated and new forecasts are made requires cyberinfrastructure to automate data processing, model fitting, prediction, model evaluation, forecast visualization, and archiving.

### 2018-02-02 manuscript/portal_iterative_forecasting_ms.Rmd
Make capitalization of Ernest 2008 consistent

@@
 While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement; few of their recommendations are in widespread use in ecology today.
-We examined what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going long-term (~40 year) ecological study that collects high-frequency data on desert rodent abundances [@brown1998; @Ernest2008].
+We examined what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going long-term (~40 year) ecological study that collects high-frequency data on desert rodent abundances [@brown1998; @ernest2008].
 We constructed our forecasting pipeline with the goal of being able to forecast rodent abundances and evaluate our predictions on a monthly basis.

### 2018-02-02 manuscript/portal_iterative_forecasting_ms.Rmd
Make sure citations aren't treated as a link

@@
 The Portal Project is a long-term ecological study situated in the Chihuahuan Desert (2 km north and 6.5 km east of Portal, Arizona, US).
-Researchers have been continuously collecting data at the site since 1977, including data on the abundance of rodent and plant species (monthly and twice yearly, respectively) and climactic factors such as air temperature and precipitation (daily)[@brown1998; @ernest2009; @ernest2016].
+Researchers have been continuously collecting data at the site since 1977, including data on the abundance of rodent and plant species (monthly and twice yearly, respectively) and climactic factors such as air temperature and precipitation (daily) [@brown1998; @ernest2009; @ernest2016].
 The site consists of 24 50m x 50m experimental plots.

### 2018-02-02 manuscript/portal_iterative_forecasting_ms.Rmd
Add link to PortalData repo
Fix typo

@@
 The local use of the QC scripts to flag problematic data greatly reduces the time spent error-checking and ensures that the quality of data is consistent.
-The data is then uploaded to the GitHub-based Portal Data repository.
+The data is then uploaded to the GitHub-based Portal Data repository (https://github.com/weecology/PortalData).
 GitHub (https://github.com/) is a software development tool for managing computer code development, but we have also found it useful for data management.
@@
 Once data is available, it needs to be processed into a form appropriate for modeling (Figure 1).
-For many ecological datasets, this requires not only simple data manipulation but also a good understanding of the data to allow data to facilitate appropriate aggregation.
+For many ecological datasets, this requires not only simple data manipulation but also a good understanding of the data to facilitate appropriate aggregation.
 Data manipulation steps are often conducted using custom one-off code to convert the raw data into the desired form [@morris2013], but this approach has several limitations.

### 2018-02-02 manuscript/portal_iterative_forecasting_ms.Rmd
Revise changes to modeling and forecasting section
This section had been overly softened in ways that make things less clear.

@@
 Iterative near-term forecasting involves regularly refitting a variety of different models (Figure 1).
-Additionally, any new models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach.
+Ideally, new models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach.
 We use CI to refit the models and make new forecasts each time the modeling code changes and when new data becomes available (Figure 2).
 We use a plugin infrastructure to allow new models to be easily added to the system.
-Our approach is to treat each model as interchangable black boxes - we feed them all the same inputs and we expect them to generate the same structure for model outputs (Figure 3).
-Each iteration of the forecasting cycle, all models are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts.
-This structure allows weighted ensemble models to be produced based on how well individual models fit the training data.
-Our hope is that by requiring all models to take the same input (though models may differ in how they use that input) and provide the same outputs, we've gained the capacity to add and compare very different types of models that we may not currently be able to foresee and design around.
-Our infrastrcture should allow easy addition of more complicated models and more sophisticated approaches to creating ensembles even though to date our models and ensemble-creating approaches are extremely basic.
+This approach treats each model as an interchangable black boxes - all models have access to the same input data and generate the same structure for model outputs (Figure 3).
+Each iteration of the forecasting cycle, all models that are present are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts.
+A weighted ensemble model is then added with weights based on how well individual models fit the training data.
+This plugin infrastructure makes it easy to add and compare very different types of models, from the basic time-series approaches currently implemented to complex state-space and machine learning models.
+As long as a model script can load the provided data and produce the appropriate output it will be run and its results incorporated into the rest of the forecasting system.
 
@@
 
-We designed our forecast output format to to store relatively generic forecasts information to allow flexibility in what the models predict (i.e.
-species abundances, biomass).
+In addition to flexibility in model internals, we also wanted to support flexibility in what the models predict.
+Allowing models to make forecasts for system properties ranging from individual species' population abundances to total community biomass facilitates exploration of differences in forecastability across different aspects of ecological systems.
+We designed a forecast output format to support this.
 Each forecast output file contains the date being forecast, the collection date of the data used for fitting the models, the date the forecast was made, the state variable being forecast (e.g., rodent biomass, the abundance of a species), and the forecast value and associated uncertainty of that forecast (Figure 3).
-Our goal is to store a variety of different forecasts in a common format with the hope of this being a useful starting point for developing a standard for storing ecological forecasts more generally.
+This allows us to store a variety of different forecasts in a common format and may serve as a useful starting point for developing a standard for storing ecological forecasts more generally.
 
-Our forecasts are currently evaluated using root mean square error (RMSE) to evaluate point forecasts and coverage to evaluate uncertainty.
+Forecasts are currently evaluated using root mean square error (RMSE) to evaluate point forecasts and coverage to evaluate uncertainty.
 We plan to add additional metrics in the future.

### 2018-02-02 manuscript/portal_iterative_forecasting_ms.Rmd
Cleanup edits in Discussion

@@
 We designed this forecasting system with the goal of making it relatively easy to build, maintain, and extend.
-We used existing technology for both running the pipeline and building individual components, which allowed is to build the system relatively cheaply in terms of both time and money.
+We used existing technology for both running the pipeline and building individual components, which allowed us to build the system relatively cheaply in terms of both time and money.
 This included the use of tools like Docker for reproducibility, the Travis CI continuous integration system for automatically running the pipeline, Rmarkdown and `knitr` for generating the website, and the already existing integration between Github and Zenodo  to archive the forecasts.
@@
 This should substantially lower the barriers to other scientists contributing models to this forecasting effort.
-We also automatically archive the resulting forecasts publicly so thatthe performance of these forecasts can be assessed by both us and other researchers as new data is collected.
-This serves as a form of pre-registration of expected results () by providing a quantitative record of the forecast before the data being predicted were collected.
+We also automatically archive the resulting forecasts publicly so that the performance of these forecasts can be assessed by both us and other researchers as new data is collected.
+This serves as a form of pre-registration by providing a quantitative record of the forecast before the data being predicted were collected.
 
@@
 Tool development to reduce the effort required for scientists to set up their own short-term forecasting systems would clearly be useful, but our efforts show that it is possible for scientists to use existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems.
-Forecasting systems such as ours can begin the process of making iterative forecasts while we wait for more integrated and easier to deploy platforms for iterative forecasting to be developed.
-Once more integrated platforms exist, it will be easier to convert current systems into fully operationalized forecasting systems that are relied on for decision making [@dietze2018, other paper by operationalization co-authors].
+Developing systems that  produce automated iterative forecasts will make it easier to convert research focused forecasts into fully operationalized forecasting systems that are relied on for decision making [@dietze2018, other paper by operationalization co-authors].
 
@@
 The forecasting infrastructure does the time-consuming work of data processing, data integration, and model assessment, allowing new research to focus on the models being developed and the inferences about the system that can be drawn from them [@dietze2018].
-We plan to use this pipeline to drive future research into understanding the processes that govern the dynamics of individual populations and the community as a whole at this long-term study site.
+We plan to use this pipeline to drive future research into understanding the processes that govern the dynamics of individual populations and the community as a whole.
 By regularly running different models for population and community dynamics, a near-term iterative pipeline such as ours should also make it possible to rapidly detect changes in how the system is operating, which should allow the rapid identification of ecological transitions or even possibly allow them to be prevented [@carpenter example].

### 2018-02-02 report.Rmd
Load getURL directly from RCurl
report.Rmd was relying on sourcing another file to import RCurl. This broke
on some platforms and lead to confusing code.

@@
 # get new moon number associated with this period
-newmoons_table = read.csv(text = getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
+newmoons_table = read.csv(text = RCurl::getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
 newmoons_table$censusdate = as.Date(newmoons_table$censusdate,format='%Y-%m-%d')

### 2018-02-02 report.Rmd
Allow most recent census date to be properly compared to prediction dates
In order to effectively use `<` with dates this value needs to be a single
date rather than a date embedded in a data frame.

--- 2017-12-01 report.Rmd
@@
 # get new moon number associated with this period
-newmoons_table = read.csv(text = getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
+newmoons_table = read.csv(text = RCurl::getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
 newmoons_table$censusdate = as.Date(newmoons_table$censusdate,format='%Y-%m-%d')
 moon_code = dplyr::filter(newmoons_table,period==most_recent_period) %>% select(newmoonnumber) %>% as.integer()
-most_recent_date = dplyr::filter(newmoons_table,period==most_recent_period) %>% dplyr::select(censusdate)
+most_recent_date = dplyr::filter(newmoons_table,period==most_recent_period) %>% dplyr::select(censusdate) %>% pull(censusdate)
 

### 2018-02-02 report.Rmd
Update report.Rmd
select & pull are redundant
adding dplyr namespace for pull

--- 2017-12-01 report.Rmd
@@
 # get new moon number associated with this period
-newmoons_table = read.csv(text = getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
+newmoons_table = read.csv(text = RCurl::getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
 newmoons_table$censusdate = as.Date(newmoons_table$censusdate,format='%Y-%m-%d')
 moon_code = dplyr::filter(newmoons_table,period==most_recent_period) %>% select(newmoonnumber) %>% as.integer()
-most_recent_date = dplyr::filter(newmoons_table,period==most_recent_period) %>% dplyr::select(censusdate)
+most_recent_date = dplyr::filter(newmoons_table,period==most_recent_period) %>% dplyr::pull(censusdate)
 

### 2018-02-02 index.Rmd

@@
 source("tools/forecast_tools.R")
+source("tools/model_functions.R")
 ```
@@
 for_data$model = "ensemble"
-for_data$forecastdate = as.Date(paste(for_data$forecastyear,for_data$forecastmonth,'15',sep='-'))
+# get dates for future new moons for plotting
+future_moons = get_moon_data() %>% get_future_moons() %>% select(c("newmoonnumber","newmoondate"))
+for_data = inner_join(for_data, future_moons, by = c("newmoonnumber" = "newmoonnumber"))
 
@@
              for_data = for_data,
-             for_date_col_name = "forecastdate",
+             for_date_col_name = "newmoondate",
              for_val_col_name = "estimate",
@@
                for_data = for_data,
-               for_date_col_name = "forecastdate",
+               for_date_col_name = "newmoondate",
                for_val_col_name = "estimate",

### 2018-02-02 evaluation.Rmd

@@
 species_names = portalr::FullPath('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
-  read.csv(stringsAsFactors=FALSE) %>%
+  read.csv(stringsAsFactors=FALSE, na.strings = '') %>%
   select(species = speciescode, full_species_name = scientificname)
 
-#Fix neotoma, and add a total entry
-species_names$species[species_names$full_species_name=='Neotoma albigula'] <- "NA"
+#add a total entry
 species_names = species_names %>%
-  mutate(species = ifelse(full_species_name=='Neotoma albigula', 'NA', species)) %>%
   add_row(species='total', full_species_name='Total Rodents')
-
-```
-
-## How have the models recently?
-
-This graph shows the forecast error of some recent forecasts for the most recent sampling dates.
-This is shown for all model types in the columns and several common rodent species (and total rodents) in the rows.
-These forecasts are for the entire site.
-```{r recent_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=15, fig.height=15}
-# The length of time to look backward
-prior_new_moons = 12
 
@@
 new_moons = read.csv(new_moon_file)
-most_recent_newmoon = max(new_moons$newmoonnumber)
-earliest_newmoon_to_graph = most_recent_newmoon - prior_new_moons
 
+# Setup rodent observations
 species_abundance = portalr::abundance(shape='flat', level='Site') %>%
@@
 
+```
+
+## How have the models recently?
+
+This graph shows the forecast error of some recent forecasts for the most recent sampling dates.
+This is shown for all model types in the columns and several common rodent species (and total rodents) in the rows.
+These forecasts are for the entire site.
+```{r recent_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=15, fig.height=15}
+# The length of time to look backward
+prior_new_moons = 12
+
+most_recent_newmoon = max(new_moons$newmoonnumber)
+earliest_newmoon_to_graph = most_recent_newmoon - prior_new_moons
+
 observation_data = species_abundance %>%
@@
 
-#Get the all the forecasts made during observation period
+#Get all the recent forecasts
 forecast_data = compile_forecasts(use_hindcasts = FALSE) %>%
@@
 
-# Keep 1 forecast per unique initial_newmoon, when there are multiple forecasts
+# Keep 1 forecast per unique initial_newmoon. When there are multiple forecasts
 # using the same initial_newmoon, use the first one.
@@
 getPalette = colorRampPalette(RColorBrewer::brewer.pal(9, "Set1"))
-#Randomize pallette so nearby regions are not similar colors
 set.seed(2)
@@
 
-new_moon_file = portalr::FullPath('PortalData/Rodents/moon_dates.csv', '~')
-new_moons = read.csv(new_moon_file)
-
-species_abundance = portalr::abundance(shape='flat', level='Site') %>%
-  rename(actual = abundance) %>%
-  mutate(level='All',currency='abundance') %>%
-  left_join(new_moons, by='period') %>%
-  filter(species %in% species_of_interest)
-
-total_abundance = portalr::abundance(shape='flat', level='Site') %>%
-  group_by(period) %>%
-  summarise(actual=sum(abundance)) %>%
-  ungroup() %>%
-  mutate(level='All',currency='abundance',species='total') %>%
-  left_join(new_moons, by='period')
-
 observation_data = species_abundance %>%

### 2018-02-02 manuscript/portal_iterative_forecasting_ms.Rmd
Vertical figure 1 a and b, reorder figure references in ms

@@
 Implementation of iterative forecasting requires the regular rebuilding of models with new raw data as it becomes available and the presentation of those forecasts in usable forms; in our case, this occurs monthly.
-Rebuilding models in an efficient and maintainable way relies on developing an automated pipeline to handle the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presenting of the forecasts (Figure 1).
-To implement the pipeline outlined in Figure 1, we used a "continuous analysis" framework [*sensu* @beaulieu2017] that automatically processes the most up-to-date data, refits the models, makes new forecasts, archives the forecasts, and updates a website with analysis of current and previous forecasts.
+Rebuilding models in an efficient and maintainable way relies on developing an automated pipeline to handle the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presenting of the forecasts (Figure 1a).
+To implement the pipeline outlined in Figure 1a, we used a "continuous analysis" framework [*sensu* @beaulieu2017] that automatically processes the most up-to-date data, refits the models, makes new forecasts, archives the forecasts, and updates a website with analysis of current and previous forecasts.
 In this section we describe our approach to streamlining and automating the multiple components of the forecasting pipeline and the tools and infrastructure we employed to execute each stage of the pipeline.
@@
 ![Figure 1.
-Stages of the forecasting pipeline.
+a) Stages of the forecasting pipeline.
 To go from raw data to forecast presentation involves a number of stages, each of which requires unique tasks, tools and infrastructure.
@@
 All stages require code written in R.
-](figs/pipeline_fig.png).
+b) Continuous integration system.
+Each box denotes the core infrastructure used for each stage of the forecasting pipeline denoted in Figure 1a.
+Continuous integration, denoted here with the Travis icon (a girl wearing safety glasses and hardhat), plugs into our pipeline at every stage.
+Travis triggers the code involved in major events that involve integration across stages of the pipeline, such as taking the output from the forecasting stage (purple box) to create an updated presentation (rose box).
+Travis also runs tasks within a stage, such as conducting tests to make sure code changes have not introduced errors (icons on arrows originating and ending on the same box).
+](figs/Figure1.png)
 
@@
 Because of the widespread use of CI in software development, alternative services that can run code on local or cloud-based computational infrastructure also exist (e.g., Drone (http://try.drone.io/)) [@beaulieu2017].
-As detailed below, we use CI to quality check data, test code using "unit tests" [@wilson2014], build models, make forecasts, and publicly present and archive the results (Figure 2).
-
-![Figure 2.
-Continuous integration system.
-Each box denotes the core infrastructure used for each stage of the forecasting pipeline denoted in Figure 1.
-Continuous integration, denoted here with the Travis icon (a girl wearing safety glasses and hardhat), plugs into our pipeline at every stage.
-Travis triggers the code involved in major events that involve integration across stages of the pipeline, such as taking the output from the forecasting stage (purple box) to create an updated presentation (rose box).
-Travis also runs tasks within a stage, such as conducting tests to make sure code changes have not introduced errors (icons on arrows originating and ending on the same box)](figs/contanaly_fig.png)
+As detailed below, we use CI to quality check data, test code using "unit tests" [@wilson2014], build models, make forecasts, and publicly present and archive the results (Figure 1b).
 
@@
 Since our data is collected monthly, ensuring that the models have access to the newest data requires a data latency period of less than 1 month from collection to availability for modeling.
-To accomplish this, we automated components of the data processing and quality assurance/quality control (QA/QC) process to reduce the time needed to add new data to the database (Figures 1 and 2).
+To accomplish this, we automated components of the data processing and quality assurance/quality control (QA/QC) process to reduce the time needed to add new data to the database (Figure 1).
 
@@
 To avoid these issues for the Portal Project data, the Portal team has been developing an R package (portalR; http://github.com/weecology/portalr) for acquiring the data and handling common data cleaning and aggregation tasks.
-As a result, our modeling and forecasting code only needs to install this package and run the data manipulation and summary functions to get the appropriate data (Figure 2).
+As a result, our modeling and forecasting code only needs to install this package and run the data manipulation and summary functions to get the appropriate data (Figure 1b).
 The package undergoes thorough automated unit tests to ensure that data manipulations are achieving the desired results.
@@
 Ideally, new models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach.
-We use CI to refit the models and make new forecasts each time the modeling code changes and when new data becomes available (Figure 2).
+We use CI to refit the models and make new forecasts each time the modeling code changes and when new data becomes available (Figure 1b).
 We use a plugin infrastructure to allow new models to be easily added to the system.
-This approach treats each model as an interchangable black boxes - all models have access to the same input data and generate the same structure for model outputs (Figure 3).
+This approach treats each model as an interchangable black boxes - all models have access to the same input data and generate the same structure for model outputs (Figure 2).
 Each iteration of the forecasting cycle, all models that are present are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts.
@@
 
-![Figure 3.
+![Figure 2.
 Demonstration of plugin infrastructure.
@@
 We designed a forecast output format to support this.
-Each forecast output file contains the date being forecast, the collection date of the data used for fitting the models, the date the forecast was made, the state variable being forecast (e.g., rodent biomass, the abundance of a species), and the forecast value and associated uncertainty of that forecast (Figure 3).
+Each forecast output file contains the date being forecast, the collection date of the data used for fitting the models, the date the forecast was made, the state variable being forecast (e.g., rodent biomass, the abundance of a species), and the forecast value and associated uncertainty of that forecast (Figure 2).
 This allows us to store a variety of different forecasts in a common format and may serve as a useful starting point for developing a standard for storing ecological forecasts more generally.
@@
 
-In addition to archiving the results of each forecast, we present them on a website that displays monthly rodent forecasts, model evaluation metrics, monthly reports, and information about the study site (Figure 4; [http://portal.naturecast.org](http://portal.naturecast.org/)).
+In addition to archiving the results of each forecast, we present them on a website that displays monthly rodent forecasts, model evaluation metrics, monthly reports, and information about the study site (Figure 3; [http://portal.naturecast.org](http://portal.naturecast.org/)).
 The website includes a graphical presentation of the most recent month's forecasts (including uncertainty) and compares the latest data to the previous forecasts.
@@
 
-![Figure 4.
+![Figure 3.
 Screen capture of the homepage of the Portal Forecasting website; http://portal.naturecast.org.
@@
 
-Following the recommendations of Dietze et al [-dietze2018], we developed an automated iterative forecasting system (Figures 1 and 2) to support repeated forecasting of an ecological system.
+Following the recommendations of Dietze et al [-dietze2018], we developed an automated iterative forecasting system (Figure 1) to support repeated forecasting of an ecological system.
 Our forecasting system automatically acquires and processes the newest data, refits the models, makes new forecasts, publicly archives those forecasts, and presents both the current forecast and information on how previous forecasts performed.
@@
 By using this "continuous analysis" approach [@beaulieu2017], where analyses are automatically rerun when changes are made to data, models, or associated code, we have reduced the time required by scientists to run and maintain the forecasting pipeline.
-To make the system extensible so that new models could be easily incorporated, we used a plugin-based infrastructure so that adding a new model to the system is as easy as adding a single file to the 'models' folder in our repository (Figure 3).
+To make the system extensible so that new models could be easily incorporated, we used a plugin-based infrastructure so that adding a new model to the system is as easy as adding a single file to the 'models' folder in our repository (Figure 2).
 This should substantially lower the barriers to other scientists contributing models to this forecasting effort.

### 2018-02-04 manuscript/portal_iterative_forecasting_ms.Rmd

@@
 In 2001, Clark et al.
-[-clark2001] called for a more central role of forecasting in ecology, and since then an increasing number of ecological forecasts are being published.
+[-@clark2001] called for a more central role of forecasting in ecology, and since then an increasing number of ecological forecasts are being published.
 However, most of these forecasts are made once, published, and never assessed or updated.
@@
 
-With the goal of making ecological forecasting more dynamic and responsive, Dietze et al [-dietze2018] recently called for an increase in iterative near-term forecasting.
+With the goal of making ecological forecasting more dynamic and responsive, Dietze et al [-@dietze2018] recently called for an increase in iterative near-term forecasting.
 Iterative near-term forecasting means making forecasts for the near future and making these forecasts repeatedly through a cycle of forecast evaluation, integration of updated data, and generation of new forecasts.
@@
 Iterative near-term forecasting has the potential to promote rapid improvement in the state of ecological forecasting by quickly identifying how models are failing, facilitating rapid testing of improved models, and incorporating updated data so models run with the most up-to-date information on the system available.
-While use of iterative near-term forecasting is often discussed in the context of policy and management [@clark2001; @Luo2011; @petchey2015], this approach to model testing can also be used to improve our basic understanding of ecological systems [@dietze2017].
+While use of iterative near-term forecasting is often discussed in the context of policy and management [@clark2001; @luo2011; @petchey2015], this approach to model testing can also be used to improve our basic understanding of ecological systems [@dietze2017].
 For example, alternative mechanistic models can be competed against each other to see which model provides the best forecasts for near-term dynamics, thus providing insights into the relative importance of different processes driving dynamics of ecological systems [@dietze2018].
@@
 
-Because iterative near-term forecasting requires a dynamic interplay of models, predictions, and data, Dietze et al [-dietze2018] highlight approaches to data management, model construction and evaluation, and cyberinfrastructure that are necessary to effectively implement this type of forecasting (Box 1).
+Because iterative near-term forecasting requires a dynamic interplay of models, predictions, and data, Dietze et al [-@dietze2018] highlight approaches to data management, model construction and evaluation, and cyberinfrastructure that are necessary to effectively implement this type of forecasting (Box 1).
 Data to be used for iterative near-term forecasting needs to be widely accessible, which requires data to be released quickly under open licenses [@dietze2018; @vargas2017] and structured so that it can be used easily by a variety of researchers and in multiple modeling approaches [@borer2009; @strasser2011].
@@
 If those dependencies change (or disappear) in the wider world, they still exist, unchanged, in your container.
-We use an existing platform, Docker [@Merkel2014], to store an exact image of the complete software environment for running the forecasts.
+We use an existing platform, Docker [@merkel2014], to store an exact image of the complete software environment for running the forecasts.
 Docker also allows a specified set of packages to be used consistently across different computer and server environments.
@@
 
-Following the recommendations of Dietze et al [-dietze2018], we developed an automated iterative forecasting system (Figure 1) to support repeated forecasting of an ecological system.
+Following the recommendations of Dietze et al [-@dietze2018], we developed an automated iterative forecasting system (Figure 1) to support repeated forecasting of an ecological system.
 Our forecasting system automatically acquires and processes the newest data, refits the models, makes new forecasts, publicly archives those forecasts, and presents both the current forecast and information on how previous forecasts performed.
@@
 
-A list of some of the key practices developed by Dietze et al [-dietze2018] for facilitating iterative near-term ecological forecasting and discussion of why these practices are important.
+A list of some of the key practices developed by Dietze et al [-@dietze2018] for facilitating iterative near-term ecological forecasting and discussion of why these practices are important.
 

### 2018-02-05 manuscript/portal_iterative_forecasting_ms.Rmd
Adds carpenter and team buiding refs

--- 2018-02-02 manuscript/portal_iterative_forecasting_ms.Rmd
@@
 It is rare to find such breadth within a single research group, and our system was developed as a collaboration between the lab collecting the data and a computational ecology lab.
-When teams have a breadth of expertise, communication can be challenging.
+When teams have a breadth of expertise, communication can be challenging [@winowiecki2011].
 We found a shared base of knowledge related to both the field research and fundamental computational skills was important for the success of the group.
@@
 This provided a shared set of knowledge and vocabulary that actively facilitated interdisciplinary interactions.
-Given the current state of tools for forecasting, forecasting teams will need people with significant experience in working with continuous integration and APIs, which means interdisciplinary teams will generally be required for creating these pipelines until tool development improves.
+Given the current state of tools for forecasting, forecasting teams will need people with significant experience in working with continuous integration and APIs.
+This means interdisciplinary teams will generally be required for creating these pipelines until tool development improves and a focus on training scientists in forecasting in 'team science' will be necessary for iterative forecasting attempts to be successful [@read2017].
 
@@
 We plan to use this pipeline to drive future research into understanding the processes that govern the dynamics of individual populations and the community as a whole.
-By regularly running different models for population and community dynamics, a near-term iterative pipeline such as ours should also make it possible to rapidly detect changes in how the system is operating, which should allow the rapid identification of ecological transitions or even possibly allow them to be prevented [@carpenter example].
+By regularly running different models for population and community dynamics, a near-term iterative pipeline such as ours should also make it possible to rapidly detect changes in how the system is operating, which should allow the rapid identification of ecological transitions or even possibly allow them to be prevented [@pace2017].
 By building an automated iterative near-term forecasting infrastructure we can improve our ability to forecast natural systems, our understanding of the biology driving ecological dynamics, and detect or even predict changes in system state that are important for conservation and management.

### 2018-02-05 manuscript/portal_iterative_forecasting_ms.Rmd
Add abstract and keywords
Co-authored-by: Morgan Ernest <morgan@weecology.org>

--- 2018-02-04 manuscript/portal_iterative_forecasting_ms.Rmd
@@
 \raggedright
+
+# Abstract
+
+* Most forecasts for the future state of ecological systems are conducted once and never updated or assessed.
+As a result, many available ecological forecasts are not based on the most up-to-date data, and the scientific progress of ecological forecasting models is slowed by a lack of feedback on how well the forecasts perform.
+* Iterative near-term ecological forecasting involves repeated daily to annual scale forecasts of an ecological system as new data becomes available and regular assessment of the resulting forecasts.
+We demonstrate how automated iterative near-term forecasting systems for ecology can be constructed by building one to conduct monthly forecasts of rodent abundances at the Portal Project, a long-term study with over 40 years of monthly data.
+This system automates most aspects of the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presentation of the forecasts.
+* The forecasting system uses R code for working with data, fitting models, making forecasts, and archiving and presenting these forecasts.
+The resulting pipeline is automated using continuous integration (a software development tool) to run the entire pipeline once a week.
+The cyberinfrastructure is designed for long-term maintainability and to allow the easy addition of new models.
+Constructing this forecasting system required a team with expertise ranging from field site experience to software development.
+* Automated near-term iterative forecasting systems will allow the science of ecological forecasting to advance more rapidly and provide the most up-to-date forecasts possible for conservation and management.
+These forecasting systems will also accelerate basic science by allowing new models of natural systems to be quickly implemented and compared to existing models.
+Using existing technology, and teams with diverse skill sets, it is possible for ecologists to build these systems and use them to advance our understanding of natural systems.
+
+Key-words: forecasting, prediction, mammals, iterative forecasting, Portal Project
 

### 2018-02-05 manuscript/portal_iterative_forecasting_ms.Rmd
Update archiving section to include external archiving
Closes #172

Co-authored-by: Morgan Ernest <morgan@weecology.org>

--- 2018-02-05 manuscript/portal_iterative_forecasting_ms.Rmd
@@
 
-Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2017; @harris2018] (Figure 1).
-Archiving serves as a form of pre-registration for model predictions, helping facilitate unbiased interpretation of model performance.
-We wanted our archives to not only be accessible to others but to also be a permanent record whose continued existence was not dependent on any particular researcher or funding stream.
-Storing forecast outputs in our forecasting GitHub repository was not sufficient for archival purposes because repositories can be deleted.
-We explored three major repositories for archiving our forecasts: FigShare (https://figshare.com/), Zenodo (https://zenodo.org/), and Open Science Framework (https://osf.io/).
+Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2018; @harris2018] (Figure 1).
+Archiving serves as a form of pre-registration for model predictions, because the forecasts cannot be modified once the data to assess them has been collected.
+This helps facilitate an unbiased interpretation of model performance.
+To serve this role archives should be publicly accessible and be a permanent record that cannot be changed or deleted.
+The second criteria means that GitHub is not sufficient for archival purposes because repositories can changed or deleted [@bergman2012].
+We explored three major repositories for archiving forecasts: FigShare (https://figshare.com/), Zenodo (https://zenodo.org/), and Open Science Framework (https://osf.io/).
 While all three repositories allowed for easy manual submissions (i.e., a human uploading files after each forecast), automating this process was substantially more difficult.
-Various combinations of repositories, APIs (i.e., interfaces that allow automated ways interacting with these websites) and associated R packages had issues with: 1) integrating authorization with continuous integration; 2) automatically making archived files public; 3) adding new files to an existing location; or 4) automatically permanently archiving the files.
-Our eventual solution was to leverage the GitHub-Zenodo integration (https://guides.github.com/activities/citable-code/) and automatically push forecasts to the GitHub repository from the CI server.
+Various combinations of repositories, APIs (i.e., interfaces for automatically interacting with these websites) and associated R packages had issues with: 1) integrating authorization with continuous integration; 2) automatically making archived files public; 3) adding new files to an existing location; or 4) automatically permanently archiving the files.
+Our eventual solution was to leverage the GitHub-Zenodo integration (https://guides.github.com/activities/citable-code/) and automatically push forecasts to a GitHub repository from the CI server and release them via the GitHub API.
 The GitHub-Zenodo integration is designed to automatically create versioned archives of GitHub repositories.
-There is an existing one-time process for linking our forecasting repository on  GitHub with Zenodo.
-Once this link is created, each time a new forecast is created, our pipeline adds the new forecasts to the GitHub repository and uses the GitHub API to create a new "release" of our repository.
-This triggers the Zenodo-GitHub integration, which automatically archives the resulting forecasts and the code that generated them under a top-level DOI that refers to all archived forecasts ([https://doi.org/10.5281/zenodo.833438](https://doi.org/10.5281/zenodo.833438)).
-Through this process, we automatically archive every forecast made with a documented history of the archive.
-While this approach is functional because everything in the repository is archived when a new forecast is made, these archives have complicated folder structures, making it more harder for users to find and access the forecasting results.
-Finding simpler approaches to automated archiving would increase the transparency of the forecase results and their ease of use by external researchers.
+We created a repository for storing forecasts (https://github.com/weecology/forecasts) and linked this repository with Zenodo (a one-time manual process).
+Each time a new forecast is created, our pipeline adds the new forecasts to the GitHub repository and uses the GitHub API to create a new "release" for that repository.
+This triggers the Zenodo-GitHub integration, which automatically archives the resulting forecasts under a top-level DOI that refers to all archived forecasts ([https://doi.org/10.5281/zenodo.839580](https://doi.org/10.5281/zenodo.839580)).
+Through this process, we automatically archive every forecast made with a documented time-stamp for every forecast in the archive.
+In addition, we also archive the full state of the repository for modeling and forecasting ([https://doi.org/10.5281/zenodo.833438](https://doi.org/10.5281/zenodo.833438)).
+This ensures that every forecast is fully reproducible since the exact code used to generate every forecast is preserved.
+Early forecasts from this system are archived as part of this modeling and forecasting code archive.
 

### 2018-02-05 manuscript/portal_iterative_forecasting_ms.Rmd

--- 2018-02-05 manuscript/portal_iterative_forecasting_ms.Rmd
@@
 ---
-title: "Automated iterative forecasting for the Portal Project"
+title: "Automated iterative near-term forecasting for the Portal Project"
 csl: mee.csl
@@
 
-Iterative forecasting requires data that is collected repeatedly, and it benefits most from data that is collected frequently, as this provides more opportunities for updating model results and assessing (and potentially improving) model performance [Box 1; @dietze2018].
+Iterative forecasting requires data that is collected repeatedly, and it benefits most from data that are collected frequently, as this provides more opportunities for updating model results and assessing (and potentially improving) model performance [Box 1; @dietze2018].
 The Portal Project is a long-term ecological study situated in the Chihuahuan Desert (2 km north and 6.5 km east of Portal, Arizona, US).
-Researchers have been continuously collecting data at the site since 1977, including data on the abundance of rodent and plant species (monthly and twice yearly, respectively) and climactic factors such as air temperature and precipitation (daily) [@brown1998; @ernest2009; @ernest2016].
+Researchers have been continuously collecting data at the site since 1977, including data on the abundance of rodent and plant species (monthly and twice yearly, respectively) and climatic factors such as air temperature and precipitation (daily) [@brown1998; @ernest2009; @ernest2016].
 The site consists of 24 50m x 50m experimental plots.
@@
 For all rodents caught during a trapping session, information on species identity, size, and reproductive condition is collected, and new individuals are given identification tags.
-This data on rodent populations is high-frequency, uses consistent trapping methodology, and has an extended time-series (469 monthly samples and counting), making this study an ideal case for near-term iterative forecasting.
+This information on rodent populations is high-frequency, uses consistent trapping methodology, and has an extended time-series (469 monthly samples and counting), making this study an ideal case for near-term iterative forecasting.
 
@@
 Each box denotes the core infrastructure used for each stage of the forecasting pipeline denoted in Figure 1a.
-Continuous integration, denoted here with the Travis icon (a girl wearing safety glasses and hardhat), plugs into our pipeline at every stage.
-Travis triggers the code involved in major events that involve integration across stages of the pipeline, such as taking the output from the forecasting stage (purple box) to create an updated presentation (rose box).
+Continuous integration, denoted here with the Travis icon (a woman wearing safety glasses and hardhat), plugs into our pipeline at every stage.
+Travis triggers the code involved in major events that link the stages of the pipeline, such as taking the output from the forecasting stage (purple box) to create an updated presentation (rose box).
 Travis also runs tasks within a stage, such as conducting tests to make sure code changes have not introduced errors (icons on arrows originating and ending on the same box).
@@
 Many of us have experienced the frustrations that can occur when software updates (e.g., changes in R package versions) create errors in previously functional code.
-We experienced this issue when a package one of our models relies on, `tscount` [@liboschik2015], was temporarily removed from CRAN (the R package repository) and, therefore, could not be installed in the usual way.
+We experienced this issue when the `tscount` package [@liboschik2015], used by one of our forecasting models, was temporarily removed from CRAN (the R package repository) and could not be installed in the usual way.
 This broke our forecasting pipeline because we could no longer run models that used that package.
@@
 Docker also allows a specified set of packages to be used consistently across different computer and server environments.
-Using containers allows us to update to new package versions after testing and making any necessary changes to the data processing and analysis code.
+Using containers allows us to control transitions to new package versions, implementing them only after we have tested them and made any necessary changes to the data processing and analysis code.
 We use a container created by the Rocker project which is a Docker image with many important R packages (i.e.
@@
 Frequent data collection and rapid processing are both important for providing timely forecasts.
-Since our data is collected monthly, ensuring that the models have access to the newest data requires a data latency period of less than 1 month from collection to availability for modeling.
+Since we collect data monthly, ensuring that the models have access to the newest data requires a data latency period of less than 1 month from collection to availability for modeling.
 To accomplish this, we automated components of the data processing and quality assurance/quality control (QA/QC) process to reduce the time needed to add new data to the database (Figure 1).
 
-New data is double-entered into Microsoft Excel using the "data validation" feature.
-The two versions are then compared in an R script to control for errors in data entry.
+New data are double-entered into Microsoft Excel using the "data validation" feature.
+The two versions are then compared using an R script to control for errors in data entry.
 Quality control (QC) checks using the `testthat` R package [@wickham2011] are run on the data to test for validity and consistency both within the new data and between the new and archived data.
 The local use of the QC scripts to flag problematic data greatly reduces the time spent error-checking and ensures that the quality of data is consistent.
-The data is then uploaded to the GitHub-based Portal Data repository (https://github.com/weecology/PortalData).
+The data are then uploaded to the GitHub-based Portal Data repository (https://github.com/weecology/PortalData).
 GitHub (https://github.com/) is a software development tool for managing computer code development, but we have also found it useful for data management.
-On GitHub, changes to data can be tracked through the Git version control system which logs all changes made to any files in the repository.
+On GitHub, changes to data can be tracked through the Git version control system which logs all changes made to any files in the repository - giving us a record of exactly of when specific lines of data were changed or added.
 All updates to data are processed through "pull requests", which are notifications that someone has a modified version of the data to contribute.
-QA/QC checks are automatically rerun on the submitted data using continuous integration to ensure that these checks have been run and that no avoidable errors reach the official version of the dataset.
+QA/QC checks are automatically rerun on the submitted data using continuous integration to ensure that these checks are run and that no avoidable errors reach the official version of the dataset.
 
@@
 
-The Portal Project has a long history of making its data publicly available, which means that anyone can use it for forecasting or other projects.
+The Portal Project has a long history of making its data publicly available, so that anyone can use it for forecasting or other projects.
 Historically the publication of the data was conducted through data papers [@ernest2009,@ernest2016], the most common approach in ecology; however, this approach caused years of data latency.
@@
 
-Once data is available, it needs to be processed into a form appropriate for modeling (Figure 1).
+Once data is available, it must be processed into a form appropriate for modeling (Figure 1).
 For many ecological datasets, this requires not only simple data manipulation but also a good understanding of the data to facilitate appropriate aggregation.
@@
 Ideally, new models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach.
-We use CI to refit the models and make new forecasts each time the modeling code changes and when new data becomes available (Figure 1b).
+We use CI to refit the models and make new forecasts each time the modeling code changes and when new data become available (Figure 1b).
 We use a plugin infrastructure to allow new models to be easily added to the system.
-This approach treats each model as an interchangable black boxes - all models have access to the same input data and generate the same structure for model outputs (Figure 2).
-Each iteration of the forecasting cycle, all models that are present are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts.
+This approach treats each model as an interchangable black box - all models have access to the same input data and generate the same structure for model outputs (Figure 2).
+Each iteration of the forecasting cycle, all existing models are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts.
 A weighted ensemble model is then added with weights based on how well individual models fit the training data.
-This plugin infrastructure makes it easy to add and compare very different types of models, from the basic time-series approaches currently implemented to complex state-space and machine learning models.
+This plugin infrastructure makes it easy to add and compare very different types of models, from the basic time-series approaches currently implemented to the more complex state-space and machine learning models we hope to implement in the future.
 As long as a model script can load the provided data and produce the appropriate output it will be run and its results incorporated into the rest of the forecasting system.
@@
 2017_12_08forecasts.csv) to allow cross-model evaluations.
-Model outpus files are housed in a folder containing all forecast outputs from all previous dates to facilitate our arhciving process and and forecast assessment activities ](figs/plugin_fig.gif)
+Model output files are housed in a folder containing all forecast outputs from all previous dates to facilitate our arhciving process and forecast assessment activities ](figs/plugin_fig.gif)
 
-In addition to flexibility in model internals, we also wanted to support flexibility in what the models predict.
+In addition to flexibility in what model structures can be supported, we also wanted to support flexibility in what the models predict.
 Allowing models to make forecasts for system properties ranging from individual species' population abundances to total community biomass facilitates exploration of differences in forecastability across different aspects of ecological systems.
@@
 
-In addition to archiving the results of each forecast, we present them on a website that displays monthly rodent forecasts, model evaluation metrics, monthly reports, and information about the study site (Figure 3; [http://portal.naturecast.org](http://portal.naturecast.org/)).
+Each month, we present our forecasts on a website that displays monthly rodent forecasts, model evaluation metrics, monthly reports, and information about the study site (Figure 3; [http://portal.naturecast.org](http://portal.naturecast.org/)).
 The website includes a graphical presentation of the most recent month's forecasts (including uncertainty) and compares the latest data to the previous forecasts.
-Information on the the species and the field site targeted to a general audience are also included.
+Information on the species and the field site targeted to a general audience are also included.
 The site is built using Rmarkdown [@rmarkdown], which naturally integrates into the pipeline, and is automatically updated after each forecast.
@@
 The files for the website are stored in a subdirectory of the forecasting repository.
-As a result, the website is also archived automatically as part of the forecast archiving.
+As a result, the website is also archived automatically as part of archiving the forecast results.
 
 ![Figure 3.
-Screen capture of the homepage of the Portal Forecasting website; http://portal.naturecast.org.
+Screen capture of the homepage of the Portal Forecasting website (http://portal.naturecast.org).
 This site contains information on the most current forecasts, evaluation of forecast performance, and general information about the species being forecast.](figs/website_fig.png)
@@
 Every week our forecasting system generates a new set of forecasts with no human intervention, except for the entry of new field data.
-This ensures that forecasts based on the most recent data are always available and allows us to rapidly assess the performance of multiple forecasting models for a number of different states of the system, including the abundances of individual species and community-level variables such as total abundance.
+Our forecasting system ensures that forecasts based on the most recent data are always available.
+Our forecasting system is also designed to allow rapid assessment of the performance of multiple forecasting models for a number of different states of the system, including the abundances of individual species and community-level variables such as total abundance.
 To create this iterative near-term forecasting system, we used R to process data and conduct analyses, and we leveraged already existing services (i.e.
@@
 Finally, we found no satisfactory existing solution for archiving our results.
-All approaches we tried had limitations when it came to automatically generating publicly versioned archives of forecasts on a repeated basis.
-Overall, we found existing technology to be sufficient to the task, but it required greater expertise and a greater investment of time than is ideal.
-Tool development to reduce the effort required for scientists to set up their own short-term forecasting systems would clearly be useful, but our efforts show that it is possible for scientists to use existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems.
-Developing systems that  produce automated iterative forecasts will make it easier to convert research focused forecasts into fully operationalized forecasting systems that are relied on for decision making [@dietze2018, other paper by operationalization co-authors].
+All approaches we tried had limitations when it came to automatically generating publicly versioned archives of forecasts on a repeated basis, and our eventual solution was difficult to configure to such a degree that it will an impediment for most researchers.
+Overall, we found existing technology to be sufficient to the task of creating an iterative forecasting pipeline, but it required greater expertise and a greater investment of time than is ideal.
+Tool development to reduce the effort required for scientists to set up their own short-term forecasting systems would clearly be useful.
+Developing systems that  produce automated iterative forecasts will make it easier to convert research focused forecasts into fully operationalized forecasting systems that are relied on for decision making [@dietze2018].
+While more user-friendly approaches would make it easier for more scientists make their own iterative forecasts, our efforts show that it is possible to use existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems.
 
@@
 Given the current state of tools for forecasting, forecasting teams will need people with significant experience in working with continuous integration and APIs.
-This means interdisciplinary teams will generally be required for creating these pipelines until tool development improves and a focus on training scientists in forecasting in 'team science' will be necessary for iterative forecasting attempts to be successful [@read2017].
+This means interdisciplinary teams will generally be required for creating these pipelines until tool development improves.
+To improve the success of these diverse groups, we believe efforts at providing 'team science' training to scientists interested in forecasting will be beneifical for the success of iterative forecasting attempts for the foreseeable future  [@read2017].
 

### 2018-02-07 manuscript/portal_iterative_forecasting_ms.Rmd
Manuscript polishing
1. Reduce redundancy in abstract
2. General wording and tightening throughout

--- 2018-02-05 manuscript/portal_iterative_forecasting_ms.Rmd
@@
 
-Forecasting the future state of ecological systems is important for management, conservation, and evaluation of our basic understanding of how ecology operates [@clark2001; @tallis2006; @diaz2015; @dietze2017].
+Forecasting the future state of ecological systems is important for management, conservation, and evaluation of how well models capture the processes governing ecological systems [@clark2001; @tallis2006; @diaz2015; @dietze2017].
 In 2001, Clark et al.
@@
 However, most of these forecasts are made once, published, and never assessed or updated.
-Without assessment, we have limited information on how much confidence to place in our predictions; without regular updates, forecasts lack the most up-to-date information as conditions change [@dietze2018].
 This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions.
-More regular updating and assessment will advance ecological forecasting as a field by speeding up the identification of the best models for forecasting specific issues in specific ecosystems and contribute overall to our understanding of how to best design forecasting approaches for ecology.
+The lack of active assessment results in limited information on how much confidence to place in forecasts and makes if difficult to determine which forecasting methods to build on.
+Without regular updates, forecasts lack the most current data and forecast accurracy will decay over time [@petch2015; @dietze2018].
+More regular updating and assessment will advance ecological forecasting as a field by accelerating the identification of the best models for individual forecasts and improving our understanding of how to best design forecasting approaches for ecology in general.
 For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].
@@
 With the goal of making ecological forecasting more dynamic and responsive, Dietze et al [-@dietze2018] recently called for an increase in iterative near-term forecasting.
-Iterative near-term forecasting means making forecasts for the near future and making these forecasts repeatedly through a cycle of forecast evaluation, integration of updated data, and generation of new forecasts.
-This approach to forecasting has a number of advantages.
-Because forecasts are made 'near-term'---daily to annual forecasting instead of multi-decadal---predictions can be assessed more quickly and frequently, leading to more rapid model improvements [@tredennick2016; @dietze2018].
-Because the forecasts are made repeatedly through time, new data can be integrated with each new iteration of the forecast cycle.
-This iterative approach to forecasting allows any changes in the state of the system that have occurred since the previous forecast to be incorporated and accounted for, thereby improving the accuracy of future forecasts [@dietze2018].
-Iterative near-term forecasting has the potential to promote rapid improvement in the state of ecological forecasting by quickly identifying how models are failing, facilitating rapid testing of improved models, and incorporating updated data so models run with the most up-to-date information on the system available.
-While use of iterative near-term forecasting is often discussed in the context of policy and management [@clark2001; @luo2011; @petchey2015], this approach to model testing can also be used to improve our basic understanding of ecological systems [@dietze2017].
-For example, alternative mechanistic models can be competed against each other to see which model provides the best forecasts for near-term dynamics, thus providing insights into the relative importance of different processes driving dynamics of ecological systems [@dietze2018].
-Whether deployed for basic or applied uses, iterative near-term forecasting incorporates a more dynamic interplay between models, predictions, and data that is clearly needed to improve ecological forecasting and our understanding of ecological systems more broadly.
+Iterative near-term forecasting is defined by making predictions for the near future and repeatedly updating those predictions through a cycle of evaluation, integration of new data, and generation of new forecasts.
+Because forecasts are made 'near-term'---daily to annual time scales instead of multi-decadal---predictions can be assessed more quickly and frequently, leading to more rapid model improvements [@tredennick2016; @dietze2018].
+In addition, since forecasts are made repeatedly through time, new data can be continuosly integrated with each iteration [@dietze2018].
+As a result, iterative near-term forecasting has the potential to promote rapid improvement in the state of ecological forecasting by quickly identifying how models are failing, facilitating rapid testing of improved models, and incorporating the most up-to-date data available.
+In addition to yielding improved information for guiding policy and management [@clark2001; @luo2011; @petchey2015], this iterative approach will help improve our basic understanding of ecological systems [@dietze2018].
+For example, alternative mechanistic models can be compared to determine which model provides the best forecasts, thus providing insights into the importance of different ecological processes [@dietze2018].
+Iterative near-term forecasting provides a more dynamic interplay between models, predictions, and data that will improve ecological forecasting and our understanding of ecological systems more broadly.
 
-Because iterative near-term forecasting requires a dynamic interplay of models, predictions, and data, Dietze et al [-@dietze2018] highlight approaches to data management, model construction and evaluation, and cyberinfrastructure that are necessary to effectively implement this type of forecasting (Box 1).
-Data to be used for iterative near-term forecasting needs to be widely accessible, which requires data to be released quickly under open licenses [@dietze2018; @vargas2017] and structured so that it can be used easily by a variety of researchers and in multiple modeling approaches [@borer2009; @strasser2011].
+Because iterative near-term forecasting requires a dynamic integration of models, predictions, and data, Dietze et al [-@dietze2018] highlight approaches to data management, model construction and evaluation, and cyberinfrastructure that are necessary to effectively implement this type of forecasting (Box 1).
+Data needs to be released quickly under open licenses [@vargas2017; @dietze2018] and structured so that it can be used easily by a variety of researchers and in multiple modeling approaches [@borer2009; @strasser2011].
 Models need to be able to deal with uncertainty, in the predictors and the predictions, to properly convey uncertainty in the resulting forecasts [@diniz-filho2009].
@@
 
-While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement; few of their recommendations are in widespread use in ecology today.
-We examined what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going long-term (~40 year) ecological study that collects high-frequency data on desert rodent abundances [@brown1998; @ernest2008].
-We constructed our forecasting pipeline with the goal of being able to forecast rodent abundances and evaluate our predictions on a monthly basis.
+While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement and few of their recommendations are in widespread use in ecology today.
+We explored what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going long-term ecological study that collects high-frequency data on desert rodent abundances [@brown1998; @ernest2008].
+We constructed an automated forecasting pipeline with the goal of being able to forecast rodent abundances and evaluate our predictions on a monthly basis.
 In this paper, we discuss our approach for creating this iterative near-term forecasting pipeline, the challenges we encountered, the tools we used, and the lessons we learned so that others can create their own iterative forecasting systems.
@@
 
-Iterative forecasting requires data that is collected repeatedly, and it benefits most from data that are collected frequently, as this provides more opportunities for updating model results and assessing (and potentially improving) model performance [Box 1; @dietze2018].
+Iterative forecasting is most effective with frequently collected data, since it provides more opportunities for updating model results and assessing (and potentially improving) model performance [Box 1; @dietze2018].
 The Portal Project is a long-term ecological study situated in the Chihuahuan Desert (2 km north and 6.5 km east of Portal, Arizona, US).
@@
 Implementation of iterative forecasting requires the regular rebuilding of models with new raw data as it becomes available and the presentation of those forecasts in usable forms; in our case, this occurs monthly.
-Rebuilding models in an efficient and maintainable way relies on developing an automated pipeline to handle the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presenting of the forecasts (Figure 1a).
+Rebuilding models in an efficient and maintainable way relies on developing an automated pipeline to handle the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presention of the forecasts (Figure 1a).
 To implement the pipeline outlined in Figure 1a, we used a "continuous analysis" framework [*sensu* @beaulieu2017] that automatically processes the most up-to-date data, refits the models, makes new forecasts, archives the forecasts, and updates a website with analysis of current and previous forecasts.
-In this section we describe our approach to streamlining and automating the multiple components of the forecasting pipeline and the tools and infrastructure we employed to execute each stage of the pipeline.
+In this section we describe our approach to streamlining and automating the multiple components of the forecasting pipeline and the tools and infrastructure we employed to execute each component.
 
@@
 To go from raw data to forecast presentation involves a number of stages, each of which requires unique tasks, tools and infrastructure.
-While each stage has unique tasks associated with it, the stages of our pipeline are also interdependent.
-Our pipeline is primarily a linear structure where outputs from one stage form the inputs for the subsequent stage.
-All stages require code written in R.
+The stages of are interdependent with outputs from one stage forming the inputs for the subsequent stage.
+Tasks in all stages are run using code written in R.
 b) Continuous integration system.
-Each box denotes the core infrastructure used for each stage of the forecasting pipeline denoted in Figure 1a.
-Continuous integration, denoted here with the Travis icon (a woman wearing safety glasses and hardhat), plugs into our pipeline at every stage.
-Travis triggers the code involved in major events that link the stages of the pipeline, such as taking the output from the forecasting stage (purple box) to create an updated presentation (rose box).
-Travis also runs tasks within a stage, such as conducting tests to make sure code changes have not introduced errors (icons on arrows originating and ending on the same box).
+Each box denotes the core infrastructure used for each stage of the forecasting pipeline.
+Continuous integration (denoted by the Travis icon; a woman wearing safety glasses and hardhat) triggers the code involved in events that link the stages of the pipeline, such as using the output from the forecasting stage (purple box) to create an updated website (rose box).
+Travis also runs tasks within a stage, such as testing code and adding weather data (icons on arrows originating and ending on the same box).
 ](figs/Figure1.png)
@@
 A core aspect of iterative near-term forecasting is the regular rerunning of the forecasting pipeline.
-We chose to have the computer run our pipeline and employed "continuous analysis" [*sensu* @beaulieu2017] to drive the automation of both the full pipeline and a number of its individual components.
+We employed "continuous analysis" [*sensu* @beaulieu2017] to drive the automation of both the full pipeline and a number of its individual components.
 Continuous analysis uses a set of tools originally designed for software development called "continuous integration" (CI).
@@
 Essentially, CI is a computer helper whose job is to watch the pipeline and, when it sees a change in the code or data, it runs all the computer scripts needed to ensure that the forecasting pipeline runs from beginning to end.
-This is useful for iterative near-term forecasting because it does not rely on humans to remember to create forecasts when new models or data are added.
+This is useful for iterative near-term forecasting because it does not rely on humans to create new forecasts whenever new models or data are added.
 These tools are common in the area of software development where they are used to automate software testing and integrate work by multiple developers working on the same code base.
@@
 Our forecasting pipeline currently runs on a publicly available continuous integration service (Travis CI; https://travis-ci.org/) that is free for open source projects (up to a limited amount of computing time).
-Because of the widespread use of CI in software development, alternative services that can run code on local or cloud-based computational infrastructure also exist (e.g., Drone (http://try.drone.io/)) [@beaulieu2017].
-As detailed below, we use CI to quality check data, test code using "unit tests" [@wilson2014], build models, make forecasts, and publicly present and archive the results (Figure 1b).
+Because of the widespread use of CI in software development, alternative services that can run code on local or cloud-based computational infrastructure also exist [@beaulieu2017].
+We use CI to quality check data, test code using "unit tests" [@wilson2014], build models, make forecasts, and publicly present and archive the results (Figure 1b).
 
@@
 All updates to data are processed through "pull requests", which are notifications that someone has a modified version of the data to contribute.
-QA/QC checks are automatically rerun on the submitted data using continuous integration to ensure that these checks are run and that no avoidable errors reach the official version of the dataset.
+QA/QC checks are automatically run on the submitted data using continuous integration to ensure that these checks are run and that no avoidable errors reach the official version of the dataset.
 
@@
 Historically the publication of the data was conducted through data papers [@ernest2009,@ernest2016], the most common approach in ecology; however, this approach caused years of data latency.
-Recently, the project has switched to posting data directly to a public GitHub repository with a CC0 (i.e.
-no rights reserved) license (Figure 1).
-This immediate posting reduces that data latency to less than one month and, therefore, makes meaningful iterative near-term forecasting possible for not only our group but other interested parties, as well.
+With the recent switch to posting data directly to a public GitHub repository (Figure 1) with a CC0 waiver (i.e.
+no restrictions on data use; https://creativecommons.org/publicdomain/zero/1.0/) data latency for everyone has been reduced to less than one month making meaningful iterative near-term forecasting possible for not only our group but other interested parties, as well.
 
@@
 Finally, this kind of code is generally poorly tested, which can lead to errors based on mistakes in data manipulation.
-To avoid these issues for the Portal Project data, the Portal team has been developing an R package (portalR; http://github.com/weecology/portalr) for acquiring the data and handling common data cleaning and aggregation tasks.
+To avoid these issues for the Portal Project data, the Portal team has been developing an R package (portalr; http://github.com/weecology/portalr) for acquiring the data and handling common data cleaning and aggregation tasks.
 As a result, our modeling and forecasting code only needs to install this package and run the data manipulation and summary functions to get the appropriate data (Figure 1b).
-The package undergoes thorough automated unit tests to ensure that data manipulations are achieving the desired results.
+The package undergoes thorough automated unit testing to ensure that data manipulations are achieving the desired results.
 Having data manipulation code maintained in a separate package that focuses on consistently providing properly summarized forms of the most recent data has made maintaining the forecasting code itself much more straightforward.
@@
 This approach treats each model as an interchangable black box - all models have access to the same input data and generate the same structure for model outputs (Figure 2).
-Each iteration of the forecasting cycle, all existing models are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts.
+During each run of the forecasting code, all existing models are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts.
 A weighted ensemble model is then added with weights based on how well individual models fit the training data.
@@
 2017_12_08forecasts.csv) to allow cross-model evaluations.
-Model output files are housed in a folder containing all forecast outputs from all previous dates to facilitate our arhciving process and forecast assessment activities ](figs/plugin_fig.gif)
+Model output files are housed in a folder containing all forecast outputs from all previous dates to facilitate archiving and forecast assessment.](figs/plugin_fig.gif)
 
@@
 We designed a forecast output format to support this.
-Each forecast output file contains the date being forecast, the collection date of the data used for fitting the models, the date the forecast was made, the state variable being forecast (e.g., rodent biomass, the abundance of a species), and the forecast value and associated uncertainty of that forecast (Figure 2).
+Each forecast output file contains the date being forecast, the collection date of the data used for fitting the models, the model name, the date the forecast was made, the state variable being forecast (e.g., rodent biomass, the abundance of a species), and the forecast value and associated uncertainty of that forecast (Figure 2).
 This allows us to store a variety of different forecasts in a common format and may serve as a useful starting point for developing a standard for storing ecological forecasts more generally.
@@
 We plan to add additional metrics in the future.
-In addition to evaluating the actual forecasts, we also use hindcasting (forecasting on already collected data) to gain additional insight into the methods that work best for forecasting this system.
+In addition to evaluating the actual forecasts, we also use hindcasting [forecasting on already collected data; @jolliffe2003] to gain additional insight into the methods that work best for forecasting this system.
 For example, a model is fit using rodent observations up to June 2005, then used to make a forecast 12 months out to May 2006.
 The observations of that 12 month period can immediately be used to evaluate the model.
-Hindcasting can be conducted using data that was already been collected, thus allowing model comparison of large numbers of hindcasts and giving insight into which models make the best forecasts without needing to wait for new data to be collected.
+Since hindcasting is conducted using data that has already been collected, it allows model comparisons to be conducted on large numbers of hindcasts and provides insight into which models make the best forecasts without needing to wait for new data to be collected [@harris2018].
 It can also be used to quickly evaluate new models instead of waiting for an adequate amount of data to accumulate.
@@
 To serve this role archives should be publicly accessible and be a permanent record that cannot be changed or deleted.
-The second criteria means that GitHub is not sufficient for archival purposes because repositories can changed or deleted [@bergman2012].
+The second criteria means that GitHub is not sufficient for archival purposes because repositories can be changed or deleted [@bergman2012; @white2015].
 We explored three major repositories for archiving forecasts: FigShare (https://figshare.com/), Zenodo (https://zenodo.org/), and Open Science Framework (https://osf.io/).
 While all three repositories allowed for easy manual submissions (i.e., a human uploading files after each forecast), automating this process was substantially more difficult.
-Various combinations of repositories, APIs (i.e., interfaces for automatically interacting with these websites) and associated R packages had issues with: 1) integrating authorization with continuous integration; 2) automatically making archived files public; 3) adding new files to an existing location; or 4) automatically permanently archiving the files.
+Various combinations of repositories, APIs (i.e., interfaces for automatically interacting with the archiving websites) and associated R packages had issues with: 1) integrating authorization with continuous integration; 2) automatically making archived files public; 3) adding new files to an existing location; or 4) automatically permanently archiving the files.
 Our eventual solution was to leverage the GitHub-Zenodo integration (https://guides.github.com/activities/citable-code/) and automatically push forecasts to a GitHub repository from the CI server and release them via the GitHub API.
@@
 This triggers the Zenodo-GitHub integration, which automatically archives the resulting forecasts under a top-level DOI that refers to all archived forecasts ([https://doi.org/10.5281/zenodo.839580](https://doi.org/10.5281/zenodo.839580)).
-Through this process, we automatically archive every forecast made with a documented time-stamp for every forecast in the archive.
-In addition, we also archive the full state of the repository for modeling and forecasting ([https://doi.org/10.5281/zenodo.833438](https://doi.org/10.5281/zenodo.833438)).
+Through this process, we automatically archive every forecast made with a documented time-stamp.
+In addition, we also archive the full state of the modeling and forecasting repository ([https://doi.org/10.5281/zenodo.833438](https://doi.org/10.5281/zenodo.833438)).
 This ensures that every forecast is fully reproducible since the exact code used to generate every forecast is preserved.
-Early forecasts from this system are archived as part of this modeling and forecasting code archive.
+Early forecasts from this system are archived in this modeling and forecasting code archive, not the newer forecasts repository.
 
@@
 The website includes a graphical presentation of the most recent month's forecasts (including uncertainty) and compares the latest data to the previous forecasts.
-Information on the species and the field site targeted to a general audience are also included.
+Information on the species and the field site are also included.
 The site is built using Rmarkdown [@rmarkdown], which naturally integrates into the pipeline, and is automatically updated after each forecast.
@@
 Our forecasting system ensures that forecasts based on the most recent data are always available.
-Our forecasting system is also designed to allow rapid assessment of the performance of multiple forecasting models for a number of different states of the system, including the abundances of individual species and community-level variables such as total abundance.
-To create this iterative near-term forecasting system, we used R to process data and conduct analyses, and we leveraged already existing services (i.e.
+It is also designed to allow rapid assessment of the performance of multiple forecasting models for a number of different states of the system, including the abundances of individual species and community-level variables such as total abundance.
+To create this iterative near-term forecasting system, we used R to process data and conduct analyses, and leveraged existing tools and services (i.e.
 GitHub, Travis, Docker) for more complicated cyberinfrastructure tasks.
-Thus, our approach to developing iterative near-term forecasting infrastructure provides an example for how short-term ecological forecasting systems can be initially developed.
+Thus, our approach to developing iterative near-term forecasting infrastructure provides an example for how short-term ecological forecasting systems can be developed.
 
@@
 We used existing technology for both running the pipeline and building individual components, which allowed us to build the system relatively cheaply in terms of both time and money.
-This included the use of tools like Docker for reproducibility, the Travis CI continuous integration system for automatically running the pipeline, Rmarkdown and `knitr` for generating the website, and the already existing integration between Github and Zenodo  to archive the forecasts.
+This included the use of tools like Docker for reproducibility, Travis CI continuous integration for automatically running the pipeline, Rmarkdown and `knitr` for generating the website, and the already existing integration between Github and Zenodo  to archive the forecasts.
 By using this "continuous analysis" approach [@beaulieu2017], where analyses are automatically rerun when changes are made to data, models, or associated code, we have reduced the time required by scientists to run and maintain the forecasting pipeline.
@@
 Overall, we found existing technology to be sufficient to the task of creating an iterative forecasting pipeline, but it required greater expertise and a greater investment of time than is ideal.
-Tool development to reduce the effort required for scientists to set up their own short-term forecasting systems would clearly be useful.
-Developing systems that  produce automated iterative forecasts will make it easier to convert research focused forecasts into fully operationalized forecasting systems that are relied on for decision making [@dietze2018].
-While more user-friendly approaches would make it easier for more scientists make their own iterative forecasts, our efforts show that it is possible to use existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems.
+Additional tool development to reduce the effort required for scientists to set up their own short-term forecasting systems would clearly be useful.
+However, our efforts show that it is possible to use existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems.
 
@@
 We found a shared base of knowledge related to both the field research and fundamental computational skills was important for the success of the group.
+The two labs are part of a joint interdisciplinary ecology group that has a mission of breaking down barriers between field and computational/theoretical ecologists  (http://weecology.org).
 Everyone on the team had received training in fundamental data management and computing skills through a combination of university courses, Software and Data Carpentry workshops [@teal2015], and informal lab training efforts.
@@
 This means interdisciplinary teams will generally be required for creating these pipelines until tool development improves.
-To improve the success of these diverse groups, we believe efforts at providing 'team science' training to scientists interested in forecasting will be beneifical for the success of iterative forecasting attempts for the foreseeable future  [@read2017].
+To improve the success of these diverse groups, we believe efforts at providing 'team science' training to scientists interested in forecasting will be beneficial for the success of iterative forecasting attempts for the foreseeable future  [@read2017].
 
-We developed this infrastructure for automatically making iterative forecasts with the goals of making accurate forecasts for this well-studied system, learning what methods work well for ecological forecasting more generally, and improving our understanding of the processes driving ecological dynamics.
-The most obvious application of automated iterative ecological forecasting is for speeding up development of forecasting models by providing the most recent data available to models and by quickly iterating to improve the models used for forecasting.
+We developed infrastructure for automatically making iterative forecasts with the goals of making accurate forecasts for this well-studied system, learning what methods work well for ecological forecasting more generally, and improving our understanding of the processes driving ecological dynamics.
+The most obvious application of automated iterative ecological forecasting is for speeding up development of forecasting models by using the most recent data available and by quickly iterating to improve the models used for forecasting.
 By learning what works best for forecasting in this and other ecological systems, we will better understand what the best approaches are for ecological forecasting more generally.
 By designing the pipeline so that it can forecast many different aspects of the ecological community, we also hope to learn about what aspects of ecology are more forecastable.
-Finally, automated forecasting infrastructures like this one also provide a core foundation for faster scientific inquiry more broadly because new models can quickly be applied to data and compared to existing models.
+Finally, automated forecasting infrastructures like this one also provide a core foundation for faster scientific inquiry because new models can quickly be applied to data and compared to existing models.
 The forecasting infrastructure does the time-consuming work of data processing, data integration, and model assessment, allowing new research to focus on the models being developed and the inferences about the system that can be drawn from them [@dietze2018].
@@
 By regularly running different models for population and community dynamics, a near-term iterative pipeline such as ours should also make it possible to rapidly detect changes in how the system is operating, which should allow the rapid identification of ecological transitions or even possibly allow them to be prevented [@pace2017].
-By building an automated iterative near-term forecasting infrastructure we can improve our ability to forecast natural systems, our understanding of the biology driving ecological dynamics, and detect or even predict changes in system state that are important for conservation and management.
+By building an automated iterative near-term forecasting infrastructure we can improve our ability to forecast natural systems, understand the biology driving ecological dynamics, and detect or even predict changes in system state that are important for conservation and management.
 
@@
 
+We thank Henry Senyondo for help with continuous integration and Hao Ye and Juniper Simonis for discussions and feedback on the manuscript.
+We thank all of the graduate students, postdocs, and volunteers who have collected the Portal Project over the last 40 years and the developers of the software and tools that made this project possible.
 This research was supported by the National Science Foundation through grant 1622425 to S.K.M.
@@
 White.
-We thank all of the graduate students, postdocs, and volunteers who have collected the Portal Project over the last 40 years and the developers of all of the software and tools that made this project possible.
 
@@
 
-To reduce the time and effort needed to incorporate this data into models, best practices in data structure need to be employed for managing and storing collected data to ensure it is easy to integrate into other systems (interoperability) [@borer2009; @strasser2011; @white2013].
+To reduce the time and effort needed to incorporate data into models, best practices in data structure should be employed for managing and storing collected data to ensure it is easy to integrate into other systems (interoperability) [@borer2009; @strasser2011; @white2013].
 
@@
 
-Understanding the uncertainty of forecasts is crucial to interpreting the forecasts and understanding their utility.
+Understanding the uncertainty of forecasts is crucial to interpreting and understanding their utility.
 Models used for forecasting should be probabilistic to properly quantify uncertainty and to convey how this uncertainty increases through time.
-Evaluation of forecast models should include assessment of how accurately they quantify uncertainty as well as point estimates.
-This can be done using "proper and local" scores [@hooten2015].
+Evaluation of forecast models should include assessment of how accurately they quantify uncertainty as well as point estimates  [@hooten2015].
 
@@
 
-To quickly learn about the best approaches to forecasting different aspects of ecology, multiple modeling approaches should be compared for forecasting tasks [@harris2018].
-Different modeling approaches should also be combined into ensemble models, which are known to outperform single models for many forecasting and prediction tasks [@weigel2008].
+To quickly learn about the best approaches to forecasting different aspects of ecology, multiple modeling approaches should be compared [@harris2018].
+Different modeling approaches should also be combined into ensemble models, which often outperform single models for prediction [@weigel2008].
 
@@
 Best practices should be followed in the development of scientific software and modeling to make it easier to maintain, integrate into pipelines, and build on by other researchers.
-Key best practices include using open licenses, good documentation, version control, and cross-platform support [@wilson2014; @hampton2015].
+Key best practices include open licenses, good documentation, version control, and cross-platform support [@wilson2014; @hampton2015].
 
@@
 
-To facilitate the comparison and ensembling of different modeling approaches, code for fitting models and making forecasts should be easily extensible, allowing models developed by different groups to be easily integrated into a single framework [@dietze2018].
+To facilitate the comparison and ensembling of different modeling approaches, code for fitting models and making forecasts should be easily extensible, to allow models developed by different groups to be integrated into a single framework [@dietze2018].
 
@@
 
-Iteratively making forecasts requires that acquiring the newest data, refitting the models, and making new forecasts is simple.
-Ideally, this should be done automatically without requiring human intervention.
+Each forecast iteration involves acquiring new data, refitting the models, and making new forecasts.
+This should be done automatically without requiring human intervention.
 Therefore, the process of making forecasts should emphasize end-to-end reproducibility, including data, models, and evaluation [@stodden2014], to allow the forecasts to be easily rerun as new data becomes available [@dietze2018].
-Ideally, the entire forecasting pipeline will be rerun automatically as new data becomes available.
 
@@
 
-Forecasts should be openly archived to demonstrate that the forecasts were made without knowledge of the outcomes (i.e., as a form of pre-registration *sensu*) and to allow the community to assess and compare the performance of different forecasting approaches both now and in the future [@mcgill2012; @tredennick2016; @dietze2018; @harris2018].
-Ideally, the forecasts and evaluation of their performance should be automatically posted publicly in a manner that is understandable by both interested scientists and other stakeholders.
+Forecasts should be openly archived to demonstrate that the forecasts were made without knowledge of the outcomes and to allow the community to assess and compare the performance of different forecasting approaches both now and in the future [@mcgill2012; @tredennick2016; @dietze2018; @harris2018].
+Ideally, the forecasts and evaluation of their performance should be automatically posted publicly in a manner that is understandable by both scientists and the broader stakeholder community.
 

### 2018-02-08 manuscript/portal_iterative_forecasting_ms.Rmd
typo fix

@@
 This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions.
-The lack of active assessment results in limited information on how much confidence to place in forecasts and makes if difficult to determine which forecasting methods to build on.
+The lack of active assessment results in limited information on how much confidence to place in forecasts and makes it difficult to determine which forecasting methods to build on.
 Without regular updates, forecasts lack the most current data and forecast accurracy will decay over time [@petch2015; @dietze2018].

### 2018-02-08 manuscript/portal_iterative_forecasting_ms.Rmd
Edits to intro paragraph 2

@@
 With the goal of making ecological forecasting more dynamic and responsive, Dietze et al [-@dietze2018] recently called for an increase in iterative near-term forecasting.
-Iterative near-term forecasting is defined by making predictions for the near future and repeatedly updating those predictions through a cycle of evaluation, integration of new data, and generation of new forecasts.
+Iterative near-term forecasting is defined as making predictions for the near future and repeatedly updating those predictions through a cycle of evaluation, integration of new data, and generation of new forecasts.
 Because forecasts are made 'near-term'---daily to annual time scales instead of multi-decadal---predictions can be assessed more quickly and frequently, leading to more rapid model improvements [@tredennick2016; @dietze2018].
-In addition, since forecasts are made repeatedly through time, new data can be continuosly integrated with each iteration [@dietze2018].
-As a result, iterative near-term forecasting has the potential to promote rapid improvement in the state of ecological forecasting by quickly identifying how models are failing, facilitating rapid testing of improved models, and incorporating the most up-to-date data available.
+Since forecasts are made repeatedly through time, new data can be continuously integrated with each iteration [@dietze2018].
+By quickly identifying how models are failing, facilitating rapid testing of improved models, and incorporating the most up-to-date data available, iterative near-term forecasting has the potential to promote rapid improvement in the state of ecological forecasting.
 In addition to yielding improved information for guiding policy and management [@clark2001; @luo2011; @petchey2015], this iterative approach will help improve our basic understanding of ecological systems [@dietze2018].
 For example, alternative mechanistic models can be compared to determine which model provides the best forecasts, thus providing insights into the importance of different ecological processes [@dietze2018].
-Iterative near-term forecasting provides a more dynamic interplay between models, predictions, and data that will improve ecological forecasting and our understanding of ecological systems more broadly.
+Iterative near-term forecasting provides the more dynamic interplay between models, predictions, and data that has been identified has necessary for improving ecological forecasting and our understanding of ecological systems more broadly.
 

### 2018-02-08 manuscript/portal_iterative_forecasting_ms.Rmd
small edit in archiving section
Small edit in discussion 1st paragraph

@@
 This ensures that every forecast is fully reproducible since the exact code used to generate every forecast is preserved.
-Early forecasts from this system are archived in this modeling and forecasting code archive, not the newer forecasts repository.
+Early forecasts from this system are archived in this modeling and forecasting code archive, not in the newer repository 'forecasts'.
 
@@
 Our forecasting system automatically acquires and processes the newest data, refits the models, makes new forecasts, publicly archives those forecasts, and presents both the current forecast and information on how previous forecasts performed.
-Every week our forecasting system generates a new set of forecasts with no human intervention, except for the entry of new field data.
-Our forecasting system ensures that forecasts based on the most recent data are always available.
-It is also designed to allow rapid assessment of the performance of multiple forecasting models for a number of different states of the system, including the abundances of individual species and community-level variables such as total abundance.
+Every week the forecasting system generates a new set of forecasts with no human intervention, except for the entry of new field data.
+Our forecasting system ensures that forecasts based on the most recent data are always available and is designed to allow rapid assessment of the performance of multiple forecasting models for a number of different states of the system, including the abundances of individual species and community-level variables such as total abundance.
 To create this iterative near-term forecasting system, we used R to process data and conduct analyses, and leveraged existing tools and services (i.e.

### 2018-02-11 manuscript/portal_iterative_forecasting_ms.Rmd

@@
 In 2001, Clark et al.
-[-@clark2001] called for a more central role of forecasting in ecology, and since then an increasing number of ecological forecasts are being published.
-However, most of these forecasts are made once, published, and never assessed or updated.
+[-@clark2001] called for a more central role of forecasting in ecology.
+Since then, an increasing number of ecological forecasts are being published.
+Most of these forecasts, however, are made once, published, and never assessed or updated.
 This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions.
-The lack of active assessment results in limited information on how much confidence to place in forecasts and makes it difficult to determine which forecasting methods to build on.
-Without regular updates, forecasts lack the most current data and forecast accurracy will decay over time [@petch2015; @dietze2018].
+The lack of active assessment results in limited information on how much confidence to place in forecasts and makes it difficult to determine on which forecasting methods to build.
+Without regular updates, forecasts lack the most current data, and forecast accurracy will decay over time [@petch2015; @dietze2018].
 More regular updating and assessment will advance ecological forecasting as a field by accelerating the identification of the best models for individual forecasts and improving our understanding of how to best design forecasting approaches for ecology in general.
@@
 For example, alternative mechanistic models can be compared to determine which model provides the best forecasts, thus providing insights into the importance of different ecological processes [@dietze2018].
-Iterative near-term forecasting provides the more dynamic interplay between models, predictions, and data that has been identified has necessary for improving ecological forecasting and our understanding of ecological systems more broadly.
+Iterative near-term forecasting provides the more dynamic interplay between models, predictions, and data that has been identified as necessary for improving ecological forecasting and our understanding of ecological systems more broadly.
 
@@
 Data needs to be released quickly under open licenses [@vargas2017; @dietze2018] and structured so that it can be used easily by a variety of researchers and in multiple modeling approaches [@borer2009; @strasser2011].
-Models need to be able to deal with uncertainty, in the predictors and the predictions, to properly convey uncertainty in the resulting forecasts [@diniz-filho2009].
+Models need to be able to deal with uncertainty, in both the predictors and the predictions, to properly convey uncertainty in the resulting forecasts [@diniz-filho2009].
 Multiple models should be developed, both to assess which models are performing best [@dietze2018] and to facilitate combining models to form ensemble predictions which tend to perform better than single models [@araujo2007; @diniz-filho2009].
@@
 
-While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement and few of their recommendations are in widespread use in ecology today.
-We explored what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going long-term ecological study that collects high-frequency data on desert rodent abundances [@brown1998; @ernest2008].
+While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement, and few of their recommendations are in widespread use in ecology today.
+We explored what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going, long-term ecological study that collects high-frequency data on desert rodent abundances [@brown1998; @ernest2008].
 We constructed an automated forecasting pipeline with the goal of being able to forecast rodent abundances and evaluate our predictions on a monthly basis.
@@
 For all rodents caught during a trapping session, information on species identity, size, and reproductive condition is collected, and new individuals are given identification tags.
-This information on rodent populations is high-frequency, uses consistent trapping methodology, and has an extended time-series (469 monthly samples and counting), making this study an ideal case for near-term iterative forecasting.
+This information on rodent populations is high-frequency, uses consistent trapping methodology, and has an extended time-series (470 monthly samples and counting), making this study an ideal case for near-term iterative forecasting.
 
@@
 To go from raw data to forecast presentation involves a number of stages, each of which requires unique tasks, tools and infrastructure.
-The stages of are interdependent with outputs from one stage forming the inputs for the subsequent stage.
+The stages are interdependent, with outputs from one stage forming the inputs for the subsequent stage.
 Tasks in all stages are run using code written in R.
@@
 Each box denotes the core infrastructure used for each stage of the forecasting pipeline.
-Continuous integration (denoted by the Travis icon; a woman wearing safety glasses and hardhat) triggers the code involved in events that link the stages of the pipeline, such as using the output from the forecasting stage (purple box) to create an updated website (rose box).
+Continuous integration (denoted by the Travis icon, a woman wearing safety glasses and hardhat) triggers the code involved in events that link the stages of the pipeline, such as using the output from the forecasting stage (purple box) to create an updated website (rose box).
 Travis also runs tasks within a stage, such as testing code and adding weather data (icons on arrows originating and ending on the same box).
@@
 CI combines computing environments for running code with monitoring systems to identify changes in data or code.
-Essentially, CI is a computer helper whose job is to watch the pipeline and, when it sees a change in the code or data, it runs all the computer scripts needed to ensure that the forecasting pipeline runs from beginning to end.
+Essentially, CI is a computer helper who watches the pipeline and, when it sees a change in the code or data, runs all the computer scripts needed to ensure that the forecasting pipeline runs from beginning to end.
 This is useful for iterative near-term forecasting because it does not rely on humans to create new forecasts whenever new models or data are added.
-These tools are common in the area of software development where they are used to automate software testing and integrate work by multiple developers working on the same code base.
+These tools are common in the area of software development, where they are used to automate software testing and integrate work by multiple developers working on the same code base.
 However, these tools can be used for any computational task that needs to be regularly repeated or run after changes to code or data [@beaulieu2017].
@@
 We experienced this issue when the `tscount` package [@liboschik2015], used by one of our forecasting models, was temporarily removed from CRAN (the R package repository) and could not be installed in the usual way.
-This broke our forecasting pipeline because we could no longer run models that used that package.
+This broke our forecasting pipeline, as we could no longer run models that used that package.
 To make our pipeline robust to changes in external software dependencies, we follow Beaulieu and Greene's [-@beaulieu2017] recommendation to use software containers.
-Software containers are standalone packages that contain copies of everything you need to run some piece of software, including the operating system.
-Once created, a software container is basically a time capsule, it contains all the software dependencies in the exact state used to develop and run the software.
-If those dependencies change (or disappear) in the wider world, they still exist, unchanged, in your container.
+Software containers are standalone packages that contain copies of everything needed to run a given piece of software, including the operating system.
+Once created, a software container is basically a time capsule, containing all the software dependencies in the exact state used to develop and run the software.
+If those dependencies change (or disappear) in the wider world, they still exist, unchanged, in the container.
 We use an existing platform, Docker [@merkel2014], to store an exact image of the complete software environment for running the forecasts.
@@
 Using containers allows us to control transitions to new package versions, implementing them only after we have tested them and made any necessary changes to the data processing and analysis code.
-We use a container created by the Rocker project which is a Docker image with many important R packages (i.e.
+We use a container created by the Rocker project, which is a Docker image with many important R packages (i.e.
 tidyverse) pre-installed  [@boettiger2017].
@@
 Iterative forecasting benefits from frequently updated data so that state changes can be quickly incorporated into new forecasts [@dietze2018].
-Frequent data collection and rapid processing are both important for providing timely forecasts.
+Both frequent data collection and rapid processing are important for providing timely forecasts.
 Since we collect data monthly, ensuring that the models have access to the newest data requires a data latency period of less than 1 month from collection to availability for modeling.
@@
 The local use of the QC scripts to flag problematic data greatly reduces the time spent error-checking and ensures that the quality of data is consistent.
-The data are then uploaded to the GitHub-based Portal Data repository (https://github.com/weecology/PortalData).
+The cleaned data are then uploaded to the GitHub-based PortalData repository (https://github.com/weecology/PortalData).
 GitHub (https://github.com/) is a software development tool for managing computer code development, but we have also found it useful for data management.
-On GitHub, changes to data can be tracked through the Git version control system which logs all changes made to any files in the repository - giving us a record of exactly of when specific lines of data were changed or added.
-All updates to data are processed through "pull requests", which are notifications that someone has a modified version of the data to contribute.
-QA/QC checks are automatically run on the submitted data using continuous integration to ensure that these checks are run and that no avoidable errors reach the official version of the dataset.
+On GitHub, changes to data can be tracked through the Git version control system which logs all changes made to any files in the repository, giving us a record of exactly of when specific lines of data were changed or added.
+All updates to data are processed through "pull requests," which are notifications that someone has a modified version of the data to contribute.
+QA/QC checks are automatically run on the submitted data using continuous integration to ensure that no avoidable errors reach the official version of the dataset.
 
@@
 
-The Portal Project has a long history of making its data publicly available, so that anyone can use it for forecasting or other projects.
-Historically the publication of the data was conducted through data papers [@ernest2009,@ernest2016], the most common approach in ecology; however, this approach caused years of data latency.
+The Portal Project has a long history of making its data publicly available so that anyone can use it for forecasting or other projects.
+Historically, the publication of the data was conducted through data papers [@ernest2009,@ernest2016], the most common approach in ecology; this approach, however, caused years of data latency.
 With the recent switch to posting data directly to a public GitHub repository (Figure 1) with a CC0 waiver (i.e.
-no restrictions on data use; https://creativecommons.org/publicdomain/zero/1.0/) data latency for everyone has been reduced to less than one month making meaningful iterative near-term forecasting possible for not only our group but other interested parties, as well.
+no restrictions on data use; https://creativecommons.org/publicdomain/zero/1.0/), data latency for everyone has been reduced to less than one month, making meaningful iterative near-term forecasting possible for not only our group but other interested parties, as well.
 
@@
 We use a plugin infrastructure to allow new models to be easily added to the system.
-This approach treats each model as an interchangable black box - all models have access to the same input data and generate the same structure for model outputs (Figure 2).
+This approach treats each model as an interchangable black box; all models have access to the same input data and generate the same structure for model outputs (Figure 2).
 During each run of the forecasting code, all existing models are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts.
@@
 This plugin infrastructure makes it easy to add and compare very different types of models, from the basic time-series approaches currently implemented to the more complex state-space and machine learning models we hope to implement in the future.
-As long as a model script can load the provided data and produce the appropriate output it will be run and its results incorporated into the rest of the forecasting system.
+As long as a model script can load the provided data and produce the appropriate output, it will be run and its results incorporated into the rest of the forecasting system.
 
@@
 For example, a model is fit using rodent observations up to June 2005, then used to make a forecast 12 months out to May 2006.
-The observations of that 12 month period can immediately be used to evaluate the model.
+The observations of that 12-month period can immediately be used to evaluate the model.
 Since hindcasting is conducted using data that has already been collected, it allows model comparisons to be conducted on large numbers of hindcasts and provides insight into which models make the best forecasts without needing to wait for new data to be collected [@harris2018].
@@
 Publicly archiving forecasts before new data is collected allows the field to assess, compare, and build on forecasts made by different groups [@mcgill2012; @tredennick2016; @dietze2018; @harris2018] (Figure 1).
-Archiving serves as a form of pre-registration for model predictions, because the forecasts cannot be modified once the data to assess them has been collected.
+Archiving serves as a form of pre-registration for model predictions because the forecasts cannot be modified once the data to assess them has been collected.
 This helps facilitate an unbiased interpretation of model performance.
-To serve this role archives should be publicly accessible and be a permanent record that cannot be changed or deleted.
-The second criteria means that GitHub is not sufficient for archival purposes because repositories can be changed or deleted [@bergman2012; @white2015].
+To serve this role, archives should be publicly accessible and be a permanent record that cannot be changed or deleted.
+This second criterion means that GitHub is not sufficient for archival purposes because repositories can be changed or deleted [@bergman2012; @white2015].
 We explored three major repositories for archiving forecasts: FigShare (https://figshare.com/), Zenodo (https://zenodo.org/), and Open Science Framework (https://osf.io/).
 While all three repositories allowed for easy manual submissions (i.e., a human uploading files after each forecast), automating this process was substantially more difficult.
-Various combinations of repositories, APIs (i.e., interfaces for automatically interacting with the archiving websites) and associated R packages had issues with: 1) integrating authorization with continuous integration; 2) automatically making archived files public; 3) adding new files to an existing location; or 4) automatically permanently archiving the files.
+Various combinations of repositories, APIs (i.e., interfaces for automatically interacting with the archiving websites), and associated R packages had issues with: 1) integrating authorization with continuous integration; 2) automatically making archived files public; 3) adding new files to an existing location; or 4) automatically permanently archiving the files.
 Our eventual solution was to leverage the GitHub-Zenodo integration (https://guides.github.com/activities/citable-code/) and automatically push forecasts to a GitHub repository from the CI server and release them via the GitHub API.
@@
 Each time a new forecast is created, our pipeline adds the new forecasts to the GitHub repository and uses the GitHub API to create a new "release" for that repository.
-This triggers the Zenodo-GitHub integration, which automatically archives the resulting forecasts under a top-level DOI that refers to all archived forecasts ([https://doi.org/10.5281/zenodo.839580](https://doi.org/10.5281/zenodo.839580)).
+This triggers the GitHub-Zenodo integration, which automatically archives the resulting forecasts under a top-level DOI that refers to all archived forecasts ([https://doi.org/10.5281/zenodo.839580](https://doi.org/10.5281/zenodo.839580)).
 Through this process, we automatically archive every forecast made with a documented time-stamp.
@@
 This ensures that every forecast is fully reproducible since the exact code used to generate every forecast is preserved.
-Early forecasts from this system are archived in this modeling and forecasting code archive, not in the newer repository 'forecasts'.
+Early forecasts from this system are archived in the modeling and forecasting code archive, not in the newer repository 'forecasts'.
 
@@
 Information on the species and the field site are also included.
-The site is built using Rmarkdown [@rmarkdown], which naturally integrates into the pipeline, and is automatically updated after each forecast.
+The site is built using Rmarkdown [@rmarkdown], which naturally integrates into the pipeline and is automatically updated after each forecast.
 The `knitr` R package [@xie2015] compiles the code into HTML, which is then published using Github Pages (https://pages.github.com/).
@@
 Our forecasting system automatically acquires and processes the newest data, refits the models, makes new forecasts, publicly archives those forecasts, and presents both the current forecast and information on how previous forecasts performed.
-Every week the forecasting system generates a new set of forecasts with no human intervention, except for the entry of new field data.
+Every week, the forecasting system generates a new set of forecasts with no human intervention, except for the entry of new field data.
 Our forecasting system ensures that forecasts based on the most recent data are always available and is designed to allow rapid assessment of the performance of multiple forecasting models for a number of different states of the system, including the abundances of individual species and community-level variables such as total abundance.
-To create this iterative near-term forecasting system, we used R to process data and conduct analyses, and leveraged existing tools and services (i.e.
+To create this iterative near-term forecasting system, we used R to process data and conduct analyses and leveraged existing tools and services (i.e.
 GitHub, Travis, Docker) for more complicated cyberinfrastructure tasks.
@@
 Finally, we found no satisfactory existing solution for archiving our results.
-All approaches we tried had limitations when it came to automatically generating publicly versioned archives of forecasts on a repeated basis, and our eventual solution was difficult to configure to such a degree that it will an impediment for most researchers.
+All approaches we tried had limitations when it came to automatically generating publicly-versioned archives of forecasts on a repeated basis, and our eventual solution was difficult to configure to such a degree that it will remain an impediment for most researchers.
 Overall, we found existing technology to be sufficient to the task of creating an iterative forecasting pipeline, but it required greater expertise and a greater investment of time than is ideal.
 Additional tool development to reduce the effort required for scientists to set up their own short-term forecasting systems would clearly be useful.
-However, our efforts show that it is possible to use existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems.
+Our efforts, however, show that it is possible to use existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems.
 
@@
 By regularly running different models for population and community dynamics, a near-term iterative pipeline such as ours should also make it possible to rapidly detect changes in how the system is operating, which should allow the rapid identification of ecological transitions or even possibly allow them to be prevented [@pace2017].
-By building an automated iterative near-term forecasting infrastructure we can improve our ability to forecast natural systems, understand the biology driving ecological dynamics, and detect or even predict changes in system state that are important for conservation and management.
+By building an automated iterative near-term forecasting infrastructure, we can improve our ability to forecast natural systems, understand the biology driving ecological dynamics, and detect or even predict changes in system state that are important for conservation and management.
 

### 2018-02-13 evaluation.Rmd
Turn duplicates test back on, typos in evaluation

@@
 
+models_of_interest = c('AutoArima','Ensemble','ESSS','nbGARCH','pevGARCH')
+
 species_of_interest = c('BA','DM','DO','PP','OT','NA','total')
@@
 new_moons = read.csv(new_moon_file)
+new_moons$censusdate = as.Date(new_moons$censusdate)
 
@@
 
+observation_data = species_abundance %>%
+  bind_rows(total_abundance)
+
 ```
 
-## How have the models recently?
+## How have the models done recently?
 
@@
 ```{r recent_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=15, fig.height=15}
-# The length of time to look backward
-prior_new_moons = 12
-
-most_recent_newmoon = max(new_moons$newmoonnumber)
-earliest_newmoon_to_graph = most_recent_newmoon - prior_new_moons
-
-observation_data = species_abundance %>%
-  bind_rows(total_abundance) %>%
-  filter(newmoonnumber %in% earliest_newmoon_to_graph:most_recent_newmoon)
-
+###############################################
 #Get all the recent forecasts
 forecast_data = compile_forecasts(use_hindcasts = FALSE) %>%
-  filter(newmoonnumber %in% earliest_newmoon_to_graph:most_recent_newmoon)
+  filter(model %in% models_of_interest)
 
@@
   select(-keep)
+
 ##################################
-
+#Combine forecasts and oservations
 forecast_errors = forecast_data %>%
@@
 
-###################################
-forecast_date_labels = forecast_errors %>%
-  group_by(date, model, species, censusdate) %>%
-  summarize(y_position = mean(rmse)) %>%
-  ungroup() %>%
-  distinct() %>%
-  filter(!is.na(censusdate)) %>%
-  group_by(date, model, species) %>%
-  filter(as.Date(censusdate) == min(as.Date(censusdate))) %>%
-  ungroup() %>%
-  distinct() %>%
-  mutate(label_text = paste0('Issued: ',date))
+#############################################################
+# Displaying the data from many forecasts in the past can create a mess of a plot. see #218.
+# So here three forecasts are chosen from 1, 6, and 12 newmoons into the past. 
+# Sometimes forecasts data is missing for a particular initial_newmoon, the recursive function here
+# attempts to find another one from the same time pluts/minus 2 month period. This ensures 3 forecasts
+# are always shown.
 
-####################################
-# Setup a nice palette for when prior issue dates become greater than 9
-getPalette = colorRampPalette(RColorBrewer::brewer.pal(9, "Set1"))
-set.seed(2)
-large_color_palette = sample(getPalette(40))
+# If there is no data for initial_newmoon 500, this will check nearby dates in this order: 499,501,498,402
+update_initial_newmoon = function(potential_initial_newmoon, attempt=1){
+  if(attempt==6){
+    return(NA)
+  }
+  lead_time_data = forecast_errors %>%
+    filter(initial_newmoon == potential_initial_newmoon)
+
+  # Alternate steps of -1, +2, -3, +4
+  next_try_step = ifelse(attempt%%2 == 0, attempt, attempt * -1)
+
+  # no data to show? then try one month back
+  if(nrow(lead_time_data)==0){
+    #print(paste('Try ',attempt, ', Nothing with ',potential_initial_newmoon,' trying ',potential_initial_newmoon+next_try_step))
+    return(update_initial_newmoon(potential_initial_newmoon+next_try_step, attempt+1))
+  } else {
+    return(potential_initial_newmoon)
+  }
+
+}
+# Adjust this to add more than 3 forecast dates, or change the lead times
+lead_times_to_show = c(1,6,12)
+most_recent_observed_newmoon = max(new_moons$newmoonnumber)
+initial_newmoons_to_keep = most_recent_observed_newmoon - lead_times_to_show
+
+initial_newmoons_to_keep = sapply(initial_newmoons_to_keep, update_initial_newmoon)
+
+forecast_errors = forecast_errors %>%
+  filter(initial_newmoon %in% initial_newmoons_to_keep)
+
 #####################################
@@
   geom_line(size=1.5) +
-  # label individual issue dates. This is easier to look at if there 
-  # gets to be a lot of them. Must install ggrepel and remove the legend if this is turned on.
-  #geom_label_repel(data=forecast_date_labels, aes(label=label_text, y=y_position),
-  #                 fill='grey90',size=4,nudge_y = 40, force=5,
-  #                 segment.alpha = 0.3, segment.size = 0.8) +
   scale_color_brewer(palette='Dark2') +
-  #scale_color_manual(values=large_color_palette) +
-  facet_grid(species~model) + 
+  scale_x_date(breaks = unique(forecast_errors$censusdate)) +
+  facet_grid(species~model, scales = 'free_y') + 
   theme_bw() +
-  theme(axis.text.x = element_text(angle=45, hjust=0.9, debug = FALSE),
+  theme(axis.text.x = element_text(angle=45, hjust=0.99, debug = FALSE),
         axis.text = element_text(size=12),
@@
 
-observation_data = species_abundance %>%
-  bind_rows(total_abundance)
-
 #Get the all the forecasts made during observation period

### 2018-02-14 index.Rmd
add heading for most abundance species plots

@@
 ```
+
+## Most Abundant Species
+
 ```{r highabund_ts_forecasts, echo=FALSE, message=FALSE, warning=FALSE}

### 2018-02-14 index.Rmd
add brief intro blurb

@@
 ---
+
+Welcome to Portal Forecasting!
+This is a website run by the [Weecology](http://weecology.org/) group, a group of interdisciplinary ecologists at the [University of Florida](http://www.wec.ufl.edu/).
+On this website, you'll find information about our ongoing efforts to forecast ecological systems from time series.
+Specifically, we are using a times series of rodent abundances from [The Portal project](http://portal.weecology.org/), a long-term experimental monitoring project in desert ecology.
+Enjoy!
 

### 2018-02-14 profiles.Rmd
add color pictures

--- 2017-09-27 profiles.Rmd
@@
 text_tbl <- data.frame(
+  Rodents = c("![](rodent_pictures/color_pictures/BA.png)",
+              "![](rodent_pictures/color_pictures/PB.png)",
+              "![](rodent_pictures/color_pictures/PH.png)",
+              "![](rodent_pictures/color_pictures/PI.png)",
+              "![](rodent_pictures/color_pictures/PP.png)",
+              "![](rodent_pictures/color_pictures/DM.png)",
+              "![](rodent_pictures/color_pictures/DO.png)",
+              "![](rodent_pictures/color_pictures/DS.png)",
+              "![](rodent_pictures/color_pictures/NA.png)",
+              "![](rodent_pictures/color_pictures/OL.png)",
+              "![](rodent_pictures/color_pictures/OT.png)",
+              "![](rodent_pictures/color_pictures/PF.png)",
+              "![](rodent_pictures/color_pictures/PE.png)",
+              "![](rodent_pictures/color_pictures/PL.png)",
+              "![](rodent_pictures/color_pictures/PM.png)",
+              "![](rodent_pictures/color_pictures/RF.png)",
+              "![](rodent_pictures/color_pictures/RM.png)",
+              "![](rodent_pictures/color_pictures/RO.png)",
+              "![](rodent_pictures/color_pictures/SF.png)",
+              "![](rodent_pictures/color_pictures/SH.png)",
+              "![](rodent_pictures/color_pictures/SO.png)"), 
   Scientific.Name = c("_Baiomys taylori_", 

### 2018-02-14 models.Rmd
adding models page to website
including models page in the site yml file
adding models Rmd file
adding refs bibtex file referenced in models Rmd file
NOT RENDERING NEW SITE

---
title: "Current Models"
bibliography: refs.bibtex
---

We currently analyze and forecast rodent data at Portal using four models:

## ESSS

ESSS (Exponential Smoothing State Space) is a flexible exponential smoothing 
state space model [@Hyndman2008] fit to the data at the composite (full site
and just control plots) spatial level and the composite (community) ecological
level. The model is selected and fitted using the `ets` and `forecast` 
functions in the **forecast** package [@Hyndman2017] with the 
`allow.multiplicative.trend` argument set to `TRUE` and the `forecast_ess`
function in our **portalPredictions** package [@PortalPredictions]. 
Models fit using `ets` implement what is known as the "innovations" approach 
to state space modeling, which assumes a single source of noise that is 
equivalent for the process and observation errors [@Hyndman2008].

In general, ESSS models are defined according to three model structure 
parameters: error type, trend type, and seasonality type [@Hyndman2008].
Each of the parameters can be an N (none), A (additive), or M (multiplicative)
state [@Hyndman2008]. However, because of the difference in period 
between seasonality and sampling of the Portal rodents combined with the 
hard-coded single period of the `ets` function, we could not include the 
seasonal components to the ESSS model. ESSS is fit flexibly, such that the 
model parameters can vary from fit to fit.

## AutoARIMA

AutoArima (Automatic Auto-Regressive Integrated Moving Average) is a 
flexible Auto-Regressive Integrated Moving Average (ARIMA) model fit to the 
data at the composite (full site and just control plots) spatial level
and the composite (community) ecological level. The model is selected and 
fitted using the `auto.arima` (with the `lambda` argument set to 0) and 
`forecast` functions in the **forecast** package [@Hyndman2013; @Hyndman2017]
and the `forecast_autoarima` function in our **portalPredictions** package 
[@PortalPredictions].

Generally, ARIMA models are defined according to three model structure 
parameters: the number of autoregressive terms (p), the degree of differencing 
(d), and the order of the moving average (q), and are represented as 
ARIMA(p, q, q) [@Box1970]. While the `auto.arima` function allows for seasonal
models, the seasonality is hard-coded to be on the same period as the sampling,
which is not the case for the Portal rodent surveys. As a result, no seasonal 
models were evaluated. AutoArima is fit flexibly, such that the model 
parameters can vary from fit to fit.

## nbGARCH

nbGARCH (Negative Binomial Auto-Regressive Conditional Heteroskedasticity) 
is a generalized autoregressive conditional heteroskedasticity (GARCH) 
model with overdispersion (*i.e.*, a negative binomial response) fit to the 
data at the composite (full site and just control plots) spatial level and both 
the composite (community) and the articulated (species) ecological levels. The 
model for each species and the community total is selected and fitted using the
`tsglm` function in the **tscount** package [@Liboschik2017a] and the 
`forecast_nbgarch` function in our **portalPredictions** package 
[@PortalPredictions]

GARCH models are generalized ARMA models and are defined according to their 
link function, response distribution, and two model structure parameters: the 
number of autoregressive terms (p) and the order of the moving average (q), 
and are represented as GARCH(p, q) [@Liboschik2017a]. The nbGARCH model is fit
using the log link and a negative binomial response (modeled as an
over-dispersed Poisson), as well as with p = 1 (first-order autoregression)
and q = 12 (approximately yearly moving average).

The `tsglm` function in the **tscount** package [@Liboschik2017a] uses a 
(conditional) quasi-likelihood based approach to inference and models the 
overdispersion as an additional parameter in a two-step approach. This 
two-stage approach has only been minimally evaluated, although preliminary 
simulation-based studies are promising [@Liboschik2017b].

## pevGARCH

pevGARCH (Poisson Environmental Variable Auto-Regressive Conditional
Heteroskedasticity) is a generalized autoregressive conditional
heteroskedasticity (GARCH) model fit to the data at the composite (full site 
and just control plots) spatial level and both the composite (community) and 
the articulated (species) ecological levels. The response variable is Poisson, 
and a variety of environmental variables are considered as covariates. The 
model for each species is selected and fitted using the `tsglm` function in the
**tscount** package [@Liboschik2017a] and the `forecast_pevgarch` function in
our **portalPredictions** package [@PortalPredictions].

GARCH models are generalized ARMA models and are defined according to their 
link function, response distribution, and two model structure parameters: 
the number of autoregressive terms (p) and the order of the moving average 
(q), and are represented as GARCH(p, q) [@Liboschik2017a]. The PETS model is 
fit using the log link and a Poisson response, as well as with p = 1 
(first-order autoregression) and q = 12 (yearly moving average). The 
environmental variables potentially included in the model are min, mean, and
max temperatures, precipitation, and NDVI.

The `tsglm` function in the **tscount** package [@Liboschik2017a] uses a
(conditional) quasi-likelihood based approach to inference. This approach has
only been minimally evaluated for models with covariates, although preliminary
simulation-based studies are promising [@Liboschik2017b].

Each species is fit using the following (nonexhaustive) sets of the 
environmental covariates:

  * max temp, mean temp, precipitation, NDVI
  * max temp, min temp, precipitation, NDVI
  * max temp, mean temp, min temp, precipitation
  * precipitation, NDVI
  * min temp, NDVI
  * min temp
  * max temp
  * mean temp
  * precipitation 
  * NDVI
  * -none-

The final model is an intercept-only model. The single best model of the 11 is 
selected based on AIC.

# References

### 2018-02-15 evaluation.Rmd

@@
 # Setup rodent observations
-species_abundance = portalr::abundance(shape='flat', level='Site') %>%
+species_abundance = portalr::abundance(shape='flat', level='Controls') %>%
   rename(actual = abundance) %>%
-  mutate(level='All',currency='abundance') %>%
+  mutate(level='Controls',currency='abundance') %>%
   left_join(new_moons, by='period') %>%
@@
 
-total_abundance = portalr::abundance(shape='flat', level='Site') %>%
+total_abundance = portalr::abundance(shape='flat', level='Controls') %>%
   group_by(period) %>%
@@
   ungroup() %>%
-  mutate(level='All',currency='abundance',species='total') %>%
+  mutate(level='Controls',currency='abundance',species='total') %>%
   left_join(new_moons, by='period')
@@
 This is shown for all model types in the columns and several common rodent species (and total rodents) in the rows.
-These forecasts are for the entire site.
+These evaluations are for the Control plots where no manipulation is done.
 ```{r recent_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=15, fig.height=15}
@@
 The error values are an average of all forecast errors using observations since 2010.
-Note that this currently uses hindcasts of the prior observations.
+Note that this currently uses hindcasts of the prior observations, and is also only for the Control plots.
 

### 2018-02-15 index.Rmd

--- 2018-02-02 index.Rmd
@@
 sp_predictions = get_sp_predicts(ensemble, 'All', lead_time = 1)
-sp_predict = plot_species_forecast(sp_predictions)
+sp_predict = plot_species_forecast(sp_predictions,title=paste0(sp_predictions$forecast_date[2], ": All plots"))
 plot(sp_predict)

### 2018-02-15 report.Rmd

@@
 ---
-Here's the most recent data (blue) compared to the forecast performed right before data collection (black, with error bars)
+Here's the most recent actual data (blue) compared to the forecast performed right before data collection (black, with error bars)
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+source("tools/forecast_tools.R")
+library(portalr)
+
+# get most recent rodent data
+rodents = abundance("repo", shape='flat', level='treatment')
+most_recent_period = max(rodents$period)
+observed = dplyr::filter(rodents, period == most_recent_period, treatment=='control')
+
+# get new moon number associated with this period
+newmoons_table = read.csv(text = RCurl::getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
+newmoons_table$censusdate = as.Date(newmoons_table$censusdate,format='%Y-%m-%d')
+moon_code = dplyr::filter(newmoons_table,period==most_recent_period) %>% select(newmoonnumber) %>% as.integer()
+most_recent_date = dplyr::filter(newmoons_table,period==most_recent_period) %>% dplyr::pull(censusdate)
+
+# get most recent forecast prior to data collection
+files = data.frame(names = list.files(path = './predictions', pattern = '*forecasts.csv'))
+files$filedates = as.Date(substr(files$names,1,10),format='%Y-%m-%d')
+latest_forecast = dplyr::filter(files,filedates<most_recent_date) %>% tail(1)
+
+data = read.csv(paste('predictions/',latest_forecast$names,sep=''),na.strings = '')
+
+ensemble = dplyr::filter(data, level == 'Controls', model == 'Ensemble', newmoonnumber==moon_code)
+sp_predictions = get_sp_predicts(ensemble, 'Controls', lead_time = 1)
+
+joined_data = left_join(sp_predictions, observed, by = "species")
+joined_data[is.na(joined_data)] = 0
+joined_data[joined_data$species=='total','abundance'] = sum(joined_data$abundance,na.rm=T)
+joined_data[joined_data$species=='total','period'] = most_recent_period
+
+sp_predict = plot_species_forecast(sp_predictions,title=paste0(sp_predictions$forecast_date[2], ": Control plots"))  
+sp_predict = sp_predict + geom_point(data = joined_data, mapping = aes(x = abundance, y = species),
+             color = "blue")
+plot(sp_predict)
+```
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-sp_predict = plot_species_forecast(sp_predictions)  
+sp_predict = plot_species_forecast(sp_predictions,title=paste0(sp_predictions$forecast_date[2], ": All plots"))  
 sp_predict = sp_predict + geom_point(data = joined_data, mapping = aes(x = abundance, y = species),

### 2018-02-15 evaluation.Rmd
Use longterm controls in website reporting

@@
 # Setup rodent observations
-species_abundance = portalr::abundance(shape='flat', level='Controls') %>%
-  rename(actual = abundance) %>%
-  mutate(level='Controls',currency='abundance') %>%
+species_abundance = portalr::abundance(shape='flat', level='treatment', length="longterm") %>%
+  filter(treatment=="control") %>%
+  rename(actual = abundance, level = treatment) %>%
+  mutate(currency='abundance') %>%
   left_join(new_moons, by='period') %>%
@@
 
-total_abundance = portalr::abundance(shape='flat', level='Controls') %>%
+total_abundance = portalr::abundance(shape='flat', level='treatment', length="longterm") %>%
+  filter(treatment=="control") %>%
   group_by(period) %>%
@@
   ungroup() %>%
-  mutate(level='Controls',currency='abundance',species='total') %>%
+  mutate(level='treatment',currency='abundance',species='total') %>%
   left_join(new_moons, by='period')

### 2018-02-15 index.Rmd
Use longterm controls in website reporting

--- 2018-02-14 index.Rmd
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-obs_data = portalr::abundance("repo")
-obs_data$total = rowSums(select(obs_data, -period))
+obs_data = portalr::abundance(level='treatment', length="longterm") %>%
+  filter(treatment=="control")
+obs_data$total = rowSums(select(obs_data, -c(period, treatment)))
 new_moons = read.csv(text = RCurl::getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
@@
 for_data = compile_forecasts() %>%
-  dplyr::filter(level == 'All', model == 'Ensemble', date == max(date))
+  dplyr::filter(level == 'Controls', model == 'Ensemble', date == max(date))
 
@@
 data = compile_forecasts()
-ensemble = dplyr::filter(data, level == 'All', model == 'Ensemble', date == max(date))
-sp_predictions = get_sp_predicts(ensemble, 'All', lead_time = 1)
+ensemble = dplyr::filter(data, level == 'Controls', model == 'Ensemble', date == max(date))
+sp_predictions = get_sp_predicts(ensemble, 'Contols', lead_time = 1)
 sp_predict = plot_species_forecast(sp_predictions)

### 2018-02-15 index.Rmd
Cleanup report.Rmd

--- 2018-02-15 index.Rmd
@@
 obs_data_newmoon$censusdate = as.Date(obs_data_newmoon$censusdate)
-#for_data = read.csv("predictions/2017-02-24Allforecasts.csv")
 

### 2018-02-15 index.Rmd

@@
 sp_predictions = get_sp_predicts(ensemble, 'Contols', lead_time = 1)
-sp_predict = plot_species_forecast(sp_predictions)
+sp_predict = plot_species_forecast(sp_predictions,
+            title=paste0(sp_predictions$forecast_date[2], ": Control plots"))
+
 plot(sp_predict)

### 2018-02-15 evaluation.Rmd
Switch to using repo data for all website reporting

--- 2018-02-15 evaluation.Rmd
@@
 # Setup rodent observations
-species_abundance = portalr::abundance(shape='flat', level='Controls') %>%
-  rename(actual = abundance) %>%
-  mutate(level='Controls',currency='abundance') %>%
+species_abundance = portalr::abundance("repo", shape='flat', level='treatment', length="longterm") %>%
+  filter(treatment=="control") %>%
+  rename(actual = abundance, level = treatment) %>%
+  mutate(currency='abundance') %>%
   left_join(new_moons, by='period') %>%
@@
 
-total_abundance = portalr::abundance(shape='flat', level='Controls') %>%
+total_abundance = portalr::abundance("repo", shape='flat', level='treatment', length="longterm") %>%
+  filter(treatment=="control") %>%
   group_by(period) %>%

### 2018-02-15 index.Rmd
Switch to using repo data for all website reporting

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-obs_data = portalr::abundance(level='treatment', length="longterm") %>%
+obs_data = portalr::abundance("repo", level='treatment', length="longterm") %>%
   filter(treatment=="control")

### 2018-02-16 evaluation.Rmd
Assorted fixes, render site

--- 2018-02-15 evaluation.Rmd
@@
 # Setup rodent observations
-species_abundance = portalr::abundance(shape='flat', level='Controls') %>%
-  rename(actual = abundance) %>%
+species_abundance = portalr::abundance("repo", shape='flat', level='treatment', 
+                                       length="longterm", time='newmoon') %>%
+  filter(treatment=="control") %>%
+  rename(actual = abundance, level = treatment) %>%
   mutate(level='Controls',currency='abundance') %>%
-  left_join(new_moons, by='period') %>%
   filter(species %in% species_of_interest)
 
-total_abundance = portalr::abundance(shape='flat', level='Controls') %>%
-  group_by(period) %>%
+total_abundance = portalr::abundance("repo", shape='flat', level='treatment', 
+                                     length="longterm", time='newmoon') %>%
+  filter(treatment=="control") %>%
+  group_by(newmoonnumber) %>%
   summarise(actual=sum(abundance)) %>%
   ungroup() %>%
-  mutate(level='Controls',currency='abundance',species='total') %>%
-  left_join(new_moons, by='period')
+  mutate(level='Controls',currency='abundance',species='total')
 
 observation_data = species_abundance %>%
-  bind_rows(total_abundance)
+  bind_rows(total_abundance) %>% 
+  left_join(new_moons, by='newmoonnumber')
 
@@
 forecast_errors = forecast_data %>%
-  left_join(observation_data, by=c('species','level','currency','newmoonnumber')) %>% 
+  left_join(observation_data, by=c('newmoonnumber','level','species','currency')) %>% 
   mutate(rmse = sqrt((estimate - actual)^2))

### 2018-02-16 index.Rmd

@@
 ensemble = dplyr::filter(data, level == 'Controls', model == 'Ensemble', date == max(date))
-sp_predictions = get_sp_predicts(ensemble, 'Contols', lead_time = 1)
+sp_predictions = get_sp_predicts(ensemble, 'Controls', lead_time = 1)
 sp_predict = plot_species_forecast(sp_predictions,

### 2018-02-16 profiles.Rmd

@@
               "![](rodent_pictures/color_pictures/NA.png)",
-              "![](rodent_pictures/color_pictures/OL.png)",
-              "![](rodent_pictures/color_pictures/OT.png)",
+              "![](rodent_pictures/color_pictures/OL.JPG)",
+              "![](rodent_pictures/color_pictures/OT.JPG)",
               "![](rodent_pictures/color_pictures/PF.png)",
@@
               "![](rodent_pictures/color_pictures/PM.png)",
-              "![](rodent_pictures/color_pictures/RF.png)",
-              "![](rodent_pictures/color_pictures/RM.png)",
-              "![](rodent_pictures/color_pictures/RO.png)",
-              "![](rodent_pictures/color_pictures/SF.png)",
-              "![](rodent_pictures/color_pictures/SH.png)",
-              "![](rodent_pictures/color_pictures/SO.png)"), 
+              "![](rodent_pictures/color_pictures/RF.jpg)",
+              "![](rodent_pictures/color_pictures/RM.jpg)",
+              "![](rodent_pictures/color_pictures/RO.JPG)",
+              "![](rodent_pictures/color_pictures/SF.jpg)",
+              "![](rodent_pictures/color_pictures/SH.jpg)",
+              "![](rodent_pictures/color_pictures/SO.jpg)"), 
   Scientific.Name = c("_Baiomys taylori_", 

### 2018-02-16 report.Rmd

@@
 # get most recent rodent data
-rodents = abundance("repo", shape='flat', level='treatment')
+rodents = abundance("repo", shape='flat', level='treatment', length='longterm')
 most_recent_period = max(rodents$period)

### 2018-02-16 evaluation.Rmd

@@
 These evaluations are for the Control plots where no manipulation is done.
-```{r recent_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=15, fig.height=15}
+```{r recent_eval, echo=FALSE, message=FALSE, warning=FALSE,, fig.width=15, fig.height=15}
 ###############################################
@@
 
-```{r hindcast_eval, echo=FALSE, message=TRUE, warning=FALSE,, fig.width=9, fig.height=15}
+```{r hindcast_eval, echo=FALSE, message=FALSE, warning=FALSE,, fig.width=9, fig.height=15}
 

### 2018-02-19 evaluation.Rmd
Correctly state that prediction intervals are 90% in evaluation
The prediction intervals have always been 90%, but they were described and shown
as 95%. This fixes this issue and in the process creates a new metadata value
for the prediction interval (thanks to @ha0ye's suggestion) that can be used
throughout the code base. Updates to the models will be necessary to use this
value for all instances of prediction interval in the code base.

@@
 library(portalr)
+library(yaml)
 source("tools/forecast_tools.R")
 
+model_metadata = yaml.load_file("data/model_metadata.yaml")
+pred_interval = model_metadata$pred_interval
 models_of_interest = c('AutoArima','Ensemble','ESSS','nbGARCH','pevGARCH')
@@
 **RMSE**: Root mean square error, this is a metric used to evaluate the point estimate of a forecast.  
-**Coverage**: This is the percentage of observations which fell within the 95% confidence interval of a forecast.
-Ideally this would be equal to 0.95.
-If it's higher than 0.95 the forecasts intervals are too wide, if it's lower then the forecast intervals are too narrow.
+**Coverage**: This is the percentage of observations which fell within the `r 100 * pred_interval`% confidence interval of a forecast.
+Ideally this would be equal to `r pred_interval`.
+If it's higher than `r pred_interval` the forecasts intervals are too wide, if it's lower then the forecast intervals are too narrow.
 
@@
   geom_line() +
-  geom_hline(yintercept = 0.95) +
+  geom_hline(yintercept = 0.9) +
   labs(x='Lead Time (New Moons)') +

### 2018-02-20 manuscript/portal_iterative_forecasting_ms.Rmd
intro wording
ellen's comma
Add author list to paper
Also update the author list in .zenodo.json to include Juniper

@@
 title: "Automated iterative near-term forecasting for the Portal Project"
+author:
+- Ethan P. White$^{1,2,3}$
+- Glenda M. Yenni$^1$
+- Shawn D. Taylor$^4$
+- Erica M. Christensen$^1$
+- Ellen K. Bledsoe$^4$
+- Juniper L. Simonis$^1$
+- S. K. Morgan Ernest$^{1,3}$
 csl: mee.csl
-fontsize: 12pt
 output:
@@
   html_document: default
+geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
 header-includes:
@@
 - \linenumbers
-geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
+fontsize: 12pt
 bibliography: refs.bib
@@
 \raggedright
+
+$^1$ Department of Wildlife Ecology and Conservation, University of Florida, Gainesville, FL, United States
+
+$^2$ Informatics Institute, University of Florida, Gainesville, FL, United States
+
+$^3$ Biodiversity Institute, University of Florida, Gainesville, FL, United States
+
+$^4$ School of Natural Resources and Environment, University of Florida Gainesville, FL, United States
 
@@
 The lack of active assessment results in limited information on how much confidence to place in forecasts and makes it difficult to determine on which forecasting methods to build.
-Without regular updates, forecasts lack the most current data, and forecast accurracy will decay over time [@petch2015; @dietze2018].
+Without regular updates, forecasts lack the most current data, and the longer a forecast remains out of date, the less accurate it becomes [@petch2015; @dietze2018].
 More regular updating and assessment will advance ecological forecasting as a field by accelerating the identification of the best models for individual forecasts and improving our understanding of how to best design forecasting approaches for ecology in general.

### 2018-02-20 manuscript/portal_iterative_forecasting_ms.Rmd
Fix broken references

@@
 output:
+  html_document: default
   pdf_document:
     keep_tex: yes
-  html_document: default
 geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
@@
 The lack of active assessment results in limited information on how much confidence to place in forecasts and makes it difficult to determine on which forecasting methods to build.
-Without regular updates, forecasts lack the most current data, and the longer a forecast remains out of date, the less accurate it becomes [@petch2015; @dietze2018].
+Without regular updates, forecasts lack the most current data, and the longer a forecast remains out of date, the less accurate it becomes [@petchey2015; @dietze2018].
 More regular updating and assessment will advance ecological forecasting as a field by accelerating the identification of the best models for individual forecasts and improving our understanding of how to best design forecasting approaches for ecology in general.
@@
 This means interdisciplinary teams will generally be required for creating these pipelines until tool development improves.
-To improve the success of these diverse groups, we believe efforts at providing 'team science' training to scientists interested in forecasting will be beneficial for the success of iterative forecasting attempts for the foreseeable future  [@read2017].
+To improve the success of these diverse groups, we believe efforts at providing 'team science' training to scientists interested in forecasting will be beneficial for the success of iterative forecasting attempts for the foreseeable future  [@read2016].
 

### 2018-02-20 manuscript/portal_iterative_forecasting_ms.Rmd
Minor cleanup and additions for manuscript submission
* Put all authors on one line
* Remove Juniper from acknowledgements now that they are an author
* Add Data Accessibility section including associated Zenodo references

@@
 title: "Automated iterative near-term forecasting for the Portal Project"
-author:
-- Ethan P. White$^{1,2,3}$
-- Glenda M. Yenni$^1$
-- Shawn D. Taylor$^4$
-- Erica M. Christensen$^1$
-- Ellen K. Bledsoe$^4$
-- Juniper L. Simonis$^1$
-- S. K. Morgan Ernest$^{1,3}$
+author: Ethan P. White$^{1,2,3}$, Glenda M. Yenni$^1$, Shawn D. Taylor$^4$, Erica
+  M. Christensen$^1$, Ellen K. Bledsoe$^4$, Juniper L. Simonis$^1$, S. K. Morgan Ernest$^{1,3}$
 csl: mee.csl
 output:
-  html_document: default
   pdf_document:
     keep_tex: yes
+  html_document: default
 geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
@@
 
-* Most forecasts for the future state of ecological systems are conducted once and never updated or assessed.
+1.
+Most forecasts for the future state of ecological systems are conducted once and never updated or assessed.
 As a result, many available ecological forecasts are not based on the most up-to-date data, and the scientific progress of ecological forecasting models is slowed by a lack of feedback on how well the forecasts perform.
-* Iterative near-term ecological forecasting involves repeated daily to annual scale forecasts of an ecological system as new data becomes available and regular assessment of the resulting forecasts.
+2.
+Iterative near-term ecological forecasting involves repeated daily to annual scale forecasts of an ecological system as new data becomes available and regular assessment of the resulting forecasts.
 We demonstrate how automated iterative near-term forecasting systems for ecology can be constructed by building one to conduct monthly forecasts of rodent abundances at the Portal Project, a long-term study with over 40 years of monthly data.
 This system automates most aspects of the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presentation of the forecasts.
-* The forecasting system uses R code for working with data, fitting models, making forecasts, and archiving and presenting these forecasts.
+3.
+The forecasting system uses R code for working with data, fitting models, making forecasts, and archiving and presenting these forecasts.
 The resulting pipeline is automated using continuous integration (a software development tool) to run the entire pipeline once a week.
@@
 Constructing this forecasting system required a team with expertise ranging from field site experience to software development.
-* Automated near-term iterative forecasting systems will allow the science of ecological forecasting to advance more rapidly and provide the most up-to-date forecasts possible for conservation and management.
+4.
+Automated near-term iterative forecasting systems will allow the science of ecological forecasting to advance more rapidly and provide the most up-to-date forecasts possible for conservation and management.
 These forecasting systems will also accelerate basic science by allowing new models of natural systems to be quickly implemented and compared to existing models.
@@
 
-We thank Henry Senyondo for help with continuous integration and Hao Ye and Juniper Simonis for discussions and feedback on the manuscript.
+We thank Henry Senyondo for help with continuous integration and Hao Ye for discussions and feedback on the manuscript.
 We thank all of the graduate students, postdocs, and volunteers who have collected the Portal Project over the last 40 years and the developers of the software and tools that made this project possible.
@@
 White.
+
+# Data Accessibility
+
+The data used in this study is from the Portal Project and is openly available (CC0) on GitHub (https://github.com/weecology/PortalData).
+Code for reproducing all analyses is available on GitHub (https://github.com/weecology/portalPredictions) and archived on Zenodo [@zenodocode].
+Forecasts made by this system are all archived to Zenodo [@zenodoforecasts].
 

### 2018-02-20 manuscript/portal_iterative_forecasting_ms.Rmd
Improve title

@@
 ---
-title: "Automated iterative near-term forecasting for the Portal Project"
+title: Developing an automated iterative near-term forecasting system for an ecological
+  study
 author: Ethan P. White$^{1,2,3}$, Glenda M. Yenni$^1$, Shawn D. Taylor$^4$, Erica
@@
 The Portal Project has a long history of making its data publicly available so that anyone can use it for forecasting or other projects.
-Historically, the publication of the data was conducted through data papers [@ernest2009,@ernest2016], the most common approach in ecology; this approach, however, caused years of data latency.
+Historically, the publication of the data was conducted through data papers [@ernest2009, @ernest2016], the most common approach in ecology; this approach, however, caused years of data latency.
 With the recent switch to posting data directly to a public GitHub repository (Figure 1) with a CC0 waiver (i.e.

### 2018-03-16 index.Rmd

@@
 source("tools/forecast_tools.R")
-source("tools/model_functions.R")
+source("tools/data_tools.R")
 ```

### 2018-03-19 models.Rmd
update the model script and description

@@
 and the composite (community) ecological level. The model is selected and 
-fitted using the `auto.arima` (with the `lambda` argument set to 0) and 
-`forecast` functions in the **forecast** package [@Hyndman2013; @Hyndman2017]
-and the `forecast_autoarima` function in our **portalPredictions** package 
-[@PortalPredictions].
+fitted using the `auto.arima` and `forecast` functions in the **forecast**
+package [@Hyndman2013; @Hyndman2017] and the `forecast_autoarima` function 
+in our **portalPredictions** package [@PortalPredictions].
 

### 2018-04-13 evaluation.Rmd
Continue to get latest data from portalr (following update in PortalData versioning scheme)

@@
 # Setup rodent observations
-species_abundance = portalr::abundance("repo", shape='flat', level='treatment', 
+species_abundance = portalr::abundance("repo", clean=FALSE, shape='flat', level='treatment', 
                                        length="longterm", time='newmoon') %>%
@@
 
-total_abundance = portalr::abundance("repo", shape='flat', level='treatment', 
+total_abundance = portalr::abundance("repo", clean=FALSE, shape='flat', level='treatment', 
                                      length="longterm", time='newmoon') %>%

### 2018-04-13 PortalHindcasts.R

@@ -14,7 +14,7 @@ forecast_date = Sys.Date()
 initial_time_newmoons=490:403
 
 #Get the most recent data loaded into the data folder
-portalr::download_observations(release_only=FALSE)
+portalr::download_observations()
 moons = get_moon_data()
 rodent_data = get_rodent_data(moons, forecast_date)
 weather_data = get_weather_data(moons, rodent_data$all, lag=6)

### 2018-04-13 tools/data_tools.R

@@ -52,7 +52,7 @@ get_rodent_data <- function(moons, forecast_date){
   #get Portal abundance data for the entire site and for control plots only.
   
   #Control plots
-  controls = portalr::abundance(level="Treatment",type="Rodents",length="Longterm", incomplete = FALSE)
+  controls = portalr::abundance(clean=FALSE,level="Treatment",type="Rodents",length="Longterm", incomplete = FALSE)
   #Drop PI
   controls = controls[ , -which(colnames(controls) == "PI")]
   #The total rodent count in each treatment
@@ -66,7 +66,7 @@ get_rodent_data <- function(moons, forecast_date){
     select(-newmoondate,-censusdate)
   
   #All plots
-  all = portalr::abundance(level="Site",type="Rodents",length="all", incomplete = FALSE)
+  all = portalr::abundance(clean=FALSE,level="Site",type="Rodents",length="all", incomplete = FALSE)
   #Drop PI
   all = all[ , -which(colnames(all) == "PI")]
   #The total rodent count across the entire site

### 2018-04-13 tools/prepare_data.R

@@ -8,7 +8,7 @@ filename_suffix = 'forecasts'
 #The date this forecast is run. Always today's date.
 forecast_date = Sys.Date()
 
-portalr::download_observations(release_only = FALSE)
+portalr::download_observations()
 moons = get_moon_data()
 moons$newmoondate <- as.Date(as.character(moons$newmoondate))
 current_moons <- select(moons, newmoonnumber, newmoondate, period, censusdate)

### 2018-04-13 index.Rmd

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-obs_data = portalr::abundance("repo", level='treatment', length="longterm") %>%
+obs_data = portalr::abundance("repo", clean=FALSE, level='treatment', length="longterm") %>%
   filter(treatment=="control")

### 2018-04-13 report.Rmd

--- 2018-02-15 report.Rmd
@@
 # get most recent rodent data
-rodents = abundance("repo", shape='flat', level='treatment')
+rodents = abundance("repo", clean=FALSE, shape='flat', level='treatment', length='longterm')
 most_recent_period = max(rodents$period)
@@
 # get most recent rodent data
-rodents = abundance("repo", shape='flat')
+rodents = abundance("repo", clean=FALSE, shape='flat')
 most_recent_period = max(rodents$period)

### 2018-04-13 index.Rmd

@@
 ensemble = dplyr::filter(data, level == 'Controls', model == 'Ensemble', date == max(date))
-sp_predictions = get_sp_predicts(ensemble, 'Controls', lead_time = 1)
+sp_predictions = get_sp_predicts(ensemble, 'Controls', 
+            lead_time = length(unique(ensemble$newmoonnumber)) - 11)
 sp_predict = plot_species_forecast(sp_predictions,

### 2018-05-03 index.Rmd
improve figure descriptions
on main page and evaluation page

@@
 This is the forecast for next month's sampling of rodents at Portal.
+The black line indicates the historic total rodent abundance.
+The blue line indicates the forecasted amount over the next 12 months, along with confidence intervals.
+This forecast is from the Ensemble model.
 
@@
 
+Point forecasts and 90% confidence intervals for specific species at the site for the next sampling period.
+
 ```{r speciesforecasts, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
+This figures are similar to the primary Total Abundance forecast above, but for the abundance of three common rodent species. 
 ```{r highabund_ts_forecasts, echo=FALSE, message=FALSE, warning=FALSE}

### 2018-06-29 evaluation.Rmd

@@
 **RMSE**: Root mean square error, this is a metric used to evaluate the point estimate of a forecast.  
-**Coverage**: This is the percentage of observations which fell within the `r 100 * pred_interval`% confidence interval of a forecast.
-Ideally this would be equal to `r pred_interval`.
-If it's higher than `r pred_interval` the forecasts intervals are too wide, if it's lower then the forecast intervals are too narrow.
+**Coverage**: This is the percentage of observations which fell within the 90% confidence interval of a forecast.
+Ideally this would be equal to 0.90.
+If it is higher than 0.90 the forecasts intervals are too wide, if it's lower then the forecast intervals are too narrow.
 

### 2018-06-29 manuscript/response_to_review.Rmd
Response to reviews
* Adds addition examples and justification
* Addresses confusion about target audience
* Makes a number of minor clarifications and improvements
* Responds to reviewers
* Switches gif image to png to improve useability with most latex installs

Co-authored-by: Morgan Ernest <morgan@weecology.org>

---
title: "Response to Review"
output: pdf_document
---

## Associate Editor

**The main issue regards the motivation (reviewer 1) and target audience (reviewer 2) of the application. If this is meant for researchers that are proficient at modeling and coding, it should be made clearer how they benefit from this automated application (wouldn't they already have similar protocols and pipelines in place in their routine research?). If instead, this is for empiricists, or environmental managers, with little coding experience, the manuscript is too technical and the application perhaps should be made more 'user-friendly' (reviewer 1). My suspicion (and reviewer 2's) is that it aims for that increasingly common middle person (the computationally literate ecologist). However, as Reviewer 2 points out, it is not clear how such audience should interact with the defined pipeline and how, for example, they could propose new models. A fourth (very attractive) possibility is that this is meant as an interface for people with different skills and motivations to collaborate (e.g. to foster the feedback between scientific development and applied use). If this is the case, the protocols of interaction need to be clarified, as well as the audience for this article (e.g. is it the computational ecologist who will read this article and propose applied scientists to work with her/him through this system?). Defining the audience clearly should guide how to improve the application and manuscript in a way that it ensures its use and long-term survival as an important short-term forecasting tool.**

We have expanded the last paragraph of the Introduction to address the question of target audience and split the opening paragraph of the Introduction into two paragraphs to provide expanded treatment of the motivation for this approach including the addition of three examples.

The AE and reviewer are correct that we have target this paper primarily at the "increasingly common middle person" and "researchers that are proficient at modeling and coding". Thanks to the expanded efforts at computational training we believe that many researchers have the necessary skills to implement these sorts of approaches and we think that the association between "empiricists" and "little coding experience" is increasingly out of date as demonstrated by the fact that this system was developed in large part by the empiricists collecting the data.

That said, we certainly recognize that not all ecologists and evolutionary biologists have the necessary skills to implement this sort of pipeline (just as many do not have the skills to implement a Bayesian statistical analysis). As a result, a single paper in this space won't be perfect for everyone. We have tried hard to reach a middle ground where users with the necessary skills to implement this sort of system will be able to understand precisely what we did while also providing introduction and motivation for researchers with less advanced computational skills to learn more and collaborate. To reach researchers with relatively basic computational skills we have recently completed a paper focused on the data management component of this pipeline that focuses on providing a ground up introduction to this kind of automation in an ecological context (https://doi.org/10.1101/344804). We hope that the combination of these two papers will successfully introduce this general class of approaches and concepts to a broad array of researchers.

We disagree with the assumption that researchers that are proficient at modeling and coding would already have similar protocols and pipelines in place in their routine research?)". We are active in this field and aware of very few efforts of this scope in ecology. These researchers will have the foundational tools to develop similar systems (or contribute to this one).

# Reviewer: 1

**Overall this is a clear (and very thorough) description of a useful set of methods and tools for addressing a compelling need in ecology.  In this age of increased skepticism of science among the general public and increased awareness among scientists of the mischief that statistics can cause, this kind of automatically updating system seems like the most honest and open approach to science possible.**

Thanks!

**My only real comment/question regards the motivation for the automatic and iterative nature of this. The full data processing to forecasting pipeline is great because it integrates data processing, model fitting and model comparison.  But I suppose the skeptical reader might wonder, can’t all of this be done without automatic and iterative part? Can’t we compare model predictions the old fashion way, using held out data, not by forecasting truly novel data? The response might be that forecasts themselves are inherently useful for planning.  Again the skeptic might question whether there really is a demand for such short-term forecasts? Maybe more examples could be given. My hunch is that the real benefit of iteratively adding truly novel data is that it makes the science as honest as possible. I guess I’m asking that the text somewhere address the skeptic’s questions above.**

There is definitely a need and demand for short-term forecasting. This was laid out in detail in Dietze et al. (2018; PNAS), but we agree that the current paper benefits from providing more direct treatment of this. We have significantly expanded the opening paragraphs of the introduction to provide specific examples and further emphasize the value of this approach.

**I might also suggest differentiating parameter estimation and model development. Forecasts can improve by adding new models, or by adding new data which then improves parameter estimation in the old models.  The automatic pipeline is great at the second part of this, continually adding data and re-fitting the old models to the new data and then producing new forecasts.  However, since there was so much data in the Portal project to begin with, I doubt that parameter estimates in the models have changed much since the automatic pipeline was established (maybe they have?).  However, in a new project that was just getting started (e.g. NEON), the main improvements in forecast skill early on would be in adding more data.  Eventually, however, the models might stop improving and reach some prediction limit imposed by the structure of the model.  As the authors point out in the discussion, finding this point will be key for improving ecological theory. I think a lot of ecologists think their models are useful, they just need better parameter estimates.  My sense is that actually, most ecological models are just crap for forecasting, it’s not for lack of data! Maybe this is an obvious point, but the role of improved parameter estimation in this whole endeavor kind of gets lost for me.  Does the pipeline even save the fitted model parameters? Would that be worthwhile?**

There are a number of interesting points in this paragraph that we feel are beyond the scope of this current paper. We have already described the importance of new models and how we have facilitated adding them. We have now further expanded this section to describe how those external to the group can contribute improved models. Introducing and discussing the different implications of differnt forms of error and how the are associated with increasing in data vs. improvements in models is an interesting subject for forecasting, but is out of scope of this paper. It would definitely be something interesting to explore in future work.

**In Figure 1, I feel that something is missing here about how other scientists are supposed to interact with the data and forecasts.  The manuscript makes it pretty clear that this is not just for internal use, but is it actually inviting other labs to add models and make predictions? How specifically would another lab contribute a model to the project? Would they fork the project on github, develop a new model and then do a pull request? Or would they create a separate prediction archive themselves? Maybe add an arrow in Fig 1. showing how new models can be plugged into the forecasting work flow.  Box 1 addresses some of this but I guess I want more details.**

We have added description of how others can contribute to the Modeling and Forecasting section of the paper. We tried to also add this to the figure, but could not find a way to do so without making the figure overly crowded and complicated.

**Some other miscellaneous questions:**
**Can models be dropped from the ensemble?**

This is not currently done at the moment, but is on the development roadmap.

**What if old data needs to be corrected or updated? That could change everything, including the interpretation of the archived predictions. Is there a way to handle that?**

This is a great question about a tricky problem in forecast archiving that we have encountered and are still trying to determine out the optimal way to address. One can update the previous forecast files and use the versioning to maintain a copy of the forecasts that contained the underlying error or leave them as they are an store supplementary data indicating that forecasts prior to the selection should not be used in later analyses. Since we aren't happy with any of the current solutions, and this is a fairly detailed point related to archiving, we think it is best to leave this for later discussion.

**Minor comments:**

**Lines 63 to 83:  Could this section include some examples that demonstrate the value of iterative near-term forecasting from other fields? Could it be argued that iterative near term forecasting has led to progress in other fields? Say meteorology and political science?**

We have now mentioned the established benefits for iterative near term forecasting in meteorology and added several citations for readers interested in exploring this further.

**Figure 1: Another point on figure one.  What about model improvement? It seems like there should be a feedback loop somewhere showing how models get refined, added, dropped etc.**

We tried to add a visual representation of this to Figure 1, but could not make it work without making the figure overly crowded. We have added some additional discussion of this "Modeling and Forecasting" section of the paper.

**There’s a lot of technical terms and software names.  Worth adding a glossary or table listing all the tools and acronyms used in the paper.  Github, Travis, CI, QA/QC, Docker, Rocker, Software containers, portalData, portalr, portalPredictions, Zenodo, data latency, etc.**

We have added a glossary.

**Figure labels are doubled.**

Fixed.

**Travis and github icons are very small in Fig. 1**

We have increased the size of the icons.

**comment on the website:  portal.naturecast.org/evaluation.html  – looks like there’s a word missing in the caption explaining the coverage graphs.  “Coverage: ... equal to . “**

Fixed. Thanks.

**comment on the website: http://portal.naturecast.org/index.html  – Figures need more information/captions.  Which model is producing the blue lines (the forecasts).  Is this the ensemble model?**

Yes. We have clarified this on the website.

**Line 163:  This section here about dependencies seems a little out of place, or at least the transition from the CI section to the details of dependencies management seems abrupt.**

There are two main components to Continuous Analysis as defined by Beaulieu-Jones & Greene (2017). We have one paragraph on each in this section on Continous Analysis. We're not sure how else to handle this, but if the reviewer has any specific suggestions we're happy to consider them.

**Figure 2:  The figure is pixelated and hard to read (might just be the draft version I have).**

The raw images are all high resolution so presumably this will be fixed in production if the paper is accepted.

## Reviewer: 2

**I enjoyed reading the manuscript by White et al, which provides a very clear description of using cutting-edge software development technologies for ecology by automating near-term forecasting. The manuscript is well thought-out, and well-written and I think should be accepted for publication. I have a few minor line-by-line suggestions and a broader question/theme that was in my mind throughout reading the article that I would suggest the authors consider.**

Thanks!

**Minor line-by-line suggestions:**

**L48: Could you cite a few here, maybe give a sentence or two overview of what they do/what questions they seek to answer (are we talking about continuous temperature sensors, camera traps, terrestrial/aquatic/marine?)**

We have split the first paragraph into two to allow us to add multiple examples and associated citations.

**L102: Like L48 above, please provide citations and examples with this statement**

We have added examples and citations to this sentence.

**Fig 1:**

**- The first time the reader sees "GitHub" and "Travis": please add citations and descriptions (and/or reshuffle when Fig 1 is cited within the text so you can describe it first in the text)**

There is a challenging order issue here. We want readers to be able to see the stages of the forecasting pipeline before delving into the details. This is Figure 1a, which is referenced in the text before the definitions of Travis & GitHub. That section of the figure/legend does not reference those terms. Travis was cited and described before Figure 1b was referenced. We have updated the text so that this is also true for GitHub. We feel that this is the best compromise in that it keeps the two directly related pieces of Figure 1 together, allows the reader to see the big picture of the pipeline before diving into the details, and introduces the associated terms before the relevant section of the figure is referenced in the text.

**- Consistency: the legend describes the Travis icon but not the GitHub icon**

Added.

**- A, B are not labeled in the figure**

Fixed.

**This paper is written to quite a narrow audience: ecologists that already code and could benefit from this kind of automation. This is fine. But many researchers do not yet have the coding skills to benefit from automation (the audiences for Software Carpentry, who the authors mention). To me, it seems that in ecology there are two computational camps at the moment: ecologists who are ready for the workflows mentioned in this paper, and those beginning to engage through Software Carpentry. This is obviously over-simplified, but as I read this manuscript I found myself thinking that it would be nice to try to speak to both camps, instead of just the one with the skills for this automation. How do people get there? How do we get ecologists to the point where they too can develop a workflow like the one in this manuscript?**

A single paper in this space won't be perfect for everyone. We have tried hard to reach a middle ground where users with the necessary skills to implement this sort of system will be able to understand precisely what we did while also providing introduction and motivation for researchers with less advanced computational skills to learn more and collaborate. Thankfully given recently expanded efforts to provide more advanced computational training to scientists we disagree that it is currently written for "quite a narrow audience". That said, we agree with the reviewer that reaching out to researchers with relatively basic computational skills is important. To address this broader need for introductory material on automation and continuous integration in ecology, we have recently completed a paper focused on the management of data that is regularly updated, which focuses on communicating at an introductory level, including the inclusion of a tutorial for helping beginners to this area get started (https://doi.org/10.1101/344804). We hope that the combination of these two papers will successfully introduce these approaches and concepts to a broad array of researchers.

**I was happy to see the discussion of team science (~L421), but training and skill-building immediately came to mind, and I didn't know who the "team" was, it seems that it is the Weecology team, not including the Portal Project. Can the Portal Project maintain this on their own? I still didn't have a sense whether Florida researchers created something for Arizona field scientists that will be maintained by the Florida scientists, or whether the Arizona group was trained to maintain it on their own. My assumption is that the Arizona team would be in the Software Carpentry camp, and I'm thinking about how to bridge these camps and empower more scientists with the skills to be able to understand the integrative process the manuscript describes. The goal is to get all ecologists into the automation camp, how do we do it?**

**I would suggest adding some sentences that address this to the intro and discussion. This could traditionally be seen as out of the scope of the manuscript, but that will continue to perpetuate the two camps instead of bringing the Software Carpentry camp towards the automation camp.**

As is already described in the paper, the field researchers are an active part of the team that develops and runs this system. Both components of the team are part of Weecology. As we've mentioned above the clear distinction between empiricists and computional folks really isn't valid anymore and the field researchers involved in this project all have strong computational skills. We have modified the language in this section to clarify how the team works. While we agree that "The goal is to get all ecologists into the automation camp", and we are working hard on multiple fronts to make this a reality, this is a larger challenge that is beyond the scope of this current paper.

### 2018-06-29 manuscript/portal_iterative_forecasting_ms.Rmd
Improvements to ms Figure 1 and legend
Response to reviews
* Adds addition examples and justification
* Addresses confusion about target audience
* Makes a number of minor clarifications and improvements
* Responds to reviewers
* Switches gif image to png to improve useability with most latex installs

Co-authored-by: Morgan Ernest <morgan@weecology.org>

@@
 [-@clark2001] called for a more central role of forecasting in ecology.
-Since then, an increasing number of ecological forecasts are being published.
-Most of these forecasts, however, are made once, published, and never assessed or updated.
+Since then, an increasing number of ecological forecasts are being published that focus on a societally important questions from daily to decadal time scales [@dietze2018].
+At daily scales, ecological forecasts predict the occurance of environmental issues like toxic algal blooms [@stumpf2009].
+At quarterly scales forecasts are used to predict the stocks of fisheries and make decisions about quotas for fishing [@NOAA2016].
+At decadal time scales, ecological forecasts are used to predict how biodiversity will change as it responds to anthropogenic influences [@harris2018].
+These forecasting examples highlight the important role that ecological forecasts play in recasting ecological knowledge in societally relevant ways and also improve our understanding of ecological systems by testing the ability of our models to predict how systems will change in the future [@dietze2018, @harris2018].
+
+While some of the examples given above (e.g., fisheries stock estimates) are regularly repeated, most ecological forecasts are made once, published, and never assessed or updated [@dietze2018].
 This lack of both regular assessment and active updating has limited the progress of ecological forecasting and hindered our ability to make useful and reliable predictions.
@@
 More regular updating and assessment will advance ecological forecasting as a field by accelerating the identification of the best models for individual forecasts and improving our understanding of how to best design forecasting approaches for ecology in general.
+This approach has helped accelerate forecasting ability in other fields including meteorology [@kalnay2003, @mcgill2012, @bauer2015].
 For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].
@@
 
-While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement, and few of their recommendations are in widespread use in ecology today.
+While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement (e.g., making quality data available in near real-time and automatically rerunning forecasts in reproducible ways), and few of their recommendations are in widespread use in ecology today [e.g., @wilson2014; @stodden2014; @yenni2018].
 We explored what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going, long-term ecological study that collects high-frequency data on desert rodent abundances [@brown1998; @ernest2008].
@@
 In this paper, we discuss our approach for creating this iterative near-term forecasting pipeline, the challenges we encountered, the tools we used, and the lessons we learned so that others can create their own iterative forecasting systems.
+For those interested in implementing iterative forecasting, either on their own or as part of a team, this paper will provide a roadmap for how to build such a system and what skills will be helpful to do so.
+For readers looking for an introduction to automation and continous integration in an ecological context, we recommend our paper on data management for continuously collected data, which includes a tutorial on how to set up some of the aspects of automation described in this paper (@yenni2018).
 
@@
 
-![Figure 1.
-a) Stages of the forecasting pipeline.
+![a) Stages of the forecasting pipeline.
 To go from raw data to forecast presentation involves a number of stages, each of which requires unique tasks, tools and infrastructure.
@@
 Travis also runs tasks within a stage, such as testing code and adding weather data (icons on arrows originating and ending on the same box).
-](figs/Figure1.png)
+The code for driving different stages of this pipeline is stored on GitHub (denoted by the GitHub icon, an "octocat").](figs/Figure1.png)
 
@@
 Our forecasting pipeline currently runs on a publicly available continuous integration service (Travis CI; https://travis-ci.org/) that is free for open source projects (up to a limited amount of computing time).
+This continuous integration integrates directly with GitHub (https://github.com), the online repository where we store the associated code and data.
 Because of the widespread use of CI in software development, alternative services that can run code on local or cloud-based computational infrastructure also exist [@beaulieu2017].
@@
 As long as a model script can load the provided data and produce the appropriate output, it will be run and its results incorporated into the rest of the forecasting system.
+This means that anyone can add a new model to the existing system by: 1) creating their own copy of the project (typically by forking the project on GitHub); 2) developing a new model; and 3) submitting a pull request to our repository.
 
-![Figure 2.
-Demonstration of plugin infrastructure.
+![Demonstration of plugin infrastructure.
 All model scripts (represented here by the example AR1.R) are housed in a single folder.
@@
 2017_12_08forecasts.csv) to allow cross-model evaluations.
-Model output files are housed in a folder containing all forecast outputs from all previous dates to facilitate archiving and forecast assessment.](figs/plugin_fig.gif)
+Model output files are housed in a folder containing all forecast outputs from all previous dates to facilitate archiving and forecast assessment.](figs/plugin_fig.png)
 
@@
 It can also be used to quickly evaluate new models instead of waiting for an adequate amount of data to accumulate.
+As the performance of different models is understood through evaluation of forecasts and hindcasts models can be refined or removed from the system or ensemble to iteratively improve the resulting forecasts.
 
@@
 
-![Figure 3.
-Screen capture of the homepage of the Portal Forecasting website (http://portal.naturecast.org).
+![Screen capture of the homepage of the Portal Forecasting website (http://portal.naturecast.org).
 This site contains information on the most current forecasts, evaluation of forecast performance, and general information about the species being forecast.](figs/website_fig.png)
@@
 Because of the breadth of expertise needed to set up our forecasting pipeline, our effort required a team with diverse skills and perspectives, ranging from software development to field site expertise.
-It is rare to find such breadth within a single research group, and our system was developed as a collaboration between the lab collecting the data and a computational ecology lab.
+It is rare to find such breadth within a single individual, and our system was developed as a collaboration between the lab collecting the data and a computational ecology lab.
 When teams have a breadth of expertise, communication can be challenging [@winowiecki2011].
-We found a shared base of knowledge related to both the field research and fundamental computational skills was important for the success of the group.
+We found a shared base of knowledge related to both the field research and computational skills was important for the success of the group.
 The two labs are part of a joint interdisciplinary ecology group that has a mission of breaking down barriers between field and computational/theoretical ecologists (http://weecology.org).
-Everyone on the team had received training in fundamental data management and computing skills through a combination of university courses, Software and Data Carpentry workshops [@teal2015], and informal lab training efforts.
+Everyone on the team had received training in fundamental data management and computing skills through a combination of university courses, Software and Data Carpentry workshops [@teal2015], and lab training efforts.
 In addition, everyone was broadly familiar with the study site and methods of data collection, and most team members had participated in field work at the site on multiple occasions.
 This provided a shared set of knowledge and vocabulary that actively facilitated interdisciplinary interactions.
-Given the current state of tools for forecasting, forecasting teams will need people with significant experience in working with continuous integration and APIs.
-This means interdisciplinary teams will generally be required for creating these pipelines until tool development improves.
+All members of the team actively participated in the development of the forecasting pipeline.
+Given the current state of tools for forecasting, forecasting teams require some experience in working with continuous integration and APIs.
+This means either interdisciplinary teams or additional training will often be required for creating these pipelines until tool development improves.
 To improve the success of these diverse groups, we believe efforts at providing 'team science' training to scientists interested in forecasting will be beneficial for the success of iterative forecasting attempts for the foreseeable future [@read2016].
@@
 
+# Box 2. Glossary of terms
+
+**CI.** 'Continuous Integration.' The practice of continuouslly building and testing a code base as it is developed. 
+**Data latency.** The time it takes for data to be available for use.
+**Docker.** An open-source Linux program for containerization (see software container).
+**git.** An open-source version control system.
+**GitHub.** A web-based host for git projects. Other options for a similar service include GitLab or Bitbucket.
+**PortalData.** The git repository for the Portal data, found on GitHub.
+**portalPredictions.** The git repository for the forecasts made using Portal data, found on GitHub.
+**portalr.** An R package for using the Portal data.
+**QA/QC.** 'Quality Assurance.' Testing the quality of a product. 'Quality Control.' The process of ensuring the quality of a product.
+**Rocker.** A project making it easy to use Docker containers in the R environment.
+**Software container.** Allows a developer to package up an application with all of the parts it needs to run reliably.
+**testthat.** R package used to set up automated testing for QA/QC.
+**Travis.** A continuous integration service that integrates easily with GitHub and R. Examples of similar programs are Jenkins or CodeShip.
+**Unit test.** A component of quality control in which each smallest testable part of software is formally tested.
+**Zenodo.** An open data archive that integrates easily with GitHub.
+
 # References

### 2018-06-29 manuscript/portal_iterative_forecasting_ms.Rmd

--- 2018-02-20 manuscript/portal_iterative_forecasting_ms.Rmd
@@
 Each box denotes the core infrastructure used for each stage of the forecasting pipeline.
+Boxes with the GitHub icon (octocat) represent git repositories.
 Continuous integration (denoted by the Travis icon, a woman wearing safety glasses and hardhat) triggers the code involved in events that link the stages of the pipeline, such as using the output from the forecasting stage (purple box) to create an updated website (rose box).
 Travis also runs tasks within a stage, such as testing code and adding weather data (icons on arrows originating and ending on the same box).
-](figs/Figure1.png)
+Note that this pipeline only describes automated tasks, not all activity within a repository.](figs/Figure1.png)
 

### 2018-07-02 manuscript/response_to_review.Rmd
Update response_to_review.Rmd

@@
 The AE and reviewer are correct that we have target this paper primarily at the "increasingly common middle person" and "researchers that are proficient at modeling and coding".
-Thanks to the expanded efforts at computational training we believe that many researchers have the necessary skills to implement these sorts of approaches and we think that the association between "empiricists" and "little coding experience" is increasingly out of date as demonstrated by the fact that this system was developed in large part by the empiricists collecting the data.
+Thanks to the expanded efforts at computational training, we believe that many researchers have the necessary skills to implement these sorts of approaches, and we think that the association between "empiricists" and "little coding experience" is increasingly out of date as demonstrated by the fact that this system was developed in large part by the empiricists collecting the data.
 
@@
 We have tried hard to reach a middle ground where users with the necessary skills to implement this sort of system will be able to understand precisely what we did while also providing introduction and motivation for researchers with less advanced computational skills to learn more and collaborate.
-To reach researchers with relatively basic computational skills we have recently completed a paper focused on the data management component of this pipeline that focuses on providing a ground up introduction to this kind of automation in an ecological context (https://doi.org/10.1101/344804).
+To reach researchers with relatively basic computational skills, we have recently completed a paper focused on the data management component of this pipeline that focuses on providing a ground up introduction to this kind of automation in an ecological context (https://doi.org/10.1101/344804).
 We hope that the combination of these two papers will successfully introduce this general class of approaches and concepts to a broad array of researchers.
 
-We disagree with the assumption that researchers that are proficient at modeling and coding would already have similar protocols and pipelines in place in their routine research?)".
-We are active in this field and aware of very few efforts of this scope in ecology.
+We disagree with the assumption that researchers that are proficient at modeling and coding would "already have similar protocols and pipelines in place in their routine research." We are active in this field and aware of very few efforts of this scope in ecology.
 These researchers will have the foundational tools to develop similar systems (or contribute to this one).
@@
 We have now further expanded this section to describe how those external to the group can contribute improved models.
-Introducing and discussing the different implications of differnt forms of error and how the are associated with increasing in data vs.
-improvements in models is an interesting subject for forecasting, but is out of scope of this paper.
+Introducing and discussing the different implications of different forms of error and how the are associated with increasing in data vs.
+improvements in models is an interesting subject for forecasting but is out of scope of this paper.
 It would definitely be something interesting to explore in future work.
@@
 
-This is not currently done at the moment, but is on the development roadmap.
+This is not currently done at the moment but is on the development roadmap.
 
@@
 
-We tried to add a visual representation of this to Figure 1, but could not make it work without making the figure overly crowded.
+We tried to add a visual representation of this to Figure 1 but could not make it work without making the figure overly crowded.
 We have added some additional discussion of this "Modeling and Forecasting" section of the paper.
@@
 We have tried hard to reach a middle ground where users with the necessary skills to implement this sort of system will be able to understand precisely what we did while also providing introduction and motivation for researchers with less advanced computational skills to learn more and collaborate.
-Thankfully given recently expanded efforts to provide more advanced computational training to scientists we disagree that it is currently written for "quite a narrow audience".
+Thankfully, given recently expanded efforts to provide more advanced computational training to scientists, we disagree that it is currently written for "quite a narrow audience".
 That said, we agree with the reviewer that reaching out to researchers with relatively basic computational skills is important.
@@
 Both components of the team are part of Weecology.
-As we've mentioned above the clear distinction between empiricists and computional folks really isn't valid anymore and the field researchers involved in this project all have strong computational skills.
+As we've mentioned above, the clear distinction between empiricists and computional folks really isn't valid anymore and the field researchers involved in this project all have strong computational skills.
 We have modified the language in this section to clarify how the team works.

### 2018-07-02 manuscript/portal_iterative_forecasting_ms.Rmd
Update portal_iterative_forecasting_ms.Rmd

--- 2018-06-29 manuscript/portal_iterative_forecasting_ms.Rmd
@@
 At daily scales, ecological forecasts predict the occurance of environmental issues like toxic algal blooms [@stumpf2009].
-At quarterly scales forecasts are used to predict the stocks of fisheries and make decisions about quotas for fishing [@NOAA2016].
+At quarterly scales, forecasts are used to predict the stocks of fisheries and make decisions about quotas for fishing [@NOAA2016].
 At decadal time scales, ecological forecasts are used to predict how biodiversity will change as it responds to anthropogenic influences [@harris2018].
@@
 
-With the goal of making ecological forecasting more dynamic and responsive, Dietze et al [-@dietze2018] recently called for an increase in iterative near-term forecasting.
+With the goal of making ecological forecasting more dynamic and responsive, Dietze et al.
+[-@dietze2018] recently called for an increase in iterative near-term forecasting.
 Iterative near-term forecasting is defined as making predictions for the near future and repeatedly updating those predictions through a cycle of evaluation, integration of new data, and generation of new forecasts.
@@
 
-Because iterative near-term forecasting requires a dynamic integration of models, predictions, and data, Dietze et al [-@dietze2018] highlight approaches to data management, model construction and evaluation, and cyberinfrastructure that are necessary to effectively implement this type of forecasting (Box 1).
+Because iterative near-term forecasting requires a dynamic integration of models, predictions, and data, Dietze et al.
+[-@dietze2018] highlight approaches to data management, model construction and evaluation, and cyberinfrastructure that are necessary to effectively implement this type of forecasting (Box 1).
 Data needs to be released quickly under open licenses [@vargas2017; @dietze2018] and structured so that it can be used easily by a variety of researchers and in multiple modeling approaches [@borer2009; @strasser2011].
@@
 
-While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al (Box 1) are not trivial to implement (e.g., making quality data available in near real-time and automatically rerunning forecasts in reproducible ways), and few of their recommendations are in widespread use in ecology today [e.g., @wilson2014; @stodden2014; @yenni2018].
+While iterative near-term forecasting is an important next step in the evolution of ecological forecasting, the requirements outlined by Dietze et al.
+(Box 1) are not trivial to implement (e.g., making quality data available in near real-time and automatically rerunning forecasts in reproducible ways), and few of their recommendations are in widespread use in ecology today [e.g., @wilson2014; @stodden2014; @yenni2018].
 We explored what it would entail to operationalize Dietze et al's recommendations by constructing our own iterative near-term forecasting pipeline for an on-going, long-term ecological study that collects high-frequency data on desert rodent abundances [@brown1998; @ernest2008].
@@
 For all rodents caught during a trapping session, information on species identity, size, and reproductive condition is collected, and new individuals are given identification tags.
-This information on rodent populations is high-frequency, uses consistent trapping methodology, and has an extended time-series (470 monthly samples and counting), making this study an ideal case for near-term iterative forecasting.
+This information on rodent populations is high-frequency, uses consistent trapping methodology, and has an extended time-series (475 monthly samples and counting), making this study an ideal case for near-term iterative forecasting.
 
@@
 It can also be used to quickly evaluate new models instead of waiting for an adequate amount of data to accumulate.
-As the performance of different models is understood through evaluation of forecasts and hindcasts models can be refined or removed from the system or ensemble to iteratively improve the resulting forecasts.
+As the performance of different models is understood through evaluation of forecasts and hindcasts, models can be refined or removed from the system or ensemble to iteratively improve the resulting forecasts.
 

### 2018-07-03 manuscript/portal_iterative_forecasting_ms.Rmd

@@
 At decadal time scales, ecological forecasts are used to predict how biodiversity will change as it responds to anthropogenic influences [@harris2018].
-These forecasting examples highlight the important role that ecological forecasts play in recasting ecological knowledge in societally relevant ways and also improve our understanding of ecological systems by testing the ability of our models to predict how systems will change in the future [@dietze2018, @harris2018].
+These forecasting examples highlight the important role that ecological forecasts play in recasting ecological knowledge in societally relevant ways and also improve our understanding of ecological systems by testing the ability of our models to predict how systems will change in the future [@dietze2018; @harris2018].
 
@@
 More regular updating and assessment will advance ecological forecasting as a field by accelerating the identification of the best models for individual forecasts and improving our understanding of how to best design forecasting approaches for ecology in general.
-This approach has helped accelerate forecasting ability in other fields including meteorology [@kalnay2003, @mcgill2012, @bauer2015].
+This approach has helped accelerate forecasting ability in other fields including meteorology [@kalnay2003; @mcgill2012; @bauer2015].
 For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].
@@
 For those interested in implementing iterative forecasting, either on their own or as part of a team, this paper will provide a roadmap for how to build such a system and what skills will be helpful to do so.
-For readers looking for an introduction to automation and continous integration in an ecological context, we recommend our paper on data management for continuously collected data, which includes a tutorial on how to set up some of the aspects of automation described in this paper (@yenni2018).
+For readers looking for an introduction to automation and continous integration in an ecological context, we recommend our paper on data management for continuously collected data, which includes a tutorial on how to set up some of the aspects of automation described in this paper [@yenni2018].
 
@@
 
-![a) Stages of the forecasting pipeline.
+![A) Stages of the forecasting pipeline.
 To go from raw data to forecast presentation involves a number of stages, each of which requires unique tasks, tools and infrastructure.
@@
 Tasks in all stages are run using code written in R.
-b) Continuous integration system.
+B) Continuous integration system.
 Each box denotes the core infrastructure used for each stage of the forecasting pipeline.
@@
 Many of us have experienced the frustrations that can occur when software updates (e.g., changes in R package versions) create errors in previously functional code.
-We experienced this issue when the `tscount` package [@liboschik2015], used by one of our forecasting models, was temporarily removed from CRAN (the R package repository) and could not be installed in the usual way.
+We experienced this issue when the `tscount` package [@liboschik2015], used by two of our forecasting models, was temporarily removed from CRAN (the R package repository) and could not be installed in the usual way.
 This broke our forecasting pipeline, as we could no longer run models that used that package.
@@
 Using containers allows us to control transitions to new package versions, implementing them only after we have tested them and made any necessary changes to the data processing and analysis code.
-We use a container created by the Rocker project, which is a Docker image with many important R packages (i.e.
-tidyverse) pre-installed  [@boettiger2017].
+We use a container created by the Rocker project, which is a Docker image with many important R packages [i.e., the tidyverse packages; @tidyverse] pre-installed  [@boettiger2017].
 We add our code and dependencies to this existing Rocker image to create a software container for our forecasting pipeline.
@@
 
-Once data is available, it must be processed into a form appropriate for modeling (Figure 1).
+Once data are available, it must be processed into a form appropriate for modeling (Figure 1).
 For many ecological datasets, this requires not only simple data manipulation but also a good understanding of the data to facilitate appropriate aggregation.

### 2018-07-03 manuscript/portal_iterative_forecasting_ms.Rmd
Polishing edits to revised manuscript

@@
 [-@clark2001] called for a more central role of forecasting in ecology.
-Since then, an increasing number of ecological forecasts are being published that focus on a societally important questions from daily to decadal time scales [@dietze2018].
+Since then, an increasing number of ecological forecasts are being published that focus on societally important questions from daily to decadal time scales [@dietze2018].
 At daily scales, ecological forecasts predict the occurance of environmental issues like toxic algal blooms [@stumpf2009].
@@
 More regular updating and assessment will advance ecological forecasting as a field by accelerating the identification of the best models for individual forecasts and improving our understanding of how to best design forecasting approaches for ecology in general.
-This approach has helped accelerate forecasting ability in other fields including meteorology [@kalnay2003; @mcgill2012; @bauer2015].
+This approach has helped accelerate forecasting ability in other fields such as meteorology [@kalnay2003; @mcgill2012; @bauer2015].
 For ecological forecasting to mature as a field, we need to change how we produce and interact with forecasts, creating a more dynamic interplay between model development, prediction generation, and incorporation of new data and information [@dietze2018].
@@
 
-In addition to automatically running software pipelines, the other key component of "continuous analysis" is making sure that the pipelines will continue to run even as software dependencies change [@beaulieu2017].
+In addition to automatically running software pipelines, the other key component of "continuous analysis" is making sure that the pipelines will continue to run even as software dependencies change [@boettiger2015; @beaulieu2017].
 Many of us have experienced the frustrations that can occur when software updates (e.g., changes in R package versions) create errors in previously functional code.
@@
 Since we collect data monthly, ensuring that the models have access to the newest data requires a data latency period of less than 1 month from collection to availability for modeling.
-To accomplish this, we automated components of the data processing and quality assurance/quality control (QA/QC) process to reduce the time needed to add new data to the database (Figure 1).
+To accomplish this, we automated components of the data processing and quality assurance/quality control (QA/QC) process to reduce the time needed to add new data to the database (@yenni2018; Figure 1).
 
@@
 
-Once data are available, it must be processed into a form appropriate for modeling (Figure 1).
+Once data are available, they must be processed into a form appropriate for modeling (Figure 1).
 For many ecological datasets, this requires not only simple data manipulation but also a good understanding of the data to facilitate appropriate aggregation.
@@
 In addition, we also archive the full state of the modeling and forecasting repository ([https://doi.org/10.5281/zenodo.833438](https://doi.org/10.5281/zenodo.833438)).
-This ensures that every forecast is fully reproducible since the exact code used to generate every forecast is preserved.
+Through a similar process, the raw data in the data repository is also archived on a Zenodo whenever data is added or changed [@yenni2018], allowing retrieval of older versions of the data used for forecasting ([https://doi.org/10.5281/zenodo.1219752](https://doi.org/10.5281/zenodo.1219752)).
+This ensures that every forecast is fully reproducible since the exact code and data used to generate every forecast is preserved.
 Early forecasts from this system are archived in the modeling and forecasting code archive, not in the newer repository 'forecasts'.
@@
 Because of the breadth of expertise needed to set up our forecasting pipeline, our effort required a team with diverse skills and perspectives, ranging from software development to field site expertise.
-It is rare to find such breadth within a single individual, and our system was developed as a collaboration between the lab collecting the data and a computational ecology lab.
+It is rare to find such breadth within a single individual, and our system was developed as a collaboration between the lab collecting and managing the data and a computational ecology lab.
 When teams have a breadth of expertise, communication can be challenging [@winowiecki2011].

### 2018-07-04 manuscript/response_to_review.Rmd

@@
 
-The AE and reviewer are correct that we have target this paper primarily at the "increasingly common middle person" and "researchers that are proficient at modeling and coding".
-Thanks to the expanded efforts at computational training, we believe that many researchers have the necessary skills to implement these sorts of approaches, and we think that the association between "empiricists" and "little coding experience" is increasingly out of date as demonstrated by the fact that this system was developed in large part by the empiricists collecting the data.
+The AE and reviewer are correct that we have targeted this paper primarily at the "increasingly common middle person" and "researchers that are proficient at modeling and coding".
+Thanks to the expanded efforts at computational training, we believe that many researchers have the necessary skills to implement these sorts of approaches, either on their own or as part of a team.
+And we think that the association between "empiricists" and "little coding experience" is increasingly out of date as demonstrated by the fact that this system was developed in large part by the empiricists collecting the data.
 
@@
 We have now further expanded this section to describe how those external to the group can contribute improved models.
-Introducing and discussing the different implications of different forms of error and how the are associated with increasing in data vs.
+Introducing and discussing the different implications of different forms of error and how they are associated with increased data vs.
 improvements in models is an interesting subject for forecasting but is out of scope of this paper.
@@
 
-We have added description of how others can contribute to the Modeling and Forecasting section of the paper.
+We have added a description of how others can contribute to the Modeling and Forecasting section of the paper.
 We tried to also add this to the figure, but could not find a way to do so without making the figure overly crowded and complicated.
@@
 
-This is a great question about a tricky problem in forecast archiving that we have encountered and are still trying to determine out the optimal way to address.
-One can update the previous forecast files and use the versioning to maintain a copy of the forecasts that contained the underlying error or leave them as they are an store supplementary data indicating that forecasts prior to the selection should not be used in later analyses.
+This is a great question about a tricky problem in forecast archiving that we have encountered and are still trying to determine the optimal way to address.
+One can update the previous forecast files and use the versioning to maintain a copy of the forecasts that contained the underlying error or leave them as they are and store supplementary data indicating that forecasts prior to the selection should not be used in later analyses.
 Since we aren't happy with any of the current solutions, and this is a fairly detailed point related to archiving, we think it is best to leave this for later discussion.
@@
 We tried to add a visual representation of this to Figure 1 but could not make it work without making the figure overly crowded.
-We have added some additional discussion of this "Modeling and Forecasting" section of the paper.
+We have added some additional discussion of this to the "Modeling and Forecasting" section of the paper.
 
@@
 
-There are two main components to Continuous Analysis as defined by Beaulieu-Jones & Greene (2017).
+There are two main components to Continuous Analysis as defined by Beaulieu-Jones & Greene (2017): continuous integration and dependency management.
 We have one paragraph on each in this section on Continous Analysis.

### 2018-07-04 manuscript/portal_iterative_forecasting_ms.Rmd
More minor edits to the revision
* Add citations to living data paper
* Move Boettiger citation to link with Containers not Continuous Analysis
* Small editorial changes
* Add acknowledgement for Heather Bradley

@@
 The Portal Project is a long-term ecological study situated in the Chihuahuan Desert (2 km north and 6.5 km east of Portal, Arizona, US).
-Researchers have been continuously collecting data at the site since 1977, including data on the abundance of rodent and plant species (monthly and twice yearly, respectively) and climatic factors such as air temperature and precipitation (daily) [@brown1998; @ernest2009; @ernest2016].
+Researchers have been continuously collecting data at the site since 1977, including data on the abundance of rodent and plant species (monthly and twice yearly, respectively) and climatic factors such as air temperature and precipitation (daily) [@brown1998; @ernest2009; @ernest2016; @ernest2018].
 The site consists of 24 50m x 50m experimental plots.
@@
 
-In addition to automatically running software pipelines, the other key component of "continuous analysis" is making sure that the pipelines will continue to run even as software dependencies change [@boettiger2015; @beaulieu2017].
+In addition to automatically running software pipelines, the other key component of "continuous analysis" is making sure that the pipelines will continue to run even as software dependencies change [@beaulieu2017].
 Many of us have experienced the frustrations that can occur when software updates (e.g., changes in R package versions) create errors in previously functional code.
@@
 To make our pipeline robust to changes in external software dependencies, we follow Beaulieu and Greene's [-@beaulieu2017] recommendation to use software containers.
-Software containers are standalone packages that contain copies of everything needed to run a given piece of software, including the operating system.
-Once created, a software container is basically a time capsule, containing all the software dependencies in the exact state used to develop and run the software.
+Software containers are standalone packages that contain copies of everything needed to run a given piece of software, including the operating system [@boettiger2015].
+Once created, a software container is basically a time capsule, containing all the software dependencies in the exact state used to develop and run the software [@boettiger2015].
 If those dependencies change (or disappear) in the wider world, they still exist, unchanged, in the container.
@@
 Since we collect data monthly, ensuring that the models have access to the newest data requires a data latency period of less than 1 month from collection to availability for modeling.
-To accomplish this, we automated components of the data processing and quality assurance/quality control (QA/QC) process to reduce the time needed to add new data to the database (@yenni2018; Figure 1).
+To accomplish this, we automated components of the data processing and quality assurance/quality control (QA/QC) process to reduce the time needed to add new data to the database [@yenni2018; Figure 1].
 
@@
 All updates to data are processed through "pull requests," which are notifications that someone has a modified version of the data to contribute.
-QA/QC checks are automatically run on the submitted data using continuous integration to ensure that no avoidable errors reach the official version of the dataset.
+QA/QC checks are automatically run on the submitted data using continuous integration to ensure that no avoidable errors reach the official version of the dataset [@yenni2018].
 
@@
 Supplementary data tables related to trapping history are updated based on the data added to the main data tables.
-Using CI for this ensures that all supplementary data tables are always up-to-date with the core data.
+Using CI for this ensures that all supplementary data tables are always up-to-date with the core data [@yenni2018].
 
@@
 The Portal Project has a long history of making its data publicly available so that anyone can use it for forecasting or other projects.
-Historically, the publication of the data was conducted through data papers [@ernest2009, @ernest2016], the most common approach in ecology; this approach, however, caused years of data latency.
+Historically, the publication of the data was conducted through data papers [@ernest2009; @ernest2016], the most common approach in ecology; this approach, however, caused years of data latency.
 With the recent switch to posting data directly to a public GitHub repository (Figure 1) with a CC0 waiver (i.e.
-no restrictions on data use; https://creativecommons.org/publicdomain/zero/1.0/), data latency for everyone has been reduced to less than one month, making meaningful iterative near-term forecasting possible for not only our group but other interested parties, as well.
+no restrictions on data use; https://creativecommons.org/publicdomain/zero/1.0/), data latency for everyone has been reduced to less than one month, making meaningful iterative near-term forecasting possible for not only our group but other interested parties, as well [@ernest2018; @yenni2018].
 
@@
 All members of the team actively participated in the development of the forecasting pipeline.
-Given the current state of tools for forecasting, forecasting teams require some experience in working with continuous integration and APIs.
+Given the current state of tools for automated iterative forecasting, forecasting teams require some experience in working with continuous integration and APIs.
 This means either interdisciplinary teams or additional training will often be required for creating these pipelines until tool development improves.
@@
 We thank all of the graduate students, postdocs, and volunteers who have collected the Portal Project over the last 40 years and the developers of the software and tools that made this project possible.
+We thank Heather Bradley for all of her logistical support that made this research possible.
 This research was supported by the National Science Foundation through grant 1622425 to S.K.M.
@@
 
-The data used in this study is from the Portal Project and is openly available (CC0) on GitHub (https://github.com/weecology/PortalData).
+The data used in this study is from the Portal Project and is openly available (CC0) on GitHub (https://github.com/weecology/PortalData) and archived on Zenodo (@portaldata).
 Code for reproducing all analyses is available on GitHub (https://github.com/weecology/portalPredictions) and archived on Zenodo [@zenodocode].

### 2018-07-21 models.Rmd

@@
 (d), and the order of the moving average (q), and are represented as 
-ARIMA(p, q, q) [@Box1970]. While the `auto.arima` function allows for seasonal
+ARIMA(p, d, q) [@Box1970]. While the `auto.arima` function allows for seasonal
 models, the seasonality is hard-coded to be on the same period as the sampling,
@@
 the number of autoregressive terms (p) and the order of the moving average 
-(q), and are represented as GARCH(p, q) [@Liboschik2017a]. The PETS model is 
-fit using the log link and a Poisson response, as well as with p = 1 
+(q), and are represented as GARCH(p, q) [@Liboschik2017a]. The pevGARCH model 
+is fit using the log link and a Poisson response, as well as with p = 1 
 (first-order autoregression) and q = 12 (yearly moving average). The 
@@
 
+## Ensemble
+
+In addition to the base models, we include an ensemble, constructed using the
+base model AIC weights. The ensemble mean is calculated as the AIC-weighted
+mean of all model means. The ensemble variance is estimated as the sum of 
+the AIC-weighted mean of all model variances and the variance of the estimated
+AIC-weighted mean, calculated using the unbiased estimate of sample variances.
+
 # References

### 2018-08-27 manuscript/portal_iterative_forecasting_ms.Rmd

@@
 These forecasting systems will also accelerate basic science by allowing new models of natural systems to be quickly implemented and compared to existing models.
-Using existing technology, and teams with diverse skill sets, it is possible for ecologists to build these systems and use them to advance our understanding of natural systems.
+Using existing technology, and teams with diverse skill sets, it is possible for ecologists to build automated forecasting systems and use them to advance our understanding of natural systems.
 
@@
 Since then, an increasing number of ecological forecasts are being published that focus on societally important questions from daily to decadal time scales [@dietze2018].
-At daily scales, ecological forecasts predict the occurance of environmental issues like toxic algal blooms [@stumpf2009].
-At quarterly scales, forecasts are used to predict the stocks of fisheries and make decisions about quotas for fishing [@NOAA2016].
+At daily scales, ecological forecasts predict the occurrence of environmental issues like toxic algal blooms [@stumpf2009] and pollen [@prank2013].
+At monthly scales, forecasts are used to predict the stocks of fisheries [@NOAA2016] and the probability of coral bleaching events [@liu2018].
 At decadal time scales, ecological forecasts are used to predict how biodiversity will change as it responds to anthropogenic influences [@harris2018].
@@
 Models need to be able to deal with uncertainty, in both the predictors and the predictions, to properly convey uncertainty in the resulting forecasts [@diniz-filho2009].
-Multiple models should be developed, both to assess which models are performing best [@dietze2018] and to facilitate combining models to form ensemble predictions which tend to perform better than single models [@araujo2007; @diniz-filho2009].
+Multiple models should be developed, both to assess which models are performing best [@dietze2018] and to facilitate combining models to form ensemble predictions which tend to perform better than single models [@araujo2007; @diniz-filho2009; @dormann2018].
 Ensuring that data and models are regularly updated and new forecasts are made requires cyberinfrastructure to automate data processing, model fitting, prediction, model evaluation, forecast visualization, and archiving.
@@
 This information on rodent populations is high-frequency, uses consistent trapping methodology, and has an extended time-series (475 monthly samples and counting), making this study an ideal case for near-term iterative forecasting.
+Forecasting of rodent population dynamics in the southwest (and more broadly) is important because of their link to zoonotic diseases such as hantavirus and plague [@parmenter1993; @gage2005; @springer2016].
+In addition, this forecasting system is being used to improve population dynamic modeling for this community and to explore the utility of incorporating experimental data into ecological forecasting models.
 
@@
 
-Implementation of iterative forecasting requires the regular rebuilding of models with new raw data as it becomes available and the presentation of those forecasts in usable forms; in our case, this occurs monthly.
-Rebuilding models in an efficient and maintainable way relies on developing an automated pipeline to handle the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presention of the forecasts (Figure 1a).
-To implement the pipeline outlined in Figure 1a, we used a "continuous analysis" framework [*sensu* @beaulieu2017] that automatically processes the most up-to-date data, refits the models, makes new forecasts, archives the forecasts, and updates a website with analysis of current and previous forecasts.
+Implementation of iterative forecasting requires the regular updating of models with new raw data as it becomes available and the presentation of those forecasts in usable forms; in our case, this occurs monthly.
+Updating models in an efficient and maintainable way relies on developing an automated pipeline to handle the six stages of converting raw data into new forecasts: data collection, data sharing, data manipulation, modeling and forecasting, archiving, and presention of the forecasts (Figure 1a).
+To implement the pipeline outlined in Figure 1a, we used a "continuous analysis" framework [*sensu* @beaulieu2017] that automatically processes the most up-to-date data, updates the models, makes new forecasts, archives the forecasts, and updates a website with analysis of current and previous forecasts.
 In this section we describe our approach to streamlining and automating the multiple components of the forecasting pipeline and the tools and infrastructure we employed to execute each component.
@@
 
-In addition to automatically running software pipelines, the other key component of "continuous analysis" is making sure that the pipelines will continue to run even as software dependencies change [@beaulieu2017].
-Many of us have experienced the frustrations that can occur when software updates (e.g., changes in R package versions) create errors in previously functional code.
-We experienced this issue when the `tscount` package [@liboschik2015], used by two of our forecasting models, was temporarily removed from CRAN (the R package repository) and could not be installed in the usual way.
-This broke our forecasting pipeline, as we could no longer run models that used that package.
-To make our pipeline robust to changes in external software dependencies, we follow Beaulieu and Greene's [-@beaulieu2017] recommendation to use software containers.
+To ensure that software pipelines continue to run automatically as software dependencies change, a key component of "continuous analysis" is the use of a reproducible computational environment [@beaulieu2017].
+We followed Beaulieu and Greene's [-@beaulieu2017] recommendation to use software containers.
 Software containers are standalone packages that contain copies of everything needed to run a given piece of software, including the operating system [@boettiger2015].
@@
 If those dependencies change (or disappear) in the wider world, they still exist, unchanged, in the container.
-We use an existing platform, Docker [@merkel2014], to store an exact image of the complete software environment for running the forecasts.
-Docker also allows a specified set of packages to be used consistently across different computer and server environments.
-Using containers allows us to control transitions to new package versions, implementing them only after we have tested them and made any necessary changes to the data processing and analysis code.
-We use a container created by the Rocker project, which is a Docker image with many important R packages [i.e., the tidyverse packages; @tidyverse] pre-installed  [@boettiger2017].
-We add our code and dependencies to this existing Rocker image to create a software container for our forecasting pipeline.
+We use an existing platform, Docker [@merkel2014], to store an exact image of our complete software environment by adding our project specific code to a container created by the Rocker project, which is a Docker image with many important R packages [i.e., the tidyverse packages; @tidyverse] pre-installed  [@boettiger2017].
+We implemented this system because we experienced issues with external dependencies breaking our pipeline (e.g., when the `tscount` package [@liboschik2015], was temporarily removed from CRAN and could not be installed in the usual way).
 In combination, the automated running of the pipeline (continuous integration) and the guarantee it will not stop working unexpectedly due to software dependencies (via a software container) allows continuous analysis to serve as the glue that connects all stages of the forecasting pipeline.
@@
 
-Iterative near-term forecasting involves regularly refitting a variety of different models (Figure 1).
+Iterative near-term forecasting involves regularly updating a variety of different models (Figure 1).
 Ideally, new models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach.
-We use CI to refit the models and make new forecasts each time the modeling code changes and when new data become available (Figure 1b).
-We use a plugin infrastructure to allow new models to be easily added to the system.
+We use CI to update the models and make new forecasts each time the modeling code changes and when new data become available (Figure 1b).
+We use a modular plugin infrastructure to allow new models to be easily added to the system.
 This approach treats each model as an interchangable black box; all models have access to the same input data and generate the same structure for model outputs (Figure 2).
+Details of how to add a new model to the system are provided in the core GitHub repository (https://github.com/weecology/portalPredictions/wiki/Adding-a-new-model).
 During each run of the forecasting code, all existing models are run and the standardized outputs are combined into a single file to store the results of the different models' forecasts.
@@
 Forecasts are currently evaluated using root mean square error (RMSE) to evaluate point forecasts and coverage to evaluate uncertainty.
-We plan to add additional metrics in the future.
+We plan to add additional metrics, like deviance, that incorporate both accuracy and uncertainty and better match the calibration method [@hooten2015; @dietze2018].
 In addition to evaluating the actual forecasts, we also use hindcasting [forecasting on already collected data; @jolliffe2003] to gain additional insight into the methods that work best for forecasting this system.
@@
 Following the recommendations of Dietze et al [-@dietze2018], we developed an automated iterative forecasting system (Figure 1) to support repeated forecasting of an ecological system.
-Our forecasting system automatically acquires and processes the newest data, refits the models, makes new forecasts, publicly archives those forecasts, and presents both the current forecast and information on how previous forecasts performed.
+Our forecasting system automatically acquires and processes the newest data, updates the models, makes new forecasts, publicly archives those forecasts, and presents both the current forecast and information on how previous forecasts performed.
 Every week, the forecasting system generates a new set of forecasts with no human intervention, except for the entry of new field data.
@@
 Our efforts, however, show that it is possible to use existing tools to develop initial iterative systems as a method for both advancing scientific understanding and developing proof of concept forecasting systems.
+
+Expanding the community of ecological forecasters using continous analysis approaches will require both an expansion of the current toolkit and the development of standards to facilitate interoperability of forecasts and forecasting systems.
+One of the major challenges for our current forecasting system is supporting computationally intensive forecasts.
+Projects involving larger datasets and/or complex modelling approaches will require either hosted solutions that provide infrastructure for running continuous integration and allow long-running distributed jobs or solutions that involve the user setting up their own continuous analysis sytem on cloud infrastructure or high performance computing centers.
+Event-driven serverless cloud platforms like OpenWhisk (https://openwhisk.apache.org/) and AWS Lambda (https://aws.amazon.com/lambda/) provide potential as hosted solutions, and open source continuous integrations systems like Jenkins (https://jenkins.io/) can be integrated with either cloud or high performance computing centers.
+However, both solutions are currently more complicated to set up than the hosted continuous integration approach we have employed using Travis.
+In addition to scalability issue for more computationally intensive projects, the toolkit for continuous analysis needs be made more researcher friendly.
+To broaden the user-base that can use continuous analysis for forecasting, we recommend the development of tools that make setting up continuous analysis easier by automating configuration steps.
+We also recommend the development of tools or data repository infrastructure to support the easy automated archiving of regularly generated data and forecasts [see @yenni2018].
+Finally, the development of standards for ecological forecasting to allow interoperability among forecasting systems will be essential for the growth of this field [see discussions of data standards, meta-data, and ontologies in ecology more broadly @jones2006; @madin2008; @michener2012].
+Now is an opportune time for developing these standards while the community of ecological forecasters is still small.
+While we have developed an initial format for storing and sharing forecasts, it is still lacking in several areas.
+Most notably, our approach to storing information on models and their associated uncertainty is insufficient for all but the simplest models.
+Improving this framework will require capturing both covariances between state variables and the full uncertainty in the models, by either storing full model objects or additional information like full ensembles of predictions (e.g., from Monte Carlo based approaches).
+This is challenging due to a lack of general standards for reporting uncertainty [@dietze2018].
 
@@
 
+# Author Contributions
+
+All authors conceived the ideas and designed methodology; All authors developed the automated forecasting system; EPW and SKME led the writing of the manuscript.
+All authors contributed critically to the drafts and gave final approval for publication.
+
 # Box 1. Key practices for automated iterative near-term ecological forecasting
@@
 Models used for forecasting should be probabilistic to properly quantify uncertainty and to convey how this uncertainty increases through time.
-Evaluation of forecast models should include assessment of how accurately they quantify uncertainty as well as point estimates  [@hooten2015].
+Evaluation of forecast models should include assessment of how accurately they quantify uncertainty as well as point estimates  [@hooten2015; @harris2018].
 
@@
 
-Each forecast iteration involves acquiring new data, refitting the models, and making new forecasts.
+Each forecast iteration involves acquiring new data, updating the models, and making new forecasts.
 This should be done automatically without requiring human intervention.

### 2018-08-27 manuscript/portal_iterative_forecasting_ms.Rmd

@@
 
-New data are double-entered into Microsoft Excel using the "data validation" feature.
+New data is double-entered into Microsoft Excel using the "data validation" feature.
 The two versions are then compared using an R script to control for errors in data entry.
@@
 The local use of the QC scripts to flag problematic data greatly reduces the time spent error-checking and ensures that the quality of data is consistent.
-The cleaned data are then uploaded to the GitHub-based PortalData repository (https://github.com/weecology/PortalData).
+The cleaned data is then uploaded to the GitHub-based PortalData repository (https://github.com/weecology/PortalData).
 GitHub (https://github.com/) is a software development tool for managing computer code development, but we have also found it useful for data management.
@@
 
-Once data are available, they must be processed into a form appropriate for modeling (Figure 1).
+Once data is available, it must be processed into a form appropriate for modeling (Figure 1).
 For many ecological datasets, this requires not only simple data manipulation but also a good understanding of the data to facilitate appropriate aggregation.
@@
 Ideally, new models should be easy to incorporate to allow for iterative improvements to the general modeling structure and approach.
-We use CI to update the models and make new forecasts each time the modeling code changes and when new data become available (Figure 1b).
+We use CI to update the models and make new forecasts each time the modeling code changes and when new data becomes available (Figure 1b).
 We use a modular plugin infrastructure to allow new models to be easily added to the system.
@@
 We also automatically archive the resulting forecasts publicly so that the performance of these forecasts can be assessed by both us and other researchers as new data is collected.
-This serves as a form of pre-registration by providing a quantitative record of the forecast before the data being predicted were collected.
+This serves as a form of pre-registration by providing a quantitative record of the forecast before the data being predicted was collected.
 

### 2018-08-27 manuscript/portal_iterative_forecasting_ms.Rmd
Update pdf and html versions

@@
 output:
-  html_document: default
   pdf_document:
     keep_tex: yes
+  html_document: default
 geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper

### 2018-11-06 evaluation.Rmd

@@
 species_of_interest = c('BA','DM','DO','PP','OT','NA','total')
-species_names = portalr::FullPath('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
+species_names = portalr:::full_path('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
   read.csv(stringsAsFactors=FALSE, na.strings = '') %>%
@@
 
-new_moon_file = portalr::FullPath('PortalData/Rodents/moon_dates.csv', '~')
+new_moon_file = portalr:::full_path('PortalData/Rodents/moon_dates.csv', '~')
 new_moons = read.csv(new_moon_file)

### 2018-12-01 PortalForecasts.R
Beginning the switch to portalcasting
Add portalcasting to package installs, use portalcasting functions to drive directory setup and modeling

@@ -1,20 +1,12 @@
-source('tools/data_tools.R')
-source('tools/forecast_tools.R')
-library(yaml)
+library(portalcasting)
 
-model_metadata = yaml.load_file("data/model_metadata.yaml")
-forecast_date = as.Date(model_metadata$forecast_date)
+#Update data and models
+setup_dir()
 
-if(!(forecast_date == Sys.Date())){ stop('Data not updated') }
+#Run all models using portalcasting defaults
+#Save predictions to predictions directory in base directory
 
-#########Run all models##################################################  
-cat("Running models", "\n")
-dir.create("tmp")
-sapply( list.files("models", full.names=TRUE), source ) ###Temporary, while only modeling in R
+portalcast()
 
-#####Collect all forecast results and save to predictions directory######
-cat("Compiling forecasts", "\n")
-newforecasts=forecastall(forecast_date)
-unlink("tmp/*")
-######Update Website#####################################################
+#Update Website
 rmarkdown::render_site()

### 2018-12-01 forecasting/PortalData/DataCleaningScripts/clean_plant_quadrat_data.R

@@ -1,0 +1,127 @@
+# Clean Plant Quadrat Data 
+# most of the code taken from EMC
+# organized by EKB
+# May 10, 2016
+# updated 4/23/18 by EMC
+
+# Notes:
+#   Section 1 
+#       a. Season and year should be changed accordingly
+#     * b. Make sure filepath is assigned to the appropriate directory *
+#       c. Use the most up-to-date version of the species list from GitHub
+#   Section 2
+#       a. From general_data_cleaning_functions.R
+#   Section 3
+#       a. This code is taken from EMC's "clean_plant_2014_2015.r
+#       b. QC code in plant_data_cleaning_functions.R
+#       c. Add new species to "Portal_plant_species.csv"
+#   Section 4
+#     * a. Make sure to check directory for appending the existing file *
+
+library(dplyr)
+source('DataCleaningScripts/general_data_cleaning_functions.R')
+source('DataCleaningScripts/plant_data_cleaning_functions.R')
+
+######################
+# 1. Load Excel file #
+######################
+
+season <-  'Summer'
+year <-  '2018'
+filepath <-  '/Users/renatadiaz/Dropbox/Portal/PORTAL_primary_data/Plant/Quadrats/Dataraw/Newdata/'
+
+newfile <-  paste(filepath, season, year, '.xlsx', sep='')
+
+##################################
+# 2. Compare double-entered data #
+##################################
+
+splist = read.csv('./Plants/Portal_plant_species.csv',as.is=T)
+
+unmatched = compare_worksheets(newfile)
+
+# iterate through mismatches and fix them
+ws1 = openxlsx::read.xlsx(newfile, sheet = 1, colNames = TRUE, na.strings = '')
+ws2 = openxlsx::read.xlsx(newfile, sheet = 2, colNames = TRUE, na.strings = '')
+
+i = 1
+
+i  = i + 1
+unmatched[i, ]
+ws1[unmatched[i, 'row'] -1 , ]
+ws2[unmatched[i, 'row'] - 1, ]
+
+# Save matching datasheet
+write.csv(ws1, '/Users/renatadiaz/Dropbox/Portal/PORTAL_primary_data/Plant/Quadrats/Dataraw/Newdata/Summer2018_matched.csv', row.names = FALSE)
+
+######################
+# 3. Quality control #
+######################
+
[…]

### 2018-12-01 forecasting/PortalData/DataCleaningScripts/old_weather_overlap.R

@@ -1,0 +1,117 @@
+# This script takes a raw .dat file downloaded directly from the Portal weather 
+# station, determines if there are any gaps, and appends the new data to the
+# weather overlap file.
+# Met445.dat can be used for testing - it has a lot of problems
+
+
+library(lubridate)
+library(dplyr)
+# ==============================================================================
+# Load file
+# ==============================================================================
+
+# Open raw .dat file of new data
+filepath = "~/Dropbox/Portal/PORTAL_primary_data/Weather/Raw_data/2002_Station/"
+
+metfile = "Met478"
+
+rawdata = read.csv(paste(filepath,metfile,'.dat',sep=''),head=F,sep=',',
+                   col.names=c('code','year','jday','hour','precipitation','airtemp','RH'))
+
+# Convert Julian day to month and day
+rawdata$date = as.Date(paste(rawdata$year,rawdata$jday),format='%Y %j')
+rawdata$month = as.integer(format(rawdata$date,'%m'))
+rawdata$day = as.integer(format(rawdata$date,'%d'))
+rawdata$timestamp = ymd_hms(paste(rawdata$year,"-",rawdata$month,"-",rawdata$day," ",rawdata$hour/100,":00:00",sep=""))
+
+# Select weather data (Code=101) from battery status data (Code=102) then add battery data as column
+weathdat = rawdata[rawdata$code==101,]
+battery= rawdata[rawdata$code==102,] %>% select(year,month,day,hour,battv=precipitation)
+weathdat=left_join(weathdat,battery,by=c("year","month","day","hour"))
+
+# Make precipitation correction 
+# calculate mL from current reading (with what the datalogger *thinks*): 4.73 mL/tip *(weathdat$precipitation mm / .1 mm/tip)
+# then calculate mm based on actual funnel on gauge: .254 mm/tip * calculated mL / 8.25 mL/tip
+weathdat$precipitation=4.73*(.254*weathdat$precipitation/.1)/8.25
+
+# ==============================================================================
+# 1. Quality control
+# ==============================================================================
+
+# check for errors in hour (should be 100,200,300,...)
+if (any(!(weathdat$hour %in% seq(from=100,to=2400,by=100)))) {
+  print('Hour error')
+  weathdat = filter(weathdat,hour %in% seq(100,2400,100))
[…]

### 2018-12-01 forecasting/PortalData/install-packages.R

@@ -1,0 +1,9 @@
+# Install pacman if it isn't already installed
+
+if ("pacman" %in% rownames(installed.packages()) == FALSE) install.packages("pacman")
+
+
+# Install analysis packages using pacman
+
+pacman::p_load(stringr, Hmisc, dplyr, openxlsx, lubridate, lunar,
+               sqldf, RCurl, EML, testthat, htmltab, zoo, tidyr, semver)

### 2018-12-01 forecasting/PortalData/testthat.R

@@ -1,0 +1,3 @@
+library(testthat)
+
+test_dir("testthat", reporter = c("check", "progress"))

### 2018-12-01 forecasting/PortalData/testthat/test-plant_quadrats.R

@@ -1,0 +1,57 @@
+library(testthat)
+context("checks new plant quadrat data and plant species list")
+
+data <- read.csv("../Plants/Portal_plant_quadrats.csv")
+species <-  read.csv('../Plants/Portal_plant_species.csv')
+
+test_that("valid year", {
+  
+  expect_true(all(data$year %in% 1981:max(data$year)))
+})
+
+test_that("valid season", {
+  
+  expect_true(all(data$season %in% c("winter","summer")))
+})
+
+test_that("valid plot", {
+  
+  expect_true(all(data$plot %in% 1:24))
+})
+
+test_that("valid quadrat", {
+  
+  expect_true(all(data$quadrat %in% c(11,13,15,17,31,33,35,37,51,53,55,57,71,73,75,77)))
+})
+
+test_that("valid species", {
+  
+  expect_true(all(data$species %in% species$speciescode))
+})
+
+test_that("valid abundance", {
+  
+  expect_true(all(na.omit(data$abundance) %in% 0:2400))
+})
+
+test_that("valid cover", {
+  
+  expect_true(all(data$cover[which(data$cover!=0.1)] %in% 1:100))
+})
+
+
+test_that("no duplicate data", {
+  
+  expect_true(sum(duplicated(data))==0)
+})
+
+
+test_that("valid duration in Portal_plant_species", {
+  
+  expect_true(all(species$duration %in% c('Perennial','Annual','Unknown')))
+})
+
+test_that("valid community in Portal_plant_species", {
+  
+  expect_true(all(species$community %in% c('Shrub','Summer Annual','Perennial Forb','Summer and Winter Annual','Unknown','Winter Annual','Subshrub','Perennial Grass')))
+})

### 2018-12-01 forecasting/PortalData/testthat/test-update_portal_censuses.R

@@ -1,0 +1,42 @@
+source('../DataCleaningScripts/update_portal_plant_censuses.R')
+
+library(testthat)
+context("checks new plant census dates are being added correctly")
+
+censuses=update_portal_plant_censuses()
+
+test_that("valid season", {
+  
+  expect_true(all(censuses$season %in% c("winter","summer")))
+})
+
+test_that("valid year", {
+  
+  expect_true(all(censuses$year %in% 1981:max(censuses$year,na.rm=T)))
+})
+
+test_that("valid plot", {
+  
+  expect_true(all(censuses$plot %in% 1:24))
+})
+
+test_that("valid quadrat", {
+  
+  expect_true(all(censuses$quadrat %in% c(11,13,15,17,31,33,35,37,51,53,55,57,71,73,75,77)))
+})
+
+test_that("valid censuses", {
+  
+  expect_true(all(censuses$censused %in% c(0,1)))
+})
+
+test_that("valid area", {
+  
+  expect_true(all(censuses$area %in% c(0,0.5,0.25)))
+})
+
+test_that("no duplicate data", {
+  
+  expect_true(sum(duplicated(censuses))==0)
+})
+

### 2018-12-01 forecasting/PortalData/testthat/test-update_portal_trapping.R

@@ -1,0 +1,17 @@
+source('../DataCleaningScripts/update_portal_rodent_trapping.r')
+
+library(testthat)
+context("checks new trapping dates are being added correctly")
+
+trappingdat=update_portal_rodent_trapping()
+
+test_that("no skipped periods, in order", {
+  
+  expect_true(all(diff(unique(trappingdat$period))==1))
+})
+
+
+test_that("no duplicate data", {
+
+  expect_true(sum(duplicated(trappingdat))==0)
+})

### 2018-12-01 forecasting/PortalData/testthat/test-weather.R

@@ -1,0 +1,85 @@
+source("../DataCleaningScripts/new_weather_data.R")
+library(testthat)
+context("checks new weather data")
+
+data=new_met_data()
+newdata = data[[1]]
+newdata_cols = colnames(newdata)
+weather = data[[2]]
+weather_cols = colnames(weather)
+stormsnew = data[[3]]
+stormsnew_cols = colnames(stormsnew)
+storms = data[[4]]
+storms_cols = colnames(storms)
+
+portal4sw = read.csv(file = "../Weather/Portal4sw_regional_weather.csv",header=T,
+                   colClasses=c("character", rep("integer",3), "character", "integer", rep("character",3), "Date"))
+sansimon = read.csv(file = "../Weather/Sansimon_regional_weather.csv",header=T,
+                        colClasses=c("character", rep("integer",3), "character", "integer", rep("character",3), "Date"))
+
+test_that("required column names in new weather df", {
+  
+  expect_identical(weather_cols, newdata_cols)
+  expect_identical(storms_cols, stormsnew_cols)
+})
+
+test_that("Hour in 100:2400", {
+  
+  expect_true(all(newdata$hour %in% seq(from=100,to=2400,by=100)))
+})
+
+test_that("Air Temperature ok", {
+
+  expect_true(all(newdata$airtemp > -30))
+  expect_true(all(newdata$airtemp <= 100))
+})
+
+test_that("Relative humidity ok", {
+
+  expect_true(all(newdata$RH > 0))
+  expect_true(all(newdata$RH <= 100))
+})
+
+test_that("battery status ok", {
+  
+  expect_true(all(newdata$battv > 9))
+  expect_true(all(stormsnew$battv > 9))
+})
+
+test_that("Precipitation ok", {
+  
+  expect_true(all(newdata$precipitation >= 0))
+  expect_true(all(newdata$precipitation < 100))
+  expect_true(all(stormsnew$precipitation >= 0))
+  expect_true(all(stormsnew$precipitation < 12))
+})
+
+test_that("Precipitation in multiples of 0.254", {
+  
+  expect_true(all(lapply(newdata$precipitation > 9, ifelse,
+    abs(newdata$precipitation / 0.254 - signif(newdata$precipitation / 0.254, digits = 3)) < 0.02,
+    newdata$precipitation%%0.254==0)))
+  expect_true(sum(stormsnew$precipitation%%0.254)==0)
+})
+
[…]

### 2018-12-01 forecasting/PortalData/update_repo.sh

@@ -1,0 +1,15 @@
+git config --global user.email "weecologydeploy@weecology.org"
+git config --global user.name "Weecology Deploy Bot"
+
+git checkout master
+git add -u
+git commit -m "Update supplemental data tables: Travis Build $TRAVIS_BUILD_NUMBER [skip ci]"
+
+git remote add deploy https://${GITHUB_TOKEN}@github.com/weecology/PortalData.git > /dev/null 2>&1
+git push --quiet deploy master > /dev/null 2>&1
+
+# Create a new release to trigger Zenodo archiving
+version=$(cat version.txt)
+git tag $version
+git push --quiet deploy --tags > /dev/null 2>&1
+curl -v -i -X POST -H "Content-Type:application/json" -H "Authorization: token $GITHUB_TOKEN" https://api.github.com/repos/weecology/PortalData/releases -d "{\"tag_name\":\"$version\"}"

### 2018-12-01 models/pevgarch.R

@@ -1,250 +1,0 @@
-# pevGARCH
-#
-# Model "pevGARCH" is a generalized autoregresive conditional 
-#  heteroscedasticity model with a Poisson response variable
-#  For this model, environmental data are included as predictors of abundance,
-#  but with a 6 month lag between the covariate values and the abundances.
-
-  library(tscount)
-  library(forecast)
-  library(yaml)
-  library(magrittr)
-  library(portalr)
-  source('tools/model_tools.R')
-
-#' Function for pevGARCH
-#'
-#' @param abundances table of rodent abundances and time measures
-#' @param forecast_date the dates to be forecast from
-#' @param forecast_months the months of the dates to be forecast
-#' @param forecast_years the years of the dates to be forecast
-#' @param forecast_newmoons the numbers of the new moons to be forecast
-#' @param level name of the type of plots included ("All" or "Controls")
-#' @param num_forecast_newmoons number of new moons to forecast
-#' @param CI_level confidence interval level used for forecast envelope
-#' @param weather_data table of already lagged weather data
-#' @param weather_forecast table of forecasted weather
-#' @param ndvi_data table of already lagged NDVI data
-#' @param ndvi_forecast table of forecasted NDVI
-#' @return list of forecast and aic tables
-#'
-  forecast_pevgarch <- function(abundances, forecast_date, forecast_months, 
-                                forecast_years, forecast_newmoons, level,
-                                num_forecast_newmoons, CI_level = 0.9,
-                                weather_data, weather_fcast,
-                                ndvi_data, ndvi_fcast){
-
-    species <- colnames(abundances)[2:(ncol(abundances) - 3)]
-
-    interpolated_abundances <- interpolate_abundance(abundances)
-
-    fit_start_newmoon <- min(abundances$newmoonnumber)
-    fit_end_newmoon <- max(abundances$newmoonnumber)
-    initial_newmoon <- max(abundances$newmoonnumber)
-
-    zero_abund_mean <- rep(0, num_forecast_newmoons)
-    zero_abund_interval <- matrix(0, ncol = 2, nrow = num_forecast_newmoons)
-    colnames(zero_abund_interval) <- c("lower", "upper")
-
-    zero_abund_forecast <- list(zero_abund_mean, zero_abund_interval)
-    names(zero_abund_forecast) <- c("pred", "interval")
-
-    cov_data <- right_join(weather_data, ndvi_data, by = "newmoonnumber")
-    cov_nm <- cov_data$newmoonnumber
[…]

### 2018-12-01 tools/forecast_tools.R

@@ -5,100 +5,6 @@ library(ggplot2)
 library(rmarkdown)
 library(RCurl)
 
-##########################Forecast processing############################
-#' Combine all new forecasts (from the tmp directory), add ensembles
-#' 
-#' @param forecast_date
-#' @param filename_suffix
-#' @return list(forecasts,all_model_aic)
-#' @example forecastall('forecasts')
-
-forecastall <- function(forecast_date,filename_suffix = 'forecasts') {
-  
-  #Append results to forecasts and AIC tables
-  forecasts = do.call(rbind,
-                      lapply(list.files("tmp",pattern = paste(filename_suffix, ".csv",sep=""), full.names = TRUE), 
-                             read.csv, na.strings = "", colClasses = c("Date", "integer", "integer", 
-                                                                       "integer", "character", "character", "character", 
-                                                                       "character", "numeric", "numeric", "numeric",
-                                                                       "integer", "integer", "integer")))
-  
-  all_model_aic = do.call(rbind,
-                          lapply(list.files("tmp",pattern = paste(filename_suffix, "_model_aic.csv",sep=""), full.names = TRUE), 
-                                 read.csv, na.strings = ""))
-  
-  forecast_filename = file.path('predictions', paste(as.character(forecast_date), filename_suffix, ".csv", sep=""))
-  model_aic_filename = file.path('predictions', paste(as.character(forecast_date), filename_suffix, "_model_aic.csv", sep=""))
-  append_csv(forecasts, forecast_filename)
-  append_csv(all_model_aic, model_aic_filename)
-  
-  ########Add ensembles to files############################################
-  ensemble=make_ensemble(forecasts) %>% 
-    subset(select=colnames(forecasts))
-  append_csv(ensemble, forecast_filename)
-  
-  return(list(forecasts,all_model_aic))
-}
-
-######Tools for writing forecasts to file and aics to separate file###############
-#Appending a csv without re-writing the header.
-append_csv=function(df, filename){
-  write.table(df, filename, sep = ',', row.names = FALSE, col.names = !file.exists(filename), append = file.exists(filename))
-}
-
-#Get all model aic values and calculate akaike weights
[…]

### 2018-12-01 tools/prepare_data.R

@@ -1,73 +1,0 @@
-'%>%' <- magrittr::"%>%"
-library(yaml)
-library(portalr)
-source('tools/data_tools.R')
-
-filename_suffix = 'forecasts'
-
-#The date this forecast is run. Always today's date.
-forecast_date = Sys.Date()
-
-portalr::download_observations()
-moons = get_moon_data()
-moons$newmoondate <- as.Date(as.character(moons$newmoondate))
-current_moons <- select(moons, newmoonnumber, newmoondate, period, censusdate)
-#get dates of 12 new moons following newmoon of interest
-future_moons <- get_future_moons(moons)
-
-not_future_moons <- future_moons[which(future_moons$newmoondate < forecast_date), ]
-if (nrow(not_future_moons) > 0){
-
-  add_nfm <- not_future_moons
-  add_nfm$year <- as.numeric(format(add_nfm$newmoondate, "%Y"))
-  add_nfm$month <- as.numeric(format(add_nfm$newmoondate, "%m"))
-  moons <- rbind(moons, add_nfm)
-  current_moons <- select(moons, newmoonnumber, newmoondate, period, censusdate)
-  future_moons <- get_future_moons(moons)
-}
-
-total_moons <- rbind(current_moons, future_moons)
-
-
-rodent_data = get_rodent_data(moons, forecast_date)
-weather_data = weather("newmoon",fill=TRUE) %>% dplyr::ungroup() %>%
-                dplyr::select("mintemp","maxtemp","meantemp","precipitation","newmoonnumber")
-incompletes <- which(is.na(weather_data$newmoonnumber))
-if(length(incompletes) > 0){
-  weather_data <- weather_data[-incompletes, ]
-}
-ndvi_data <- ndvi("newmoon", fill = TRUE)
-
-# Beginning and end of the forecast timeperiod
-most_recent_newmoon <- moons$newmoonnumber[max(which(total_moons$newmoondate < forecast_date))]
-most_recent_rodent_period <- max(tail(rodent_data$all, 1)$period, tail(rodent_data$control, 1)$period)
-most_recent_rodent_newmoon <- moons$newmoonnumber[which(moons$period == most_recent_rodent_period)]
-most_recent_covariate_newmoon <- min(tail(weather_data, 1)$newmoonnumber, tail(ndvi_data, 1)$newmoonnumber)
-last_forecast_newmoon <- most_recent_newmoon + 12
-first_forecast_covariate_newmoon <- most_recent_covariate_newmoon + 1
[…]

### 2018-12-01 evaluation.Rmd
Beginning the switch to portalcasting
Add portalcasting to package installs, use portalcasting functions to drive directory setup and modeling

--- 2018-06-29 evaluation.Rmd
@@
 
-model_metadata = yaml.load_file("data/model_metadata.yaml")
+model_metadata = yaml.load_file("forecasting/data/model_metadata.yaml")
 pred_interval = model_metadata$pred_interval
@@
 species_of_interest = c('BA','DM','DO','PP','OT','NA','total')
-species_names = portalr::FullPath('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
+species_names = portalr:::full_path('forecasting/PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
   read.csv(stringsAsFactors=FALSE, na.strings = '') %>%
@@
 
-new_moon_file = portalr::FullPath('PortalData/Rodents/moon_dates.csv', '~')
+new_moon_file = portalr:::full_path('forecasting/PortalData/Rodents/moon_dates.csv', '~')
 new_moons = read.csv(new_moon_file)

### 2018-12-01 report.Rmd
Beginning the switch to portalcasting
Add portalcasting to package installs, use portalcasting functions to drive directory setup and modeling

@@
 # get most recent forecast prior to data collection
-files = data.frame(names = list.files(path = './predictions', pattern = '*forecasts.csv'))
+files = data.frame(names = list.files(path = './forecasting/predictions', pattern = '*forecasts.csv'))
 files$filedates = as.Date(substr(files$names,1,10),format='%Y-%m-%d')
@@
 
-data = read.csv(paste('predictions/',latest_forecast$names,sep=''),na.strings = '')
+data = read.csv(paste('forecasting/predictions/',latest_forecast$names,sep=''),na.strings = '')
 

### 2018-12-01 archive.sh

@@ -13,7 +13,7 @@ if [ "$TRAVIS_EVENT_TYPE" == "cron" ]; then
 
     # Commit changes to portalPredictions repo
     git checkout master
-    git add predictions/* docs/* data/*
+    git add forecasting/* docs/* 
     git commit -m "Update forecasts: Travis Build $TRAVIS_BUILD_NUMBER [ci skip]"
 
     git remote add deploy https://${GITHUB_TOKEN}@github.com/weecology/portalPredictions.git > /dev/null 2>&1
@@ -28,7 +28,7 @@ if [ "$TRAVIS_EVENT_TYPE" == "cron" ]; then
     # Clone forecasts archive repo
     cd ../../
     git clone https://github.com/weecology/forecasts
-    cp portalPredictions/portalPredictionsResult/predictions/*.* forecasts/portal/
+    cp portalPredictions/portalPredictionsResult/forecasting/predictions/*.* forecasts/portal/
     cd forecasts
 
     # Commit to forecasts repo

### 2018-12-01 forecasting/PortalData/DataCleaningScripts/clean_shrub_transect_data.R

@@ -1,0 +1,143 @@
+# Clean Shrub Transect Data 
+# most of the code taken from EMC
+# organized by EKB
+# May 11, 2016
+# refactored 4/24/18 EMC
+
+# Notes:
+#   Section 1
+#     * a. Ensure season, year, and filepath are correct *
+#   Section 2
+#       a. check double-entered data [general_data_cleaning_functions.R]
+#   Section 3
+#     * a. Ensure that species list path is correct *
+#   Section 4
+#       a. ONLY DO THIS ONCE!
+#     * b. Ensure the file path is correct *
+
+#####################################################################################
+
+library(openxlsx)
+library(dplyr)
+source('DataCleaningScripts/general_data_cleaning_functions.R')
+source('DataCleaningScripts/plant_data_cleaning_functions.R')
+
+######################
+# 1. Load Excel file #
+######################
+season <-  'Summer'
+year <-  '2018'
+filepath <-  '/Users/renatadiaz/Dropbox/Portal/PORTAL_primary_data/Plant/TRANSECTS/ShrubTransects(2015-present)/RawData/'
+
+excel_file <-  paste(filepath, "ShrubTransect_", season, year, '.xlsx', sep='')
+
+##################################
+# 2. Compare double-entered data #
+##################################
+
+unmatched = compare_worksheets(excel_file)
+
+# iterate through mismatches and fix them
+ws1 = openxlsx::read.xlsx(excel_file, sheet = 1, colNames = TRUE, na.strings = '')
+ws2 = openxlsx::read.xlsx(excel_file, sheet = 2, colNames = TRUE, na.strings = '')
+
+i = 1
+
+i  = i + 1
+unmatched[i, ]
+ws1[unmatched[i, 'row'] -1 , ]
+ws2[unmatched[i, 'row'] - 1, ]
+
+# Save matching datasheet
+write.csv(ws1, '/Users/renatadiaz/Dropbox/Portal/PORTAL_primary_data/Plant/TRANSECTS/ShrubTransects(2015-present)/RawData/ShrubTransect_Summer2018_clean.csv', row.names = FALSE)
+
+######################
+# 3. Quality control #
+######################
+
+ws = read.csv('/Users/renatadiaz/Dropbox/Portal/PORTAL_primary_data/Plant/TRANSECTS/ShrubTransects(2015-present)/RawData/ShrubTransect_Summer2018_clean.csv', stringsAsFactors= F)
+
+splist = read.csv('Plants/Portal_plant_species.csv',as.is=T)
+
[…]

### 2018-12-01 forecasting/PortalData/DataCleaningScripts/get_gimms_NDVI.R

@@ -1,0 +1,124 @@
+#Functions for retrieving and processing NDVI data for Portal.
+#Currently only GIMMS is supported, with years 1981-2013 available.
+# This code is modified from get_ndvi_data.R script in the Weecology repo: BBS-forecasting
+#(https://github.com/weecology/bbs-forecasting/blob/master/R/get_ndvi_data.R)
+# Modified for Portal by Morgan Ernest and Shawn Taylor
+
+#GIMMS: Call get_gimms_ndvi() to retrieve a dataframe of (year, month, ndvi).
+#-It will take a while to download the 14gb of data and extract values the 1st time around. 
+#-To redo it just delete the database file.
+#-na values are due to missing periods from filtering out unwanted quality flags.
+library(dplyr)
+library(gimms)
+library(sp)
+library(rgdal)
+
+#################################################
+#This assumes we want all gimms files that are available. It queries for files
+#that are available for download and compares against files already in the gimms
+#data directory.
+#Returns a list of files to download, which may be length 0.
+##################################################
+get_gimms_download_list=function(gimms_folder = './GIMMS'){
+  available_files_download_path=gimms::updateInventory(version=0)
+  available_files_name=basename(available_files_download_path)
+
+  files_present=list.files(gimms_folder)
+  #hdr files are created from some of the gimms processing that we don't want to
+  #use here.
+  files_present=files_present[!grepl('hdr', files_present)]
+
+  to_download=available_files_download_path[! available_files_name %in% files_present]
+
+  return(to_download)
+}
+
+################################################
+#Extract values from a single gimms file given a set of coordinates.
+#Excludes values which don't meet NDVI quality flags.
+#From the GIMMS readme:
+#FLAG = 7 (missing data)
+#FLAG = 6 (NDVI retrieved from average seasonal profile, possibly snow)
+#FLAG = 5 (NDVI retrieved from average seasonal profile)
+#FLAG = 4 (NDVI retrieved from spline interpolation, possibly snow)
+#FLAG = 3 (NDVI retrieved from spline interpolation)
+#FLAG = 2 (Good value)
+#FLAG = 1 (Good value)
+################################################
[…]

### 2018-12-01 forecasting/PortalData/DataCleaningScripts/new_moon_numbers.r

@@ -1,0 +1,115 @@
+# some code to match new moon dates to period sampling dates
+
+library(dplyr)
+library(lubridate)
+library(lunar)
+
+#' Find date of first trapping night for each period in a data frame
+#' Returns data frame of Period (unique period numbers) and CensusDate
+#'
+#'
+#'
+#' @param dat data frame containing at least columns "Period" and "CensusDate"
+#' 
+#' @example find_first_trap_night(trappingdat)
+#' 
+find_first_trap_night = function(dat) {
+  trap_dates = c()
+  for (p in unique(dat$period)) {
+    perioddate <- dat %>% filter(period==p) %>% filter(censusdate == min(censusdate)) %>% head(1) %>% select(period,censusdate)
+    trap_dates = rbind(trap_dates,perioddate)
+  }
+  return(trap_dates)
+}
+
+#' Finds the date in a vector of dates (new moons) that is closest to a target date (a census)
+#' Returns the index of the vector
+#'
+#'
+#' @param target_date single date object
+#' @param newmoondates vector of date objects
+#' 
+#' @example closest_newmoon(as.Date('2017-02-26'),moon_dates$newmoondate)
+#' 
+closest_newmoon = function(target_date,newmoondates) {
+  loc = which.min(abs(target_date-newmoondates))
+  closest = newmoondates[loc]
+  return(closest)
+}
+
+#' Updates existing version of moon_dates.csv with latest census numbers
+#'
+#'
+#'
+#' 
+#' @example update_moon_dates()
+
+update_moon_dates = function() {
+  # load existing moon_dates.csv file
+  moon_dates=read.csv("../Rodents/moon_dates.csv",stringsAsFactors = FALSE)
+  moon_dates$newmoondate = as_date(moon_dates$newmoondate)
+  moon_dates$censusdate = as_date(moon_dates$censusdate)
+  # load rodent trapping data
+  trappingdat=read.csv("../Rodents/Portal_rodent_trapping.csv")  
+  trappingdat$censusdate = as_date(paste(trappingdat$year,trappingdat$month,trappingdat$day,sep='-'))
+  
+  
+  # proceed only if trappingdat has more recent dates than moon_dates
+  if (max(abs(trappingdat$period),na.rm=T) > max(abs(moon_dates$period),na.rm=TRUE)) {  
+   
+  #Get dates and period numbers of new data only  
[…]

### 2018-12-01 forecasting/PortalData/DataCleaningScripts/rodent_data_cleaning_functions.R

@@ -1,0 +1,141 @@
+library(dplyr)
+
+currentdir = getwd()
+
+if(substr(currentdir, nchar(currentdir) - 8, nchar(currentdir)) == '/testthat') {
+source('../DataCleaningScripts/general_data_cleaning_functions.R')
+} else {
+  source('DataCleaningScripts/general_data_cleaning_functions.R')
+}
+
+
+#' Main function: runs a few rodent data quality checks
+#' 
+#'
+#' @param ws data frame read in from raw data excel file
+#' @param scannerfile path to txt file of tag numbers downloaded from tag scanner
+#' 
+#' @return none, unless there are errors in data
+
+rodent_data_quality_checks = function(ws,scannerfile) {
+  
+  # Compare tag numbers entered on sheets to scanner download
+  unpaired = compare_tags(ws,scannerfile)
+  if (nrow(unpaired) > 0) {
+    print('tag number problem:')
+    print(unpaired)
+  }
+  
+  # Check that reproductive charactaristics match M/F designation
+  MFcheck = male_female_check(ws)
+  if (length(MFcheck) > 0) {
+    print(paste('check M/F in rows:',paste(MFcheck,collapse='  ')))
+  }
+  
+  # Check for duplicate stake numbers within a plot
+  dups = suspect_stake(ws)
+  if (nrow(dups) > 0) {
+    print('suspect stakes:')
+    print(dups)
+  }
+  
+  # Check all plots present in data
+  missingplots = all_plots(ws)
+  if (length(missingplots)>0) {print(paste('missing plots:',paste(missingplots,collapse='  ')))}
+  
+  # Flag missing data
+  #   -fields all lines of data should have
+  m1 = check_missing_data(ws,fields = c('month','day','year','period','plot'))
+  if (length(m1)>0) {print(paste('missing data in row: ',paste(m1,collapse='  ')))}
+  #   -fields that should be filled, excluding cases that already have a flag
+  #    in the note1 field
+  m2 = check_missing_data(ws[is.na(ws$note1),],fields=c('stake','species','sex','hfl','wgt'))
+  if (length(m2)>0) {print(paste('missing data in row: ',paste(m2,collapse='  ')))}
+  #   - NOTE: does not check for missing tag or sexual characteristics
+  
+}
+
+
+
+
+
+
+
+
+#' This function compares pit tags entered on data sheets to tags downloaded from scanner
+#' 
[…]

### 2018-12-01 forecasting/PortalData/DataCleaningScripts/update_portal_plant_censuses.R

@@ -1,0 +1,56 @@
+# This creates the Portal_plant_censuses.csv file in the Plants folder
+# Record of which quadrats were or were not counted in each census and exact dates
+# **This assumes all quadrats were censused in each season listed in the data**
+# If a quadrat is flagged as missing in the data cleaning process, 
+# you must manually change it's value to zero
+# If a season gets skipped entirely, it should also be manually added
+
+library(dplyr)
+
+#' Appends new census dates to Portal_plant_censuses
+#'
+#'
+#' 
+#' @example update_portal_plant_censuses()
+#' 
+#'
+#'
+
+update_portal_plant_censuses = function() {
+  # load quadrat data
+  plantdat = read.csv("../Plants/Portal_plant_quadrats.csv",stringsAsFactors = FALSE,as.is=TRUE,na.strings = '')  
+  # load rodent trapping data
+  censuses=read.csv("../Plants/Portal_plant_censuses.csv",stringsAsFactors = FALSE)  
+  
+  # proceed only if plantdat has more recent data than censuses
+  
+  #find new rows
+  newdat = unique(anti_join(plantdat[,1:3],censuses[,1:3], by = c("year", "season", "plot")))
+  
+  if (nrow(newdat) != 0) {
+    
+    # extract quadrat data beyond what's already in censuses
+    quadrats = c(11,13,15,17,31,33,35,37,51,53,55,57,71,73,75,77)
+    newdat = merge(newdat,quadrats) %>% rename(quadrat=y) %>% arrange(year,desc(season),plot,quadrat) 
+    newdat$censused = 1
+    newdat$area = 0.25
+
+    # append to censuses
+    censuses = rbind(censuses,newdat)
+  }
+  return(censuses)
+}
+
+#' Rewrites file Portal_plant_censuses.csv with latest census info
+#'
+#'
+#' 
+#' @example writecensustable()
+#' 
+#'
+#'
+writecensustable <- function() {
+  
+  censuses=update_portal_plant_censuses()
+  # write updated data frame to csv
+  write.csv(censuses, file="../Plants/Portal_plant_censuses.csv", row.names=FALSE, quote = FALSE) }

### 2018-12-01 forecasting/PortalData/EML/eml_portal_data.R

@@ -1,0 +1,582 @@
+# title: "EML for Streaming Data"
+# author: "Anna Liu"
+# date: "October 5, 2016"
+
+#devtools::install_github("ropensci/EML")
+library(EML)
+library(RCurl)
+
+## Data tables
+# Ants
+
+# read files 
+ant_bait = getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Ants/Portal_ant_bait.csv")
+Encoding(ant_bait) = "latin1"
+ant_bait = iconv(ant_bait, "latin1", "UTF-8", sub="")
+
+ant_colony = getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Ants/Portal_ant_colony.csv")
+Encoding(ant_colony) = "latin1"
+ant_colony = iconv(ant_colony, "latin1", "UTF-8", sub="")
+
+ant_dataflags = read.csv("https://raw.githubusercontent.com/weecology/PortalData/master/Ants/Portal_ant_dataflags.csv")
+ant_dataflags = as.data.frame(sapply(ant_dataflags, function(x){
+  x = as.character(x)
+  Encoding(x) = "latin1"
+  iconv(x, "latin1", "UTF-8", sub="")
+}))
+
+ant_species = read.csv("https://raw.githubusercontent.com/weecology/PortalData/master/Ants/Portal_ant_species.csv")
+ant_species = as.data.frame(sapply(ant_species, function(x){
+  x = as.character(x)
+  Encoding(x) = "latin1"
+  iconv(x, "latin1", "UTF-8", sub="")
+}))
+
+# eml for bait
+ant_bait_attributes = data.frame(
+  attributeName=c("Month","Year","Plot","Stake","Species","Abundance"), 
+  formatString=c(NA,"YYYY",NA,NA,NA,NA), 
+  definition=c("July",NA,"plot number","stake number","species name","Number of individuals"), 
+  unit=c("dimensionless","nominalYear","number",
+         "number","dimensionless","number"),
+  attributeDefinition=c("July of each year",
+                        "Year. Range: 1977-2009",
+                        "Plot number. Range: 1-24",
+                        "Stake number on each plot. Range: 1-49",
+                        "Species name (string)",
+                        "Number of individuals of a certain species within the 10 cm diameter bait circle"),
+  numberType=c(NA,NA,"integer","integer",NA,"integer"),
+  stringsAsFactors=FALSE)
+
[…]

### 2018-12-01 forecasting/PortalData/testthat/test-rodent_functions.R

@@ -1,0 +1,41 @@
+# Uses a test dataset to make sure rodent data cleaning functions are catching errors
+
+library(openxlsx)
+library(testthat)
+source('../DataCleaningScripts/general_data_cleaning_functions.R')
+source('../DataCleaningScripts/rodent_data_cleaning_functions.R')
+
+
+context("checks rodent data cleaning functions")
+
+testfile = '../DataCleaningScripts/rodent_test_data.xlsx'
+
+testdat = read.xlsx(testfile,sheet=1,colNames = T,na.strings = '')
+scannerfile = '../DataCleaningScripts/test_tags.txt'
+
+
+test_that("worksheets 1 and 2 match", {
+  expect_equal(compare_worksheets(testfile),data.frame(row=40,column='hfl'))
+})
+
+test_that("Check that tag numbers match between the sheets and scanner", {
+  expect_equal(compare_tags(testdat,scannerfile),data.frame(where=c('scan','sheet'),tag=c('B267E8','B267EB')))
+})
+
+test_that("Check for conflict between M/F and reproductive characteristics", {
+  expect_equal(male_female_check(testdat),146)
+})
+
+test_that("Check for duplicate plot/stake pairs", {
+  expect_equal(suspect_stake(testdat),data.frame(plot=17,stake=65))
+})
+
+test_that("check that all 24 plots present in data", {
+  expect_equal(all_plots(testdat),c(1,10))
+})
+
+test_that("look for missing data ", {
+  expect_equal(check_missing_data(testdat,c('month','day','year','period','plot')),c(5,15))
+  expect_equal(check_missing_data(testdat[is.na(testdat$note1),],fields=c('stake','species','sex','hfl','wgt')),11)
+})
+

### 2018-12-01 forecasting/PortalData/testthat/test-shrub_transects.R

@@ -1,0 +1,81 @@
+library(testthat)
+context("checks shrub transect data")
+
+data <- read.csv("../Plants/Portal_plant_transects_2015_present.csv")
+old_transects <- read.csv("../Plants/Portal_plant_transects_1989_2009.csv")
+species <-  read.csv('../Plants/Portal_plant_species.csv')
+
+test_that("valid year", {
+  
+  expect_true(all(data$year %in% 1981:max(data$year)))
+})
+
+test_that("valid plot", {
+  
+  expect_true(all(data$plot %in% 1:24))
+})
+
+test_that("valid transect", {
+  
+  expect_true(all(data$transect %in% c(11,71)))
+})
+
+test_that("valid species", {
+  
+  expect_true(all(data$species %in% species$speciescode))
+})
+
+test_that("valid start", {
+  
+  expect_true(all(data$start %in% 0:7500))
+})
+
+
+test_that("valid stop", {
+  
+  expect_true(all(data$stop %in% 0:7500))
+})
+
+
+test_that("stop greater than start", {
+  
+  expect_true(all(data$stop >= data$start))
+})
+
+test_that("valid height", {
+  
+  expect_true(min(data$height, na.rm = T) > 0)
+  expect_true(max(data$height, na.rm = T) <= 400)
+})
+
+test_that("no duplicate data", {
+  
+  expect_true(sum(duplicated(data))==0)
+})
+
+test_that("valid year", {
+  
+  expect_true(all(old_transects$year %in% c(1989,1992,1995,1998,2001,2004,2009)))
+})
+
+test_that("valid plot", {
+  
+  expect_true(all(old_transects$plot %in% 1:24))
+})
+
+test_that("valid transect", {
+  
+  expect_true(all(old_transects$transect %in% c("NW","SW","NE","SE","")))
+})
+
+test_that("valid species", {
+  
+  expect_true(all(old_transects$species %in% species$speciescode))
+})
+
+test_that("valid point", {
+  
+  expect_true(all(old_transects$point %in% c(1:250,NA)))
+})
+
+

### 2018-12-01 forecasting/PortalData/update_version.R

@@ -1,0 +1,24 @@
+#' Update the version number
+#'
+#' By default update the minor version number since most changes are new data.
+#' If [major] or [patch] is in the last commit summary increment the matching
+#' version instead.
+
+repo <- git2r::repository(".")
+last_commit <- git2r::commits(repo)[[1]]
+current_ver <- semver::parse_version(readLines("version.txt"))
+if (grepl("Merge pull request", last_commit['summary'])){
+    last_commit <- git2r::commits(repo)[[2]]
+}
+
+if (grepl("\\[no version bump\\]", last_commit['summary'])) {
+  new_ver <- current_ver
+} else if (grepl("\\[major\\]", last_commit['summary'])) {
+  new_ver <- semver::increment_version(current_ver, "major", 1L)
+} else if (grepl("\\[patch\\]", last_commit['summary'])) {
+  new_ver <- semver::increment_version(current_ver, "patch", 1L)
+} else {
+  new_ver <- semver::increment_version(current_ver, "minor", 1L)
+}
+
+writeLines(as.character(new_ver), "version.txt")

### 2018-12-01 forecasting/models/AutoArima.R

@@ -1,0 +1,7 @@
+tree <- dirtree("~", "forecasting", c("predictions", "models", "PortalData", "data", "tmp"));
+all <- read.csv(file_path(tree, "data/all.csv"));
+controls <- read.csv(file_path(tree, "data/controls.csv"));
+metadata <- yaml::yaml.load_file(file_path(tree, "data/metadata.yaml"));
+f_a <- AutoArima(all, metadata);
+f_c <- AutoArima(controls, metadata, level = "Controls");
+save_forecast_output(f_a, f_c, "AutoArima", metadata, sub_path(tree, "tmp"))

### 2018-12-01 forecasting/models/ESSS.R

@@ -1,0 +1,7 @@
+tree <- dirtree("~", "forecasting", c("predictions", "models", "PortalData", "data", "tmp"));
+all <- read.csv(file_path(tree, "data/all.csv"));
+controls <- read.csv(file_path(tree, "data/controls.csv"));
+metadata <- yaml::yaml.load_file(file_path(tree, "data/metadata.yaml"));
+f_a <- ESSS(all, metadata);
+f_c <- ESSS(controls, metadata, level = "Controls");
+save_forecast_output(f_a, f_c, "ESSS", metadata, sub_path(tree, "tmp"))

### 2018-12-01 forecasting/models/nbGARCH.R

@@ -1,0 +1,7 @@
+tree <- dirtree("~", "forecasting", c("predictions", "models", "PortalData", "data", "tmp"));
+all <- read.csv(file_path(tree, "data/all.csv"));
+controls <- read.csv(file_path(tree, "data/controls.csv"));
+metadata <- yaml::yaml.load_file(file_path(tree, "data/metadata.yaml"));
+f_a <- nbGARCH(all, metadata);
+f_c <- nbGARCH(controls, metadata, level = "Controls");
+save_forecast_output(f_a, f_c, "nbGARCH", metadata, sub_path(tree, "tmp"))

### 2018-12-01 forecasting/models/pevGARCH.R

@@ -1,0 +1,8 @@
+tree <- dirtree("~", "forecasting", c("predictions", "models", "PortalData", "data", "tmp"));
+all <- read.csv(file_path(tree, "data/all.csv"));
+controls <- read.csv(file_path(tree, "data/controls.csv"));
+covariates <- read.csv(file_path(tree, "data/covariates.csv")); 
+metadata <- yaml::yaml.load_file(file_path(tree, "data/metadata.yaml"));
+f_a <- pevGARCH(all, covariates, metadata, lag = 6);
+f_c <- pevGARCH(controls, covariates, metadata, level = "Controls", lag = 6);
+save_forecast_output(f_a, f_c, "pevGARCH", metadata, sub_path(tree, "tmp"))

### 2018-12-01 install-packages.R

@@ -8,3 +8,4 @@ if ("pacman" %in% rownames(installed.packages()) == FALSE) install.packages("pac
 pacman::p_load(devtools, dplyr, forecast, ggplot2, htmltab, lubridate, ltsa, magrittr, 
                parallel, RCurl, readr, rmarkdown, testthat, tidyverse, tscount, yaml, zoo)
 pacman::p_load_gh('weecology/portalr')
+pacman::p_load_gh('weecology/portalcasting')

### 2018-12-01 models/nbgarch.R

@@ -1,152 +1,0 @@
-# nbGARCH
-#
-# Model "nbGARCH" is a generalized autoregresive conditional 
-#  heteroscedasticity model with overdispersion (i.e. a negative binomial 
-#  response variable)
-#  
-
-  library(tscount)
-  library(forecast)
-  source('tools/model_tools.R')
-
-#' Function for nbGARCH
-#'
-#' @param abundances table of rodent abundances and time measures
-#' @param forecast_date the dates to be forecast from
-#' @param forecast_months the months of the dates to be forecast
-#' @param forecast_years the years of the dates to be forecast
-#' @param forecast_newmoons the numbers of the new moons to be forecast
-#' @param level name of the type of plots included ("All" or "Controls")
-#' @param num_forecast_newmoons number of new moons to forecast
-#' @param CI_level confidence interval level used for forecast envelope
-#' @return list of forecast and aic tables
-#'
-  forecast_nbgarch <- function(abundances, forecast_date, forecast_months, 
-                               forecast_years, forecast_newmoons, level,
-                               num_forecast_newmoons, CI_level = 0.9){
-
-    species <- colnames(abundances)[2:(ncol(abundances) - 3)]
-
-    interpolated_abundances <- interpolate_abundance(abundances)
-
-    fit_start_newmoon <- min(abundances$newmoonnumber)
-    fit_end_newmoon <- max(abundances$newmoonnumber)
-    initial_newmoon <- max(abundances$newmoonnumber)
-
-    zero_abund_mean <- rep(0, num_forecast_newmoons)
-    zero_abund_interval <- matrix(0, ncol = 2, nrow = num_forecast_newmoons)
-    colnames(zero_abund_interval) <- c("lower", "upper")
-
-    zero_abund_forecast <- list(zero_abund_mean, zero_abund_interval)
-    names(zero_abund_forecast) <- c("pred", "interval")
-
-    output_fcast <- data.frame()
-    output_aic <- data.frame()
-
-    for(s in species){
-
-      ss <- s
-      if(ss == "NA."){
-        ss <- "NA"
-      }
-      cat("Fitting negative binomial GARCH model for", ss, "\n")
-
-      species_abundance <- interpolated_abundances %>% extract2(s)
-    
-      if(sum(species_abundance) == 0){
-        spec_forecast <- zero_abund_forecast
-        spec_aic <- 1e6
-      } else{
-        nbgarch_mod <- tsglm(species_abundance, 
[…]

### 2018-12-01 tests/testthat/test-new_forecasts.R

@@ -2,11 +2,11 @@ library(testthat)
 library(yaml)
 library(zoo)
 context("checks that new forecasts are being added correctly")
-model_metadata = yaml.load_file("../../data/model_metadata.yaml")
+model_metadata = yaml.load_file("../../forecasting/data/model_metadata.yaml")
 
-forecasts=read.csv(file.path('../../predictions', paste(as.character(as.Date(model_metadata$forecast_date)), 
+forecasts=read.csv(file.path('../../forecasting/predictions', paste(as.character(as.Date(model_metadata$forecast_date)), 
                                         model_metadata$filename_suffix, ".csv", sep="")), na.strings = "")
-forecastaics=read.csv(file.path('../../predictions', paste(as.character(as.Date(model_metadata$forecast_date)), 
+forecastaics=read.csv(file.path('../../forecasting/predictions', paste(as.character(as.Date(model_metadata$forecast_date)), 
                                         model_metadata$filename_suffix, "_model_aic.csv", sep="")), na.strings = "")
 
 forecastnames = c("date", "forecastmonth",  "forecastyear",  "newmoonnumber",

### 2018-12-01 tools/data_tools.R

@@ -25,7 +25,7 @@ FullPath <- function( ReferencePath, BasePath=getwd()){
 #' 
 get_moon_data <- function(){
   #Get the newmoon number of the  most recent sample
-  moons <- read.csv(FullPath('PortalData/Rodents/moon_dates.csv', '~'), header=T)
+  moons <- read.csv(FullPath('forecasting/PortalData/Rodents/moon_dates.csv', '~'), header=T)
   
   #Add in year and month to join with the rest of the data
   moons$year=year(moons$newmoondate)

### 2018-12-03 evaluation.Rmd
Remove unused tools

@@
 
-model_metadata = yaml.load_file("forecasting/data/model_metadata.yaml")
+model_metadata = yaml.load_file("forecasting/data/metadata.yaml")
 pred_interval = model_metadata$pred_interval

### 2018-12-03 report.Rmd
Reorganize, cleanup function writing

--- 2018-02-16 report.Rmd
@@
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+'%>%' <- magrittr::'%>%'
+source("tools/forecast_tools.R")
+```
+
 Here's the most recent actual data (blue) compared to the forecast performed right before data collection (black, with error bars)
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-source("tools/forecast_tools.R")
-library(portalr)
 
 # get most recent rodent data
-rodents = abundance("repo", shape='flat', level='treatment', length='longterm')
-most_recent_period = max(rodents$period)
-observed = dplyr::filter(rodents, period == most_recent_period, treatment=='control')
+observed = portalr::abundance(clean=FALSE, shape='flat', level='treatment', plots='longterm', time = 'newmoon') %>%
+            dplyr::filter(newmoonnumber == max(newmoonnumber), treatment=='control')
 
 # get new moon number associated with this period
-newmoons_table = read.csv(text = RCurl::getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
-newmoons_table$censusdate = as.Date(newmoons_table$censusdate,format='%Y-%m-%d')
-moon_code = dplyr::filter(newmoons_table,period==most_recent_period) %>% select(newmoonnumber) %>% as.integer()
-most_recent_date = dplyr::filter(newmoons_table,period==most_recent_period) %>% dplyr::pull(censusdate)
+most_recent_date = read.csv('PortalData/Rodents/moon_dates.csv') %>%
+                   dplyr::mutate(censusdate = as.Date(censusdate,format='%Y-%m-%d')) %>%
+                   dplyr::filter(newmoonnumber==unique(observed$newmoonnumber)) %>% 
+                   dplyr::pull(censusdate)
 
@@
 
-ensemble = dplyr::filter(data, level == 'Controls', model == 'Ensemble', newmoonnumber==moon_code)
+ensemble = dplyr::filter(data, level == 'Controls', model == 'Ensemble', newmoonnumber==unique(observed$newmoonnumber))
 sp_predictions = get_sp_predicts(ensemble, 'Controls', lead_time = 1)
 
-joined_data = left_join(sp_predictions, observed, by = "species")
+joined_data = dplyr::left_join(sp_predictions, observed, by = c('species','newmoonnumber'))
 joined_data[is.na(joined_data)] = 0
 joined_data[joined_data$species=='total','abundance'] = sum(joined_data$abundance,na.rm=T)
-joined_data[joined_data$species=='total','period'] = most_recent_period
 
 sp_predict = plot_species_forecast(sp_predictions,title=paste0(sp_predictions$forecast_date[2], ": Control plots"))  
-sp_predict = sp_predict + geom_point(data = joined_data, mapping = aes(x = abundance, y = species),
+sp_predict = sp_predict + ggplot2::geom_point(data = joined_data, mapping = ggplot2::aes(x = abundance, y = species),
              color = "blue")
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-source("tools/forecast_tools.R")
-library(portalr)
 
 # get most recent rodent data
-rodents = abundance("repo", shape='flat')
-most_recent_period = max(rodents$period)
-observed = dplyr::filter(rodents, period == most_recent_period)
+observed = portalr::abundance(clean=FALSE, shape='flat', time = 'newmoon') %>%
+            dplyr::filter(newmoonnumber == max(newmoonnumber))
 
 # get new moon number associated with this period
-newmoons_table = read.csv(text = RCurl::getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
-newmoons_table$censusdate = as.Date(newmoons_table$censusdate,format='%Y-%m-%d')
-moon_code = dplyr::filter(newmoons_table,period==most_recent_period) %>% select(newmoonnumber) %>% as.integer()
-most_recent_date = dplyr::filter(newmoons_table,period==most_recent_period) %>% dplyr::pull(censusdate)
+most_recent_date = read.csv('PortalData/Rodents/moon_dates.csv') %>%
+                   dplyr::mutate(censusdate = as.Date(censusdate,format='%Y-%m-%d')) %>%
+                   dplyr::filter(newmoonnumber==unique(observed$newmoonnumber)) %>% 
+                   dplyr::pull(censusdate)
 
@@
 
-ensemble = dplyr::filter(data, level == 'All', model == 'Ensemble', newmoonnumber==moon_code)
+ensemble = dplyr::filter(data, level == 'All', model == 'Ensemble', newmoonnumber==unique(observed$newmoonnumber))
 sp_predictions = get_sp_predicts(ensemble, 'All', lead_time = 1)
 
-joined_data = left_join(sp_predictions, observed, by = "species")
+joined_data = dplyr::left_join(sp_predictions, observed, by = c('species','newmoonnumber'))
 joined_data[is.na(joined_data)] = 0
 joined_data[joined_data$species=='total','abundance'] = sum(joined_data$abundance,na.rm=T)
-joined_data[joined_data$species=='total','period'] = most_recent_period
 
 sp_predict = plot_species_forecast(sp_predictions,title=paste0(sp_predictions$forecast_date[2], ": All plots"))  
-sp_predict = sp_predict + geom_point(data = joined_data, mapping = aes(x = abundance, y = species),
+sp_predict = sp_predict + ggplot2::geom_point(data = joined_data, mapping = ggplot2::aes(x = abundance, y = species),
              color = "blue")

### 2018-12-03 archive.sh

@@ -28,7 +28,7 @@ if [ "$TRAVIS_EVENT_TYPE" == "cron" ]; then
     # Clone forecasts archive repo
     cd ../../
     git clone https://github.com/weecology/forecasts
-    cp portalPredictions/portalPredictionsResult/forecasting/predictions/*.* forecasts/portal/
+    cp portalPredictions/portalPredictionsResult/predictions/*.* forecasts/portal/
     cd forecasts
 
     # Commit to forecasts repo

### 2018-12-03 forecasting/PortalData/DataCleaningScripts/clean_shrub_transect_data.R

@@ -1,143 +1,0 @@
-# Clean Shrub Transect Data 
-# most of the code taken from EMC
-# organized by EKB
-# May 11, 2016
-# refactored 4/24/18 EMC
-
-# Notes:
-#   Section 1
-#     * a. Ensure season, year, and filepath are correct *
-#   Section 2
-#       a. check double-entered data [general_data_cleaning_functions.R]
-#   Section 3
-#     * a. Ensure that species list path is correct *
-#   Section 4
-#       a. ONLY DO THIS ONCE!
-#     * b. Ensure the file path is correct *
-
-#####################################################################################
-
-library(openxlsx)
-library(dplyr)
-source('DataCleaningScripts/general_data_cleaning_functions.R')
-source('DataCleaningScripts/plant_data_cleaning_functions.R')
-
-######################
-# 1. Load Excel file #
-######################
-season <-  'Summer'
-year <-  '2018'
-filepath <-  '/Users/renatadiaz/Dropbox/Portal/PORTAL_primary_data/Plant/TRANSECTS/ShrubTransects(2015-present)/RawData/'
-
-excel_file <-  paste(filepath, "ShrubTransect_", season, year, '.xlsx', sep='')
-
-##################################
-# 2. Compare double-entered data #
-##################################
-
-unmatched = compare_worksheets(excel_file)
-
-# iterate through mismatches and fix them
-ws1 = openxlsx::read.xlsx(excel_file, sheet = 1, colNames = TRUE, na.strings = '')
-ws2 = openxlsx::read.xlsx(excel_file, sheet = 2, colNames = TRUE, na.strings = '')
-
-i = 1
-
-i  = i + 1
-unmatched[i, ]
-ws1[unmatched[i, 'row'] -1 , ]
-ws2[unmatched[i, 'row'] - 1, ]
-
-# Save matching datasheet
-write.csv(ws1, '/Users/renatadiaz/Dropbox/Portal/PORTAL_primary_data/Plant/TRANSECTS/ShrubTransects(2015-present)/RawData/ShrubTransect_Summer2018_clean.csv', row.names = FALSE)
-
-######################
-# 3. Quality control #
-######################
-
-ws = read.csv('/Users/renatadiaz/Dropbox/Portal/PORTAL_primary_data/Plant/TRANSECTS/ShrubTransects(2015-present)/RawData/ShrubTransect_Summer2018_clean.csv', stringsAsFactors= F)
-
-splist = read.csv('Plants/Portal_plant_species.csv',as.is=T)
-
[…]

### 2018-12-03 forecasting/PortalData/DataCleaningScripts/general_data_cleaning_functions.R

@@ -1,83 +1,0 @@
-# These functions are used in the data cleaning process for both rodent data and plant data
-
-
-#library(openxlsx)
-
-
-#' @title compare worksheets
-#' @description compares sheet 1 and sheet 2 of an excel worksheet of double-entered data (i.e. the two sheets should be identical)
-#'              Row order matters: data sheets must be entered in same order in both sheets of the excel file
-#'              prints "worksheets identical" if the two worksheets match
-#'
-#' @param excel_file path to excel file 
-#' @return data frame containing row number and column name where any mismatches occur
-#' 
-compare_worksheets = function(excel_file) {
-  
-  # this function compares the two excel worksheet to identify inconsistencies
-  
-  # load data from excel workbook
-  ws1 = openxlsx::read.xlsx(excel_file, sheet = 1, colNames = TRUE, na.strings = c('', 'NA', ' '))
-  ws2 = openxlsx::read.xlsx(excel_file, sheet = 2, colNames = TRUE, na.strings = c('', 'NA', ''))
-  
-  
-  # if the two worksheets are identical, exit function
-  if (identical(ws1,ws2)) {                                   
-    print('Worksheets identical')
-  } 
-  
-  # otherwise, loop through rows one at a time
-  else {
-      unmatched = data.frame(row = c(),column = c())         # empty data frame for storing output
-      num_rows = length(ws1$month)
-      curr_row = 1
-      while (curr_row<=num_rows) {
-        v1 = as.character(as.vector(ws1[curr_row,]))          # extract row from worksheet 1
-        v2 = as.character(as.vector(ws2[curr_row,]))          # extract row from worksheet 2
-        
-        # if the two versions of the row are not identical
-        if (!identical(v1,v2)) {
-          # loop through each element in the row
-          col_error = vector()
-          for (n in seq(length(v1))) {                        
-            if (!identical(v1[n],v2[n])) {
-              # add the column name to output vector
-              col_error = append(col_error,colnames(ws1)[n])
-            }
-          }
-          # append row and column info to output data frame (curr_row+1 to skip header in excel file)
-          unmatched = rbind(unmatched,data.frame(row = curr_row+1,column = col_error))
-        }
-        curr_row = curr_row+1             # increment index and continue loop
-      }
-      return(unmatched)
-    }
-}
-
-#' @title check for presence of all plots in data
[…]

### 2018-12-03 forecasting/PortalData/DataCleaningScripts/get_landsat_data.R

@@ -1,249 +1,0 @@
-# written by Erica Christensen 1/8/15; modified 1/25/18
-# This script processes Landsat8 data from compressed file to NDVI calculation
-
-library(raster)
-library(rgdal)
-library(dplyr)
-
-
-# Steps each file goes through:
-#    1. extract raster files from compressed file
-#    2. crop ndvi and QA rasters to area around Portal
-#    3. delete large raster files, save small raster files to separate folder
-#    4. use QA raster to mask bad pixels
-#    5. apply scale factor (.0001)
-#    6. summarize masked and cropped ndvi image to get mean ndvi value per image
-#  then compile summarized data for all files into a single csv
-
-############################################################################
-# Functions
-############################################################################
-
-#' @title create_portal_box
-#' 
-#' @description creates a square SpatialPolygons object centered at a given lon/lat point and a given radius from that point.
-#'              CRS must be specified
-#'              center of Portal: 31.937769, -109.08029
-#'              corners of Portal: NW = 31.939568, -109.083177
-#'                    NE = 31.939569, -109.077460
-#'                    SW = 31.935860, -109.082948
-#'                    SE = 31.935948, -109.077317
-#' 
-#' @param centroid vector of coordinates in lon/lat of center of box [lon,lat]
-#' @param radius desired radius, in m
-#' @param portal_crs CRS of desired SpatialPolygons object, needs to match rasters
-#' 
-#' @return portal_box is a SpatialPolygons object that can be used for cropping rasters
-#' 
-#' @example create_portal_box(centroid = c(-109.08029, 31.937769), radius = 1125, 
-#'                            portal_crs = CRS("+proj=utm +zone=12 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0 "))
-#' 
-create_portal_box = function(centroid=c(-109.08029, 31.937769), 
-                             radius=1125, 
-                             portal_crs=CRS("+proj=utm +zone=12 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0 ")) {
-  # create SpatialPointsDataFrame of centroid of site
[…]

### 2018-12-03 forecasting/PortalData/DataCleaningScripts/get_regional_weather.R

@@ -1,108 +1,0 @@
-`%>%` <- magrittr::`%>%`
-
-#' Reports regional weather data with the data from
-#' DAILY GLOBAL HISTORICAL CLIMATOLOGY NETWORK (GHCN-DAILY). 
-#' 
-#' @example get_regional_weather()
-
-get_regional_weather <- function() {
-  
-portal4sw_station=readLines('https://www1.ncdc.noaa.gov/pub/data/ghcn/daily/all/USC00026716.dly')
-sansimon_station=readLines('https://www1.ncdc.noaa.gov/pub/data/ghcn/daily/all/US1AZCH0005.dly')
-
-portal4sw = clean_station_data(portal4sw_station)
-sansimon = clean_station_data(sansimon_station)
-
-#Keep only new data
-all_4sw = read.csv(file = "../Weather/Portal4sw_regional_weather.csv",header=T,
-                   colClasses=c("character", rep("integer",3), "character", "integer", rep("character",3), "Date"))
-all_sansimon = read.csv(file = "../Weather/Sansimon_regional_weather.csv",header=T,
-                        colClasses=c("character", rep("integer",3), "character", "integer", rep("character",3), "Date"))
-
-new_4sw=dplyr::setdiff(portal4sw,all_4sw)
-new_sansimon=dplyr::setdiff(sansimon,all_sansimon)
-
-return(list(new_4sw,new_sansimon))
-
-}
-
-
-#' Creates a dataframe from the mess DAILY GLOBAL HISTORICAL CLIMATOLOGY NETWORK (GHCN-DAILY) gives you. 
-#' The metadata specify substrings that correspond to columns.
-#' However, within a month, the timeseries grows by adding days to the string (as 'columns') rather than as new rows.
-#' Pull out values from string and assign to rows, adding a 'day' column, to get the data in a normal long format.
-#' Data from regional stations are available at ncdc.noaa.gov/pub/data/ghcn/daily (the GHCN-DAILY timeseries).
-#' Metadata are described here: https://www1.ncdc.noaa.gov/pub/data/ghcn/daily/readme.txt
-#' 
-#' @param stationdata A vector of data in the format of GHCN-DAILY
-
-
-clean_station_data <- function(stationdata) {
-
-data_out  = data.frame(id=as.character(substr(stationdata,1,11)), stringsAsFactors=FALSE)
-data_out$year = as.integer(substr(stationdata,12,15))
[…]

### 2018-12-03 forecasting/PortalData/DataCleaningScripts/new_weather_data.R

@@ -1,96 +1,0 @@
-`%>%` <- magrittr::`%>%`
-
-#' Selects new weather data
-#'
-#'
-#'
-#' @example new_met_data()
-#'
-#'
-#'
-# This function checks for new data at the datalogger-hosted website
-
-new_met_data <- function() {
-  
-  #httr::set_config(httr::timeout(seconds = 120))
-
-# Pull raw data (latest week of records, plus some overlap for saftey) & rename columns
-
-rawdata = htmltab::htmltab(doc='http://166.153.133.121/?command=TableDisplay&table=MET&records=1000', sep = "")  %>% 
-
-  dplyr::rename(airtemp=AirTC_Avg,precipitation=Rain_mm_Tot,timestamp=TimeStamp,record=Record,battv=BattV)
-
-# Pull raw storms data (latest 2500 records) & rename columns
-
-stormsnew = htmltab::htmltab(doc="http://166.153.133.121/?command=TableDisplay&table=Storms&records=2500", sep = "")  %>%
- 
-  dplyr::rename(timestamp = TimeStamp, record = Record, battv = BattV_Min, precipitation = Rain_mm_Tot)
-
-# Convert Timestamp
-rawdata$timestamp = lubridate::ymd_hms(rawdata$timestamp)
-stormsnew$timestamp = lubridate::ymd_hms(stormsnew$timestamp)
-
-#Get Year, Month, Day, Hour
-rawdata=cbind(year = lubridate::year(rawdata$timestamp),
-              month = lubridate::month(rawdata$timestamp),
-              day = lubridate::day(rawdata$timestamp),
-              hour = lubridate::hour(rawdata$timestamp),rawdata)
-
-#Fix hour and day so midnight=2400
-rawdata$hour[rawdata$hour==0] = 24 ; rawdata$hour = 100*rawdata$hour
-rawdata$day[rawdata$hour==2400] = rawdata$day[which(rawdata$hour==2400)-1]
-rawdata$month[rawdata$hour==2400] = rawdata$month[which(rawdata$hour==2400)-1]
-rawdata$year[rawdata$hour==2400] = rawdata$year[which(rawdata$hour==2400)-1]
-
-#Fix column classes
-rawdata$record = as.integer(rawdata$record)
-rawdata[,7:25] = lapply(rawdata[,7:25],as.numeric)
-
-class(stormsnew$record)="numeric"
-class(stormsnew$battv)="numeric"
-class(stormsnew$precipitation)="numeric"
-
[…]

### 2018-12-03 forecasting/PortalData/DataCleaningScripts/old_weather_overlap.R

@@ -1,117 +1,0 @@
-# This script takes a raw .dat file downloaded directly from the Portal weather 
-# station, determines if there are any gaps, and appends the new data to the
-# weather overlap file.
-# Met445.dat can be used for testing - it has a lot of problems
-
-
-library(lubridate)
-library(dplyr)
-# ==============================================================================
-# Load file
-# ==============================================================================
-
-# Open raw .dat file of new data
-filepath = "~/Dropbox/Portal/PORTAL_primary_data/Weather/Raw_data/2002_Station/"
-
-metfile = "Met478"
-
-rawdata = read.csv(paste(filepath,metfile,'.dat',sep=''),head=F,sep=',',
-                   col.names=c('code','year','jday','hour','precipitation','airtemp','RH'))
-
-# Convert Julian day to month and day
-rawdata$date = as.Date(paste(rawdata$year,rawdata$jday),format='%Y %j')
-rawdata$month = as.integer(format(rawdata$date,'%m'))
-rawdata$day = as.integer(format(rawdata$date,'%d'))
-rawdata$timestamp = ymd_hms(paste(rawdata$year,"-",rawdata$month,"-",rawdata$day," ",rawdata$hour/100,":00:00",sep=""))
-
-# Select weather data (Code=101) from battery status data (Code=102) then add battery data as column
-weathdat = rawdata[rawdata$code==101,]
-battery= rawdata[rawdata$code==102,] %>% select(year,month,day,hour,battv=precipitation)
-weathdat=left_join(weathdat,battery,by=c("year","month","day","hour"))
-
-# Make precipitation correction 
-# calculate mL from current reading (with what the datalogger *thinks*): 4.73 mL/tip *(weathdat$precipitation mm / .1 mm/tip)
-# then calculate mm based on actual funnel on gauge: .254 mm/tip * calculated mL / 8.25 mL/tip
-weathdat$precipitation=4.73*(.254*weathdat$precipitation/.1)/8.25
-
-# ==============================================================================
-# 1. Quality control
-# ==============================================================================
-
-# check for errors in hour (should be 100,200,300,...)
-if (any(!(weathdat$hour %in% seq(from=100,to=2400,by=100)))) {
-  print('Hour error')
-  weathdat = filter(weathdat,hour %in% seq(100,2400,100))
[…]

### 2018-12-03 forecasting/PortalData/DataCleaningScripts/update_portal_rodent_trapping.r

@@ -1,57 +1,0 @@
-# This creates the Portal_rodent_trapping.csv file in the Rodents folder
-# Record of which plots were or were not trapped in each census and exact dates
-
-library(dplyr)
-library(zoo)
-
-#' Appends new trapping dates to Portal_rodent_trapping
-#'
-#'
-#' 
-#' @example update_portal_rodent_trapping()
-#' 
-#'
-#'
-
-update_portal_rodent_trapping = function() {
-  # load rodent data
-  rodentdat = read.csv("../Rodents/Portal_rodent.csv",stringsAsFactors = FALSE,as.is=TRUE,na.strings = '')  
-  # load rodent trapping data
-  trappingdat=read.csv("../Rodents/Portal_rodent_trapping.csv")  
-  
-  # proceed only if rodentdat has more recent data than trappingdat
-  if (max(abs(rodentdat$period)) > max(abs(trappingdat$period))) {
-    
-    # extract plot data beyond what's already in trappingdat
-    newdat = filter(rodentdat,period > max(trappingdat$period)) %>% filter(!is.na(plot)) %>% select(month,day,year,period,plot,note1)
-    newdat$sampled = rep(1)
-    newdat$effort = rep(49)
-    newdat$sampled[newdat$note1==4] = 0
-    newdat$effort[newdat$note1==4] = 0
-    newdat$qcflag = 0
-    # select unique rows and rearrange columns
-    newdat = newdat[!duplicated(select(newdat,period,plot)),] %>% select(day,month,year,period,plot,sampled,effort,qcflag)
-    # put in order of period, plot
-    newdat = newdat[order(newdat$period,newdat$plot),]
-    # append to trappingdat
-    trappingdat = rbind(trappingdat,newdat)
-    
-    dates = as.yearmon(paste(trappingdat$month,"/",trappingdat$year,sep=""), "%m/%Y")
-    trappingdat$qcflag[dates<as.yearmon(Sys.Date())-1] = 1
-  }
-  return(trappingdat)
-}
-
-#' Rewrites file Portal_rodent_trapping.csv with latest trapping dates
-#'
-#'
-#' 
-#' @example writetrappingtable()
-#' 
-#'
-#'
-writetrappingtable <- function() {
-  
-  trappingdat=update_portal_rodent_trapping()
-  # write updated data frame to csv
-  write.csv(trappingdat, file="../Rodents/Portal_rodent_trapping.csv", row.names=FALSE, quote = FALSE) }

### 2018-12-03 forecasting/PortalData/EML/eml_portal_data.R

@@ -1,582 +1,0 @@
-# title: "EML for Streaming Data"
-# author: "Anna Liu"
-# date: "October 5, 2016"
-
-#devtools::install_github("ropensci/EML")
-library(EML)
-library(RCurl)
-
-## Data tables
-# Ants
-
-# read files 
-ant_bait = getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Ants/Portal_ant_bait.csv")
-Encoding(ant_bait) = "latin1"
-ant_bait = iconv(ant_bait, "latin1", "UTF-8", sub="")
-
-ant_colony = getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Ants/Portal_ant_colony.csv")
-Encoding(ant_colony) = "latin1"
-ant_colony = iconv(ant_colony, "latin1", "UTF-8", sub="")
-
-ant_dataflags = read.csv("https://raw.githubusercontent.com/weecology/PortalData/master/Ants/Portal_ant_dataflags.csv")
-ant_dataflags = as.data.frame(sapply(ant_dataflags, function(x){
-  x = as.character(x)
-  Encoding(x) = "latin1"
-  iconv(x, "latin1", "UTF-8", sub="")
-}))
-
-ant_species = read.csv("https://raw.githubusercontent.com/weecology/PortalData/master/Ants/Portal_ant_species.csv")
-ant_species = as.data.frame(sapply(ant_species, function(x){
-  x = as.character(x)
-  Encoding(x) = "latin1"
-  iconv(x, "latin1", "UTF-8", sub="")
-}))
-
-# eml for bait
-ant_bait_attributes = data.frame(
-  attributeName=c("Month","Year","Plot","Stake","Species","Abundance"), 
-  formatString=c(NA,"YYYY",NA,NA,NA,NA), 
-  definition=c("July",NA,"plot number","stake number","species name","Number of individuals"), 
-  unit=c("dimensionless","nominalYear","number",
-         "number","dimensionless","number"),
-  attributeDefinition=c("July of each year",
-                        "Year. Range: 1977-2009",
-                        "Plot number. Range: 1-24",
-                        "Stake number on each plot. Range: 1-49",
-                        "Species name (string)",
-                        "Number of individuals of a certain species within the 10 cm diameter bait circle"),
-  numberType=c(NA,NA,"integer","integer",NA,"integer"),
-  stringsAsFactors=FALSE)
-
[…]

### 2018-12-03 forecasting/PortalData/testthat/test-plant_quadrats.R

@@ -1,57 +1,0 @@
-library(testthat)
-context("checks new plant quadrat data and plant species list")
-
-data <- read.csv("../Plants/Portal_plant_quadrats.csv")
-species <-  read.csv('../Plants/Portal_plant_species.csv')
-
-test_that("valid year", {
-  
-  expect_true(all(data$year %in% 1981:max(data$year)))
-})
-
-test_that("valid season", {
-  
-  expect_true(all(data$season %in% c("winter","summer")))
-})
-
-test_that("valid plot", {
-  
-  expect_true(all(data$plot %in% 1:24))
-})
-
-test_that("valid quadrat", {
-  
-  expect_true(all(data$quadrat %in% c(11,13,15,17,31,33,35,37,51,53,55,57,71,73,75,77)))
-})
-
-test_that("valid species", {
-  
-  expect_true(all(data$species %in% species$speciescode))
-})
-
-test_that("valid abundance", {
-  
-  expect_true(all(na.omit(data$abundance) %in% 0:2400))
-})
-
-test_that("valid cover", {
-  
-  expect_true(all(data$cover[which(data$cover!=0.1)] %in% 1:100))
-})
-
-
-test_that("no duplicate data", {
-  
-  expect_true(sum(duplicated(data))==0)
-})
-
-
-test_that("valid duration in Portal_plant_species", {
-  
-  expect_true(all(species$duration %in% c('Perennial','Annual','Unknown')))
-})
-
-test_that("valid community in Portal_plant_species", {
-  
-  expect_true(all(species$community %in% c('Shrub','Summer Annual','Perennial Forb','Summer and Winter Annual','Unknown','Winter Annual','Subshrub','Perennial Grass')))
-})

### 2018-12-03 forecasting/PortalData/testthat/test-rodent_functions.R

@@ -1,41 +1,0 @@
-# Uses a test dataset to make sure rodent data cleaning functions are catching errors
-
-library(openxlsx)
-library(testthat)
-source('../DataCleaningScripts/general_data_cleaning_functions.R')
-source('../DataCleaningScripts/rodent_data_cleaning_functions.R')
-
-
-context("checks rodent data cleaning functions")
-
-testfile = '../DataCleaningScripts/rodent_test_data.xlsx'
-
-testdat = read.xlsx(testfile,sheet=1,colNames = T,na.strings = '')
-scannerfile = '../DataCleaningScripts/test_tags.txt'
-
-
-test_that("worksheets 1 and 2 match", {
-  expect_equal(compare_worksheets(testfile),data.frame(row=40,column='hfl'))
-})
-
-test_that("Check that tag numbers match between the sheets and scanner", {
-  expect_equal(compare_tags(testdat,scannerfile),data.frame(where=c('scan','sheet'),tag=c('B267E8','B267EB')))
-})
-
-test_that("Check for conflict between M/F and reproductive characteristics", {
-  expect_equal(male_female_check(testdat),146)
-})
-
-test_that("Check for duplicate plot/stake pairs", {
-  expect_equal(suspect_stake(testdat),data.frame(plot=17,stake=65))
-})
-
-test_that("check that all 24 plots present in data", {
-  expect_equal(all_plots(testdat),c(1,10))
-})
-
-test_that("look for missing data ", {
-  expect_equal(check_missing_data(testdat,c('month','day','year','period','plot')),c(5,15))
-  expect_equal(check_missing_data(testdat[is.na(testdat$note1),],fields=c('stake','species','sex','hfl','wgt')),11)
-})
-

### 2018-12-03 forecasting/PortalData/testthat/test-shrub_transects.R

@@ -1,81 +1,0 @@
-library(testthat)
-context("checks shrub transect data")
-
-data <- read.csv("../Plants/Portal_plant_transects_2015_present.csv")
-old_transects <- read.csv("../Plants/Portal_plant_transects_1989_2009.csv")
-species <-  read.csv('../Plants/Portal_plant_species.csv')
-
-test_that("valid year", {
-  
-  expect_true(all(data$year %in% 1981:max(data$year)))
-})
-
-test_that("valid plot", {
-  
-  expect_true(all(data$plot %in% 1:24))
-})
-
-test_that("valid transect", {
-  
-  expect_true(all(data$transect %in% c(11,71)))
-})
-
-test_that("valid species", {
-  
-  expect_true(all(data$species %in% species$speciescode))
-})
-
-test_that("valid start", {
-  
-  expect_true(all(data$start %in% 0:7500))
-})
-
-
-test_that("valid stop", {
-  
-  expect_true(all(data$stop %in% 0:7500))
-})
-
-
-test_that("stop greater than start", {
-  
-  expect_true(all(data$stop >= data$start))
-})
-
-test_that("valid height", {
-  
-  expect_true(min(data$height, na.rm = T) > 0)
-  expect_true(max(data$height, na.rm = T) <= 400)
-})
-
-test_that("no duplicate data", {
-  
-  expect_true(sum(duplicated(data))==0)
-})
-
-test_that("valid year", {
-  
-  expect_true(all(old_transects$year %in% c(1989,1992,1995,1998,2001,2004,2009)))
-})
-
-test_that("valid plot", {
-  
-  expect_true(all(old_transects$plot %in% 1:24))
-})
-
-test_that("valid transect", {
-  
-  expect_true(all(old_transects$transect %in% c("NW","SW","NE","SE","")))
-})
-
-test_that("valid species", {
-  
-  expect_true(all(old_transects$species %in% species$speciescode))
-})
-
-test_that("valid point", {
-  
-  expect_true(all(old_transects$point %in% c(1:250,NA)))
-})
-
-

### 2018-12-03 forecasting/models/nbGARCH.R

@@ -1,7 +1,0 @@
-tree <- dirtree("~", "forecasting", c("predictions", "models", "PortalData", "data", "tmp"));
-all <- read.csv(file_path(tree, "data/all.csv"));
-controls <- read.csv(file_path(tree, "data/controls.csv"));
-metadata <- yaml::yaml.load_file(file_path(tree, "data/metadata.yaml"));
-f_a <- nbGARCH(all, metadata);
-f_c <- nbGARCH(controls, metadata, level = "Controls");
-save_forecast_output(f_a, f_c, "nbGARCH", metadata, sub_path(tree, "tmp"))

### 2018-12-03 models/ESSS.R

@@ -1,0 +1,7 @@
+tree <- dirtree("~", "portalPredictions", c("predictions", "models", "PortalData", "data", "tmp"));
+all <- read.csv(file_path(tree, "data/all.csv"));
+controls <- read.csv(file_path(tree, "data/controls.csv"));
+metadata <- yaml::yaml.load_file(file_path(tree, "data/metadata.yaml"));
+f_a <- ESSS(all, metadata);
+f_c <- ESSS(controls, metadata, level = "Controls");
+save_forecast_output(f_a, f_c, "ESSS", metadata, sub_path(tree, "tmp"))

### 2018-12-03 models/nbGARCH.R

@@ -1,0 +1,7 @@
+tree <- dirtree("~", "portalPredictions", c("predictions", "models", "PortalData", "data", "tmp"));
+all <- read.csv(file_path(tree, "data/all.csv"));
+controls <- read.csv(file_path(tree, "data/controls.csv"));
+metadata <- yaml::yaml.load_file(file_path(tree, "data/metadata.yaml"));
+f_a <- nbGARCH(all, metadata);
+f_c <- nbGARCH(controls, metadata, level = "Controls");
+save_forecast_output(f_a, f_c, "nbGARCH", metadata, sub_path(tree, "tmp"))

### 2018-12-03 tests/testthat/test-new_forecasts.R

@@ -1,12 +1,11 @@
 library(testthat)
-library(yaml)
-library(zoo)
-context("checks that new forecasts are being added correctly")
-model_metadata = yaml.load_file("../../forecasting/data/metadata.yaml")
 
-forecasts=read.csv(file.path('../../forecasting/predictions', paste(as.character(as.Date(model_metadata$forecast_date)), 
+context("checks that new forecasts are being added correctly")
+model_metadata = yaml::yaml.load_file("../data/metadata.yaml")
+
+forecasts=read.csv(file.path('../predictions', paste(as.character(as.Date(model_metadata$forecast_date)), 
                                         model_metadata$filename_suffix, ".csv", sep="")), na.strings = "")
-forecastaics=read.csv(file.path('../../forecasting/predictions', paste(as.character(as.Date(model_metadata$forecast_date)), 
+forecastaics=read.csv(file.path('../predictions', paste(as.character(as.Date(model_metadata$forecast_date)), 
                                         model_metadata$filename_suffix, "_model_aic.csv", sep="")), na.strings = "")
 
 forecastnames = c("date", "forecastmonth",  "forecastyear",  "newmoonnumber",

### 2018-12-03 tools/data_tools.R

@@ -1,82 +1,0 @@
-library(lubridate)
-library(dplyr)
-library(magrittr)
-library(htmltab)
-
-#' Return normalized path for all operating systems
-#' @param ReferencePath a path to join with current working directory
-#' @param BasePath Current working directory else path given
-#'
-#' @return
-#' @export
-#' @examples
-#' FullPath('PortalData/Rodents/Portal_rodent.csv')
-#' FullPath('PortalData/Rodents/Portal_rodent.csv', '~')
-FullPath <- function( ReferencePath, BasePath=getwd()){
-  BasePath = normalizePath(BasePath)
-  Path = normalizePath(file.path(BasePath, ReferencePath), mustWork = FALSE)
-  return (Path)
-}
-
-#' Get data for newmoonnumbers and related trapping period codes
-#' @return newmoons table
-#' @examples
-#' get_moon_data()
-#' 
-get_moon_data <- function(){
-  #Get the newmoon number of the  most recent sample
-  moons <- read.csv(FullPath('forecasting/PortalData/Rodents/moon_dates.csv', '~'), header=T)
-  
-  #Add in year and month to join with the rest of the data
-  moons$year=year(moons$newmoondate)
-  moons$month=month(moons$newmoondate)
-  return(moons)
-}
-
-####################################################################################
-#' Get rodent data, tailored for forecasting (all plots and controls only)
-#' @param moons current newmoonnumber table
-#' @param forecast_date date the forecast is run
-#' 
-#' @return a list of two dataframes, all plots and control plots
-#' @examples
-#' get_rodent_data(moons, forecast_date)
-#'
-
-get_rodent_data <- function(moons, forecast_date){
-  #Period 203/newmoonnumber 217 will be when the training data timeseries
-  #begins. Corresponding to Jan 1995
-  historic_start_period=203
-  historic_start_newmoon=217
-  
-  #get Portal abundance data for the entire site and for control plots only.
-  
-  #Control plots
-  controls = portalr::abundance(clean=FALSE,level="Treatment",type="Rodents",length="Longterm",min_plots=24)
-  #Drop PI
-  controls = controls[ , -which(colnames(controls) == "PI")]
-  #The total rodent count in each treatment
-  controls$total = rowSums(controls[,-(1:2)])
-  #Drop non-control treatments and add in newmoonnumber
-  controls = controls %>%
-    filter(treatment == 'control') %>%
[…]

### 2018-12-03 tools/forecast_tools.R

@@ -1,64 +1,90 @@
-library(tidyverse)
-library(lubridate)
-library(zoo)
-library(ggplot2)
-library(rmarkdown)
-library(RCurl)
-
 ##########Tools for forecast presentation##############
 
-#' 
-#' @param data
-#' @param lvl
-#' @param lead_time
-get_sp_predicts = function(data, lvl, lead_time) {
-  data = transform(data, forecast_date = as.yearmon(paste(forecastmonth, "/", forecastyear, sep =
-                                                            ""), format = "%m/%Y")) %>% transform(date = as.Date(date, "%Y-%m-%d"))
-  data1 = filter(data, level == lvl,
-                 date == max(as.Date(date)))
-  target_moon = min(data1$newmoonnumber) + (lead_time - 1)
-  data2 = filter(data1, newmoonnumber == target_moon)
+#' Visualize a time-series forecast
+#' Plots the observed time-series and the 1-step forecasts within it
+#' Plots the forecast time-series along with the prediction interval for future observations
+#' @param obs_data is a data.frame (observed data)
+#' @param obs_date_col_name is a string: name of the date column from obs_data
+#' @param obs_val_col_name is a string: name of the column of the value being forecast
+#' @param for_data is a data.frame (forecast data)
+#' @param for_date_col_name is a string: name of the date column from for_data
+#' @param for_val_col_name is a string: name of the column of value being forecast, from for_data
+#' @param for_model_name is a string: name of the model to be used from model column in for_data
+#' @param for_lowerpi_col_name is a string: name of the column of the lower confidence interval from for_data
+#' @param for_upperpi_col_name is a string: name of the column of the upper confidence interval from for_data
+#' @param start_newmoon is numeric: first new moon number to be plotted
+#' @param ylabel is a string: title for y-axis
+forecast_viz <- function(obs_data, obs_date_col_name, obs_val_col_name, for_data,
+                         for_date_col_name, for_val_col_name, for_model_name,
+                         for_lowerpi_col_name, for_upperpi_col_name, start_newmoon,
+                         ylabel){
+  for_data_sub = dplyr::filter(for_data, species == obs_val_col_name, model == for_model_name)
+  obs_data_sub = dplyr::filter(obs_data, newmoonnumber >= start_newmoon)
[…]

### 2018-12-03 evaluation.Rmd
Reorganize, cleanup function writing

--- 2018-11-06 evaluation.Rmd
@@
 )
-library(dplyr)
-library(portalr)
-library(yaml)
+'%>%' <- magrittr::'%>%'
 source("tools/forecast_tools.R")
 
-model_metadata = yaml.load_file("data/model_metadata.yaml")
+model_metadata = yaml::yaml.load_file("data/metadata.yaml")
 pred_interval = model_metadata$pred_interval
@@
 species_of_interest = c('BA','DM','DO','PP','OT','NA','total')
-species_names = portalr:::full_path('PortalData/Rodents/Portal_rodent_species.csv', '~') %>%
-  read.csv(stringsAsFactors=FALSE, na.strings = '') %>%
-  select(species = speciescode, full_species_name = scientificname)
+species_names = read.csv('PortalData/Rodents/Portal_rodent_species.csv',
+                         stringsAsFactors=FALSE, na.strings = '') %>%
+  dplyr::select(species = speciescode, full_species_name = scientificname) %>%
+  dplyr::add_row(species='total', full_species_name='Total Rodents')
 
-#add a total entry
-species_names = species_names %>%
-  add_row(species='total', full_species_name='Total Rodents')
-
-new_moon_file = portalr:::full_path('PortalData/Rodents/moon_dates.csv', '~')
-new_moons = read.csv(new_moon_file)
-new_moons$censusdate = as.Date(new_moons$censusdate)
+new_moons = read.csv('PortalData/Rodents/moon_dates.csv') %>%
+            dplyr::mutate(censusdate = as.Date(censusdate))
 
 # Setup rodent observations
-species_abundance = portalr::abundance("repo", clean=FALSE, shape='flat', level='treatment', 
-                                       length="longterm", time='newmoon') %>%
-  filter(treatment=="control") %>%
-  rename(actual = abundance, level = treatment) %>%
-  mutate(level='Controls',currency='abundance') %>%
-  filter(species %in% species_of_interest)
+species_abundance = portalr::abundance(clean=FALSE, shape='flat', level='treatment', 
+                                       plots="longterm", time='newmoon') %>%
+  dplyr::filter(treatment=="control") %>%
+  dplyr::rename(actual = abundance, level = treatment) %>%
+  dplyr::mutate(level='Controls',currency='abundance') %>%
+  dplyr::filter(species %in% species_of_interest)
 
-total_abundance = portalr::abundance("repo", clean=FALSE, shape='flat', level='treatment', 
-                                     length="longterm", time='newmoon') %>%
-  filter(treatment=="control") %>%
-  group_by(newmoonnumber) %>%
-  summarise(actual=sum(abundance)) %>%
-  ungroup() %>%
-  mutate(level='Controls',currency='abundance',species='total')
+total_abundance = portalr::abundance(clean=FALSE, shape='flat', level='treatment', 
+                                     plots="longterm", time='newmoon') %>%
+  dplyr::filter(treatment=="control") %>%
+  dplyr::group_by(newmoonnumber) %>%
+  dplyr::summarise(actual=sum(abundance)) %>%
+  dplyr::ungroup() %>%
+  dplyr::mutate(level='Controls',currency='abundance',species='total')
 
 observation_data = species_abundance %>%
-  bind_rows(total_abundance) %>% 
-  left_join(new_moons, by='newmoonnumber')
+  dplyr::bind_rows(total_abundance) %>% 
+  dplyr::left_join(new_moons, by='newmoonnumber')
 
@@
 forecast_data = compile_forecasts(use_hindcasts = FALSE) %>%
-  filter(model %in% models_of_interest)
+  dplyr::filter(model %in% models_of_interest)
 
@@
 forecast_dates_to_keep = forecast_data %>%
-  select(date, initial_newmoon) %>%
-  distinct() %>%
-  group_by(initial_newmoon) %>%
-  filter(date == min(date)) %>%
-  ungroup() %>%
-  mutate(keep='yes')
+  dplyr::select(date, initial_newmoon) %>%
+  dplyr::distinct() %>%
+  dplyr::group_by(initial_newmoon) %>%
+  dplyr::filter(date == min(date)) %>%
+  dplyr::ungroup() %>%
+  dplyr::mutate(keep='yes')
 
 forecast_data = forecast_data %>%
-  left_join(forecast_dates_to_keep, by=c('date','initial_newmoon')) %>%
-  filter(keep=='yes') %>%
-  select(-keep)
+  dplyr::left_join(forecast_dates_to_keep, by=c('date','initial_newmoon')) %>%
+  dplyr::filter(keep=='yes') %>%
+  dplyr::select(-keep)
 
@@
 forecast_errors = forecast_data %>%
-  left_join(observation_data, by=c('newmoonnumber','level','species','currency')) %>% 
-  mutate(rmse = sqrt((estimate - actual)^2))
+  dplyr::left_join(observation_data, by=c('newmoonnumber','level','species','currency')) %>% 
+  dplyr::mutate(rmse = sqrt((estimate - actual)^2))
 
@@
 forecast_errors = forecast_errors %>%
-  distinct()
+  dplyr::distinct()
 
@@
 forecast_errors = forecast_errors %>%
-  filter(!is.na(actual))
+  dplyr::filter(!is.na(actual))
 
@@
 forecast_errors = forecast_errors %>%
-  filter(species %in% species_of_interest) %>%
-  left_join(species_names, by='species') %>%
-  select(-species) %>%
-  rename(species=full_species_name)
+  dplyr::filter(species %in% species_of_interest) %>%
+  dplyr::left_join(species_names, by='species') %>%
+  dplyr::select(-species) %>%
+  dplyr::rename(species=full_species_name)
 
@@
 # So here three forecasts are chosen from 1, 6, and 12 newmoons into the past. 
-# Sometimes forecasts data is missing for a particular initial_newmoon, the recursive function here
-# attempts to find another one from the same time pluts/minus 2 month period. This ensures 3 forecasts
-# are always shown.
-
-# If there is no data for initial_newmoon 500, this will check nearby dates in this order: 499,501,498,402
-update_initial_newmoon = function(potential_initial_newmoon, attempt=1){
-  if(attempt==6){
-    return(NA)
-  }
-  lead_time_data = forecast_errors %>%
-    filter(initial_newmoon == potential_initial_newmoon)
-
-  # Alternate steps of -1, +2, -3, +4
-  next_try_step = ifelse(attempt%%2 == 0, attempt, attempt * -1)
-
-  # no data to show? then try one month back
-  if(nrow(lead_time_data)==0){
-    #print(paste('Try ',attempt, ', Nothing with ',potential_initial_newmoon,' trying ',potential_initial_newmoon+next_try_step))
-    return(update_initial_newmoon(potential_initial_newmoon+next_try_step, attempt+1))
-  } else {
-    return(potential_initial_newmoon)
-  }
-
-}
-# Adjust this to add more than 3 forecast dates, or change the lead times
+# Sometimes forecasts data is missing for a particular initial_newmoon, the recursive function 
+# update_initial_newmoon attempts to find another one from the same time pluts/minus 2 month period.
+#  This ensures 3 forecasts are always shown.
 lead_times_to_show = c(1,6,12)
@@
 forecast_errors = forecast_errors %>%
-  filter(initial_newmoon %in% initial_newmoons_to_keep)
+  dplyr::filter(initial_newmoon %in% initial_newmoons_to_keep)
 
 #####################################
-ggplot(forecast_errors, aes(x=censusdate, y=rmse, group=as.character(date), color=as.character(date))) +
-  geom_point(size=4)+
-  geom_line(size=1.5) +
-  scale_color_brewer(palette='Dark2') +
-  scale_x_date(breaks = unique(forecast_errors$censusdate)) +
-  facet_grid(species~model, scales = 'free_y') + 
-  theme_bw() +
-  theme(axis.text.x = element_text(angle=45, hjust=0.99, debug = FALSE),
-        axis.text = element_text(size=12),
-        axis.title = element_text(size=20),
-        strip.text = element_text(size=12),
-        legend.text = element_text(size=15),
-        legend.title = element_text(size=18)) +
-  labs(x='Sample Date',y='RMSE', color='Forecast Issue\nDate')
+ggplot2::ggplot(forecast_errors, ggplot2::aes(x=censusdate, y=rmse, group=as.character(date), color=as.character(date))) +
+  ggplot2::geom_point(size=4)+
+  ggplot2::geom_line(size=1.5) +
+  ggplot2::scale_color_brewer(palette='Dark2') +
+  ggplot2::scale_x_date(breaks = unique(forecast_errors$censusdate)) +
+  ggplot2::facet_grid(species~model, scales = 'free_y') + 
+  ggplot2::theme_bw() +
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle=45, hjust=0.99, debug = FALSE),
+        axis.text = ggplot2::element_text(size=12),
+        axis.title = ggplot2::element_text(size=20),
+        strip.text = ggplot2::element_text(size=12),
+        legend.text = ggplot2::element_text(size=15),
+        legend.title = ggplot2::element_text(size=18)) +
+  ggplot2::labs(x='Sample Date',y='RMSE', color='Forecast Issue\nDate')
 
@@
 forecast_errors = calculate_forecast_error(observation_data, forecast_data, error_metric = 'RMSE') %>%
-  filter(error_value < 200) %>% #Drop RMSE greater than this because it throws off all the graphs
-  bind_rows(calculate_forecast_error(observation_data, forecast_data, error_metric = 'coverage'))
+  dplyr::filter(error_value < 200) %>% #Drop RMSE greater than this because it throws off all the graphs
+  dplyr::bind_rows(calculate_forecast_error(observation_data, forecast_data, error_metric = 'coverage'))
 
 forecast_errors = forecast_errors %>%
-  left_join(species_names, by='species')
+  dplyr::left_join(species_names, by='species')
 
-ggplot(forecast_errors, aes(x=lead_time, y=error_value, group=model, color=model)) +
-  geom_point()+
-  geom_line() +
-  geom_hline(yintercept = 0.9) +
-  labs(x='Lead Time (New Moons)') +
-  facet_wrap(full_species_name~error_metric, scales = 'free_y', ncol=2) + 
-  theme_bw() +
-  labs(y = "Error Value", colour = "Model")
+ggplot2::ggplot(forecast_errors, ggplot2::aes(x=lead_time, y=error_value, group=model, color=model)) +
+  ggplot2::geom_point()+
+  ggplot2::geom_line() +
+  ggplot2::geom_hline(yintercept = 0.9) +
+  ggplot2::labs(x='Lead Time (New Moons)') +
+  ggplot2::facet_wrap(full_species_name~error_metric, scales = 'free_y', ncol=2) + 
+  ggplot2::theme_bw() +
+  ggplot2::labs(y = "Error Value", colour = "Model")
 

### 2018-12-03 index.Rmd
Reorganize, cleanup function writing

@@
 knitr::opts_chunk$set(echo = TRUE)
-library(dplyr)
-library(portalr)
+'%>%' <- magrittr::'%>%'
 source("tools/forecast_tools.R")
-source("tools/data_tools.R")
 ```
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-obs_data = portalr::abundance("repo", clean=FALSE, level='treatment', length="longterm") %>%
-  filter(treatment=="control")
-obs_data$total = rowSums(select(obs_data, -c(period, treatment)))
-new_moons = read.csv(text = RCurl::getURL("https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv"))
+obs_data = portalr::abundance(clean=FALSE, level='treatment', plots="longterm") %>%
+  dplyr::filter(treatment=="control")
+obs_data$total = rowSums(dplyr::select(obs_data, -c(period, treatment)))
+new_moons = read.csv('PortalData/Rodents/moon_dates.csv') %>%
+                   dplyr::mutate(censusdate = as.Date(censusdate),
+                          year = lubridate::year(newmoondate),
+                          month = lubridate::month(newmoondate))
 
-obs_data_newmoon = inner_join(obs_data, new_moons, by = c("period" = "period"))
-obs_data_newmoon$censusdate = as.Date(obs_data_newmoon$censusdate)
+obs_data_newmoon = dplyr::inner_join(obs_data, new_moons, by = c("period" = "period"))
 
@@
 
-most_recent_forecast = max(for_data$date)
-for_data = filter(for_data, date == most_recent_forecast)
-
-for_data$model = "ensemble"
 # get dates for future new moons for plotting
-future_moons = get_moon_data() %>% get_future_moons() %>% select(c("newmoonnumber","newmoondate"))
-for_data = inner_join(for_data, future_moons, by = c("newmoonnumber" = "newmoonnumber"))
+future_moons = new_moons %>% portalr::get_future_moons() %>%    
+                         dplyr::select(c("newmoonnumber","newmoondate"))
+for_data = dplyr::inner_join(for_data, future_moons, by = c("newmoonnumber" = "newmoonnumber"))
 
@@
              for_val_col_name = "estimate",
-             for_model_name = "ensemble",
+             for_model_name = "Ensemble",
              for_lowerpi_col_name = "LowerPI",
@@
 ```{r speciesforecasts, echo=FALSE, message=FALSE, warning=FALSE}
-source("tools/forecast_tools.R")
 
-data = compile_forecasts()
-ensemble = dplyr::filter(data, level == 'Controls', model == 'Ensemble', date == max(date))
-sp_predictions = get_sp_predicts(ensemble, 'Controls', 
-            lead_time = length(unique(ensemble$newmoonnumber)) - 11)
+sp_predictions = get_sp_predicts(for_data, 'Controls', 
+            lead_time = length(unique(for_data$newmoonnumber)) - 11)
 sp_predict = plot_species_forecast(sp_predictions,
@@
 most_abund_sp = sp_predictions %>% 
-  filter(species != 'total') %>% 
-  arrange(desc(estimate)) %>% 
+  dplyr::filter(species != 'total') %>% 
+  dplyr::arrange(desc(estimate)) %>% 
   head(3) %>% 
-  select(species)
+  dplyr::select(species)
 
 # load in rodent species table to get scientific names to display on plots
-species_table = read.csv(
-  text = RCurl::getURL(
-      "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_species.csv"),stringsAsFactors = F,na.strings = '')
-most_abund_sp_names = filter(species_table,speciescode %in% most_abund_sp$species) %>% select(speciescode,scientificname)
+species_table = read.csv('PortalData/Rodents/Portal_rodent_species.csv',
+                          stringsAsFactors = F,na.strings = '')
+most_abund_sp_names = dplyr::filter(species_table,speciescode %in% most_abund_sp$species) %>%      
+                      dplyr::select(speciescode,scientificname)
 
@@
                for_val_col_name = "estimate",
-               for_model_name = "ensemble",
+               for_model_name = "Ensemble",
                for_lowerpi_col_name = "LowerPI",

### 2018-12-03 PortalForecasts.R

@@ -1,12 +1,12 @@
 library(portalcasting)
 
 #Update data and models
-setup_dir()
+setup_dir(options_all = all_options(main = "portalPredictions", download_existing_predictions = FALSE))
 
 #Run all models using portalcasting defaults
 #Save predictions to predictions directory in base directory
 
-portalcast()
+portalcast(options_all = all_options(main = "portalPredictions"))
 
 #Update Website
 rmarkdown::render_site()

### 2018-12-03 forecasting/PortalData/DataCleaningScripts/FullPath.R

@@ -1,15 +1,0 @@
-#' Return normalized path for all operating systems
-#'
-#' @param ReferencePath a path to join with current working directory
-#' @param BasePath Current working directory else path given
-#'
-#' @return
-#' @export
-#' @examples
-#' FullPath('PortalData/Rodents/Portal_rodent.csv')
-#' FullPath('PortalData/Rodents/Portal_rodent.csv', '~')
-FullPath <- function( ReferencePath, BasePath=getwd()){
-  BasePath = normalizePath(BasePath)
-  Path = normalizePath(file.path(BasePath, ReferencePath), mustWork = FALSE)
-  return (Path)
-}

### 2018-12-03 forecasting/PortalData/DataCleaningScripts/new_moon_numbers.r

@@ -1,115 +1,0 @@
-# some code to match new moon dates to period sampling dates
-
-library(dplyr)
-library(lubridate)
-library(lunar)
-
-#' Find date of first trapping night for each period in a data frame
-#' Returns data frame of Period (unique period numbers) and CensusDate
-#'
-#'
-#'
-#' @param dat data frame containing at least columns "Period" and "CensusDate"
-#' 
-#' @example find_first_trap_night(trappingdat)
-#' 
-find_first_trap_night = function(dat) {
-  trap_dates = c()
-  for (p in unique(dat$period)) {
-    perioddate <- dat %>% filter(period==p) %>% filter(censusdate == min(censusdate)) %>% head(1) %>% select(period,censusdate)
-    trap_dates = rbind(trap_dates,perioddate)
-  }
-  return(trap_dates)
-}
-
-#' Finds the date in a vector of dates (new moons) that is closest to a target date (a census)
-#' Returns the index of the vector
-#'
-#'
-#' @param target_date single date object
-#' @param newmoondates vector of date objects
-#' 
-#' @example closest_newmoon(as.Date('2017-02-26'),moon_dates$newmoondate)
-#' 
-closest_newmoon = function(target_date,newmoondates) {
-  loc = which.min(abs(target_date-newmoondates))
-  closest = newmoondates[loc]
-  return(closest)
-}
-
-#' Updates existing version of moon_dates.csv with latest census numbers
-#'
-#'
-#'
-#' 
-#' @example update_moon_dates()
-
-update_moon_dates = function() {
-  # load existing moon_dates.csv file
-  moon_dates=read.csv("../Rodents/moon_dates.csv",stringsAsFactors = FALSE)
-  moon_dates$newmoondate = as_date(moon_dates$newmoondate)
-  moon_dates$censusdate = as_date(moon_dates$censusdate)
-  # load rodent trapping data
-  trappingdat=read.csv("../Rodents/Portal_rodent_trapping.csv")  
-  trappingdat$censusdate = as_date(paste(trappingdat$year,trappingdat$month,trappingdat$day,sep='-'))
-  
-  
-  # proceed only if trappingdat has more recent dates than moon_dates
-  if (max(abs(trappingdat$period),na.rm=T) > max(abs(moon_dates$period),na.rm=TRUE)) {  
-   
-  #Get dates and period numbers of new data only  
[…]

### 2018-12-03 forecasting/PortalData/DataCleaningScripts/process_raw_transect_data.R

@@ -1,312 +1,0 @@
-##Used in the original cleaning and reshaping of transect data from 1989 - 2009
-##Modified from https://github.com/sdtaylor/portalShrubs/blob/master/processRawTransectData.R
-##Final data saved in the file: PortalData/Plants/Portal_plant_transects_1989_2009.csv
-
-library(tidyr)
-library(dplyr)
-library(stringr)
-library(readxl)
-library(testthat)
-csvFolder='~/portalShrubs/shrubData/excelExports/'
-
-species=read.csv(paste('../Plants/Portal_plant_species.csv',sep='')) %>%
-  mutate(Species.Name = paste(genus,species,sep=" ")) %>%
-  distinct(Species.Name,.keep_all = TRUE)
-
-##Table of all transect names over time
-transect_names=data.frame(previous=c("A","B","C","D","NW","NE","SE","SW"),now=c("NW","NE","SE","SW","NW","NE","SE","SW"))
-
-#Each year of transect data is slightly different, so requires it's own processing
-#################################
-#################################
-### 1989
-speciesList89=read.csv(paste(csvFolder, '1989_species_list_ShawnsEdits.csv',sep='')) %>%
-  mutate_if(is.factor,as.character) %>%
-  mutate(Species.Name=replace(Species.Name,Code.Number==24,"Tragus berteronianus"))
-
-#Read in and reshape data
-year89=read.csv(paste(csvFolder,'89in.csv',sep='')) %>%
-  gather(Plot, SpeciesID, -Transect, -Point) %>%
-  mutate(Plot=substring(Plot, 5), year=1989,
-         SpeciesID = replace(SpeciesID,SpeciesID==19,'6,40'),
-         SpeciesID = replace(SpeciesID,SpeciesID==16,10)) %>%
-  mutate_if(is.factor,as.character) %>%
-  
-  #repeat columns for multiple entries
-  mutate(SpeciesID = strsplit(as.character(SpeciesID), ",")) %>%
-  unnest(SpeciesID) %>%
-  
-  #Assign species names
-  merge(speciesList89, by.x='SpeciesID',by.y='Code.Number') %>%
-  select(Transect, Point, Plot, year, Species.Name)
-
-which(is.na(year89$Species.Name))
-
-########
-### 1992
-speciesList92=read.csv(paste(csvFolder, '1992_species_list_ShawnsEdits.csv',sep='')) %>%
[…]

### 2018-12-03 forecasting/PortalData/DataCleaningScripts/rodent_data_cleaning_functions.R

@@ -1,141 +1,0 @@
-library(dplyr)
-
-currentdir = getwd()
-
-if(substr(currentdir, nchar(currentdir) - 8, nchar(currentdir)) == '/testthat') {
-source('../DataCleaningScripts/general_data_cleaning_functions.R')
-} else {
-  source('DataCleaningScripts/general_data_cleaning_functions.R')
-}
-
-
-#' Main function: runs a few rodent data quality checks
-#' 
-#'
-#' @param ws data frame read in from raw data excel file
-#' @param scannerfile path to txt file of tag numbers downloaded from tag scanner
-#' 
-#' @return none, unless there are errors in data
-
-rodent_data_quality_checks = function(ws,scannerfile) {
-  
-  # Compare tag numbers entered on sheets to scanner download
-  unpaired = compare_tags(ws,scannerfile)
-  if (nrow(unpaired) > 0) {
-    print('tag number problem:')
-    print(unpaired)
-  }
-  
-  # Check that reproductive charactaristics match M/F designation
-  MFcheck = male_female_check(ws)
-  if (length(MFcheck) > 0) {
-    print(paste('check M/F in rows:',paste(MFcheck,collapse='  ')))
-  }
-  
-  # Check for duplicate stake numbers within a plot
-  dups = suspect_stake(ws)
-  if (nrow(dups) > 0) {
-    print('suspect stakes:')
-    print(dups)
-  }
-  
-  # Check all plots present in data
-  missingplots = all_plots(ws)
-  if (length(missingplots)>0) {print(paste('missing plots:',paste(missingplots,collapse='  ')))}
-  
-  # Flag missing data
-  #   -fields all lines of data should have
-  m1 = check_missing_data(ws,fields = c('month','day','year','period','plot'))
-  if (length(m1)>0) {print(paste('missing data in row: ',paste(m1,collapse='  ')))}
-  #   -fields that should be filled, excluding cases that already have a flag
-  #    in the note1 field
-  m2 = check_missing_data(ws[is.na(ws$note1),],fields=c('stake','species','sex','hfl','wgt'))
-  if (length(m2)>0) {print(paste('missing data in row: ',paste(m2,collapse='  ')))}
-  #   - NOTE: does not check for missing tag or sexual characteristics
-  
-}
-
-
-
-
-
-
-
-
-#' This function compares pit tags entered on data sheets to tags downloaded from scanner
-#' 
[…]

### 2018-12-03 forecasting/PortalData/DataCleaningScripts/update_portal_plant_censuses.R

@@ -1,56 +1,0 @@
-# This creates the Portal_plant_censuses.csv file in the Plants folder
-# Record of which quadrats were or were not counted in each census and exact dates
-# **This assumes all quadrats were censused in each season listed in the data**
-# If a quadrat is flagged as missing in the data cleaning process, 
-# you must manually change it's value to zero
-# If a season gets skipped entirely, it should also be manually added
-
-library(dplyr)
-
-#' Appends new census dates to Portal_plant_censuses
-#'
-#'
-#' 
-#' @example update_portal_plant_censuses()
-#' 
-#'
-#'
-
-update_portal_plant_censuses = function() {
-  # load quadrat data
-  plantdat = read.csv("../Plants/Portal_plant_quadrats.csv",stringsAsFactors = FALSE,as.is=TRUE,na.strings = '')  
-  # load rodent trapping data
-  censuses=read.csv("../Plants/Portal_plant_censuses.csv",stringsAsFactors = FALSE)  
-  
-  # proceed only if plantdat has more recent data than censuses
-  
-  #find new rows
-  newdat = unique(anti_join(plantdat[,1:3],censuses[,1:3], by = c("year", "season", "plot")))
-  
-  if (nrow(newdat) != 0) {
-    
-    # extract quadrat data beyond what's already in censuses
-    quadrats = c(11,13,15,17,31,33,35,37,51,53,55,57,71,73,75,77)
-    newdat = merge(newdat,quadrats) %>% rename(quadrat=y) %>% arrange(year,desc(season),plot,quadrat) 
-    newdat$censused = 1
-    newdat$area = 0.25
-
-    # append to censuses
-    censuses = rbind(censuses,newdat)
-  }
-  return(censuses)
-}
-
-#' Rewrites file Portal_plant_censuses.csv with latest census info
-#'
-#'
-#' 
-#' @example writecensustable()
-#' 
-#'
-#'
-writecensustable <- function() {
-  
-  censuses=update_portal_plant_censuses()
-  # write updated data frame to csv
-  write.csv(censuses, file="../Plants/Portal_plant_censuses.csv", row.names=FALSE, quote = FALSE) }

### 2018-12-03 forecasting/PortalData/DataCleaningScripts/update_portal_plots.R

@@ -1,61 +1,0 @@
-library(zoo)
-library(dplyr)
-
-#' Appends new dates to Portal_plots
-#'
-#'
-#' 
-#' @example update_portal_plots()
-#' 
-#'
-#'
-
-
-update_portal_plots <- function() {
-  #load plot data
-  portal_plots = read.csv("../SiteandMethods/Portal_plots.csv",stringsAsFactors = FALSE)
-  # load rodent data
-  rodentdat = read.csv("../Rodents/Portal_rodent.csv",stringsAsFactors = FALSE,as.is=TRUE,na.strings = '')  
-  
-  # define currrent plot treatments
-  
-  controls = c(4,5,6,7,11,13,14,17,18,24)
-  removals = c(1,9,10,12,16,23)
-  exclosures = c(2,3,8,15,19,20,21,22)
-  
-  # proceed only if rodentdat has more recent data than plot data
-  #find new rows
-  newrows=which(paste(rodentdat$year,rodentdat$month) %in% paste(portal_plots$year,portal_plots$month)
-                  ==FALSE)
-  newdat = unique(na.omit(rodentdat[newrows,c(2,4)])) 
-  
-  if (nrow(newdat)>0) {
-    
-    plot=1:24
-    newplots=merge(newdat,plot,by=NULL) %>% rename(plot=y) %>% arrange(year,month,plot) 
-    
-    newplots$treatment = NA
-    
-    newplots = newplots %>% mutate(treatment = ifelse((plot %in% removals),
-                                                      "removal", treatment)) %>% 
-      mutate(treatment = ifelse((plot %in% exclosures),"exclosure", treatment)) %>% 
-      mutate(treatment = ifelse((plot %in% controls),"control", treatment))
-    
-    portal_plots = bind_rows(portal_plots,newplots)
-  }
-  
-  return(portal_plots)
-}
-
-#' Rewrites file Portal_plots.csv with latest trapping dates
-#'
-#'
-#' 
-#' @example writeportalplots()
-#' 
-#'
-#'
-writeportalplots <- function() {
-  portal_plots = update_portal_plots()
-  write.csv(portal_plots, file='../SiteandMethods/Portal_plots.csv', row.names = FALSE, quote = FALSE)
-}

### 2018-12-03 forecasting/PortalData/Reports/Permit_data.R

@@ -1,73 +1,0 @@
-create_df = function(dat) {
-  
-  #convert 'age' column to Adult/Juvenile
-  dat$lifestage = ifelse(dat$age=='J','Juvenile','Adult')
-  
-  #convert 'note5' column to alive/dead
-  dat$disposition = ifelse(dat$note5=='D','Dead','Released Alive')
-  
-  #convert date to yyyy-mm-dd format
-  date_info = with(dat,paste(year,month,day,sep='-'))
-  dates = as.Date(date_info)
-  
-  dataframe = data.frame(species = dat$species,
-                         dates,
-                         sex = dat$sex,
-                         lifestage = dat$lifestage,
-                         disposition = dat$disposition)
-  
-  return(dataframe)
-}
-
-create_permit_frame = function(new_df) {
-  #convert 2 letter species code to scientific name and common name
-  spdata = read.csv("Rodents/Portal_rodent_species.csv", head = T, sep = ',', na.strings = "", as.is=T)
-  new_df = merge(new_df,spdata,by.x='name',by.y='speciescode')
-  
-  # convert Tag to PIT if animal was released alive, blank if dead
-  new_df$Tag = ifelse(new_df$disposition == 'Released Alive','PIT','')
-  
-  #convert M/F to Male/Female
-  new_df$sex = as.character(new_df$sex)
-  new_df$sex[new_df$sex == 'M'] = 'Male'
-  new_df$sex[new_df$sex == 'F'] = 'Female'
-  new_df$sex[new_df$sex == ''] = 'na'
-  
-  outframe = data.frame(ScientificName = new_df$scientificname,
-                        CommonName = new_df$commonname,
-                        x = new_df$x,
-                        dates = new_df$dates,
-                        county = rep('Cochise',length(new_df$sex)),
-                        waterbody = rep('',length(new_df$sex)),
-                        easting = rep(681200.79,length(new_df$sex)),
-                        northing = rep(3535287.25,length(new_df$sex)),
-                        UTMZone = rep('12S',length(new_df$sex)),
-                        datum = rep('',length(new_df$sex)),
-                        lifestage = new_df$lifestage,
-                        sex = new_df$sex,
-                        disposition = new_df$disposition,
-                        museum = rep('',length(new_df$sex)),
-                        tag = new_df$Tag,
-                        fieldtag = rep('',length(new_df$sex)),
[…]

### 2018-12-03 forecasting/PortalData/install-packages.R

@@ -1,9 +1,0 @@
-# Install pacman if it isn't already installed
-
-if ("pacman" %in% rownames(installed.packages()) == FALSE) install.packages("pacman")
-
-
-# Install analysis packages using pacman
-
-pacman::p_load(stringr, Hmisc, dplyr, openxlsx, lubridate, lunar,
-               sqldf, RCurl, EML, testthat, htmltab, zoo, tidyr, semver)

### 2018-12-03 forecasting/PortalData/testthat.R

@@ -1,3 +1,0 @@
-library(testthat)
-
-test_dir("testthat", reporter = c("check", "progress"))

### 2018-12-03 forecasting/PortalData/testthat/test-new_moon_numbers.R

@@ -1,18 +1,0 @@
-source('../DataCleaningScripts/new_moon_numbers.r')
-
-library(testthat)
-context("checks that new moon numbers are being added correctly")
-moon_dates=update_moon_dates()
-
-test_that("no moondates are skipped or duplicated", {
-  
-  expect_identical(row(as.matrix(moon_dates$newmoonnumber))[,1],moon_dates$newmoonnumber)
-  expect_true(all(diff(moon_dates$newmoondate) %in% c(29,30)))
-})
-
-
-test_that("no periods skipped or duplicated", {
-
-  expect_true(all(diff(moon_dates$period)==1,na.rm=T))
-  expect_true(sum(duplicated(moon_dates$censusdate,incomparables=NA))==0)
-})

### 2018-12-03 forecasting/PortalData/testthat/test-rodent_data.R

@@ -1,92 +1,0 @@
-library(testthat)
-library(dplyr)
-context("checks that rodent data values are valid")
-
-data = read.csv("../Rodents/Portal_rodent.csv",
-  na.strings = c(""), colClasses = c('tag' = 'character'),
-  stringsAsFactors = FALSE)
-
-stakes = c(11:17,21:27,31:37,41:47,51:57,61:67,71:77)
-stakedata=data %>% filter(period > 0, complete.cases(stake))
-
-species = read.csv("../Rodents/Portal_rodent_species.csv",na.strings = c(""), 
-          colClasses=c(speciescode="character"))
-
-bodysizedata = data %>% filter(!(note1 == 12),!is.na(species)) %>% group_by(species) %>% 
-  left_join(species,by=c("species"="speciescode")) %>% filter(censustarget==1)
-  
-
-test_that("year, month and day are valid", {
-  
-  expect_true(all(data$year %in% 1977:format(Sys.Date(),"%Y")))
-  expect_true(all(data$month %in% 1:12))
-  expect_true(all(data$day %in% 1:31))
-  
-})
-
-test_that("plot values are valid", {
-  
-  expect_true(all(na.omit(data$plot) %in% 1:24))
-  
-})
-
-test_that("stake values are valid", {
-  
-  expect_true(all(stakedata$stake %in% stakes))
-  
-})
-
-test_that("species codes are valid", {
-  
-  expect_true(all(na.omit(data$species) %in% species$speciescode))
-  
-})
-
-test_that("sex codes are valid", {
-  
-  expect_true(all(na.omit(data$sex) %in% c("M","F")))
-  
-})
-
-test_that("reproduction codes are valid", {
-  
-  expect_true(all(na.omit(data$reprod) == "Z"))
-  expect_true(all(na.omit(data$age) == "J"))
-  expect_true(all(na.omit(data$testes) %in% c("S","R","M")))
-  expect_true(all(na.omit(data$vagina) %in% c("S","P","B")))
-  expect_true(all(na.omit(data$pregnant) == "P"))
-  expect_true(all(na.omit(data$nipples) %in% c("R","E","B")))
-  expect_true(all(na.omit(data$lactation) == "L"))
-  
-})
-
-test_that("body size measurements are reasonable for species", {
-  
-  expect_true(all(bodysizedata$hfl >= bodysizedata$minhfl, na.rm=TRUE))
[…]

### 2018-12-03 forecasting/PortalData/testthat/test-update_portal_censuses.R

@@ -1,42 +1,0 @@
-source('../DataCleaningScripts/update_portal_plant_censuses.R')
-
-library(testthat)
-context("checks new plant census dates are being added correctly")
-
-censuses=update_portal_plant_censuses()
-
-test_that("valid season", {
-  
-  expect_true(all(censuses$season %in% c("winter","summer")))
-})
-
-test_that("valid year", {
-  
-  expect_true(all(censuses$year %in% 1981:max(censuses$year,na.rm=T)))
-})
-
-test_that("valid plot", {
-  
-  expect_true(all(censuses$plot %in% 1:24))
-})
-
-test_that("valid quadrat", {
-  
-  expect_true(all(censuses$quadrat %in% c(11,13,15,17,31,33,35,37,51,53,55,57,71,73,75,77)))
-})
-
-test_that("valid censuses", {
-  
-  expect_true(all(censuses$censused %in% c(0,1)))
-})
-
-test_that("valid area", {
-  
-  expect_true(all(censuses$area %in% c(0,0.5,0.25)))
-})
-
-test_that("no duplicate data", {
-  
-  expect_true(sum(duplicated(censuses))==0)
-})
-

### 2018-12-03 forecasting/PortalData/testthat/test-update_portal_trapping.R

@@ -1,17 +1,0 @@
-source('../DataCleaningScripts/update_portal_rodent_trapping.r')
-
-library(testthat)
-context("checks new trapping dates are being added correctly")
-
-trappingdat=update_portal_rodent_trapping()
-
-test_that("no skipped periods, in order", {
-  
-  expect_true(all(diff(unique(trappingdat$period))==1))
-})
-
-
-test_that("no duplicate data", {
-
-  expect_true(sum(duplicated(trappingdat))==0)
-})

### 2018-12-03 forecasting/PortalData/testthat/test-weather.R

@@ -1,85 +1,0 @@
-source("../DataCleaningScripts/new_weather_data.R")
-library(testthat)
-context("checks new weather data")
-
-data=new_met_data()
-newdata = data[[1]]
-newdata_cols = colnames(newdata)
-weather = data[[2]]
-weather_cols = colnames(weather)
-stormsnew = data[[3]]
-stormsnew_cols = colnames(stormsnew)
-storms = data[[4]]
-storms_cols = colnames(storms)
-
-portal4sw = read.csv(file = "../Weather/Portal4sw_regional_weather.csv",header=T,
-                   colClasses=c("character", rep("integer",3), "character", "integer", rep("character",3), "Date"))
-sansimon = read.csv(file = "../Weather/Sansimon_regional_weather.csv",header=T,
-                        colClasses=c("character", rep("integer",3), "character", "integer", rep("character",3), "Date"))
-
-test_that("required column names in new weather df", {
-  
-  expect_identical(weather_cols, newdata_cols)
-  expect_identical(storms_cols, stormsnew_cols)
-})
-
-test_that("Hour in 100:2400", {
-  
-  expect_true(all(newdata$hour %in% seq(from=100,to=2400,by=100)))
-})
-
-test_that("Air Temperature ok", {
-
-  expect_true(all(newdata$airtemp > -30))
-  expect_true(all(newdata$airtemp <= 100))
-})
-
-test_that("Relative humidity ok", {
-
-  expect_true(all(newdata$RH > 0))
-  expect_true(all(newdata$RH <= 100))
-})
-
-test_that("battery status ok", {
-  
-  expect_true(all(newdata$battv > 9))
-  expect_true(all(stormsnew$battv > 9))
-})
-
-test_that("Precipitation ok", {
-  
-  expect_true(all(newdata$precipitation >= 0))
-  expect_true(all(newdata$precipitation < 100))
-  expect_true(all(stormsnew$precipitation >= 0))
-  expect_true(all(stormsnew$precipitation < 12))
-})
-
-test_that("Precipitation in multiples of 0.254", {
-  
-  expect_true(all(lapply(newdata$precipitation > 9, ifelse,
-    abs(newdata$precipitation / 0.254 - signif(newdata$precipitation / 0.254, digits = 3)) < 0.02,
-    newdata$precipitation%%0.254==0)))
-  expect_true(sum(stormsnew$precipitation%%0.254)==0)
-})
-
[…]

### 2018-12-03 forecasting/models/AutoArima.R

@@ -1,7 +1,0 @@
-tree <- dirtree("~", "forecasting", c("predictions", "models", "PortalData", "data", "tmp"));
-all <- read.csv(file_path(tree, "data/all.csv"));
-controls <- read.csv(file_path(tree, "data/controls.csv"));
-metadata <- yaml::yaml.load_file(file_path(tree, "data/metadata.yaml"));
-f_a <- AutoArima(all, metadata);
-f_c <- AutoArima(controls, metadata, level = "Controls");
-save_forecast_output(f_a, f_c, "AutoArima", metadata, sub_path(tree, "tmp"))

### 2018-12-03 forecasting/models/pevGARCH.R

@@ -1,8 +1,0 @@
-tree <- dirtree("~", "forecasting", c("predictions", "models", "PortalData", "data", "tmp"));
-all <- read.csv(file_path(tree, "data/all.csv"));
-controls <- read.csv(file_path(tree, "data/controls.csv"));
-covariates <- read.csv(file_path(tree, "data/covariates.csv")); 
-metadata <- yaml::yaml.load_file(file_path(tree, "data/metadata.yaml"));
-f_a <- pevGARCH(all, covariates, metadata, lag = 6);
-f_c <- pevGARCH(controls, covariates, metadata, level = "Controls", lag = 6);
-save_forecast_output(f_a, f_c, "pevGARCH", metadata, sub_path(tree, "tmp"))

### 2018-12-03 models/pevGARCH.R

@@ -1,0 +1,8 @@
+tree <- dirtree("~", "portalPredictions", c("predictions", "models", "PortalData", "data", "tmp"));
+all <- read.csv(file_path(tree, "data/all.csv"));
+controls <- read.csv(file_path(tree, "data/controls.csv"));
+covariates <- read.csv(file_path(tree, "data/covariates.csv")); 
+metadata <- yaml::yaml.load_file(file_path(tree, "data/metadata.yaml"));
+f_a <- pevGARCH(all, covariates, metadata, lag = 6);
+f_c <- pevGARCH(controls, covariates, metadata, level = "Controls", lag = 6);
+save_forecast_output(f_a, f_c, "pevGARCH", metadata, sub_path(tree, "tmp"))

### 2018-12-10 evaluation.Rmd
Pull data from repo for website building

@@
 
-model_metadata = yaml::yaml.load_file("data/metadata.yaml")
+model_metadata = yaml::yaml.load_file('data/metadata.yaml')
 pred_interval = model_metadata$pred_interval
@@
 species_of_interest = c('BA','DM','DO','PP','OT','NA','total')
-species_names = read.csv('PortalData/Rodents/Portal_rodent_species.csv',
-                         stringsAsFactors=FALSE, na.strings = '') %>%
+read.csv(
+  "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_species.csv",
+                          stringsAsFactors = F,na.strings = '') %>%
   dplyr::select(species = speciescode, full_species_name = scientificname) %>%
@@
 
-new_moons = read.csv('PortalData/Rodents/moon_dates.csv') %>%
+new_moons = read.csv(
+  "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv") %>%
             dplyr::mutate(censusdate = as.Date(censusdate))
@@
 # Setup rodent observations
-species_abundance = portalr::abundance(clean=FALSE, shape='flat', level='treatment', 
+species_abundance = portalr::abundance(path = "repo", clean=FALSE, shape='flat', level='treatment', 
                                        plots="longterm", time='newmoon') %>%
@@
 
-total_abundance = portalr::abundance(clean=FALSE, shape='flat', level='treatment', 
+total_abundance = portalr::abundance(path = "repo", clean=FALSE, shape='flat', level='treatment', 
                                      plots="longterm", time='newmoon') %>%

### 2018-12-10 index.Rmd
Pull data from repo for website building

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-obs_data = portalr::abundance(clean=FALSE, level='treatment', plots="longterm") %>%
+obs_data = portalr::abundance(path = "repo", clean=FALSE, level='treatment', plots="longterm") %>%
   dplyr::filter(treatment=="control")
 obs_data$total = rowSums(dplyr::select(obs_data, -c(period, treatment)))
-new_moons = read.csv('PortalData/Rodents/moon_dates.csv') %>%
+new_moons = read.csv(
+  "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv") %>%
                    dplyr::mutate(censusdate = as.Date(censusdate),
@@
 # load in rodent species table to get scientific names to display on plots
-species_table = read.csv('PortalData/Rodents/Portal_rodent_species.csv',
+species_table = read.csv(
+  "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_species.csv",
                           stringsAsFactors = F,na.strings = '')

### 2018-12-10 PortalForecasts.R

@@ -6,6 +6,7 @@ setup_dir(options_all = all_options(main = "portalPredictions", download_existin
 #Run all models using portalcasting defaults
 
 portalcast(options_all = all_options(main = "portalPredictions"))
+message("models done")
 
 #Update Website
 rmarkdown::render_site()

### 2018-12-12 evaluation.Rmd
Switch to portalcasting (#295)
Add portalcasting to package installs, use portalcasting functions to drive directory setup and modeling

* Remove unused tools

* Reorganize, cleanup function writing

* Cleanup website building

* Fix paths for testing

@@
 species_of_interest = c('BA','DM','DO','PP','OT','NA','total')
-read.csv(
+species_names = read.csv(
   "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/Portal_rodent_species.csv",

### 2018-12-12 report.Rmd
Switch to portalcasting (#295)
Add portalcasting to package installs, use portalcasting functions to drive directory setup and modeling

* Remove unused tools

* Reorganize, cleanup function writing

* Cleanup website building

* Fix paths for testing

@@
 # get most recent rodent data
-observed = portalr::abundance(clean=FALSE, shape='flat', level='treatment', plots='longterm', time = 'newmoon') %>%
+observed = portalr::abundance(path = "repo", clean=FALSE, shape='flat', level='treatment', plots='longterm', time = 'newmoon') %>%
             dplyr::filter(newmoonnumber == max(newmoonnumber), treatment=='control')
@@
 # get new moon number associated with this period
-most_recent_date = read.csv('PortalData/Rodents/moon_dates.csv') %>%
+most_recent_date = read.csv(
+  "https://raw.githubusercontent.com/weecology/PortalData/master/Rodents/moon_dates.csv") %>%
                    dplyr::mutate(censusdate = as.Date(censusdate,format='%Y-%m-%d')) %>%
@@
 # get most recent rodent data
-observed = portalr::abundance(clean=FALSE, shape='flat', time = 'newmoon') %>%
+observed = portalr::abundance(path = "repo", clean=FALSE, shape='flat', time = 'newmoon') %>%
             dplyr::filter(newmoonnumber == max(newmoonnumber))

### 2018-12-12 PortalForecasts.R

@@ -1,20 +1,12 @@
-source('tools/data_tools.R')
-source('tools/forecast_tools.R')
-library(yaml)
+library(portalcasting)
 
-model_metadata = yaml.load_file("data/model_metadata.yaml")
-forecast_date = as.Date(model_metadata$forecast_date)
+#Update data and models
+setup_dir(options_all = all_options(main = "portalPredictions", download_existing_predictions = FALSE))
 
-if(!(forecast_date == Sys.Date())){ stop('Data not updated') }
+#Run all models using portalcasting defaults
 
-#########Run all models##################################################  
-cat("Running models", "\n")
-dir.create("tmp")
-sapply( list.files("models", full.names=TRUE), source ) ###Temporary, while only modeling in R
+portalcast(options_all = all_options(main = "portalPredictions"))
+message("models done")
 
-#####Collect all forecast results and save to predictions directory######
-cat("Compiling forecasts", "\n")
-newforecasts=forecastall(forecast_date)
-unlink("tmp/*")
-######Update Website#####################################################
+#Update Website
 rmarkdown::render_site()

### 2018-12-12 PortalHindcasts.R

@@ -1,108 +1,1 @@
-source('tools/forecast_tools.R')
-source('tools/data_tools.R')
-library(yaml)
-
-filename_suffix = 'hindcasts'
-
-#The date this hindcast is run. Always today's date.
-forecast_date = Sys.Date()
-
-
-#Hindcast will set the time based on these NewMoonNumbers. For each one
-#a hindcast will be made which pretends that sampling period had just happened. 
-#403 to 490 is Jan,2010 - Jan,2017. 
-initial_time_newmoons=490:403
-
-#Get the most recent data loaded into the data folder
-portalr::download_observations()
-moons = get_moon_data()
-rodent_data = get_rodent_data(moons, forecast_date)
-weather_data = get_weather_data(moons, rodent_data$all, lag=6)
-
-write.csv(rodent_data$all,"data/rodent_all.csv",row.names = FALSE)
-write.csv(rodent_data$controls,"data/rodent_controls.csv",row.names = FALSE)
-write.csv(weather_data,"data/weather_data.csv",row.names = FALSE)
-
-trappings = read.csv(FullPath('PortalData/Rodents/Portal_rodent_trapping.csv', '~'))
-incomplete_samples = portalr::find_incomplete_censuses(trappings)
-
-failed_newmoons=c()
-
-for(this_newmoon in initial_time_newmoons){
-
-  moons = get_moon_data()
-  
-  this_newmoon_sampling_date = moons %>%
-    filter(newmoonnumber == this_newmoon) %>%
-    pull(censusdate)
-  
-  this_newmoon_period = moons %>%
-    filter(newmoonnumber == this_newmoon) %>%
-    pull(period)
-  
-  #Don't do hindcasting from newmoons that had incomplete samplings
-  #or were not sampled at all
-  if(this_newmoon_period %in% incomplete_samples$period | is.na(this_newmoon_sampling_date)){
-    next
-  }
-  
-  ####Setup data for hindcasting####
-
-  moons = get_moon_data() %>%
-    filter(newmoonnumber<=this_newmoon)
-  #get dates of 12 new moons following newmoon of interest
-  future_moons = get_future_moons(moons)
-  
-  #Beginning and end of the forecast timeperiod
-  most_recent_newmoon = moons$newmoonnumber[which.max(moons$period)]
-  most_recent_newmoon_date = as.Date(moons$newmoondate[which.max(moons$period)])
-  first_forecast_newmoon=most_recent_newmoon+1
[…]

### 2018-12-12 archive.sh

@@ -13,7 +13,7 @@ if [ "$TRAVIS_EVENT_TYPE" == "cron" ]; then
 
     # Commit changes to portalPredictions repo
     git checkout master
-    git add predictions/* docs/* data/*
+    git add predictions/* docs/* data/* 
     git commit -m "Update forecasts: Travis Build $TRAVIS_BUILD_NUMBER [ci skip]"
 
     git remote add deploy https://${GITHUB_TOKEN}@github.com/weecology/portalPredictions.git > /dev/null 2>&1

### 2018-12-12 models/esss.R

@@ -1,113 +1,0 @@
-# ESSS
-#
-# Model "ESSS" is a flexible exponential smoothing state space model
-#  fit using the forecast and ets functions with the possibility of 
-#  multiplicative trends. Unfortunately since the seasonality and 
-#  sampling occurring with different frequencies, which the ets function 
-#  cannot accommodate, seasonal models are not included.
-
-  library(forecast)
-  source('tools/model_tools.R')
-
-#' Function for ESSS
-#'
-#' @param abundances table of rodent abundances and time measures
-#' @param forecast_date the dates to be forecast from
-#' @param forecast_months the months of the dates to be forecast
-#' @param forecast_years the years of the dates to be forecast
-#' @param forecast_newmoons the numbers of the new moons to be forecast
-#' @param level name of the type of plots included ("All" or "Controls")
-#' @param num_forecast_newmoons number of new moons to forecast
-#' @param CI_level confidence interval level used for forecast envelope
-#' @return list of forecast and aic tables
-#'
-  forecast_esss <- function(abundances, forecast_date, forecast_months, 
-                            forecast_years, forecast_newmoons, level,
-                            num_forecast_newmoons, CI_level = 0.9){
-
-    cat("Fitting ESSS model for total \n")
-    interpolated_abundances <- interpolate_abundance(abundances)
-
-    ets_model <- ets(interpolated_abundances$total)
-    ets_forecast <- forecast(ets_model, h = num_forecast_newmoons,
-                             level = CI_level, 
-                             allow.multiplicative.trend = TRUE)
-
-    fit_start_newmoon <- min(abundances$newmoonnumber)
-    fit_end_newmoon <- max(abundances$newmoonnumber)
-    initial_newmoon <- max(abundances$newmoonnumber)
-    CI_match <- which(ets_forecast$level == CI_level * 100)
-    estimate <- as.numeric(ets_forecast$mean)
-    LowerPI <- as.numeric(ets_forecast$lower[ , CI_match]) 
-    UpperPI <- as.numeric(ets_forecast$upper[ , CI_match])
-
-    output_fcast <- data.frame(date = forecast_date, 
-                               forecastmonth = forecast_months,
-                               forecastyear = forecast_years, 
-                               newmoonnumber = forecast_newmoons,
-                               currency = "abundance",
-                               model = "ESSS", 
-                               level = level, 
-                               species = "total", 
-                               estimate = estimate,
-                               LowerPI = LowerPI, 
[…]

### 2018-12-12 tests/testthat/test-new_forecasts.R

@@ -1,13 +1,12 @@
 library(testthat)
-library(yaml)
-library(zoo)
+
 context("checks that new forecasts are being added correctly")
-model_metadata = yaml.load_file("../../data/model_metadata.yaml")
+model_metadata = yaml::yaml.load_file("../../data/metadata.yaml")
 
 forecasts=read.csv(file.path('../../predictions', paste(as.character(as.Date(model_metadata$forecast_date)), 
-                                        model_metadata$filename_suffix, ".csv", sep="")), na.strings = "")
+                                          model_metadata$filename_suffix, ".csv", sep="")), na.strings = "")
 forecastaics=read.csv(file.path('../../predictions', paste(as.character(as.Date(model_metadata$forecast_date)), 
-                                        model_metadata$filename_suffix, "_model_aic.csv", sep="")), na.strings = "")
+                                  model_metadata$filename_suffix, "_model_aic.csv", sep="")), na.strings = "")
 
 forecastnames = c("date", "forecastmonth",  "forecastyear",  "newmoonnumber",
                 "currency", "model", "level", "species", "estimate",  "LowerPI", 

### 2018-12-12 tools/data_tools.R

@@ -1,82 +1,0 @@
-library(lubridate)
-library(dplyr)
-library(magrittr)
-library(htmltab)
-
-#' Return normalized path for all operating systems
-#' @param ReferencePath a path to join with current working directory
-#' @param BasePath Current working directory else path given
-#'
-#' @return
-#' @export
-#' @examples
-#' FullPath('PortalData/Rodents/Portal_rodent.csv')
-#' FullPath('PortalData/Rodents/Portal_rodent.csv', '~')
-FullPath <- function( ReferencePath, BasePath=getwd()){
-  BasePath = normalizePath(BasePath)
-  Path = normalizePath(file.path(BasePath, ReferencePath), mustWork = FALSE)
-  return (Path)
-}
-
-#' Get data for newmoonnumbers and related trapping period codes
-#' @return newmoons table
-#' @examples
-#' get_moon_data()
-#' 
-get_moon_data <- function(){
-  #Get the newmoon number of the  most recent sample
-  moons <- read.csv(FullPath('PortalData/Rodents/moon_dates.csv', '~'), header=T)
-  
-  #Add in year and month to join with the rest of the data
-  moons$year=year(moons$newmoondate)
-  moons$month=month(moons$newmoondate)
-  return(moons)
-}
-
-####################################################################################
-#' Get rodent data, tailored for forecasting (all plots and controls only)
-#' @param moons current newmoonnumber table
-#' @param forecast_date date the forecast is run
-#' 
-#' @return a list of two dataframes, all plots and control plots
-#' @examples
-#' get_rodent_data(moons, forecast_date)
-#'
-
-get_rodent_data <- function(moons, forecast_date){
-  #Period 203/newmoonnumber 217 will be when the training data timeseries
-  #begins. Corresponding to Jan 1995
-  historic_start_period=203
-  historic_start_newmoon=217
-  
-  #get Portal abundance data for the entire site and for control plots only.
-  
-  #Control plots
-  controls = portalr::abundance(clean=FALSE,level="Treatment",type="Rodents",length="Longterm",min_plots=24)
-  #Drop PI
-  controls = controls[ , -which(colnames(controls) == "PI")]
-  #The total rodent count in each treatment
-  controls$total = rowSums(controls[,-(1:2)])
-  #Drop non-control treatments and add in newmoonnumber
-  controls = controls %>%
-    filter(treatment == 'control') %>%
[…]

### 2018-12-12 tools/forecast_tools.R

@@ -1,158 +1,93 @@
-library(tidyverse)
-library(lubridate)
-library(zoo)
-library(ggplot2)
-library(rmarkdown)
-library(RCurl)
-
-##########################Forecast processing############################
-#' Combine all new forecasts (from the tmp directory), add ensembles
-#' 
-#' @param forecast_date
-#' @param filename_suffix
-#' @return list(forecasts,all_model_aic)
-#' @example forecastall('forecasts')
-
-forecastall <- function(forecast_date,filename_suffix = 'forecasts') {
-  
-  #Append results to forecasts and AIC tables
-  forecasts = do.call(rbind,
-                      lapply(list.files("tmp",pattern = paste(filename_suffix, ".csv",sep=""), full.names = TRUE), 
-                             read.csv, na.strings = "", colClasses = c("Date", "integer", "integer", 
-                                                                       "integer", "character", "character", "character", 
-                                                                       "character", "numeric", "numeric", "numeric",
-                                                                       "integer", "integer", "integer")))
-  
-  all_model_aic = do.call(rbind,
-                          lapply(list.files("tmp",pattern = paste(filename_suffix, "_model_aic.csv",sep=""), full.names = TRUE), 
-                                 read.csv, na.strings = ""))
-  
-  forecast_filename = file.path('predictions', paste(as.character(forecast_date), filename_suffix, ".csv", sep=""))
-  model_aic_filename = file.path('predictions', paste(as.character(forecast_date), filename_suffix, "_model_aic.csv", sep=""))
-  append_csv(forecasts, forecast_filename)
-  append_csv(all_model_aic, model_aic_filename)
-  
-  ########Add ensembles to files############################################
-  ensemble=make_ensemble(forecasts) %>% 
-    subset(select=colnames(forecasts))
-  append_csv(ensemble, forecast_filename)
-  
-  return(list(forecasts,all_model_aic))
-}
-
-######Tools for writing forecasts to file and aics to separate file###############
-#Appending a csv without re-writing the header.
-append_csv=function(df, filename){
-  write.table(df, filename, sep = ',', row.names = FALSE, col.names = !file.exists(filename), append = file.exists(filename))
-}
-
-#Get all model aic values and calculate akaike weights
[…]

### 2018-12-12 tools/model_tools.R

@@ -1,225 +1,0 @@
-library(lubridate)
-library(dplyr)
-source('tools/data_tools.R')
-#' Interpolate missing (rodent) abundance data
-#' 
-#' Interpolation of missing data in the rodent abundance data set. Each 
-#'   species is individually linearly interpolated, then the total number of 
-#'   of rodents is calculated from the sum of the individual species.
-#'
-#' @param abundance data table with new moon column
-#' @return data table of interpolation-inclusive counts for each species and 
-#'  total
-#' 
-interpolate_abundance <- function(abundance){
-
-  moons <- (min(abundance$newmoonnumber)):(max(abundance$newmoonnumber))
-  nmoons <- length(moons)
-
-  species <- colnames(abundance)[2:(ncol(abundance) - 4)]
-  nspecies <- length(species)
-
-  abunds <- matrix(NA, nrow = nmoons, ncol = nspecies)
-
-  for(i in 1:nmoons){
-    if(length(which(abundance$newmoonnumber == moons[i])) > 0){
-      temp <- abundance[which(abundance$newmoonnumber == moons[i]),
-                        which(colnames(abundance) %in% species)]
-      abunds[i, ] <- as.numeric(temp)
-    }
-  }
-
-  interpolated_abunds <- abunds
-  colnames(interpolated_abunds) <- species
-
-  for(j in 1:nspecies){
-    interpolated_abunds[ , j] <- round(na.interp(abunds[ , j]))
-  }
-
-  interpolated_total <- apply(interpolated_abunds, 1, sum)
-
-  out <- data.frame(moons, interpolated_abunds, total = interpolated_total)
-
-  return(out)
-}
-
-###################################################################################
-#' Lag the covariate (weather or ndvi) data for use with analyses
-#'
-#' Lag the weather data together based on the new moons
-#'
-#' @param data dataframe of weather or ndvi data to be lagged
-#' @param lag lag between census and weather data, in new moons
-#' @param tail logical, if the data lagged to the tail end should be retained
-#' @return a dataframe of weather data with newmoonnumber now reflecting the
-#'  lag
-#'
-#' @examples
-#' lag_data(weather, lag=6)
-#'
-lag_data <- function(data, lag, tail = FALSE){
-  
[…]

### 2019-02-19 index.Rmd
tidying the website Rmds
this will not pass a full build (including website rendering) until v 0.2.3 of portalcasting is merged in and a new docker image is made

--- 2018-12-03 index.Rmd
@@
 Welcome to Portal Forecasting!
-This is a website run by the [Weecology](http://weecology.org/) group, a group of interdisciplinary ecologists at the [University of Florida](http://www.wec.ufl.edu/).
-On this website, you'll find information about our ongoing efforts to forecast ecological systems from time series.
-Specifically, we are using a times series of rodent abundances from [The Portal project](http://portal.weecology.org/), a long-term experimental monitoring project in desert ecology.
+This is a website run by the [Weecology](http://weecology.org/) group, interdisciplinary ecologists at the [University of Florida](http://www.wec.ufl.edu/).
+On this website, you'll find information about our ongoing efforts to forecast a time series of rodent abundances from [The Portal project](http://portal.weecology.org/), a long-term experimental monitoring project in desert ecology.
 Enjoy!
@@
 
-This is the forecast for next month's sampling of rodents at Portal.
+This is the forecast for next year's sampling of rodents in the control plots at Portal.
 The black line indicates the historic total rodent abundance.
-The blue line indicates the forecasted amount over the next 12 months, along with confidence intervals.
-This forecast is from the Ensemble model.
+The blue line indicates the Ensemble model's forecast for total rodent abundances over the next 12 months, along with 90% confidence intervals.
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-obs_data = portalr::abundance(clean=FALSE, level='treatment', plots="longterm") %>%
-  dplyr::filter(treatment=="control")
-obs_data$total = rowSums(dplyr::select(obs_data, -c(period, treatment)))
-new_moons = read.csv('PortalData/Rodents/moon_dates.csv') %>%
-                   dplyr::mutate(censusdate = as.Date(censusdate),
-                          year = lubridate::year(newmoondate),
-                          month = lubridate::month(newmoondate))
-
-obs_data_newmoon = dplyr::inner_join(obs_data, new_moons, by = c("period" = "period"))
-
-for_data = compile_forecasts() %>%
-  dplyr::filter(level == 'Controls', model == 'Ensemble', date == max(date))
-
-# get dates for future new moons for plotting
-future_moons = new_moons %>% portalr::get_future_moons() %>%    
-                         dplyr::select(c("newmoonnumber","newmoondate"))
-for_data = dplyr::inner_join(for_data, future_moons, by = c("newmoonnumber" = "newmoonnumber"))
-
-forecast_viz(obs_data = obs_data_newmoon,
-             obs_date_col_name = "censusdate",
-             obs_val_col_name = "total",
-             for_data = for_data,
-             for_date_col_name = "newmoondate",
-             for_val_col_name = "estimate",
-             for_model_name = "Ensemble",
-             for_lowerpi_col_name = "LowerPI",
-             for_upperpi_col_name = "UpperPI",
-             start_newmoon = 300,
-             ylabel = 'Total Abundance')
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
+plot_cast()
 ```
@@
 ```{r speciesforecasts, echo=FALSE, message=FALSE, warning=FALSE}
-
-sp_predictions = get_sp_predicts(for_data, 'Controls', 
-            lead_time = length(unique(for_data$newmoonnumber)) - 11)
-sp_predict = plot_species_forecast(sp_predictions,
-            title=paste0(sp_predictions$forecast_date[2], ": Control plots"))
-
-plot(sp_predict)
+plot_species_casts()
 ```
@@
 
-This figures are similar to the primary Total Abundance forecast above, but for the abundance of three common rodent species. 
-```{r highabund_ts_forecasts, echo=FALSE, message=FALSE, warning=FALSE}
-most_abund_sp = sp_predictions %>% 
-  dplyr::filter(species != 'total') %>% 
-  dplyr::arrange(desc(estimate)) %>% 
-  head(3) %>% 
-  dplyr::select(species)
-
-# load in rodent species table to get scientific names to display on plots
-species_table = read.csv('PortalData/Rodents/Portal_rodent_species.csv',
-                          stringsAsFactors = F,na.strings = '')
-most_abund_sp_names = dplyr::filter(species_table,speciescode %in% most_abund_sp$species) %>%      
-                      dplyr::select(speciescode,scientificname)
-
-for (n in seq(dim(most_abund_sp_names)[1])) {
-  species_forecast = forecast_viz(obs_data = obs_data_newmoon,
-               obs_date_col_name = "censusdate",
-               obs_val_col_name = most_abund_sp_names$speciescode[n],
-               for_data = for_data,
-               for_date_col_name = "newmoondate",
-               for_val_col_name = "estimate",
-               for_model_name = "Ensemble",
-               for_lowerpi_col_name = "LowerPI",
-               for_upperpi_col_name = "UpperPI",
-               start_newmoon = 300,
-               ylabel = most_abund_sp_names$scientificname[n])
-  plot(species_forecast)
+These figures are similar to the primary Total Abundance Forecast above, but for the abundance of three common rodent species. 
+```{r highabund_ts_forecasts, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
+most_ab <- select_most_ab_spp()
+for(i in 1:3){
+  plot_cast(species = most_ab[i])
 }

### 2019-02-19 report.Rmd
tidying the website Rmds
this will not pass a full build (including website rendering) until v 0.2.3 of portalcasting is merged in and a new docker image is made

--- 2018-12-03 report.Rmd
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
-# get most recent rodent data
-observed = portalr::abundance(clean=FALSE, shape='flat', level='treatment', plots='longterm', time = 'newmoon') %>%
-            dplyr::filter(newmoonnumber == max(newmoonnumber), treatment=='control')
-
-# get new moon number associated with this period
-most_recent_date = read.csv('PortalData/Rodents/moon_dates.csv') %>%
-                   dplyr::mutate(censusdate = as.Date(censusdate,format='%Y-%m-%d')) %>%
-                   dplyr::filter(newmoonnumber==unique(observed$newmoonnumber)) %>% 
-                   dplyr::pull(censusdate)
-
-# get most recent forecast prior to data collection
-files = data.frame(names = list.files(path = './predictions', pattern = '*forecasts.csv'))
-files$filedates = as.Date(substr(files$names,1,10),format='%Y-%m-%d')
-latest_forecast = dplyr::filter(files,filedates<most_recent_date) %>% tail(1)
-
-data = read.csv(paste('predictions/',latest_forecast$names,sep=''),na.strings = '')
-
-ensemble = dplyr::filter(data, level == 'Controls', model == 'Ensemble', newmoonnumber==unique(observed$newmoonnumber))
-sp_predictions = get_sp_predicts(ensemble, 'Controls', lead_time = 1)
-
-joined_data = dplyr::left_join(sp_predictions, observed, by = c('species','newmoonnumber'))
-joined_data[is.na(joined_data)] = 0
-joined_data[joined_data$species=='total','abundance'] = sum(joined_data$abundance,na.rm=T)
-
-sp_predict = plot_species_forecast(sp_predictions,title=paste0(sp_predictions$forecast_date[2], ": Control plots"))  
-sp_predict = sp_predict + ggplot2::geom_point(data = joined_data, mapping = ggplot2::aes(x = abundance, y = species),
-             color = "blue")
-plot(sp_predict)
+plot_species_casts(with_census = TRUE)
 ```
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-
-# get most recent rodent data
-observed = portalr::abundance(clean=FALSE, shape='flat', time = 'newmoon') %>%
-            dplyr::filter(newmoonnumber == max(newmoonnumber))
-
-# get new moon number associated with this period
-most_recent_date = read.csv('PortalData/Rodents/moon_dates.csv') %>%
-                   dplyr::mutate(censusdate = as.Date(censusdate,format='%Y-%m-%d')) %>%
-                   dplyr::filter(newmoonnumber==unique(observed$newmoonnumber)) %>% 
-                   dplyr::pull(censusdate)
-
-# get most recent forecast prior to data collection
-files = data.frame(names = list.files(path = './predictions', pattern = '*forecasts.csv'))
-files$filedates = as.Date(substr(files$names,1,10),format='%Y-%m-%d')
-latest_forecast = dplyr::filter(files,filedates<most_recent_date) %>% tail(1)
-
-data = read.csv(paste('predictions/',latest_forecast$names,sep=''),na.strings = '')
-
-ensemble = dplyr::filter(data, level == 'All', model == 'Ensemble', newmoonnumber==unique(observed$newmoonnumber))
-sp_predictions = get_sp_predicts(ensemble, 'All', lead_time = 1)
-
-joined_data = dplyr::left_join(sp_predictions, observed, by = c('species','newmoonnumber'))
-joined_data[is.na(joined_data)] = 0
-joined_data[joined_data$species=='total','abundance'] = sum(joined_data$abundance,na.rm=T)
-
-sp_predict = plot_species_forecast(sp_predictions,title=paste0(sp_predictions$forecast_date[2], ": All plots"))  
-sp_predict = sp_predict + ggplot2::geom_point(data = joined_data, mapping = ggplot2::aes(x = abundance, y = species),
-             color = "blue")
-plot(sp_predict)
+plot_species_casts(with_census = TRUE, level = "All")
 ```

### 2019-02-20 index.Rmd
updating with new function names

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
-plot_cast()
+plot_cast_ts()
 ```
@@
 ```{r speciesforecasts, echo=FALSE, message=FALSE, warning=FALSE}
-plot_species_casts()
+plot_cast_point()
 ```
@@
 for(i in 1:3){
-  plot_cast(species = most_ab[i])
+  plot_cast_ts(species = most_ab[i])
 }

### 2019-02-20 report.Rmd
updating with new function names

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-plot_species_casts(with_census = TRUE)
+plot_cast_point(with_census = TRUE)
 ```
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-plot_species_casts(with_census = TRUE, level = "All")
+plot_cast_point(with_census = TRUE, level = "All")
 ```

### 2019-02-20 install-packages.R

@@ -11,4 +11,5 @@ pacman::p_load(devtools, dplyr, forecast, ggplot2, htmltab, lubridate, ltsa,
                magrittr, parallel, RCurl, readr, rmarkdown, testthat, tidyverse,
                tscount, yaml, zoo)
 pacman::p_load_gh('weecology/portalr')
-pacman::p_load_gh('weecology/portalcasting')
+#pacman::p_load_gh('weecology/portalcasting')
+devtools::install_github("weecology/portalcasting", ref = "website")

### 2019-02-27 evaluation.Rmd
testing the processing in of forecasts for evaluation

--- 2018-12-03 evaluation.Rmd
@@
 )
-'%>%' <- magrittr::'%>%'
-source("tools/forecast_tools.R")
-
-model_metadata = yaml::yaml.load_file("data/metadata.yaml")
-pred_interval = model_metadata$pred_interval
-models_of_interest = c('AutoArima','Ensemble','ESSS','nbGARCH','pevGARCH')
-
-species_of_interest = c('BA','DM','DO','PP','OT','NA','total')
-species_names = read.csv('PortalData/Rodents/Portal_rodent_species.csv',
-                         stringsAsFactors=FALSE, na.strings = '') %>%
-  dplyr::select(species = speciescode, full_species_name = scientificname) %>%
-  dplyr::add_row(species='total', full_species_name='Total Rodents')
-
-new_moons = read.csv('PortalData/Rodents/moon_dates.csv') %>%
-            dplyr::mutate(censusdate = as.Date(censusdate))
-
-# Setup rodent observations
-species_abundance = portalr::abundance(clean=FALSE, shape='flat', level='treatment', 
-                                       plots="longterm", time='newmoon') %>%
-  dplyr::filter(treatment=="control") %>%
-  dplyr::rename(actual = abundance, level = treatment) %>%
-  dplyr::mutate(level='Controls',currency='abundance') %>%
-  dplyr::filter(species %in% species_of_interest)
-
-total_abundance = portalr::abundance(clean=FALSE, shape='flat', level='treatment', 
-                                     plots="longterm", time='newmoon') %>%
-  dplyr::filter(treatment=="control") %>%
-  dplyr::group_by(newmoonnumber) %>%
-  dplyr::summarise(actual=sum(abundance)) %>%
-  dplyr::ungroup() %>%
-  dplyr::mutate(level='Controls',currency='abundance',species='total')
-
-observation_data = species_abundance %>%
-  dplyr::bind_rows(total_abundance) %>% 
-  dplyr::left_join(new_moons, by='newmoonnumber')
-
+  species <- rodent_spp(set = "evalplot")
+  fcasts <- read_casts() %>%
+            select_casts(species = species, level = "Controls") %>%
+            append_observed_to_cast()
+  fcast_errors <- measure_cast_error(fcasts)
+  hcasts <- read_casts(cast_type = "hindcasts") %>%
+            select_casts(species = species, level = "Controls") %>%
+            append_observed_to_cast()
+  hcast_errors <- measure_cast_error(hcasts)
 ```
 
+## Metrics
+**Error**: Error is the raw difference between a predicted value ($\hat{y}_{i}$) and observed value ($y_{i}$): 
+**RMSE**: Root Mean Square Error (RMSE) is a metric used to evaluate the point estimate of a forecast.  
+**Coverage**: Coverage is used to evaluate the uncertainty in a forecast and is the percentage of observations which fell within the confidence interval of the prediction.
+Because we use a 90% CI, coverage would ideally be equal to 0.90.
+If coverage is higher than 0.90, the forecast's intervals are too wide; if it's lower than 0.90, the forecast intervals are too narrow.
+
 ## How have the models done recently?
-
-This graph shows the forecast error of some recent forecasts for the most recent sampling dates.
-This is shown for all model types in the columns and several common rodent species (and total rodents) in the rows.
-These evaluations are for the Control plots where no manipulation is done.
-```{r recent_eval, echo=FALSE, message=FALSE, warning=FALSE,, fig.width=15, fig.height=15}
-###############################################
-#Get all the recent forecasts
-forecast_data = compile_forecasts(use_hindcasts = FALSE) %>%
-  dplyr::filter(model %in% models_of_interest)
-
-# Round so that large decimal place differences do not cause
-# similar forecasts to be seen as distinct. 
-forecast_data$estimate = round(forecast_data$estimate, 3)
-forecast_data$LowerPI = round(forecast_data$LowerPI, 3)
-forecast_data$UpperPI = round(forecast_data$UpperPI, 3)
-
-# Keep 1 forecast per unique initial_newmoon. When there are multiple forecasts
-# using the same initial_newmoon, use the first one.
-forecast_dates_to_keep = forecast_data %>%
-  dplyr::select(date, initial_newmoon) %>%
-  dplyr::distinct() %>%
-  dplyr::group_by(initial_newmoon) %>%
-  dplyr::filter(date == min(date)) %>%
-  dplyr::ungroup() %>%
-  dplyr::mutate(keep='yes')
-
-forecast_data = forecast_data %>%
-  dplyr::left_join(forecast_dates_to_keep, by=c('date','initial_newmoon')) %>%
-  dplyr::filter(keep=='yes') %>%
-  dplyr::select(-keep)
-
-##################################
-#Combine forecasts and oservations
-forecast_errors = forecast_data %>%
-  dplyr::left_join(observation_data, by=c('newmoonnumber','level','species','currency')) %>% 
-  dplyr::mutate(rmse = sqrt((estimate - actual)^2))
-
-# Sometimes there are model runs on the same day and with all the same info,
-# this gets rid of those
-forecast_errors = forecast_errors %>%
-  dplyr::distinct()
-
-# Drop any entries that don't have an observation
-# (ie. a rainy sample period)
-forecast_errors = forecast_errors %>%
-  dplyr::filter(!is.na(actual))
-
-# Filter to common species and apply the full name
-forecast_errors = forecast_errors %>%
-  dplyr::filter(species %in% species_of_interest) %>%
-  dplyr::left_join(species_names, by='species') %>%
-  dplyr::select(-species) %>%
-  dplyr::rename(species=full_species_name)
-
-# Add a newline to the longer species names
-forecast_errors$species = with(forecast_errors, ifelse(nchar(species)>15, stringr::str_replace(species,' ','\n'), species))
-
-#############################################################
-# Displaying the data from many forecasts in the past can create a mess of a plot. see #218.
-# So here three forecasts are chosen from 1, 6, and 12 newmoons into the past. 
-# Sometimes forecasts data is missing for a particular initial_newmoon, the recursive function 
-# update_initial_newmoon attempts to find another one from the same time pluts/minus 2 month period.
-#  This ensures 3 forecasts are always shown.
-lead_times_to_show = c(1,6,12)
-most_recent_observed_newmoon = max(new_moons$newmoonnumber)
-initial_newmoons_to_keep = most_recent_observed_newmoon - lead_times_to_show
-
-initial_newmoons_to_keep = sapply(initial_newmoons_to_keep, update_initial_newmoon)
-
-forecast_errors = forecast_errors %>%
-  dplyr::filter(initial_newmoon %in% initial_newmoons_to_keep)
-
-#####################################
-ggplot2::ggplot(forecast_errors, ggplot2::aes(x=censusdate, y=rmse, group=as.character(date), color=as.character(date))) +
-  ggplot2::geom_point(size=4)+
-  ggplot2::geom_line(size=1.5) +
-  ggplot2::scale_color_brewer(palette='Dark2') +
-  ggplot2::scale_x_date(breaks = unique(forecast_errors$censusdate)) +
-  ggplot2::facet_grid(species~model, scales = 'free_y') + 
-  ggplot2::theme_bw() +
-  ggplot2::theme(axis.text.x = ggplot2::element_text(angle=45, hjust=0.99, debug = FALSE),
-        axis.text = ggplot2::element_text(size=12),
-        axis.title = ggplot2::element_text(size=20),
-        strip.text = ggplot2::element_text(size=12),
-        legend.text = ggplot2::element_text(size=15),
-        legend.title = ggplot2::element_text(size=18)) +
-  ggplot2::labs(x='Sample Date',y='RMSE', color='Forecast Issue\nDate')
-
-```
 
@@
 Note that this currently uses hindcasts of the prior observations, and is also only for the Control plots.
-
-**RMSE**: Root mean square error, this is a metric used to evaluate the point estimate of a forecast.  
-**Coverage**: This is the percentage of observations which fell within the 90% confidence interval of a forecast.
-Ideally this would be equal to 0.90.
-If it is higher than 0.90 the forecasts intervals are too wide, if it's lower then the forecast intervals are too narrow.
-
-```{r hindcast_eval, echo=FALSE, message=FALSE, warning=FALSE,, fig.width=9, fig.height=15}
-
-#Get the all the forecasts made during observation period
-forecast_data = compile_forecasts(use_hindcasts = TRUE)
-forecast_errors = calculate_forecast_error(observation_data, forecast_data, error_metric = 'RMSE') %>%
-  dplyr::filter(error_value < 200) %>% #Drop RMSE greater than this because it throws off all the graphs
-  dplyr::bind_rows(calculate_forecast_error(observation_data, forecast_data, error_metric = 'coverage'))
-
-forecast_errors = forecast_errors %>%
-  dplyr::left_join(species_names, by='species')
-
-ggplot2::ggplot(forecast_errors, ggplot2::aes(x=lead_time, y=error_value, group=model, color=model)) +
-  ggplot2::geom_point()+
-  ggplot2::geom_line() +
-  ggplot2::geom_hline(yintercept = 0.9) +
-  ggplot2::labs(x='Lead Time (New Moons)') +
-  ggplot2::facet_wrap(full_species_name~error_metric, scales = 'free_y', ncol=2) + 
-  ggplot2::theme_bw() +
-  ggplot2::labs(y = "Error Value", colour = "Model")
-
-```

### 2019-02-27 evaluation.Rmd
Update evaluation.Rmd

@@
 ## Metrics
-**Error**: Error is the raw difference between a predicted value ($\hat{y}_{i}$) and observed value ($y_{i}$): 
-**RMSE**: Root Mean Square Error (RMSE) is a metric used to evaluate the point estimate of a forecast.  
+**Error**: Error is the raw difference between a predicted point value ($\hat{y}_{i}$) and observed value ($y_{i}$): $e_{i} = \hat{y}_{i} - y_{i}$.
+Although we generally summarize errors in a fashion that results in non-negative values only (*e.g.*, via squaring), the raw error provides information about prediction bias (if the models tend to over- or under-predict in a systematic fashion).
+
+**RMSE**: Root Mean Square Error (RMSE) is used to evaluate the absolute accuracy of the point estimates of a forecast.
+Across $N$ predicted values in a time series, $RMSE = \sqrt{\frac{\sum_{i=1}^{N}{e_{i}^2}}{N}}$.
+
 **Coverage**: Coverage is used to evaluate the uncertainty in a forecast and is the percentage of observations which fell within the confidence interval of the prediction.

### 2019-02-27 evaluation.Rmd
Update evaluation.Rmd

@@
 ## How have the models done historically?
-
-These graphs show errors as a function of lead time.
-The lead time is the number of months into the future that forecast is made.
-The error values are an average of all forecast errors using observations since 2010.
-Note that this currently uses hindcasts of the prior observations, and is also only for the Control plots.

### 2019-02-27 evaluation.Rmd
Update evaluation.Rmd
adding in the evalution plot

@@
 )
-  species <- rodent_spp(set = "evalplot")
-  fcasts <- read_casts() %>%
-            select_casts(species = species, level = "Controls") %>%
-            append_observed_to_cast()
-  fcast_errors <- measure_cast_error(fcasts)
-  hcasts <- read_casts(cast_type = "hindcasts") %>%
-            select_casts(species = species, level = "Controls") %>%
-            append_observed_to_cast()
-  hcast_errors <- measure_cast_error(hcasts)
 ```
@@
 
-## How have the models done historically?
+```{r recent_eval, echo=FALSE, fig.width=7, fig.height=7}
+plot_err_lead_spp_mods()
+```
+
+## How would have the models done historically?
+
+Here, we evaluate the models' abilities to hindcast the historic data.

### 2019-02-27 index.Rmd
adding in the evalution plot

@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
+```{r, echo=FALSE, fig.width=8, fig.height=6}
 plot_cast_ts()
@@
 
-```{r speciesforecasts, echo=FALSE, message=FALSE, warning=FALSE}
+```{r speciesforecasts, echo=FALSE}
 plot_cast_point()

### 2019-02-27 report.Rmd
adding in the evalution plot

@@
 Here's the most recent actual data (blue) compared to the forecast performed right before data collection (black, with error bars)
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE}
 plot_cast_point(with_census = TRUE)
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE}
 plot_cast_point(with_census = TRUE, level = "All")

### 2019-02-27 evaluation.Rmd
Update evaluation.Rmd

@@
 
-```{r recent_eval, echo=FALSE, fig.width=7, fig.height=7}
+This figure shows the raw error as a function of lead time (how far in advance the prediction was made) for three forecast dates (colors of lines) across each combination of model (column) and species (row) for a selection of historically common species and the total rodents, for the control (unmanipulated) plots only.
+
+```{r recent_eval, echo=FALSE, fig.width=10, fig.height=10}
 plot_err_lead_spp_mods()

### 2019-03-05 evaluation.Rmd
removing old top r chunks in rmds

@@
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-	echo = FALSE,
-	message = FALSE,
-	warning = FALSE
-)
-```
 
@@
 Here, we evaluate the models' abilities to hindcast the historic data.
+This figure shows coverage (left column) and RMSE (right column) for each of a selection of historically common species and total rodents on the control (unmanipulated) plots only.
+The x-axis in each plot corresponds to the different models, including the ensemble.
+Each model's raw metric values are displayed for all included hindcasts (with jittering to show overlapping points) in addition to a summarizing quartile box-whisker (ends of the whiskers show the exteme values, box shows the 25%-75% range, horizontal line shows the median).
+The 90% coverage target is indicated by the horizontal dotted line.
+
+```{r, hindcast_eval, echo=FALSE, fig.width=9, fig.height=15}
+plot_cov_RMSE_mod_spp()
+```

### 2019-03-05 report.Rmd
removing old top r chunks in rmds

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-'%>%' <- magrittr::'%>%'
-source("tools/forecast_tools.R")
-```
-
-Here's the most recent actual data (blue) compared to the forecast performed right before data collection (black, with error bars)
+Here's the most recent forecast for which observations have been made.
+The actual data (blue) are plotted in comparison to the forecast performed right before data collection (white, with error bars) for the control (unmanipulated) plots (top) and all plots together (bottom):
 ```{r, echo=FALSE}

### 2019-03-05 index.Rmd
tidying up the website markdowns

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-'%>%' <- magrittr::'%>%'
-source("tools/forecast_tools.R")
-```
-
 ## Total Abundance Forecast
@@
 
-```{r, echo=FALSE, fig.width=8, fig.height=6}
+```{r, total_forecast, echo=FALSE, fig.width=8, fig.height=6}
 plot_cast_ts()
@@
 
-```{r speciesforecasts, echo=FALSE}
+```{r, species_forecasts, echo=FALSE}
 plot_cast_point()
@@
 These figures are similar to the primary Total Abundance Forecast above, but for the abundance of three common rodent species. 
-```{r highabund_ts_forecasts, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
+```{r, highabund_ts_forecasts, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
 most_ab <- select_most_ab_spp()

### 2019-03-05 report.Rmd
tidying up the website markdowns

@@
 The actual data (blue) are plotted in comparison to the forecast performed right before data collection (white, with error bars) for the control (unmanipulated) plots (top) and all plots together (bottom):
-```{r, echo=FALSE}
+```{r, with_census_controls, echo=FALSE}
 plot_cast_point(with_census = TRUE)
@@
 
-```{r, echo=FALSE}
+```{r, with_census_all, echo=FALSE}
 plot_cast_point(with_census = TRUE, level = "All")

### 2019-03-05 index.Rmd
more website updates

@@
 
+<br>
+
 Welcome to Portal Forecasting!
-This is a website run by the [Weecology](http://weecology.org/) group, interdisciplinary ecologists at the [University of Florida](http://www.wec.ufl.edu/).
+This is a website run by the [Weecology](http://weecology.org/) team, comprised of [Ethan White's](http://whitelab.weecology.org/) and [Morgan Ernest's](http://ernestlab.weecology.org/) lab groups at the [University of Florida](http://www.wec.ufl.edu/).
+We are a group of interdisciplinary ecologists broadly interested in collaborative approaches to empirical and computation ecology, open science, and open data.
+
 On this website, you'll find information about our ongoing efforts to forecast a time series of rodent abundances from [The Portal project](http://portal.weecology.org/), a long-term experimental monitoring project in desert ecology.
 Enjoy!
+
+<br>
 
@@
 
+<br>
+
 ## Species-Level Forecasts
@@
 ```
+
+<br>
 

### 2019-03-05 report.Rmd
more website edits
more website updates

@@
 ---
+<br>
 
@@
 The actual data (blue) are plotted in comparison to the forecast performed right before data collection (white, with error bars) for the control (unmanipulated) plots (top) and all plots together (bottom):
+
+<br>
+
 ```{r, with_census_controls, echo=FALSE}
@@
 
+<br>
+
 ```{r, with_census_all, echo=FALSE}

### 2019-03-20 models.Rmd
more website edits
Website updates (#319)
* tidying up the website markdowns

* more website edits

* more website updates

* website edits

* Update report.Rmd

* Update about.md

* Update about.md

@@
 
-We currently analyze and forecast rodent data at Portal using four models:
+We currently analyze and forecast rodent data at Portal using five models:
+
+<br>
 
@@
 state space model [@Hyndman2008] fit to the data at the composite (full site
-and just control plots) spatial level and the composite (community) ecological
-level. The model is selected and fitted using the `ets` and `forecast` 
-functions in the **forecast** package [@Hyndman2017] with the 
-`allow.multiplicative.trend` argument set to `TRUE` and the `forecast_ess`
-function in our **portalPredictions** package [@PortalPredictions]. 
+and just control plots) spatial level and both the composite (community) and 
+the articulated (species) ecological levels. The model is selected and fitted 
+using the `ets` and `forecast` functions in the **forecast** package 
+[@Hyndman2017] with the `allow.multiplicative.trend` argument set to `TRUE` 
+and the `ESSS` function in our 
+[**portalcasting** package](https://github.com/weecology/portalcasting). 
 Models fit using `ets` implement what is known as the "innovations" approach 
@@
 
+<br>
+
 ## AutoARIMA
@@
 data at the composite (full site and just control plots) spatial level
-and the composite (community) ecological level. The model is selected and 
-fitted using the `auto.arima` and `forecast` functions in the **forecast**
-package [@Hyndman2013; @Hyndman2017] and the `forecast_autoarima` function 
-in our **portalPredictions** package [@PortalPredictions].
+and both the composite (community) and the articulated (species) ecological 
+levels. The model is selected and fitted using the `auto.arima` and `forecast`
+ functions in the **forecast** package [@Hyndman2013; @Hyndman2017] and the 
+`AutoArima` function in our 
+[**portalcasting** package](https://github.com/weecology/portalcasting).
 
@@
 
+<br>
+
 ## nbGARCH
@@
 `tsglm` function in the **tscount** package [@Liboschik2017a] and the 
-`forecast_nbgarch` function in our **portalPredictions** package 
-[@PortalPredictions]
+`nbGARCH` function in our
+[**portalcasting** package](https://github.com/weecology/portalcasting).
 
@@
 
+<br>
+
+## nbsGARCH
+
+nbsGARCH (Negative Binomial Seasonal Auto-Regressive Conditional 
+Heteroskedasticity) is a generalized autoregressive conditional 
+heteroskedasticity (GARCH) model with overdispersion (*i.e.*, a negative 
+binomial response) with seasonal predictors modeled using two Fourier 
+series terms (sin and cos of the fraction of the year) fit to the 
+data at the composite (full site and just control plots) spatial level and both 
+the composite (community) and the articulated (species) ecological levels. The 
+model for each species and the community total is selected and fitted using the
+`tsglm` function in the **tscount** package [@Liboschik2017a] and the 
+`nbsGARCH` function in our
+[**portalcasting** package](https://github.com/weecology/portalcasting).
+
+GARCH models are generalized ARMA models and are defined according to their 
+link function, response distribution, and two model structure parameters: the 
+number of autoregressive terms (p) and the order of the moving average (q), 
+and are represented as GARCH(p, q) [@Liboschik2017a]. The nbsGARCH model is fit
+using the log link and a negative binomial response (modeled as an
+over-dispersed Poisson), as well as with p = 1 (first-order autoregression)
+and q = 12 (approximately yearly moving average).
+
+The `tsglm` function in the **tscount** package [@Liboschik2017a] uses a 
+(conditional) quasi-likelihood based approach to inference and models the 
+overdispersion as an additional parameter in a two-step approach. This 
+two-stage approach has only been minimally evaluated, although preliminary 
+simulation-based studies are promising [@Liboschik2017b].
+
+<br>
+
 ## pevGARCH
@@
 model for each species is selected and fitted using the `tsglm` function in the
-**tscount** package [@Liboschik2017a] and the `forecast_pevgarch` function in
-our **portalPredictions** package [@PortalPredictions].
+**tscount** package [@Liboschik2017a] and the `pevGARCH` function in
+our [**portalcasting** package](https://github.com/weecology/portalcasting).
 
@@
 
+<br>
+
 ## Ensemble
@@
 
+<br>
+
 # References

### 2019-10-30 models.Rmd
Update models.Rmd

@@
 
-## AutoARIMA
+## AutoArima
 
@@
 parameters can vary from fit to fit.
+
+<br>
+
+## NaiveArima
+
+NaiveArima (Naive Auto-Regressive Integrated Moving Average) is a fixed
+Auto-Regressive Integrated Moving Average (ARIMA) model of order (0,1,0) 
+fit to the data at the composite (full site and just control plots) spatial
+level and both the composite (community) and the articulated (species) 
+ecological levels. The model is selected and fitted using the `Arima` 
+and `forecast` functions in the **forecast** package 
+[@Hyndman2013; @Hyndman2017] and the 
+`NaiveArima` function in our 
+[**portalcasting** package](https://github.com/weecology/portalcasting).
 

### 2019-10-30 models.Rmd
more model docs

@@
 
+## simplexEDM
+
+simplexEDM (simplex projection using Empirical Dynamic Modeling) is a state-space reconstruction model adapted for forecasting and fit to the interpolated data at the composite (full site and just control plots) spatial level and both the composite (community) and the articulated (species) ecological levels.
+The method uses time-delay embedding to reconstruct a state-space for the dynamics underlying a time series [@Packard1980, @Takens1981].
+A forecast from a point in the state space is then computed as a weighted average of the trajectories of nearest neighbors of that point, a minimal algorithm known as "simplex projection" [@Sugihara1990].
+
+In applications to ecological time series, many of the parameters are set automatically, with the exception of the dimension of the time-delay embedding.
+Here, the embedding dimension ($E$) is selected as the value (between `1` and the `max_E` argument to `simplexEDM()`) that minimizes the mean absolute error over the in-sample portion of the data.
+
+<br>
+
+## jags_RW
+
+jags_RW fits a hierarchical log-scale density random walk model with a Poisson observation process using the JAGS (Just Another Gibbs Sampler) infrastructure [@Plummer2003] fit to the  data at the composite (full site and just control plots) spatial level and both the composite (community) and the articulated (species) ecological levels.
+Similar to the NaiveArima model, jags_RW has an ARIMA order of (0,1,0), but in jags_RW, it is the underlying density that takes a random walk on the log scale, whereas in NaiveArima, it is the raw counts that take a random walk on the observation scale.
+The jags_RW model is rather simple, but provides a starting template and underlying machinery for more articulated models using the JAGS infrastructure.
+
+There are two process parameters: mu (the density of the species at the beginning of the time series) and tau (the precision (inverse variance) of the random walk, which is Gaussian on the log scale).
+The observation model has no additional parameters.
+The prior distributions for mu and tau are informed by the available data collected prior to the start of the data used in the time series.
+mu is normally distributed with a mean equal to the average log-scale density and a variance that is twice as large as the observed variance.
+Due to the presence of 0s in the data and the modeling on the log scale, an offset of `count + 0.1` is used prior to taking the log and then is removed after the reconversion (exponentiation) as `density - 0.1` (where `density` is on the same scale as `count`, but can take non-integer values).
+
+<br>
+
 ## Ensemble
 
-In addition to the base models, we include an ensemble, constructed using the
-base model AIC weights. The ensemble mean is calculated as the AIC-weighted
-mean of all model means. The ensemble variance is estimated as the sum of 
-the AIC-weighted mean of all model variances and the variance of the estimated
-AIC-weighted mean, calculated using the unbiased estimate of sample variances.
+In addition to the base models, we include a starting-point ensemble.
+In versions before November 2019, the ensemble was based on AIC weights, but in the shift to separating the interpolated from non-interpolated data in model fitting, we had to transfer to an unweighted average ensemble model.
+The ensemble mean is calculated as the mean of all model means and the ensemble variance is estimated as the sum of the mean of all model variances and the variance of the estimated mean, calculated using the unbiased estimate of sample variances.
 

### 2019-10-31 evaluation.Rmd
tidying up the website markdowns
more website edits
more website updates
Website updates (#319)
* tidying up the website markdowns

* more website edits

* more website updates

* website edits

* Update report.Rmd

* Update about.md

* Update about.md
hide plotting in rmds

@@
 ---
+
+Here we evaluate how each of the models has done compared to observed values.
+
+<br>
 
@@
 
+<br>
+
 ## How have the models done recently?
@@
 
-```{r recent_eval, echo=FALSE, fig.width=10, fig.height=10}
-plot_err_lead_spp_mods()
+```{r, recent_eval, echo=FALSE, fig.width=10, fig.height=10}
+#plot_err_lead_spp_mods()
 ```
+
+<br>
 
@@
 ```{r, hindcast_eval, echo=FALSE, fig.width=9, fig.height=15}
-plot_cov_RMSE_mod_spp()
+#plot_cov_RMSE_mod_spp()
 ```

### 2019-10-31 index.Rmd
website edits

@@
 ---
-
-<br>
 
@@
 The black line indicates the historic total rodent abundance.
-The blue line indicates the Ensemble model's forecast for total rodent abundances over the next 12 months, along with 90% confidence intervals.
+The blue line indicates the Ensemble model's forecast for total rodent abundances over the next 12 months, along with 95% confidence intervals.
 
@@
 ```{r, highabund_ts_forecasts, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
-most_ab <- select_most_ab_spp()
+most_ab <- most_abundant_species()
 for(i in 1:3){

### 2019-10-31 index.Rmd
Update index.Rmd

--- 2019-03-05 index.Rmd
@@
 The black line indicates the historic total rodent abundance.
-The blue line indicates the Ensemble model's forecast for total rodent abundances over the next 12 months, along with 90% confidence intervals.
+The blue line indicates the Ensemble model's forecast (unweighted average of the individual models' forecasts) for total rodent abundances over the next 12 months, along with 95% confidence intervals.
+
+<br>
 
 ```{r, total_forecast, echo=FALSE, fig.width=8, fig.height=6}
-plot_cast_ts()
+plot_cast_ts(data_set = "controls")
 ```
@@
 
-Point forecasts and 90% confidence intervals for specific species at the site for the next sampling period.
+Point forecasts and 95% confidence intervals for specific species at the site for the next sampling period.
+
+<br>
 
 ```{r, species_forecasts, echo=FALSE}
-plot_cast_point()
+plot_cast_point(data_set = "controls")
 ```
@@
 These figures are similar to the primary Total Abundance Forecast above, but for the abundance of three common rodent species.
+
+<br>
+
 ```{r, highabund_ts_forecasts, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
-most_ab <- select_most_ab_spp()
+most_ab <- most_abundant_species(data_set = "controls")
 for(i in 1:3){
-  plot_cast_ts(species = most_ab[i])
+  plot_cast_ts(data_set = "controls", species = most_ab[i])
 }

### 2019-11-14 evaluation.Rmd

@@
 **Coverage**: Coverage is used to evaluate the uncertainty in a forecast and is the percentage of observations which fell within the confidence interval of the prediction.
-Because we use a 90% CI, coverage would ideally be equal to 0.90.
-If coverage is higher than 0.90, the forecast's intervals are too wide; if it's lower than 0.90, the forecast intervals are too narrow.
+Because we use a 95% CI, coverage would ideally be equal to 0.95.
+If coverage is higher than 0.95, the forecast's intervals are too wide; if it's lower than 0.95, the forecast intervals are too narrow.
 
@@
 
-This figure shows the raw error as a function of lead time (how far in advance the prediction was made) for three forecast dates (colors of lines) across each combination of model (column) and species (row) for a selection of historically common species and the total rodents, for the control (unmanipulated) plots only.
+This figure shows the raw error as a function of lead time (how far in advance the prediction was made) across model runs from different forecast origins for multiple models and multiple species (or total) for the control (unmanipulated) plots only.
 
 ```{r, recent_eval, echo=FALSE, fig.width=10, fig.height=10}
-#plot_err_lead_spp_mods()
+#plot_casts_err_lead()
 ```
@@
 Each model's raw metric values are displayed for all included hindcasts (with jittering to show overlapping points) in addition to a summarizing quartile box-whisker (ends of the whiskers show the exteme values, box shows the 25%-75% range, horizontal line shows the median).
-The 90% coverage target is indicated by the horizontal dotted line.
+The 95% coverage target is indicated by the horizontal dotted line.
 
 ```{r, hindcast_eval, echo=FALSE, fig.width=9, fig.height=15}
-#plot_cov_RMSE_mod_spp()
+#plot_casts_cov_RMSE()
 ```

### 2019-11-14 report.Rmd
Update report.Rmd

@@
 ---
-<br>
 
@@
 ```{r, with_census_controls, echo=FALSE}
-plot_cast_point(with_census = TRUE)
+#plot_cast_point(with_census = TRUE)
 ```
@@
 ```{r, with_census_all, echo=FALSE}
-plot_cast_point(with_census = TRUE, level = "All")
+#plot_cast_point(with_census = TRUE, level = "All")
 ```

### 2020-01-16 models.Rmd
adding gpedm portalcasting 0.17.0

@@
 
-We currently analyze and forecast rodent data at Portal using five models:
+We currently analyze and forecast rodent data at Portal using ten models:
 
@@
 
+## GPEDM
+
+GPEDM (Gaussian processes using Empirical Dynamic Modeling) is a state-space reconstruction model adapted for forecasting and fit to the interpolated data at the composite (full site and just control plots) spatial level and both the composite (community) and the articulated (species) ecological levels.
+.
+The method uses time-delay embedding to reconstruct a state-space for the dynamics underlying a time series [@Packard1980, @Takens1981].
+The forecast function is approximate using Gaussian processes.
+
+As with `simplexEDM()`, many of the parameters are fit automatically, such as the length-scale, and variance parameters (see `rEDM::block_gp()` for details).
+One exception is the dimension of the time-delay embedding.
+Here, the embedding dimension ($E$) is selected as the value (between `1` and the `max_E` argument to `GPEDM()`) that minimizes the mean absolute error over the in-sample portion of the data.
+
+<br>
+
 ## jags_RW

### 2023-04-26 app/models.Rmd
setup

---
title: "Models"
bibliography: refs.bibtex
output: 
  rmarkdown::html_document 
---

```{r setup, include = FALSE}

  knitr::opts_chunk$set(collapse = TRUE,
                        comment  = "#>")

  models_controls <- prefab_models_controls( )
  models_names    <- unlist(lapply(lapply(models_controls, getElement, "metadata"), getElement, "print_name"))
  nmodels         <- length(models_names)
  nmodels_text    <- as.character(english(nmodels))

```

<br>

We currently analyze and forecast rodent data at Portal using `r nmodels_text` models: `r models_names` [@PortalPredictions].

```{r echo = FALSE, results = "asis"}

  template <- "## %s \n \n %s \n \n   \n"

  for (i in 1:nmodels) {
    model_i_controls <- models_controls[[i]]
    cat(sprintf(template, model_i_controls$metadata$print_name, model_i_controls$metadata$text))
  }
```

# References

### 2023-04-26 PortalHindcasts.R

@@ -1,1 +1,1 @@
-portalcast(options_all = all_options(cast_type = "hindcasts"))
+# not presently available 

### 2023-04-26 resources/portalPredictions/PortalForecasts.R

@@ -1,0 +1,13 @@
+library(portalcasting)
+
+#Downloading the full production archive is slow
+options(timeout=600)
+
+#Update data and models
+setup_production()
+
+#Run all models using portalcasting defaults 
+portalcast()
+
+#Evaluate all un-evaluated forecasts
+evaluate_casts()

### 2023-04-26 resources/portalPredictions/PortalForecasts_dryrun.R

@@ -1,0 +1,12 @@
+library(portalr)
+sessionInfo()
+library(portalcasting)
+
+#Update data and models
+setup_production()
+
+#Run fastest model for testing in dry run
+portalcast(models = c("ESSS"))
+
+#Evaluate model forecast
+evaluate_cast()

### 2023-04-26 resources/portalPredictions/archive_hipergator.sh

@@ -1,0 +1,66 @@
+# Archive forecasts by pushing weekly forecasts to GitHub and tagging a
+# release so that the GitHub-Zenodo integration archives the forecasts to
+# Zenodo
+
+# Only releases on cron driven events so that only weekly forecasts and not
+# simple changes to the codebase triggers archiving.
+
+current_date=`date -I | head -c 10`
+
+# Copy version of portal_weekly_forecast.py used to run the forecast into the repo so we know what was run
+cp ../portal_weekly_forecast.sh .
+cp ../portal_dryrun_forecast.sh .
+
+# Setup on weecologydeploy user
+git config user.email "weecologydeploy@weecology.org"
+git config user.name "Weecology Deploy Bot"
+
+# Commit changes to portalPredictions repo
+git checkout main
+git add data/* models/* forecasts/* portal_weekly_forecast.sh portal_dryrun_forecast.sh
+git commit -m "Update forecasts: HiperGator Build $current_date [ci skip]"
+
+# Add deploy remote
+# Needed to grant permissions through the deploy token
+git remote remove deploy # Removing the remote ensures that updates to the GitHub Token are added to the remote
+git remote add deploy https://${GITHUB_TOKEN}@github.com/weecology/portalPredictions.git
+
+# Create a new portalPredictions tag for release
+git tag $current_date
+
+# If this is a cron event deploy, otherwise just check if we can
+
+# Push updates to upstream
+git push --quiet deploy main
+
+# Create a new portalPredictions release to trigger Zenodo archiving
+git push --quiet deploy --tags
+curl -v -i -X POST -H "Content-Type:application/json" -H "Authorization: token $GITHUB_RELEASE_TOKEN" https://api.github.com/repos/weecology/portalPredictions/releases -d "{\"tag_name\":\"$current_date\"}"
+
+
+# Clone forecasts archive repo
+cd ../
+git clone https://github.com/weecology/forecasts
+cp portalPredictions/forecasts/*.* forecasts/portal/
+cp portalPredictions/fits/*.* forecasts/portal/
+cd forecasts
+
+# Setup on weecologydeploy user
+git config user.email "weecologydeploy@weecology.org"
+git config user.name "Weecology Deploy Bot"
+
+# Commit to forecasts repo
+git add .
+git commit -m "Update Portal forecasts: Build $current_date"
+git remote remove deploy # Removing the remote ensures that updates to the GitHub Token are added to the remote
[…]

### 2023-04-26 resources/portalPredictions/tests/testthat/test-forecasts_committed.R

@@ -1,0 +1,12 @@
+library(testthat)
+
+context("checks that github repository is updated correctly")
+
+# test that the remote github main branch has the newest forecasts
+new_commit <- git2r::last_commit()
+
+test_that("placeholder", {
+
+  expect_true(grepl("Update forecasts: HiperGator Build",new_commit$summary))
+  expect_true(grepl(Sys.Date(),new_commit$summary))
+})

### 2023-04-26 resources/portalPredictions/tests/testthat/test-metadata.R

@@ -1,0 +1,8 @@
+library(testthat)
+
+context("checks that forecasts exist for use")
+
+test_that("cast metadata", {
+  metadata <- read.csv("forecasts/casts_metadata.csv")
+  expect_true(is.integer(metadata$cast_id))
+})

### 2023-04-26 tests/testthat/test-metadata.R

@@ -3,6 +3,6 @@ library(testthat)
 context("checks that forecasts exist for use")
 
 test_that("cast metadata", {
-  metadata <- read.csv("forecasts/casts_metadata.csv")
-  expect_true(is.integer(metadata$cast_id))
+  metadata <- read.csv("forecasts/forecasts_metadata.csv")
+  expect_true(is.numeric(metadata$forecast_id))
 })

### 2023-04-26 resources/portalPredictions/manuscript/response_to_review.Rmd

--- 2018-07-04 manuscript/response_to_review.Rmd
@@
 Thanks to the expanded efforts at computational training, we believe that many researchers have the necessary skills to implement these sorts of approaches, either on their own or as part of a team.
-And we think that the association between "empiricists" and "little coding experience" is increasingly out of date as demonstrated by the fact that this system was developed in large part by the empiricists collecting the data.
+We also think that the association between "empiricists" and "little coding experience" is increasingly out of date as demonstrated by the fact that this system was developed in large part by the "empiricists" collecting the data.
 

### 2023-04-26 resources/portalPredictions/manuscript/response_to_review_2.Rmd

---
title: "Response to Review Round 2"
output: pdf_document
---

## Associate Editor

**One of the previous reviewers and a third reviewer have evaluated your
resubmission. They both seem very pleased with this version and see this as an
important first step to the standardization and "cyber-implementation" of
iterative near-term forecasting. Both reviewers and authors argue well that,
while many challenges remain to be solved, this is a great first version and
discussion template. As such, it can be a very valuable contribution to the
field. To best fulfill this purpose, I invite the authors to read the last
reviewer's list of issues and challenges that remain to be solved. Enumerating
them in the discussion would constitute a valuable roadmap of future
developments and likely make this article highly influential.**

Thanks! We agree that the last reviewer has made a number of valuable
suggestions for helping lay out a roadmap for the future and we have added
discussion of many of them. See details in the response to this reviewer below.

## Reviewer 2

**I think the authors have addressed what they needed for this to be accepted
for publication.**

Thanks.

**But I think that they misinterpreted my comments about audience. If the point is
to get more people using this workflow (for themselves and for the good of
science), thinking more about where people are coming and adding sentences to
usher them along is key.**

We certainly agree that this is the goal and have done our best to accomplish it
both here and in our paper on using continuous analysis for data management.

## Reviewer 3

**In this paper the authors introduce a prototype automated workflow for producing
iterative ecological forecasts for the Portal project’s long-term rodent
monitoring. At a high level I think this is a really important paper. The
interest in iterative ecological forecasting is growing rapidly and new projects
keep coming online. However, setting up the cyberinfrastructure (CI) surrounding
such forecasts can be nontrivial and, to my knowledge, there has been no
previously published discussion outlining the challenges, potential solutions,
and emerging best practices for how to do this. While this manuscript is not
going to be the last word on how to set up a forecast architecture, it is a very
useful starting point as it highlights and makes explicit many of the issues
faced and identified approaches for dealing with these issues that leverage
modern CI advances. These emerging CI tools make the ecological forecasting
process more robust than hacking a workflow together, while at the same time
being more nimble and easier to use for smaller forecasts than the enormous
mainframe solutions employed by numerical weather forecast centers. As someone
who has thought about these issues a lot, I think the authors have done a great
job of covering a lot of issues and in bringing a robust system into operations.
In my detailed comments below I do have some suggestions for their system, but I
want to make clear that I don’t think these are problems they need to solve (and
code implemented) for this paper to be published. More often they are things
that I think would be good to bring up in the Discussion – I think there’s a bit
more to be said about the limitations current approaches and the future
directions for this area of Methods research.**

We thank the reviewer both for their insightful comments (our responses to which
are detailed below) and for being explicit that these are issues to be discussed
and worked toward in future development rather than things that need to be
implemented immediately.

Detailed comments:

**L51: Might be useful to add a few more examples of ecological forecasts
occurring on short time scales (daily to quarterly)**

**L76: You might consider a slight restructuring of the introduction. I feel
like some of the ideas being ‘introduced’ in this paragraph were already being
referenced earlier in the introduction.**

The goal of this paragraph is distinct from the prior paragraphs in that to
it introduces the specific recommendations made by Dietze et al. There is some
overlap with the general concepts that motivate these recommendations, but since
this paper builds directly on Dietze et al. we consider it important to clearly
link what we are doing directly to their ideas. So, the first two paragraphs
introduce the general space the this paragraph introduces the important
intellectual contribution of Dietze et al. We have tried to make this clear in
the topic sentence.

**L107: On the topic of model averaging, I’d recommend checking out the new
Ecological Monograph by Dormann et al https://doi.org/10.1002/ecm.1309**

We agree that the Dormann et al paper is a valuable contribution in this space
and have added a citation to it to the paper.

**L148: I agree that the Portal project is a great example for prototyping an
ecological forecast. Perhaps say a bit more about how these forecasts are being
used (or how the Portal team plans to use them) to inform their basic science.
In general I think forecasting experiments is a great idea, but I also wonder if
there are any conservation/management implications of this specific forecast.**

We have added a sentence on the applied importance of forecasting rodent
population dynamics, related to their importance in several major zoonotic
diseases, as well as a sentence describing current and future research that we are
conducting using this system.

**L151 and elsewhere:  When forecasting iteratively, this manuscript tends to
talk about “rebuilding” or “refitting” the models. It’s worth noting that one of
the most common approaches to iterative forecasting, data assimilation, does not
refit/rebuild the model at all, but instead just updates the state of the system
(and sometimes the parameters) given the previous forecast and the new data. You
might think about a slight tweak to your word choice to be more inclusive of
these approaches.**

This is an excellent point. We have change the language to "updating" throughout
the manuscript to capture the range of possible approaches.

**L180 and elsewhere: I find your use of Travis CI to facilitate Continuous
Analysis to be clever, but I also think it's worth acknowledging to the reader,
either here or in the Discussion, that this is a bit of a hack and not a system
that's designed for Continuous Analysis. Yes, it can get the job done but it
sees like an abuse of the system, it places non-trivial computational
constraints on what you can do, and it fundamentally isn’t designed to return
quantitative outputs. The authors don't really explain how they get their output
files out of Travis CI and into their database, and looking at the code it
appears that this occurs by having Travis CI itself trigger a push into various
Github repositories. By contrast, there are other containerized systems like
Clipper (UC Berkeley) and OpenWisk (IBM) that are designed for these types of
workflows that seem worth mentioning in the Discussion as possible future
directions.**

We disagree with the reviewer that this is a "hack". Travis is used and
discussed in the main continuous analysis paper (Beaulieu-Jones & Greene 2017).
It is also used in the software development community for non-trivial
computational tasks and for returning and deploying outputs. In addition, we did
fully explain our method for deploying the outputs (see lines 349-359 of the
previous submission).

That said, as previously noted (lines 422-423 in the previous submission), we
agree that Travis places meaningful computational limitations on the analysis
that can be conducted. We have added a new paragraph on future directions to the
Discussion that addresses this and mentions alternatives and routes forward for
making this kind of Continuous Analysis more scalable.

**L188: This paragraph feels like it rambles a bit and would benefit from being
tightened up and shortened.**

This pararaph has been tightened and shortened substantially in the process.

**L245: You mention how you grab previous, observed weather data but don't
mention what you’re using as the weather forecast data to drive the forecasts.
In the repo portalPredictions/data you appear to provide models with both
weather data and NDVI to use as part of model calibration, but I don’t see the
inclusion of forecasts of these quantities, which makes me wonder if they’re
being used in the forecast at all**

We do use these data in the forecasts. The pevGARCH model uses temperature,
precipitation, and NDVI as predictors. Forecasts for temperature and
precipitation are obtained from the North American Multi Model Ensemble (NMME;
Kirtman et al. 2014) downscaled to the Portal site. Forecasts for NDVI are not
currently available and so we use a seasonal auto ARIMA model to forecast future
values. However, we have avoided discussion of the specific models in this paper
since exploration of the infrastructure is already a sufficiently lengthy topic
for discussion.

**L290: The idea of “plugin infrastructure” is important, but a bit vague. Seems
like a good place to point readers to documentation on exactly what you mean
here and how to plug new things in.**

We have added a page to the GitHub repository expanding on the previous
description from the README about how to do this and added a link to that
page from this paragraph.

**Figure 2: With regards to the predefined structure of the forecasts, for many
forecast methods and scoring metrics the mean and interval are insufficient to
capture the full, probabilistic nature of the forecasts or to score forecasts
against future observations.  For example, with ensemble and Monte Carlo
methods, should you be storing the full ensemble of predictions? In a
multivariate forecast, the current format has no way of recording covariances
between state variables. It is also interesting that none of the proposed
approaches seem to take the previous forecasts as inputs to the next prediction,
and it’s not clear whether the current architecture would support that. Again,
there are not specific requests where I’m asking the authors to change their
system – this is a great first start – but things they and the community should
think about when building upon this prototype or developing alternatives.
Perhaps worth mentioning briefly in the Discussion?**

This is an excellent point and one we've been discussing extensively to try to
figure out how to improve going forward. We have now added a section on future
directions to the Discussion that addresses this and other forward looking
suggestions by the review.

**L315: I think the idea of developing standards for archiving ecological
forecasts is really important. We’re at a point where this community is still
relatively small and we should tackle these issues now, rather than face painful
interoperability issues in the future that we could have avoided. There’s an
extensive literature on data standards, meta-data, and ontologies that could be
briefly mentioned here.**

We have now added a section on future directions to the Discussion that
addresses this.

**L316: Is this the same rule being used in your calibration? (probably if
assuming Gaussian error).  If not, then this score is not “proper”, in the sense
that the best fitting model might not give the highest score. Also what about
more probabilistic scores? RMSE doesn’t penalize a model for being
overconfident.**

We have expanded the second sentence of this paragraph to address this point by
highlighting deviance as one of the next scores to be added to our assessment.

**L335: I really like the idea of thinking of ecological forecasts as a form of
pre-registration**

Thanks!

**L373: I think all of this is great. In looking at the webpage one suggestion
I'd make would be to add actual model-data comparisons under the evaluation, not
just summary statistics (RMSE, coverage) -- the patterns to how forecasts fail
are very informative. That said, if I were to add one more metric it would be
bias. Almost all calibration approaches force this to be zero, but the forecasts
could easily over- or under-predict consistently.**

Thanks for the suggestions. We do feature direct model-data comparisons for each
previous forecast on the [Rodent
Report](https://portal.naturecast.org/report.html) page, but it's a good idea to
show these over time. We've added it to the [issue
queue](https://github.com/weecology/portalPredictions/issues/287).

**L424: Add discussion of alternatives?**

Yes, this is definitely an important oversight. We did initially have some
discussion of this and aren't actually sure why we cut it. We have added a new
paragraph on future directions that addresses this point and integrates the
reviewers suggestions above related to possible alternatives.

**L427: If current options are imperfect, could you state very explicitly what
the key features are that you think we need to see in new/updated community
tools?**

We have added a new paragraph on future directions that addresses this point.
